\begin{abstract}
Simulation frameworks play a pivotal role in the safe development of robotic applications. 
However, often different components of an envisioned robotic system are best simulated in different environments/simulators. This poses a significant challenge in simulating the entire project into a single integrated robotic framework. Specifically, for partially-open or closed-source simulators, often two core limitations arise. i) Actors in the scene other than the designated robots cannot be controlled during runtime via interfaces such as ROS and ii) retrieving real-time state information (such as pose, velocity etc.) of objects in the scene is prevented. 
In this work, we address these limitations and describe our solution for the use case of integrating aerial drones simulated by the powerful simulator Sphinx (provided by Parrot Drone) into the Gazebo simulator. We achieve this by means of a mirrored instance of a drone that is included into existing Gazebo-based environments. A promising application of our integrated simulation environment is the  task of target tracking that is common in aerial multi-robot scenarios. Therefore, to demonstrate the effectiveness our our integrated simulation,  we also implement a model predictive controller (MPC) that outperforms the default PID-based controller framework provided with the Parrot's popular Anafi drone in various dynamic tracking scenarios thus enhancing the utility of the overall system.  
We test our solution by including the Anafi drone in an existing Gazebo-based simulation and evaluate the performance of the MPC through rigorous testing in simulated and real-world tracking experiments against a customized PID controller baseline. Source code is published on \url{https://github.com/robot-perception-group/anafi_sim}.


\keywords{Simulation  \and Aerial robotics \and Real world experiments.}

\end{abstract}
