\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2024


% ready for submission
\usepackage[nonatbib,final]{neurips_2024}


% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2024}


% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2024}


% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2024}


\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{graphicx}  % 加载 graphicx 包
% \usepackage{natbib}
\usepackage{array}   % 表格格式支持
\usepackage{tabularx} % 自动调整列宽
\usepackage{multirow} % 合并行
\usepackage{xcolor}   % 颜色支持
\usepackage{subcaption}    % 用于子图环境
\usepackage{calc}          % 用于计算尺寸（可选）
\renewcommand{\thefootnote}{}  % This disables footnote numbers or symbols




\title{AgentSociety: Large-Scale Simulation of\\LLM-Driven Generative Agents Advances\\Understanding of Human Behaviors and Society}



% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.


\author{%
 Jinghua Piao$^{1\dagger}$\\
  \And
  Yuwei Yan$^{1\dagger}$\\
  \And
  Jun Zhang$^{1\dagger}$\\
  \And
  Nian Li$^{1}$\\
  \And
  Junbo Yan$^{1}$\\
  \And
  Xiaochong Lan$^{1}$\\
  \And
  Zhihong Lu$^{1}$\\
  \And
  Zhiheng Zheng$^{1}$\\
 \And
  Jing Yi Wang$^{1}$\\
  \And
  Di Zhou$^{2}$\\
  \And
  Chen Gao$^{3}$\\
  \And
  Fengli Xu$^{1}$\\
  \And
  Fang Zhang$^{4*}$\\
  \And
  Ke Rong$^{2*}$\\
  \And
  Jun Su$^{4*}$\\
  \And
  Yong Li$^{1*}$\\
}


\begin{document}


\maketitle




\begin{abstract}
Understanding human behavior and society is a central focus in social sciences, with the rise of generative social science marking a significant paradigmatic shift. By leveraging bottom-up simulations, it replaces costly and logistically challenging traditional experiments with scalable, replicable, and systematic computational approaches for studying complex social dynamics. Recent advances in large language models (LLMs) have further transformed this research paradigm, enabling the creation of human-like generative social agents and realistic simulacra of society. In this paper, we propose AgentSociety, a large-scale social simulator that integrates LLM-driven agents, a realistic societal environment, and a powerful large-scale simulation engine. Based on the proposed simulator, we generate social lives for over 10k agents, simulating their 5 million interactions both among agents and between agents and their environment. Furthermore, we explore the potential of AgentSociety as a testbed for computational social experiments, focusing on four key social issues: polarization, the spread of inflammatory messages, the effects of universal basic income policies, and the impact of external shocks such as hurricanes. These four issues serve as valuable cases for assessing AgentSociety's support for typical research methods -- such as surveys, interviews, and interventions -- as well as for investigating the patterns, causes, and underlying mechanisms of social issues. The alignment between AgentSociety's outcomes and real-world experimental results not only demonstrates its ability to capture human behaviors and their underlying mechanisms, but also underscores its potential as an important platform for social scientists and policymakers. 

\end{abstract}
\footnote{\\
$^1$ Department of Electronic Engineering, Beijing National Research Center for Information Science and Technology (BNRist), Tsinghua University\\
$^2$ Institute of Economics, School of Social Sciences, Tsinghua University\\
$^3$ BNRist, Tsinghua University\\
$^4$ School of Public Policy and Management, Tsinghua University\\
$^\dagger$ These authors contributed equally to this work. \\
$^*$ Corresponding authors. E-mail: liyong07@tsinghua.edu.cn.
}
\clearpage
% Understanding human behavior and society has long been a central focus in social sciences, with increasing attention on generative social science, which leverages simulations to explore complex social dynamics. Recent advances in large language models (LLMs) have transformed this approach, enabling the creation of human-like social agents that simulate societies from a bottom-up perspective. 



% Recent advances in large language models (LLMs) have not only achieved remarkable across various tasks, but also reshaped the research paradigm for understanding human behaviors and society. Growing studies have highlighted the potential of leveraging LLMs to create human-like social agents, enabling the simulation of complex society in a bottom-up manner. Here we propose AgentSociety, a large-scale social simulator that integrates LLM-driven agents, a realistic environment, and a powerful large-scale simulation engine. Based on the proposed simulator, we generate social lives for over 10,000 agents, simulating [xxx@Zhangjun] interactions, both among agents and between agents and their environment. Furthermore, we investigate the patterns, causes, mechanisms, and mitigation strategies of various social phenomena, including polarization, the spread of inflammatory messages, universal basic income, and external shocks of hurricanes. The alignment between the simulated outcomes and real-world experimental results not only demonstrates its ability to replicate human behaviors and societal dynamics but also underscores its potential as a valuable tool for social scientists and policymakers.



\input{1_Introduction}
\input{1_Overview}
\input{3_Agents}
\input{4_Environment}
\input{5_Archtecture}
\input{6_Experiments}
\input{2_Related}
\input{7_Conclusion}

\bibliographystyle{plain}
\bibliography{sn-bibliography}% common bib file
%% if required, the content of .bbl file can be included here once bbl is generated
%%\input sn-article.bbl


\end{document}
