\section{Conclusion}
\label{sec:conclusion}

In this work, we introduced \ourattackfull (\ourattack), a membership inference attack targeting Retrieval-Augmented Generation (RAG) systems. Unlike prior methods, \ourattack leverages natural, topic-specific queries that remain undetectable by existing defense mechanisms while maintaining high effectiveness. Through extensive experiments across diverse datasets and RAG configurations, we demonstrated the robustness of our attack, achieving superior inference performance with minimal cost and low detection rates. Our analysis highlights the vulnerabilities inherent in RAG systems, emphasizing the need for more sophisticated defenses that balance security and utility
Additionally, our exploration of failure cases provides valuable insights into the limitations of both RAG systems and membership inference attacks, paving the way for future research on privacy-preserving retrieval systems.


\shortsection{Future Directions}
In this work, we proposed a new black-box MIA against RAG systems, focusing on both attack success and detectability. While our attack consistently demonstrates high AUC scores across all settings and high TPR@low FPR in most cases, there are instances where its TPR is lower than one of the baselines. This indicates room for further improvement. Exploring more complex query forms, such as multiple-choice questions, could enhance the performance and serves as a promising direction for future work.
Additionally, we evaluated our attack in a realistic setting where the RAG system rewrites the input query. Other variations of RAG systems, which involve different forms of input modification, remain unexplored. Extending evaluations to such settings would provide a broader understanding of the attack's effectiveness and robustness.
