\section{Related Work}
\paragraph{Cultural Evaluation in Language and Vision Models.}

Research in language-based models is advancing rapidly in capturing cultural nuances through large multilingual evaluation benchmarks____. In the language-vision domain, recent benchmarks like CVQA____ and GlobalRG____ focus on culturally aware question answering, retrieval, and visual grounding. Novel methods leveraging multi-agent frameworks of large multimodal models____ have shown further promise in enhancing cross-cultural understanding. For instance, MosAIC____ employs a multi-agent framework for cross-cultural understanding but focuses on image captioning in single-culture contexts rather than text-to-image generation. Our work addresses this gap by examining how state-of-the-art text-to-image models handle multicultural representations within the same image.
\vspace{-0.8em}
\paragraph{Text-to-Image Generation Models and Benchmarks.} Text-to-image generative capabilities have advanced rapidly in recent years, as evidenced by models such as Stable Diffusion-XL____, DALLE-3____, and FLUX____. Evaluation benchmarks like TIFA____, GenEval____, and GenAIBench____ traditionally emphasize technical factors such as realism, text faithfulness, and compositional accuracy. More recent work, i.e., HEIM____, extends these metrics to include socially situated aspects like toxicity, bias, and aesthetics, reflecting growing concern for the social impact of generative models____.
\vspace{-0.8em}
\paragraph{Cultural Gap and Language Limitations in Text-to-Image Generation.} Despite advancements, existing efforts predominantly focus on a narrow set of languages (e.g., English, Chinese, Japanese), leaving large user communities underserved. 
Recent multilingual models, such as Taiyi-Diffusion-XL____, target Chinese text input, while AltDiffusion____ expands language coverage to eighteen languages. However, a broader ``cultural gap'' persists____, as most models and benchmarks insufficiently capture diverse cultural settings and interactions.
\vspace{-0.8em}
\paragraph{Data Diversity and Cultural Competence.} Only recently have researchers begun to evaluate cultural competence in text-to-image models. For instance, CUBE____ assesses cultural awareness and diversity, yet still focuses on single-culture depictions per image. To our knowledge, no existing work systematically addresses multicultural scenarios—where multiple cultures may be represented in a single image—and rigorously evaluates the performance of state-of-the-art text-to-image systems under such conditions. Our approach aims to fill this gap by exploring how these models handle more complex, multicultural representations.