\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

\pdfminorversion=4
\IEEEoverridecommandlockouts 

\overrideIEEEmargins

\hyphenation{op-tical net-works semi-conduc-tor}

% \pagestyle{plain}

\usepackage{graphicx}
\usepackage{glossaries}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
\usepackage{tabularx}
\usepackage{algorithm2e}
\usepackage{textcomp}
\usepackage{tikz}   % tikz
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{subcaption}
\captionsetup[subfigure]{labelformat=parens, labelsep=none}
\captionsetup{belowskip=-10pt}

% \usepackage{algorithm}
\usepackage{algpseudocode}
\RestyleAlgo{ruled}
\newcommand*{\defeq}{\stackrel{\text{def}}{=}}
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\mb}[1]{\mathbf{#1}}

%User Math Command
\newcommand{\rank}[1]{\text{rank}\left\{#1\right\}}
\newcommand{\img}[1]{\text{img}\left\{#1\right\}}
\newcommand{\nul}[1]{\text{null}\left\{#1\right\}}
\newcommand{\col}[1]{\text{col}\left\{#1\right\}}
\newcommand{\card}[1]{\text{card}\left\{#1\right\}}
\newcommand{\de}[1]{\text{d}#1}
\newcommand{\ov}[1]{\overline{#1}}
\newcommand{\un}[1]{\underline{#1}}
\newcommand{\lie}{\mathcal{L}}
\newcommand{\ad}{\text{ad}}
\newcommand{\Ad}{\text{Ad}}
\newcommand{\tr}{\text{tr}}
\newcommand{\sk}{\text{skew}}
\newcommand{\spa}{\text{span}}
\newcommand{\diag}{\text{diag}}
\newcommand{\vect}{\text{vec}}
\newcommand{\e}[0]{\text{e}}    % exp
\newcommand{\ieref}[1]{(\ref{#1})}

\newcommand {\bs}[1]{\boldsymbol{#1}}
\newcommand{\uvec}[1]{\hat{\mathbf{#1}}}
\newcommand{\uvecf}[3]{\,^{#1}\hat{\mathbf{#2}}_{#3}}
\newcommand{\T}{^{\mathrm{T}}}  %shortcut for transpose
\newcommand{\Rot}[2]{{}^{#1}\mb{R}_{#2}}

% Acronyms
\newacronym{mra}{MRA}{Magnetic Resonance Angiography}
\newacronym{mri}{MRI}{Magnetic Resonance Imaging}
\newacronym{cta}{CTA}{Computed Tomography Angiography}
\newacronym{ct}{CT}{Computed Tomography Angiography}
\newacronym{rrt}{RRT}{Rapidly exploring Random Tree}
\newacronym{lcca}{LCCA}{Left Common Carotid Artery}
\newacronym{lvo}{LVO}{Large Vessel Occlusion}
\newacronym{lsa}{LSA}{Left Subclavian Artery}




\newcommand*\numcircled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=0.75pt] (char) {#1};}} 

\newcommand*\circled[1]{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {#1}}}}

\begin{document}

\title{\LARGE \bf Towards Autonomous Navigation of Neuroendovascular Tools for Timely Stroke Treatment via Contact-aware Path Planning}


\author{Aabha Tamhankar, Giovanni~Pittiglio%,~\IEEEmembership{Member,~IEEE}  
\thanks{FuTURE Lab, Department of Robotics Engineering, Worcester Polytechnic Insitute (WPI), Worcester, MA 01605, USA. Email: {\tt\small \{astamhankar, gpittiglio\}@wpi.edu}
\newline This work was supported by the Worcester Polytechnic Institute (WPI), Department of Robotics Engineering.}%
}

\maketitle

\begin{abstract}
In this paper, we propose a model-based contact-aware motion planner for autonomous navigation of neuroendovascular tools in acute ischemic stroke. The planner is designed to find the optimal control strategy for telescopic pre-bent catheterization tools such as guidewire and catheters, currently used for neuroendovascular procedures. A kinematic model for the telescoping tools and their interaction with the surrounding anatomy is derived to predict tools steering. By leveraging geometrical knowledge of the anatomy, obtained from pre-operative segmented 3D images, and the mechanics of the telescoping tools, the planner finds paths to the target enabled by interacting with the surroundings. We propose an actuation platform for insertion and rotation of the telescopic tools and present experimental results for the navigation from the base of the descending aorta to the \gls{lcca}. We demonstrate that, by leveraging the pre-operative plan, we can consistently navigate the \gls{lcca} with 100\% success of over 50 independent trials. We also study the robustness of the planner towards motion of the aorta and errors in the initial positioning of the robotic tools. The proposed plan can successfully reach the \gls{lcca} for rotations of the aorta of up to 10$^\circ$, and displacement of up to 10\,mm, on the coronal plane. 
\glsresetall
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{keywords}
Automation in Health Care, Steerable Catheters/Needles, Motion Planning and Control.
\end{keywords}

\IEEEpeerreviewmaketitle
\section{Introduction}
\label{sec:introduction}
\input{sections/introduction}

\section{Catheterization Tools Static Model}
\label{sec:model}
\input{sections/model}

\section{RRT Planner for Neuroendovascular Tools}
\label{sec:planning}
\input{sections/planner}

\section{Experimental Validation}
\label{sec:experiments}
\input{sections/experiments}


\section{Conclusions}
\label{sec:conclusions}
\input{sections/conclusions}

% \section*{Acknowledgment}
% The authors wish to thank Andrew Daudelin for the support with collecting the fluoroscopic images used in this paper and its supplemental video. 

\bibliographystyle{IEEEtran}
\bibliography{references}											
\end{document}


