\pdfoutput=1
\documentclass[sigconf]{acmart}

\setcopyright{acmlicensed}

\usepackage{placeins}
\usepackage{cleveref}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{groupplots} % Enables grouped plots
\pgfplotsset{compat=1.18} % Adjust version as needed



\usetikzlibrary{positioning}

\title{\model : Mixer-Networks for Irregular Multivariate Time Series Forecasting}
\author{Christian Klötergens}
\affiliation{%
    \institution{University of Hildesheim}
    \department{ISMLL}
    \city{Hildesheim}
    \country{Germany}}
\email{kloetergens@ismll.de}

\author{Tim Dernedde}
\affiliation{%
    \institution{University of Hildesheim}
    \department{ISMLL}
    \city{Hildesheim}
    \country{Germany}}
\email{dernedde@ismll.de}

\author{Lars Schmidt-Thieme}
\affiliation{%
    \institution{University of Hildesheim}
    \department{ISMLL}
    \city{Hildesheim}
    \country{Germany}}
\email{schmidt-thieme@ismll.de}

\input{macros.sty}

\begin{document}

\begin{abstract}
Forecasting Irregular Multivariate Time Series (IMTS) has recently emerged as a distinct research field, necessitating specialized models to address its unique challenges.
While most forecasting literature assumes regularly spaced observations without missing values, many real-world datasets—particularly in healthcare, climate research, and biomechanics—violate these assumptions.
Time Series (TS)-mixer models have achieved remarkable success in regular multivariate time series forecasting. 
However, they remain unexplored for IMTS due to their requirement for complete and evenly spaced observations. 
To bridge this gap, we introduce \model{}, a novel forecasting architecture designed specifically for IMTS.\@
Our approach retains the core principles of TS mixer models while introducing innovative methods to transform IMTS into fixed-size matrix representations, enabling their seamless integration with mixer modules. 
We evaluate \model{} on a benchmark of four real-world datasets from various domains. 
Our results demonstrate that \model{} establishes a new state-of-the-art in forecasting accuracy while also improving computational efficiency.    
   % While deep learning for regular time series modeling is a well-established research field, the modeling of Irregular Multivariate Time Series (IMTS) remains relatively underexplored. 
   % IMTS require specialized models and frequently arise in diverse domains such as healthcare, astronomy, and biomechanics.
   % The literature initially proposed to use neural ordinary differential equation based models, but recent studies showed that attention based models are superior in terms of accuracy and especially 
   % efficiency. Following that development we introduce Temporal Continuous Convolution Network (\model) , an attention-less model which is much simpler than the previous state-of-the-art while being on par and 
   % better in terms of forecasting accuracy. We provide experiments on four real-world datasets, which demonstrates \model 's superiority. 
\end{abstract}


\begin{CCSXML}
    <ccs2012>
       <concept>
           <concept_id>10010147.10010257.10010258.10010259</concept_id>
           <concept_desc>Computing methodologies~Supervised learning</concept_desc>
           <concept_significance>500</concept_significance>
           </concept>
     </ccs2012>
\end{CCSXML}
    
\ccsdesc[500]{Computing methodologies~Supervised learning}

\keywords{Irregular Multivariate Time Series Forecasting}

\maketitle

\input{content/introduction.tex}
\input{content/related_work.tex}
\input{content/problem.tex}
\input{content/method.tex}
\input{content/experiments.tex}
\input{content/ablations.tex}
\input{content/limitations.tex}
\input{content/conclusions.tex}
\FloatBarrier
\newpage
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}
\newpage
\appendix
\input{content/appendix.tex}

\end{document}