\section{Conclusion and Future Work}\label{sec:conclusion}
% \vspace{-0.5em}
Process discovery aims to automatically generate process models representing the underlying information system from event logs. 
Despite the development of various discovery algorithms and quality metrics, no single algorithm dominates in terms of model quality measures~\cite{AugustoCDRMMMS19PDreviewbenchmark}. 
Consequently, users often face the challenge of manually selecting suitable discovery algorithms, a process that is time-consuming and error-prone, even for experts in process mining. 
In response, this paper introduces \texttt{ProReco}, a process discovery recommender system designed to recommend the most appropriate process discovery algorithm based on user preferences and event log characteristics. 
ProReco expands upon previous work by incorporating state-of-the-art algorithms and providing transparent explanations for its recommendations.

As future work, several directions can be investigated. 
First, we plan to explore various parameter settings for the discovery algorithms. 
Due to the vast search space, the challenge is finding an efficient method to determine the best value per technique. 
Additionally, we would like to include additional measures for optimization such as runtime. 
Last but not least, we plan to conduct user studies to understand the effectiveness and usability of \texttt{ProReco} as well as to validate the benefits. 
For example, the understandability of the provided explanation could be evaluated by a dedicated user study. 