\documentclass[10pt,journal,compsoc]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{amsthm,amssymb}
\usepackage{amsmath,amssymb,amsfonts,dsfont,pifont,bm,bbm,mathrsfs,mathtools,nicefrac}
\usepackage{xcolor,algorithm,listings}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{booktabs,multirow,adjustbox,diagbox,threeparttable,multicol,colortbl,tabularray}
\usepackage{stfloats}
\usepackage{url,setspace}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{cite}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\usepackage[pagebackref,breaklinks,colorlinks,linkcolor=blue,citecolor=blue, bookmarks=false]{hyperref}
\usepackage[capitalize]{cleveref}  % Should be loaded after 'hyperref', and works perfectly with 'subfigure'.'
\Crefname{section}{Section}{Sections}
\Crefname{table}{Table}{Tables}
\Crefname{figure}{Figure}{Figures}
\Crefname{equation}{Equation}{Equations}
\hyphenpenalty=1200

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{1.0,1.0,1.0}
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\scriptsize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}
\lstset{style=mystyle}

\newcommand{\balpha}[1]{\bar{\alpha}_{#1}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\newcommand{\revise}[1]{\textcolor{black}{#1}}
\newcommand{\newrevise}[1]{\textcolor{black}{#1}}
\newcommand{\method}{DMT\xspace}
\newcommand{\supp}{\textit{Supplementary Material}\xspace}
\newcommand{\sqq}[1]{\textcolor{black}{#1}}
\newcommand{\yr}[1]{\textcolor{black}{#1}}

\begin{document}

\title{A Diffusion Model Translator for Efficient Image-to-Image Translation}

\author{
Mengfei Xia, Yu Zhou, Ran Yi, Yong-Jin Liu,~\IEEEmembership{Senior Member,~IEEE}, Wenping Wang,~\IEEEmembership{Fellow,~IEEE}
    % <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem  Mengfei Xia and Yu Zhou are with the MOE-Key Laboratory of Pervasive Computing, Department of Computer Science and Technology, Tsinghua University, Beijing 100084, China (email: \{xmf20, yzhou20\}@mails.tsinghua.edu.cn).% <-this % stops a space
\IEEEcompsocthanksitem Ran Yi is with the Department of Computer Science and Engineering, Shanghai Jiao Tong University, Shanghai, China (email: ranyi@sjtu.edu.cn).% <-this % stops a space
\IEEEcompsocthanksitem Yong-Jin Liu is with the MOE-Key Laboratory of Pervasive Computing, Department of Computer Science and Technology, Tsinghua University, Beijing 100084, China (email: liuyongjin@tsinghua.edu.cn).
\IEEEcompsocthanksitem Wenping Wang is with the Department of Computer Science and Computer Engineering at Texas A\&M University, Texas, The United States (email: wenping@tamu.edu).% <-this % stops a space}
\IEEEcompsocthanksitem Corresponding authors: Ran Yi and Yong-Jin Liu.}% <-this % stops a space}
}

% The paper headers
\markboth{IEEE TRANSACTIONS ON PATTERN ANALYSIS AND MACHINE INTELLIGENCE, VOL. 46, NO. 12, DECEMBER 2024}%
{Xia \MakeLowercase{\textit{et al.}}: A Diffusion Model Translator for Efficient Image-to-Image Translation}

% \IEEEpubid{0000--0000/00\$00.00~\copyright~2021 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.


\IEEEtitleabstractindextext{
\begin{abstract}


Applying diffusion models to image-to-image translation (I2I) has recently received increasing attention due to its practical applications.
%
Previous attempts \sqq{inject information} from the source image into each denoising step for an iterative refinement, \sqq{thus resulting in a} time-consuming implementation.
%
We propose an efficient method that equips a diffusion model with a lightweight translator, dubbed a Diffusion Model Translator (\method), to accomplish I2I. 
%
Specifically, we first \sqq{offer} theoretical justification that \sqq{in employing} the pioneer\sqq{ing} DDPM work \sqq{for} the I2\revise{I} task, it is \sqq{both} feasible and sufficient to transfer the distribution from one domain to another only at some intermediate step.
%
We \sqq{further} observe that the translation performance highly depends on the \sqq{chosen timestep for} domain transfer, and therefore propose a practical strategy to automatically select an appropriate timestep for a given task.
%
We evaluate our approach on a range of I2I applications, \sqq{including} image stylization, image colorization, segmentation to image, and sketch to image, to validate its efficacy and general utility.
%
\sqq{The c}omparisons show that our \method surpasses existing methods in both quality and efficiency.
%
Code will be made publicly available.

\end{abstract}

\begin{IEEEkeywords}
Diffusion models, image translation, deep learning, generative models.
\end{IEEEkeywords}}

\maketitle

% \IEEEdisplaynontitleabstractindextext

% \IEEEpeerreviewmaketitle

\input{TPAMI/1.intro.tex}
\input{TPAMI/2.related.tex}
\input{TPAMI/3.method.tex}
\input{TPAMI/4.exp.tex}
\input{TPAMI/5.conclusion.tex}
\input{TPAMI/9.acknowledgement}
\input{TPAMI/6.ref.tex}
\input{TPAMI/7.bio.tex}

\clearpage

\input{TPAMI/8.supp.tex}



\end{document}


