% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\usepackage[preprint]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}

%Including images in your LaTeX document requires adding
%additional package(s)
\usepackage{graphicx}
\usepackage{times}
\usepackage{latexsym}
\usepackage{times}
\usepackage{latexsym}
\usepackage{booktabs}
\usepackage{boxedminipage}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{todonotes} 
\usepackage{graphicx}


\usepackage{hyperref}
\usepackage{inconsolata}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}
\usepackage{tabu}
\usepackage{fontawesome}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xstring}
\usepackage{graphicx}
\usepackage{pbox}
\usepackage{subcaption}
\usepackage{epstopdf}
\usepackage{xstring}
\usepackage{multirow}
\newcommand{\sq}{\faCheckSquare}
\newcommand{\cq}{\faCheck}
\usepackage{soul}
\usepackage{color}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{comment}
\usepackage{todonotes}
% \usepackage{polyglossia}
\usepackage{listings} % Add this for using the listings package
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{verbatim}
\renewcommand{\UrlFont}{\ttfamily\small}

% \usepackage{arabtex}
% \usepackage{utf8}
% \setcode{utf8}

\lstset{
  basicstyle=\small\ttfamily, %\small,
  breaklines=true,
  aboveskip=4mm,
  belowskip=4mm,  
  captionpos=b, % Position the caption at the bottom
  %frame=single, % Adds a frame around the listing
  columns=fullflexible
}

\lstdefinestyle{smalllisting}{
    basicstyle=\small\ttfamily
}

\usepackage{xcolor} % For coloring text
\definecolor{lightblue}{rgb}{.50,.90,0.51}
\definecolor{tri}{rgb}{.25,.88,.82}
\definecolor{lilac}{rgb}{0.85,0.64,0.85}
\definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
% \newcommand\firoj[1]{\hlc[blue]{{\bf Firoj}: #1}}
\newcommand{\firoj}[1]{{\color{blue}\textbf{Firoj}: #1}}
\newcommand{\memex}{\emph{MemeXplain}}
\newcommand{\com}[1]{{\color{red}\textbf{Firoj}: #1}}


% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

% \title{Can Explanations Enhance the Detection of Manipulative and Hateful Content in Multilingual Memes?}
% \title{Detection of Manipulative and Hateful Content in Multilingual Memes}
\title{MemeIntel: Explainable Detection of Propagandistic and Hateful Memes}



\author{Mohamed Bayan Kmainasi$^1$\thanks{~The contribution was made while the author was interning at the Qatar Computing Research Institute.}\textsuperscript{$\dagger$}, Abul Hasnat$^2$$^,$$^3$\thanks{~Equal contribution.}, Md Arid Hasan$^4$,\\
{\bf Ali Ezzat Shahroor$^5$, Firoj Alam$^5$\textsuperscript{$\dagger$}\thanks{~Corresponding author.}}\\
  $^1$Qatar University, Qatar,   
  $^2$Blackbird.AI, USA, 
  $^3$APAVI.AI, France \\
  $^4$University of New Brunswick, Canada \\
  $^5$Qatar Computing Research Institute, Qatar\\  
  {\tt mk2314890@qu.edu.qa, \{mhasnat, ashahrrour\}@gmail.com},\\ {\tt arid.hasan@unb.ca fialam@hbku.edu.qa}\\
\\}


\begin{document}
\maketitle
\begin{abstract}

The proliferation of multimodal content on social media presents significant challenges in understanding and moderating complex, context-dependent issues such as misinformation, hate speech, and propaganda. While efforts have been made to develop resources and propose new methods for automatic detection, limited attention has been given to label detection and the generation of explanation-based rationales for predicted labels. To address this challenge, we introduce \memex{}, an explanation-enhanced dataset for \textbf{propaganda memes in Arabic and hateful memes in English}, making it the \textit{first} large-scale resource for these tasks. To solve these tasks, we propose a \textbf{multi-stage optimization approach} and train \textbf{Vision-Language Models (VLMs)}. Our results demonstrate that this approach significantly improves performance over the base model for both \textbf{label detection} and \textbf{explanation generation}, outperforming the current state-of-the-art with an \textbf{absolute improvement of $\sim3$\% on ArMeme and $\sim7$\% on Hateful Memes}.
For reproducibility and future research, we aim to make the \memex{} dataset and experimental resources publicly available.\footnote{Please contact the corresponding author.
% \url{anonymous.com}
}

\end{abstract}

\input{sections/introduction}
\input{sections/related_work}
\input{sections/dataset}
\input{sections/exp_expl_generation}
\input{sections/experiments}
\input{sections/results}
\input{sections/conclusion}


\section{Limitations}
Due to the complex nature of manual explanation creation, we have relied on GPT-4o for explanation generation. To ensure the reliability of the explanation we have manually evaluated in four criteria such as informativeness, clarity, plausiblity, and faithfulness on a small sample for each set of explanation. The preliminary evaluation scores suggest that we can rely on the gold explanation as the reference. As a part of ongoing work we plan to conduct manual evaluation on a larger set. An important aspect of the ArMeme dataset is that it is highly imbalanced, which affects overall performance. One possible approach to address this issue is to increase the number of memes labeled as propaganda, other, and not-meme. This can be achieved through data augmentation or by collecting additional memes. 


\section*{Ethics and Broader Impact}
We extended existing datasets by adding explanations. To the best of our knowledge, the dataset does not contain any personally identifiable information, making privacy risks nonexistent. Regarding the explanations, we provided clear annotation instructions and cautioned annotators that some memes might be offensive. It is important to note that annotations are inherently subjective, which can introduce biases into the overall evaluation results. We encourage researchers and users of this dataset to remain critical when developing models or conducting further research. Models built using this dataset could be highly valuable for fact-checkers, journalists, and social media platforms.

\section*{Acknowledgments}
The work of F. Alam is supported by the NPRP grant 14C-0916-210015 from the Qatar National Research Fund part of Qatar Research Development and Innovation Council (QRDI). The findings achieved herein are solely the responsibility of the authors.


\bibliography{bibliography/main}

\appendix

% \section{Example Appendix}
% \label{sec:appendix}

\input{sections/appendix}

\end{document}
