\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage[utf8]{inputenc}
\usepackage{braket}
\usepackage{amsthm}
\usepackage{amsmath, amssymb}
\usepackage[pdftex]{graphicx}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{mathtools}
\usepackage{algpseudocode}
\usepackage{latexsym}
\usepackage{optidef}

\usepackage[pagebackref]{hyperref}
\hypersetup{
    linktocpage,
    colorlinks=true,
    citecolor=blue,
    linkcolor=magenta,
    urlcolor=orange,
}

\newcommand{\BQP}{\mathsf{BQP}}
\newcommand{\NP}{\mathsf{NP}}
\newcommand{\MA}{\mathsf{MA}}
\newcommand{\QMA}{\mathsf{QMA}}
\newcommand{\Q}{\mathsf{Q}}
\newcommand{\C}{\mathsf{C}}
\newcommand{\D}{\mathsf{D}}
\newcommand{\R}{\mathsf{R}}
\newcommand{\EQ}{\mathsf{EQ}}
\newcommand{\poly}{\mathrm{poly}}
\newcommand{\GT}{\mathsf{GT}}
\newcommand{\MAX}{\mathsf{MAX}}
\newcommand{\tr}{\mathrm{tr}}
\newcommand{\TV}{\mathrm{TV}}
\newcommand{\HS}{\mathrm{HS}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\I}{\mathrm{I}}
\newcommand{\Ent}{\mathrm{H}}
\newcommand{\Min}[2]{#1 \wedge #2}
\newcommand{\ceil}[1]{\left\lceil #1 \right\rceil}

\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}

\usepackage[capitalise,nameinlink]{cleveref}
\Crefname{lemma}{Lemma}{Lemmas}
\Crefname{fact}{Fact}{Facts}
\Crefname{theorem}{Theorem}{Theorems}
\Crefname{corollary}{Corollary}{Corollaries}
\Crefname{claim}{Claim}{Claims}
\Crefname{example}{Example}{Examples}
\Crefname{problem}{Problem}{Problems}
\Crefname{definition}{Definition}{Definitions}
\Crefname{notation}{Notation}{Notations}
\Crefname{assumption}{Assumption}{Assumptions}
\Crefname{subsection}{Section}{Sections}
\Crefname{section}{Section}{Sections}
% \Crefname{equation}{Equation}{Equations}
\Crefformat{equation}{(#2#1#3)}
%\Crefname{figure}{Figure}{Figures}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{fact}{Fact}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}{Conjecture}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{question}{Question}

\theoremstyle{remark}
\newtheorem{remark}{Remark}

\title{Does there exist a quantum fingerprinting protocol \\ without coherent measurements?}

\author{Atsuya Hasegawa\thanks{Graduate School of Mathematics, Nagoya University. \ Email: \email{atsuya.hasegawa@math.nagoya-u.ac.jp}} \and Srijita Kundu\thanks{Institute for Quantum Computing and Department of Combinatorics and Optimization, University of Waterloo. \ Email: \email{srijita.kundu@uwaterloo.ca}} \and Fran{\c{c}}ois Le Gall\thanks{Graduate School of Mathematics, Nagoya University. \ Email: \email{legall@math.nagoya-u.ac.jp}} \and Harumichi Nishimura\thanks{Graduate School of Informatics, Nagoya University. \ Email: \email{hnishimura@i.nagoya-u.ac.jp}} \and Qisheng Wang\thanks{School of Informatics, University of Edinburgh. \ Email: \email{QishengWang1994@gmail.com}}}

\date{}

\begin{document}

\maketitle

\begin{abstract}
Buhrman, Cleve, Watrous, and de Wolf (PRL 2001) discovered the quantum fingerprinting protocol, which is the quantum SMP protocol with $O(\log n)$ qubits communication for the equality problem. In the protocol, Alice and Bob create some quantum fingerprints of their inputs, and the referee conducts the SWAP tests for the quantum fingerprints. Since $\Omega(\sqrt{n})$ bits communication is required with the classical SMP scheme for the equality problem first shown by Newman and Szegedy (STOC 1996), there exists an exponential quantum advantage in the amount of communication.

In this paper, we consider a setting in which the referee can do only incoherent measurements rather than coherent measurements including the SWAP tests. We first show that, in the case of one-way LOCC measurements, $\Omega(\sqrt{n})$ qubits communication is required. To prove the result, we derive a new method to replace quantum messages by classical messages and consider a reduction to the optimal lower bound in the hybrid SMP model where one message is quantum and the other is classical, which was first shown by Klauck and Podder (MFCS 2014). Our method uses the result of Oszmaniec, Guerini, Wittek, and Ac{\'i}n (PRL 2017), who showed that general POVM measurements can be simulated by randomized projective measurements with small ancilla qubits, and Newman's theorem.

We further investigate the setting of quantum SMP protocols with two-way LOCC measurements, and derive a lower bound against some restricted two-way LOCC measurements. To prove it, we revisit the technique to replace quantum messages by classical deterministic messages introduced by Aaronson (ToC 2005) and generalized by Gavinsky, Regev, and de Wolf (CJTCS 2008), and show that, using the deterministic message, the referee can simulate the two-way LOCC measurements.
\end{abstract}

\clearpage

\tableofcontents

\input{introduction}
\input{preliminary}
\input{one-way_LOCC}
\input{two-way_LOCC}
\input{separation}

\section*{Acknowledgment}
AH thanks Richard Cleve, Uma Girish, Alex B. Grilo, Kohdai Kuroiwa, Alex May, Masayuki Miyamoto, Ryuhei Mori, Ashwin Nayak, Daiki Suruga, Yuki Takeuchi, Eyuri Wakakuwa, Yibin Wang for helpful discussions. AH is also grateful to Richard Cleve and Ken-ichi Kawarabayashi for their generous support. Part of the work was done while AH was visiting University of Waterloo, and AH is grateful for their hospitality.

AH was supported by JSPS KAKENHI grants Nos.~JP22J22563, 24H00071, and  JST ASPIRE Grant No.~JPMJAP2302. SK was supported by the Natural Sciences and Engineering Research Council of Canada (NSERC) Discovery Grants Program, and Fujitsu Labs America. FLG was supported by JSPS KAKENHI grants Nos.~JP20H05966, 20H00579, 24H00071, MEXT Q-LEAP grant No.~JPMXS0120319794 and JST CREST grant No.~JPMJCR24I4. HN was supported by JSPS KAKENHI grants Nos.~JP20H05966, 22H00522, 24H00071, 24K22293, MEXT Q-LEAP grant No.~PMXS0120319794 and JST CREST grant No.~JPMJCR24I4. QW was supported by the Engineering and Physical Sciences Research Council under Grant No.~EP/X026167/1.

\bibliographystyle{alpha}
\bibliography{ref}

\end{document}
