\documentclass[11pt]{article}
\usepackage{lmodern}
\usepackage{dsfont}
\usepackage{environ}
\linespread{1.06}

\input{definitions.sty}

\input{macros.tex}

\begin{document}

\title{The Early Days of the Ethereum Blob Fee Market and Lessons Learnt}


\author{
 	    \textbf{Lioba Heimbach}\footnote{Work performed in part during an internship at a16z crypto.}
        \\
 		\small ETH Zurich \\
 		\small \emailhref{hlioba@ethz.ch}
        \and
 	    \textbf{Jason Milionis}$^*$\footnote{Jason's research was supported in part by NSF awards CNS-2212745, CCF-2212233, DMS-2134059, and CCF-1763970, by an Onassis Foundation Scholarship, and an A.G. Leventis educational grant.}
        \\
        \small Department of Computer Science \\
 		\small Columbia University \\
 		\small \emailhref{jm@cs.columbia.edu}
}
\date{Initial version: October 6, 2024 \\
      Current version: February 18, 2025
}
\maketitle

\thispagestyle{empty}

\begin{abstract}
Ethereum has adopted a rollup-centric roadmap to scale by making rollups (layer 2 scaling solutions) the primary method for handling transactions. The first significant step towards this goal was EIP-4844, which introduced blob transactions that are designed to meet the data availability needs of layer 2 protocols. This work constitutes the first rigorous and comprehensive empirical analysis of transaction- and mempool-level data since the institution of blobs on Ethereum on March 13, 2024. We perform a longitudinal study of the early days of the \emph{blob fee market} analyzing the landscape and the behaviors of its participants. We identify and measure the inefficiencies arising out of suboptimal block packing, showing that at times it has resulted in up to 70\% relative fee loss. We hone in and give further insight into two (congested) peak demand periods for blobs. Finally, we document a market design issue relating to subset bidding due to the inflexibility of the transaction structure on packing data as blobs and suggest possible ways to fix it. The latter market structure issue also applies more generally for any discrete objects included within transactions.
\end{abstract}

\hspace{3cm}
\begin{center}
    ~~\textbf{Keywords:} Ethereum, Layer 2, blobs, EIP-4844, rollup
\end{center}




\input{intro}

\input{analysis}

\input{discussion}

\section*{Acknowledgments}
We thank Edward Felten, Pranav Garimidi, Scott Duke Kominers, and Tim Roughgarden for helpful discussions on the subject matter of our paper.



\newpage   
\printbibliography
\newpage    
\appendix
\input{appendix}


\end{document}
