\section{Related Work}
\label{sec:headings}


\subsection{Data Curation for Web-crawled Datasets}

Recent research underscores the critical role of data curation in enhancing model performance with large-scale image-text datasets. Various studies focus on improving dataset quality through curation methods, such as enhancing the descriptiveness and cross-modal feature alignment of image-text pairs, and reducing redundancy ____.

In a broader context, DataComp is a benchmark designed to evaluate the performance of multimodal models on large-scale, real-world datasets____. Recent advancements in the DataComp benchmark highlight notable progress in data curation techniques. Yokoo et al.____ advanced data filtering using image-text similarity and caption modification, achieving notable progress in the Filtering and Bring Your Own Data (BYOD) Tracks. Yu et al.____ evaluated data selection criteria's impact on model performance, while Chen et al.____ introduced DataJuicer for managing large-scale datasets. Nguyen et al.____ enhanced image captioning for better modality alignment, and Maini et al.____ presented T-MARS for improved visual representation by bypassing text feature learning. Additionally, significant contributions have been made in the areas of synthetic data, contrastive learning, image-text alignment, and few-shot learning____.


Despite the significant advancements in data curation achieved by Radenovic et al.____ and Nguyen et al.____ through enhancing data relevance, existing automated filtering methods may still exclude valuable but less conventional data points or introduce biases by focusing too narrowly on specific aspects, potentially overlooking broader contextual information. 

\subsection{Ensemble Learning}
Ensemble learning, which combines multiple models to improve performance and generalization, has long been a foundational approach in machine learning. Classic methods such as Bagging____ and Boosting ____ initially demonstrated how model aggregation could reduce variance and improve accuracy.

Afterwards, ensemble learning has been increasingly applied to specialized tasks. Zimek, Schubert, \& Kriegel ____  highlighted how ensemble methods enhance outlier detection by aggregating results from multiple models. Beluch et al.____ demonstrated that ensemble-based uncertainty sampling can significantly improve efficiency in active learning. Rasmus et al. ____ demonstrated that ensemble techniques enhance semi-supervised learning by effectively utilizing both labeled and unlabeled data. Song et al.____ used ensemble methods to improve data quality by detecting and filtering noisy or mislabeled data.

These advancements illustrate how ensemble techniques refine data preprocessing and improve model inputs.