\section{Related Work}
\vspace{-0.4em}
\paragraph {Multi-Agent System.}
Contemporary LLM-based multi-agent systems (MAS) can be broadly categorized into two paradigms: \textbf{(1) Fixed agentic networks} with pre-established, manually crafted architectures, from debate~\citep{chateval,debate2-thu}, collaboration~\citep{eot,NAACL2024_Agent-Self-Collaboration} to competetive~\citep{zhao2023competeai}. MacNet~\citep{qian2024scaling} systematically analyzed typical multi-agent collaboration topologies such as chain, tree, graph, \textit{etc.}; \textbf{(2) Dynamic agentic networks} that configure their structure and communication strategies based on real-time feedback and observations. ADAS~\citep{ADAS} and its follow-up works~\citep{zhang_aflow_2024,shang2024agentsquareautomaticllmagent} leverage search methods such as Monte Carlo Tree Search (MCTS) and evolutionary algorithm~\citep{zhang2025evoflow} to discover effective agent strategies. Other works like DyLAN~\citep{arXiv2023_Dynamic-LLM-Agent}, GPTSwarm~\citep{zhuge2024gptswarm} and AgentPrune~\citep{zhang2024cutcrapeconomicalcommunication} dynamically optimize the inter-agent topologies. Nevertheless, contemporary MAS is often LLM-homogeneous, \textit{i.e.}, relying exclusively on the same LLM backbone, failing to collectively organize heterogenous LLM-agents.


\vspace{-0.4em}
\paragraph{Single LLM Routing}
Efficient routing strategies for single LLMs have been extensively explored to balance computational cost and model performance.
%and have evolved significantly from static approaches to dynamic, task-aware systems. 
Early attempts on LLM routing include HybridLLM~\citep{ding2024hybridllmcostefficientqualityaware}, RouteLLM~\citep{ong2024routellmlearningroutellms} and FrugalGPT~\citep{chen2023frugalgptuselargelanguage}, which primarily focus on binary routing and leverage techniques like sequential pipelines or preference-driven routing to enhance decision-making performance. More recent practices, including Rootoo~\cite{dai2024costeffectiveonlinemultillmselection}, C2MAB-V~\cite{mohammadshahi2024routoolearningroutelarge}, GraphRouter~\citep{feng2024graphroutergraphbasedrouterllm} and RouterDC~\cite{chen2024routerdc} have shifted towards multi-choice selection frameworks. However, existing routing methodologies mainly focus on single-agent scenarios, and their unawareness of inter-agent topology constrains their applicability to more complex tasks and limits scalability in larger systems.