\section{Deferred proofs in Section \ref{sec:implications}}\label{app:implications}
\subsection{Proof of Proposition \ref{prop:swap_reg_breg_div}}
\begin{proof}
    For simplicity, we only prove the result for $\psreg_{\sigma} ^ {\ell}$ since the result for $\sreg_{\sigma} ^ {\ell}$ follows by simply replacing $\cP_{t}(p)$ with $\ind{p_{t} = p}$. We have the following chain of equalities: \begin{align*}
        \psreg_{\sigma} ^ {\ell} &= \sum_{p \in \cZ} \sum_{t = 1} ^ {T} \cP_{t}(p) (\ell(p, y_{t}) - \ell(\sigma(p), y_{t})) \\
        &= \sum_{p \in \cZ} \sum_{t = 1} ^ {T} \cP_{t}(p) \bigc{\ell(p) + \ip{\partial \ell(p)}{y_{t} - p} - \ell(\sigma(p)) - \ip{\partial \ell(\sigma(p))}{y_{t} - \sigma(p)}} \\
        &=  \sum_{p \in \cZ} \bigc{\sum_{t = 1} ^ {T} \cP_{t}(p)} \bigc{\ell(p) + \ip{\partial \ell(p)}{\tilde{\rho}_{p} - p} - \ell(\sigma(p)) - \ip{\partial \ell(\sigma(p))}{\tilde{\rho}_{p} - \sigma(p)}} \\
        &= \sum_{p \in \cZ} \bigc{\sum_{t = 1} ^ {T} \cP_{t}(p)} \bigc{\breg_{-\ell}(\tilde{\rho}_{p}, p) - \breg_{-\ell}(\tilde{\rho}_{p}, \sigma(p))},
    \end{align*}
where the second equality follows from Lemma \ref{lem:characterization_proper_loss}, while the final equality follows by adding and subtracting $\ell(\tilde{\rho}_{p})$. Taking supremum over $\sigma: [0, 1] \to [0, 1]$, we obtain \begin{align*}
    \sup_{\sigma: [0, 1] \to [0, 1]} \psreg_{\sigma} ^ {\ell} = \sum_{p \in \cZ} \bigc{\sum_{t = 1} ^ {T} \cP_{t}(p)} \bigc{\breg_{-\ell}(\tilde{\rho}_{p}, p) - \inf_{\sigma: [0, 1] \to [0, 1]} \breg_{-\ell}(\tilde{\rho}_{p}, \sigma(p))}.
\end{align*}
Next, we realize that $\breg_{\phi}(x, y) \ge 0$ since $\phi$ is convex, and the choice of $\sigma(p) = \tilde{\rho}_{p}$ leads to $\breg_{-\ell}(\tilde{\rho}_{p}, \sigma(p)) = 0$. Therefore, \begin{align*}
    \psreg^{\ell} = \sum_{p \in \cZ} \bigc{\sum_{t = 1} ^ {T} \cP_{t}(p)} \breg_{-\ell}(\tilde{\rho}_{p}, p)
\end{align*}
which completes the proof.
\end{proof}


\subsection{Proof of Proposition \ref{prop:breg_div_decomposable}}
\begin{proof}
    For simplicity, we only consider the case when $p \le \hat{p}$, since the other case follows exactly similarly. Applying the result of Lemma \ref{lem:relate_breg_vbreg}, we obtain \begin{align*}
        \breg_{-\ell}(\hat{p}, p) &= \int_{p} ^ {\hat{p}} \abs{\ell''(\mu)} (\hat{p} - \mu) d\mu \le c \cdot \int_{p} ^ {\hat{p}} \bigc{\frac{1}{\mu} + \frac{1}{1 - \mu}} \cdot (\hat{p} - \mu) d\mu,
    \end{align*}
    where the inequality follows from Lemma \ref{lem:hessian_growth}. By direct computation, the integral above evaluates to \begin{align*}
        \hat{p} \cdot \int_{p} ^ {\hat{p}} \frac{d\mu}{\mu} + (1 - \hat{p}) \cdot \int_{p} ^ {\hat{p}} \frac{d\mu}{\mu - 1} = \hat{p} \cdot \log \frac{\hat{p}}{p} + (1 - \hat{p}) \cdot \log \frac{1 - \hat{p}}{1 - p} = \KL(\hat{p}, p).
    \end{align*}
    Therefore, we have $\breg_{-\ell}(\hat{p}, p) \le c \cdot \KL(\hat{p}, p)$, which completes the proof of the first part of the Proposition. The second part follows by combining the result of Proposition \ref{prop:swap_reg_breg_div} with the result obtained above, and taking a supremum over $\ell \in \cL_{2}$. This completes the proof.
\end{proof}

\subsection{Proof of Proposition \ref{prop:bound_breg_div_quadratically}}

\begin{proof}
    For simplicity, we only consider the case when $p \le \hat{p}$, since the other case follows exactly similarly. Applying the result of Lemma \ref{lem:relate_breg_vbreg}, we obtain 
    $$\breg_{-\ell}(\hat{p}, p) \le G \int_{p} ^ {\hat{p}} (\hat{p} - \mu) d\mu = G \bigc{\hat{p}(\hat{p} - p) - \frac{\hat{p} ^ {2} - p ^ {2}}{2}} = \frac{G}{2} (\hat{p} - p) ^ {2}.$$ 
            The case when $\hat{p} \le p$ follows similarly. Applying the result of Proposition \ref{prop:swap_reg_breg_div}, taking a supremum over $\ell \in \cL_{G}$, and bounding $\cal_{2}, \pcal_{2}$ in terms of $\kcal, \pkcal$ completes the proof.
\end{proof}
