\section{Concluding Remarks}\label{sec:conclusion}

This paper presents Micro Blossom, a heterogeneous architecture for exact MWPM decoding in QEC. Micro Blossom employs $O(|V|)$ parallel processing units to accelerate the decoding of common cases while leaving the complex decoding logic and data structures to software.
Using Micro Blossom, we show for the first time that exact MWPM decoding can reach sub-$\mu s$ decoding latency for $d = 13$ and $p = 0.1\%$, an important step towards useful fault-tolerant quantum computation.

We note that the key ideas employed in Micro Blossom can be useful to accelerate QEC decoders targeting more general quantum LDPC codes, which represent decoding graphs with hypergraphs.
Our insight is that these decoders can also exploit locality~\cite{delfosse2022toward,hillmann2024localized,wu2024hypergraph}.
By handling the common and simple cases using parallel PUs and delegating only rare and complex cases to software, these decoders can achieve higher throughput and lower latency.

There are several ways to further improve the Micro Blossom decoder. First, one can improve resource efficiency via context switching (\S\ref{sec:implementation}) and new microarchitectural designs such as coarse-grained PUs~\cite{ziad2024local}.
Second, in addition to offloading isolated \conf handling to hardware as described in \S\ref{sec:pre-matching}, one can further offload simple alternating tree structures to hardware to reduce latency and CPU usage.

More importantly, Micro Blossom as reported here supports a single logical qubit.
To perform useful quantum computing, we must extend it to support multiple logical qubits and logical gates~\cite{beverland2022assessing}.
In particular, one must address several system-level research challenges.
Operating on top of the Micro Blossom decoder, an operating system-like controller must dynamically configure the decoder to handle a dynamic decoding graph with growing rounds of measurements  in a streaming manner.
A promising approach is to translate quantum gate instructions into decoding blocks~\cite{wu2024lego} and support dynamic inter-block fusions, which complements the intra-block fusion described in \S\ref{sec:fusion}.
Operating below the Micro Blossom decoder, the quantum control stack must interface with the decoder efficiently to load and route syndrome data at Terabit/s~\cite{delfosse2020hierarchical}.
This requires an optimized data plane with low-latency communication protocols in order to sustain real-time error correction.
