\section{Related Work}
\paragraph{Reasoning and Generalization}
    Over the years the study of reasoning with learning based approaches has often focused on the domain of games such as chess, shogi and Go____, Poker ____ or board games ____.
    While these mainly focus on correct play or in-distribution performance, the CLRS Algorithmic Reasoning Benchmark introduced by ____ puts emphasis on generalizable reasoning. It consists of a diverse collection of well known algorithms collected from the textbook ``Introduction to Algorithms'' by____ providing a resource to assess algorithmic reasoning for learning based approaches. Moreover, there exists a CLRS-Text ____ to better assess the reasoning capabilities from a language perspective as well. 
    ____ provide a more theoretically supported view on the generalization performance of some classes of neural networks, trained with SGD.
    They specifically focus on Boolean functions, learned in a supervised training regime.
    

\paragraph{Graph Neural Networks} First introduced by the works of ____, Graph Neural Networks have seen a recent emergence through a variety of new architecture types and applications ____. The graph-based representation underlying these models is particularly powerful as it provides a natural framework for capturing relational information and structural dependencies between entities. This has made GNNs especially interesting for tackling combinatorial optimization problems ____ and reasoning tasks that require understanding relationships between multiple elements ____. A key advantage of graph-based approaches is their capability to handle problems of varying sizes and complexities. One specific research direction focuses how these models generalize across different problem instances and sizes ____, with benchmarks like CLRS ____ providing a more systematic evaluation frameworks for assessing algorithmic reasoning capabilities with a focus on size generalization. This in turn has sparked more investigations into developing appropriate tools and architectures for such reasoning ____.