%\documentclass[acmtog]{formats/acmart} 
\documentclass[acmtog]{acmart}
\makeatletter
\let\@authorsaddresses\@empty
\makeatother

%\documentclass[10pt, format=manuscript,nonacm]{acmart}

%\documentclass[acmtog,anonymous,review]{acmart}


\usepackage{booktabs} % For formal tables
\citestyle{acmauthoryear}
\setcitestyle{square}
\usepackage{ifthen}
\usepackage{enumitem}
\usepackage{algorithm}
%\usepackage{algorithmicx}
%\usepackage{algcompatible}
\usepackage[noend]{algpseudocode}
%\usepackage{algorithm2e}
\usepackage{float}
\usepackage{syntax}
\usepackage{amsfonts}
\let\Bbbk\relax
\usepackage{listings}
\usepackage{fancyvrb}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{xspace}
\usepackage{colortbl}
\usepackage{gensymb}
\usepackage{verbatim}
\usepackage{colortbl}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{cleveref}
\usepackage{pifont}
\usepackage{circledsteps}



%\lstset{style=mystyle}
\newcommand{\etal}{et al.}
\newcommand\TODO[1]{{\color{red} A: #1}}
\newcommand{\yuexin}[1]{{\color{blue} YX: #1}}
\newcommand{\ZY}[1]{{\color{cyan} ZY: #1}}
\newcommand{\manyi}[1]{{\color{brown} MY: #1}}
\newcommand{\haoran}[1]{{\color{olive} HR: #1}}
\newcommand{\yiteng}[1]{{\color{red} YT: #1}}
\newcommand{\yiming}[1]{{\color{orange} YM: #1}}
\newcommand{\changhe}[1]{{\color{orange} CH: #1}}


\usepackage{amsthm,amsmath,amssymb}
\usepackage{mathrsfs}
\usepackage[normalem]{ulem}
\begin{document}
\settopmatter{printacmref=false}
\setcopyright{none}
\renewcommand\footnotetextcopyrightpermission[1]{}
\pagestyle{empty} 

%\setcopyright{acmlicensed}
%\acmJournal{TOG}
%\acmYear{XXX} \acmVolume{XX} \acmNumber{XX} \acmArticle{XX} \acmMonth{XX} %\acmPrice{15.00}\acmDOI{10.1145/3618324}
\title{SymbioSim: Human-in-the-loop Simulation Platform for Bidirectional Continuing Learning in Human-Robot Interaction}


\author{Haoran Chen}
\authornote{Joint first authors.}
\affiliation{%
  \institution{Shandong University}
  \country{China}
}
\author{Yiteng Xu}
\authornotemark[1]
\affiliation{%
  \institution{ShanghaiTech University}
  \country{China}
}
\author{Yiming Ren, Yaoqin Ye}
\affiliation{%
  \institution{ShanghaiTech University}
  \country{China}
}
%\author{Yaoqin Ye}
%\affiliation{%
%  \institution{ShanghaiTech University}
%}
\author{Xinran Li, Ning Ding}
\affiliation{%
  \institution{Shandong University}
  \country{China}
}
%\author{Ning Ding}
%\affiliation{%
%  \institution{Shandong University}
%}
\author{Peishan Cong, Ziyi Wang, Bushi Liu, Yuhan Chen}
\affiliation{%
  \institution{ShanghaiTech University}
  \country{China}
}
%\author{Ziyi Wang}
%\affiliation{%
%  \institution{ShanghaiTech University}
%}
%\author{Bushi Liu}
%\affiliation{%
%  \institution{ShanghaiTech University}
%}
%\author{Yuhan Chen}
%\affiliation{%
%  \institution{ShanghaiTech University}
%}
\author{Zhiyang Dou}
\affiliation{%
  \institution{The University of Hong Kong}
  \country{China}
}
\author{Xiaokun Leng}
\affiliation{%
  \institution{LEJU(Shenzhen) Robotics Co., Ltd}
  \country{China}
}
\author{Manyi Li}
\authornote{Corresponding authors.}
\affiliation{%
  \institution{Shandong University}
  \country{China}
}
\author{Yuexin Ma}
\authornotemark[2]
\affiliation{%
  \institution{ShanghaiTech University}
  \country{China}
}
\author{Changhe Tu}
\affiliation{%
  \institution{Shandong University}
  \country{China}
}

\begin{abstract}
The development of intelligent robots seeks to seamlessly integrate them into the human world, providing assistance and companionship in daily life and work, with the ultimate goal of achieving human-robot symbiosis. To realize this vision, robots must continuously learn and evolve through consistent interaction and collaboration with humans, while humans need to gradually develop an understanding of and trust in robots through shared experiences. However, training and testing algorithms directly on physical robots involve substantial costs and safety risks. Moreover, current robotic simulators fail to support real human participation, limiting their ability to provide authentic interaction experiences and gather valuable human feedback.
In this paper, we introduce \textit{SymbioSim}, a novel human-in-the-loop robotic simulation platform designed to enable the safe and efficient development, evaluation, and optimization of human-robot interactions. By leveraging a carefully designed system architecture and modules, \textit{SymbioSim} delivers a natural and realistic interaction experience, facilitating bidirectional continuous learning and adaptation for both humans and robots. Extensive experiments and user studies demonstrate the platform's promising performance and highlight its potential to significantly advance research on human-robot symbiosis.
\end{abstract}


%\ccsdesc[500]{Computing methodologies~Shape modeling}
%\ccsdesc[300]{Computing methodologies~Graphics systems and interfaces}

%\acmJournal{TOG}

\keywords{Robotic Simulation, Augmented Reality, Human-robot Interaction, Real-time Motion Generation}

\input{figures/fig-teaser}

%\input{macros}
\maketitle


\input{sections/1-intro}
\input{sections/2-relatedWork}
\input{sections/3-system-new}
\input{sections/4-1-HR-interaction}
\input{sections/4-2-HR-collaboration}
\input{sections/5-ablation}
\input{sections/6-systemeval}
\input{sections/7-conclusion}

\section{Acknowledgement}
This work is supported by the National Natural Science Foundation of China (U23A20312, No.62302269, No.62206173), Shandong Provincial Natural Science Foundation (No.ZR2023QF077), Shandong Province Excellent Young Scientists Fund Program (Overseas) (No.2023HWYQ-034), Shanghai Frontiers Science Center of Human-centered Artificial Intelligence (ShangHAI), MoE Key Laboratory of Intelligent Perception and Human-Machine Collaboration (KLIP-HuMaCo).


\bibliographystyle{formats/ACM-Reference-Format}
\bibliography{Reference} 
% \appendix
%\input{appendix}
%\input{sections/appendix}

\end{document}
