\subsection{Dataset}

In our study of \bfps, we collected a dataset that encompasses 19,379 commits, 9,708 merged pull requests, and 3,532 issues for Apollo~\cite{apollo}, as well as 22,825 commits, 5,391 merged pull requests, and 559 issues for Autoware. 
All these commits, issues, and pull requests were established on or prior to November 13, 2023, as detailed in \autoref{tab:dataset}.

\begin{wraptable}[7]{r}{0.52\textwidth}
\vspace{-1ex}
\centering
\caption{Statistics of ADS Dataset}
\begin{adjustbox}{width=\linewidth}
\begin{tabular}{ccccccc}
\Xhline{3\arrayrulewidth}
    \textbf{System} & \textbf{\makecell{Start Date--\\End Date}} & \textbf{Commits} & \textbf{Issues} & \textbf{\makecell{Pull\\Requests}} & \textbf{\makecell{Bug\\Fixes}} \\
\hline
Apollo & \makecell{07/04/2017--\\11/13/2023} & 19,379 & 3,532 & 9,708 & 338 \\
\hline
Autoware & \makecell{08/25/2015--\\11/13/2023} & 22,825 & 559 & 5,391 & 993 \\
\Xhline{3\arrayrulewidth}
\end{tabular}
\end{adjustbox}
\label{tab:dataset}
\end{wraptable}

Considering the objective to identify \bfps in ADSes, we collected closed and merged single-file pull requests that fix bugs. Such pull requests enable us to 
(1) scrutinize the modified source code, associated issues, and developer discussions; 
and (2) avoid too many bugs and \bfps that occurred in one pull request to interfere with the evaluation and analysis. 
Note that GitHub's pull requests serve diverse functions, such as introducing new features, bug fixes, or refactoring. 
We implemented a technique to maximize the capture of bug-fix pull requests. In doing so, we employed an approach similar to those of previous research~\cite{GarciaF0AXC20,IslamNPR19,VasilescuYWDF15,ZhangCCXZ18} by curating a list of bug-associated keywords like ``fix'', ``repair'', ``defect'', ``error'', ``bug'', ``issue'', ``mistake'', ``incorrect'', ``fault'', and ``flaw'', subsequently searching for these terms in both titles and content.
Pull requests with any of these keywords were classified as initial bug-fix requests. We further filtered out 62 pull requests that were not real bug fixes even with keywords and repeated pull requests derived from different branches during labeling by analyzing their description and intentions.
This methodology yielded 338 and 993 integrated pull requests for Apollo and Autoware, respectively, that fit our criteria.
