\section{Extension to other methods}
\label{sec:s_extension}
% SDS image + FPR
% SDS NeRF + FPR
% CDS + FPR
% PDS + FPR
Since our method optimizes the source latent to estimate a more accurate score, it can be applied to other methods that are based on SDS despite that we report the results using our method to DDS.

During SDS optimization, FPR can be used to preserve the original content and reduce the blurry effect. As shown in \cref{fig:existingSDS}, the conserved rate of the information of the source image is controllable by the number of FPR iteration.

When the proposed FPR is integrated into CDS, the texture of the source image is further maintained, as illustrated in \cref{fig:sup_fpr_cds}. In addition, FPR promoted reducing the over-boosting of color often found in the translated images of CDS. This confirms that the proposed FPR can be a universal regularization to preserve the identity of the source image for text-guided image editing.
\input{Fig./Qual/Editing/existingSDS}
\input{sec/X_supp/Fig/comp/fpr_cds}

% \paragraph{FPR + SDS in 2D} FPR can be applied to SDS optimization for a given source image and prompt to preserve the original contents and reduce the blurry effect. As shown in \cref{fig:existingSDS}, the conserved rate of the information of the source image is controllable by the number of iterations of FPR.
% \input{Fig./Qual/Editing/existingSDS}

Furthermore, FPR can help optimize not only pixel space but also the parametric editor such as PDS~\cite{koo2024posterior}. As demonstrated in \cref{fig:ids-vs-pds}, \cref{fig:pds_svg}, and \cref{tab:nerf_svg}, the edited results with our method show that FPR assists in maintaining the original contents. By comparing the first and second rows of \cref{fig:ids-vs-pds}, the use of FPR results in the preservation of source components more effectively compared to PDS. Similarly, in the third and fourth rows, the results obtained using FPR retain key original features, such as the shape and color of the face as well as the color of the clothing. Furthermore, the gradient weights, FPR assigns minimal weight to the structure of the source image, such as the background, while primarily focusing the weights on the editing points. For 3D and 2D editing, we implement the experiments based on official code of PDS\footnote{\url{https://github.com/KAIST-Visual-AI-Group/PDS}}. We use the subset of Instruct-NeRF2NeRF \cite{haque2023instruct} for 3D editing and Scalable Vector Graphics (SVGs) with their text description used in \cite{jain2023vectorfusion} for 2D editing. 
% \paragraph{FPR + SDS in 3D.}

% \paragraph{FPR + CDS} As shown in \cref{fig:sup_fpr_cds}, FPR can regulate the  control the 
% As shown in CDS's translation of \cref{fig:sup_fpr_cds}, CDS has the effect of boosting color. FPR makes the translated result of CDS 
% \input{sec/X_supp/Fig/comp/fpr_cds}
% \paragraph{FPR + PDS.}
\input{sec/X_supp/Fig/comp/ids-vs-pds}
\input{sec/X_supp/Fig/comp/pds_SVG}
\input{sec/X_supp/Tab/PDS_SVGquan}