\section{Related Work}
% evolving deepfake methods
% countermeasures for countermeasures - malafide, malacopula (need acces to the systems)

%% Some introduction (can be removed or moved)
Audio deepfake detection, or anti-spoofing, has seen an impressive increase in the number of scientific publications and commercial tool releases over the last few years. There are dedicated challenges (e.g. ASVspoof____ or ADD____) and special sessions at the major conferences (e.g., Interspeech or ICASSP). These challenges and sessions reveal the growing interest in deepfake countermeasures, which is strongly correlated to the threats to cybersecurity that deepfake generation methods pose. Audio deepfake generation is performed by Text-to-Speech (TTS) or Voice Conversion (VC) methods, and systems such as 
ElevenLabs, Resemble AI or Respeecher  can now copy a target speaker's voice with only a few seconds of audio____.

Some of the early stages of deepfake detection systems have relied on building dedicated machine learning architectures that process raw waveform____, spectral features____ or learnable deep features____.
However, with the advent of large pre-trained self-supervised learning (SSL) audio models (e.g. wav2vec 2.0,  wavLM, Whisper and their variants), the deepfake detection methods have shifted focus and exploit the inherent audio representations that these models learn.
The SSL models are used as feature extractors (or \textit{frontends}), and simple____ or more complex____ classifier modules (or \textit{backends}) are appended to the pre-trained____ or finetuned____ SSL component.

% do not generalise
One problem that most of these systems encounter is the fact that they do not properly generalise to out-of-distribution samples -- meaning that they do not learn meaningful audio attributes, but rather the classification is performed by mostly
%\footnote{Out-of-distribution samples can be either samples from unknown generation systems or from speakers with particular speech characteristics.} 
exploiting potential data processing oversights in the available samples. 
For example, M\"uller et al.____ and Zhang et al.____ showed that the length of silence segments can be an accurate indicator of fake versus real data for some datasets. In a separate study M\"uller et al.____ determined a correlation between the length of the audio sample and the final decision, 
while Negroni et al.____ exploited low frequency artefacts at splicing points in the audio signal for partial deepfake detection.

% to overcome lack of genralisation - augmentation
To overcome the lack of generalisation, 
data augmentation____ can be set in place. The audio training set can be extended with real____, vocoded____ or synthesised samples____; or by simply performing signal-based modifications to the existing samples (e.g. adding noise, reverberation, truncating the signal, scrambling, etc.)____. The major issue of data augmentation is the proper selection of additional samples and signal processing methods which are informative for the final decision, which is not a trivial task____.

To make things even more complex, a very recent security threat is that of adversarial attacks targeting the deepfake detection systems themselves____. The attacks imperceptibly alter a generated audio sample to fool a particular or several deepfake detectors. 

% might need 1-2 sentences for the transition from related work to DeePen.

% \subsection{Data Augmentation}
% Write about data augmentation papers, such as rawboost____, etc.

% \subsection{Fooling ASV}
% Write about other papers such as MalaFide (Adv. Attacks) and do research on similar papers such as ours.

% \subsection{Other}
% Other related work, possible "silence is golden" paper.
% New Approach to Voice Auth