\section{Conclusion and Future Work}
In this paper, inspired by the principles of search and learning~\cite{Sutton2019BitterLesson,snell2024scaling}, we systematically investigate the integration of learning into search. We first explore how learning benefits the search process via LLMs, demonstrating that LLMs improve search efficiency by reducing search space.
Then, building on these insights, we introduce a novel framework, \method, and its variant \cmethod, designed to combine the reasoning capabilities of LLMs with search strategies to achieve efficient and accurate problem-solving. 
Extensive experiments conducted on three real-world planning tasks demonstrate that \method achieves near-perfect pass rates across various settings while significantly reducing search spaces, showcasing the effectiveness and efficiency of \method. Furthermore, we also explore how search can benefit LLMs, evaluating whether LLMs can develop self-search capabilities. 
We show that search significantly enhances their reasoning and learning performance. These findings highlight the bidirectional synergy between search and learning, emphasizing the potential of integrating search into LLMs.

Our research 
% bridges the gap between systematic search and learning-driven reasoning, 
paves the way for further exploration into the convergence of search and learning. Future work will focus on investigating how LLMs can better conduct self-searches, further unlocking their potential for complex problem-solving.

\section{Limitations}
One potential limitation of this work is the necessity of encoding our search strategies into prompts to enable LLMs to perform self-search. In the future, it is worth exploring how to allow LLMs to autonomously conduct self-search during reasoning without explicit supervision within the prompts. Additionally, our work primarily focuses on LLMs; an important direction for future investigation is to assess the applicability of our search strategy to multi-modal LLMs.

\section{Impact Statements}
This paper introduces \method and \cmethod, frameworks designed to enhance search processes by integrating the reasoning capabilities of LLMs with structured search strategies. 
% Our work aims to advance the field of AI by addressing the challenges of efficiency and completeness in search algorithms, enabling LLMs to solve complex problems more effectively.
The potential societal impacts of this research are broad but largely align with the established consequences of improving computational problem-solving techniques. Enhanced search efficiency and the ability to leverage LLMs for self-search could benefit applications in diverse fields such as healthcare, logistics, education, and robotics, where intelligent decision-making is crucial. 
However, as with all advancements in AI, there are ethical considerations. For example, the misuse of improved search strategies in domains such as automated surveillance or adversarial systems could lead to privacy or security concerns. We encourage researchers and practitioners to apply this work responsibly and ensure it aligns with ethical guidelines.
We see no immediate risks or unintended negative consequences specific to this work that require urgent attention. This paper primarily contributes to foundational research in search and learning integration. Future exploration of self-search capabilities in LLMs will include careful assessment of ethical implications to ensure responsible development and deployment of these technologies.