\documentclass[11pt,final]{article}

%Added after EC
\usepackage{xcolor}


%Make sure the order of begin document, title, author, abstract, make title is correct
%%%%%%%%%%%%%%%%


\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{lmodern}



\usepackage[numbers,sort&compress]{natbib} 
\setcitestyle{numbers}

\usepackage[colorlinks=true]{hyperref}
\hypersetup{
     colorlinks  = true,
     urlcolor    = teal,
	 citecolor   = teal,
	 linkcolor   = red
}




\usepackage[ruled,noend,linesnumbered]{algorithm2e} % For algorithms
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}



\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}
\usepackage{url}            % simple URL typesetting
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{booktabs}       % professional-quality tables
%\usepackage[linesnumbered,ruled,noend]{algorithm2e}

%Make sure this appears AFTER ams packages, etc.
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\newtheorem{definition}{Definition}
%\newtheorem{example}{Example}
\newtheorem{remark}{Remark}
\newtheorem{property}{Property}
\newtheorem{claim}{Claim}
\newtheorem{observation}{Observation}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition} % Ensures normal (non-italic) text
\newtheorem{example}{Example}


\newcount\Includeappendix %include appendices or not
\Includeappendix=1
\newcommand{\appnx}[1]{{\ifnum\Includeappendix=1{#1}\else{the appendix}\fi}}


% Choose a citation style by commenting/uncommenting the appropriate line:
%\setcitestyle{acmnumeric}
\input{def}
% Title. Note the optional short title for running heads. In the interest of anonymization, please do not include any acknowledgements.
\begin{document}

\title{Envious Explore and Exploit}

% Anonymized submission.
\author{
Omer Ben{-}Porat%
\thanks{%
    {Technion---Israel Institute of Technology (\url{omerbp@technion.ac.il})}, corresponding author}
\and Yotam Ganfi%
\thanks{%
    {Weizmann Institute (\url{yotam.gafni@gmail.com})}}
\and Or Markovetzki%
\thanks{%
    {Technion---Israel Institute of Technology (\url{ormar@campus.technion.ac.il})}}
}
\maketitle
% Abstract. Note that this must come before \maketitle.
\begin{abstract}
Explore-and-exploit tradeoffs play a key role in recommendation systems (RSs), aiming at serving users better by learning from previous interactions. Despite their commercial success, the societal effects of explore-and-exploit mechanisms are not well understood, especially regarding the utility discrepancy they generate between different users.
In this work, we measure such discrepancy using the economic notion of envy. We present a multi-armed bandit-like model in which every round consists of several sessions, and rewards are realized once per round. We call the latter property \emph{reward consistency}, and show that the RS can leverage this property for better societal outcomes. On the downside, doing so also generates envy, as late-to-arrive users enjoy the information gathered by early-to-arrive users. We examine the generated envy under several arrival order mechanisms and virtually any \emph{anonymous} algorithm, i.e., any algorithm that treats all similar users similarly without leveraging their identities. We provide tight envy bounds on uniform arrival and upper bound the envy for nudged arrival, in which the RS can affect the order of arrival by nudging its users. Furthermore, we study the efficiency-fairness trade-off by devising an algorithm that allows constant envy and approximates the optimal welfare in restricted settings. Finally, we validate our theoretical results empirically using simulations.
\end{abstract}


% Paper body
\input{sections/intro} %WIP
\input{sections/model} 
\input{sections/uniform} 
\input{sections/nudge}
\input{sections/efc}
\input{sections/discussion} 


\section*{Acknowledgments}\label{sec:Acknowledgments}
The work of O. Ben-Porat was supported by the Israel Science Foundation (ISF; Grant No. 3079/24). 

\bibliographystyle{plainnat}
\input{main.bbl}

% Appendix
\ifnum\Includeappendix=1{
\appendix
%\addtocontents{toc}{\setcounter{tocdepth}{1}}
\input{sections/appendix-uniform}
\input{sections/appendix-nudge}
\input{sections/appendix-nudge-models}
\input{sections/appendix-efc}
\input{sections/appendix-average}
\input{sections/appendix-socially-opt}
\input{sections/appendix-simulations}

}%
\fi

\end{document}
