\section{Experimental Results}\label{sec:evaluation}

This paper investigates \bfps in ADSes to address the following research questions:

\vspace{-1.5mm}
\framed
    \vspace{-2mm}
    \noindent
    \textbf{RQ1 (\Bfp frequency):} What are the most common \bfps in the ADS?
    \vspace{-2mm}
\endframed
\vspace{-1.5mm}

\yuntianyihl{This question aims to unravel the prevalent methodologies employed to address syntactic bugs—those related to the structure or format of the code—and semantic bugs, which concern the meaning or behavior of the code within ADS.}


\vspace{-1.5mm}
\framed
    \vspace{-2mm}
    \noindent
    \textbf{RQ2 (\Bfa frequency):} What are the most common \bfas in the ADS?
    \vspace{-2mm}
\endframed
\vspace{-1.5mm}

This question investigates the frequency of specific \bfas and strategies in the ADS. By categorizing these actions into domain-specific and domain-independent fixes, the study aims to provide actionable insights into which strategies are most frequent in bug fixing.

\vspace{-1.5mm}
\framed
    \vspace{-2mm}
    \noindent
    \textbf{RQ3 (\Bfps across root causes):} How do semantic \bfps vary across different root causes of bugs in the ADS?
    \vspace{-2mm}
\endframed
\vspace{-1.5mm}

\yuntianyihl{This analysis delves into the diversity of bug origins within ADS and examines the correspondence between these root causes and the \bfps applied. Understanding this correlation is important for tailoring debugging strategies to the specific nature of a bug’s root cause.}

\vspace{-1.5mm}
\framed
    \vspace{-2mm}
    \noindent
    \textbf{RQ4 (\Bfps across symptoms):} How do \yuntianyihl{semantic} \bfps vary across different symptoms in the ADS?
    \vspace{-2mm}
\endframed
\vspace{-1.5mm}

\yuntianyihl{This question focuses on the manifestation of bugs through various symptoms within ADS.
By mapping \bfps to specific symptoms, the study aims to elucidate targeted strategies for symptom-based debugging, fostering symptom-oriented guidance to ADS optimization.}
    
\vspace{-1.5mm}
\framed
    \vspace{-2mm}
    \noindent
    \textbf{RQ5 (\Bfps across ADS modules):} Are \yuntianyihl{semantic} \bfps different across various modules in the ADS? 
    \vspace{-2mm}
\endframed
\vspace{-1.5mm}

\yuntianyihl{ADS encompasses a spectrum of modules, 
this question explores whether the \bfps employed are module-specific, reflecting the unique challenges and requirements of each module.}

\input{sections/evaluation/rq1}
\input{sections/evaluation/rq2}
\input{sections/evaluation/rq3}
\input{sections/evaluation/rq4}
\input{sections/evaluation/rq5}