 Social influence profoundly impacts individual choices and collective behaviors in politics. In this work, driven by the goal of protecting elections from improper influence, we consider the following scenario: an individual, who has vested interests in political party $Y$, is aware through reliable surveys that parties $X$ and $Y$ are likely to get 50.1\% and 49.9\% of the vote, respectively. Could this individual employ   strategies to alter public opinions and consequently invert these polling numbers in favor of party $Y$?

We address this question by employing: (i) the  Friedkin-Johnsen (FJ) opinion dynamics model, which is mathematically sophisticated and effectively captures the way individual biases and social interactions shape opinions, making it crucial for examining social influence, and (ii) interventions similar to those in Asch's experiments, which involve selecting a group of stooges within the network to spread a specific opinion. We mathematically formalize the aforementioned motivation as an optimization framework and establish that it is NP-hard and inapproximable within any constant factor. We introduce three efficient polynomial-time algorithms. The first two utilize a continuous approach: one employs gradient descent with Huber's estimator to approximate the median, and the other uses a sigmoid threshold influence function. The third utilizes a combinatorial greedy algorithm for targeted interventions. Through comparative analysis against various natural baselines and using real-world data, our results demonstrate that in numerous cases a small fraction of nodes chosen as stooges can significantly sway election outcomes under the Friedkin-Johnsen model.



% Let's consider the following type of graphs: trees where information only flows towards the root (from nodes with higher depth towards nodes with lower depth).

% A natural representation is companies, where managers just talk with their direct subordinates, or government, and we want to influence a certain internal election (aka change opinion of some employees, so that if there was a vote the motion would pass). This DP should work on some similar problems (like each employee has a price, but then the prices should be integers and the complexity depends on the sum of the prices). Or what's the minimum number of stooges to convince the CEO to adopt the motion. Main this is the price should be integers and the sum of prices should be small. It's also worth noting that this will NOT work on our previous problems of polarization. It will also not work on just maximizing opinions of people. This only works because of the discrete nature of voting for / against the motion (continuous opinions get converted to discrete ones)  

% We can use dynamic programming to find the optimal choice of stooges to bring the opinion of more than half the nodes above $0.5$.

% Let $dp[i][j][k] = $ maximum amount of change that node $i$ accumulated from its subtree and can pass forward to its parent so that j nodes in its subtree have opinion $> 0.5$ and k nodes were turned into stooges.

% The optimal number of stooges will be the smallest k so that $dp[root][(N+1)/2][k]$ is a valid state.

% In order to compute the states of the dynamic programming, we first look at the base case: the leaves. We have 2 possibilities: making the node itself a stooge (in which case j=1, k=1 and the accumulated value is also 1), or not making it a stooge (in which case k=0, the accumulated value is the opinion of the node and j is either 0 or 1 depending on whether the opinion was above or below $0.5$

% For a node that does have children, we need to find the best way to split the k stooges across them. We do so by creating an auxiliary dynamic programming: $cdp[i][j][k] = $ largest sum of opinions across the first i children with a total of j nodes $> 0.5$ and using a total of k nodes. This is very similar to knapsack, since when we add the $(i+1)^{th}$ child and have to combine a $cdp[i][j][k]$ with $dp[c][j'][k']$ we create the state $dp[i+1][j+j'][k+k'] = cdp[i][j][k] + w_c \cdot dp[c][j'][k']$.