\documentclass{article}


\usepackage[preprint]{neurips_2024}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes.multipart}
\usepackage{nicematrix}
\usepackage{bm}
\usepackage{setspace}
\usepackage{xspace}
\usepackage[percent]{overpic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{booktabs} 

\input{header}
\input{defs}


\title{Variational Phylogenetic Inference with Products over Bipartitions}


\author{
  Evan Sidrow \\
  Department of Statistics and Actuarial Science\\
  Simon Fraser University\\
  Burnaby, Canada \\
  \texttt{esidrow@sfu.ca} \\
  \And
  Alexandre Bouchard-C\^ot\'e \\
  Department of Statistics \\
  University of British Columbia \\
  Vancouver, Canada \\
  \AND
  Lloyd T.\ Elliott \\
  Department of Statistics and Actuarial Science\\
  Simon Fraser University\\
  Burnaby, Canada
}


\begin{document}


\maketitle


\begin{abstract}

Bayesian phylogenetics requires accurate and efficient approximation of posterior distributions over trees. In this work, we develop a variational Bayesian approach for ultrametric phylogenetic trees. We present a novel variational family based on coalescent times of a single-linkage clustering and derive a closed-form density of the resulting distribution over trees. Unlike existing methods for ultrametric trees, our method performs inference over all of tree space, it does not require any Markov chain Monte Carlo subroutines, and our variational family is differentiable. Through experiments on benchmark genomic datasets and an application to SARS-CoV-2, we demonstrate that our method achieves competitive accuracy while requiring significantly fewer gradient evaluations than existing state-of-the-art techniques.

\end{abstract}

\renewcommand\ttdefault{cmvtt}

\section{Introduction}
\input intro.tex
%
\section{Background}
\input background.tex
%
\section{Methods}
%
\input model.tex
%
\section{Experiments}
\input case_study.tex
%
\section{Discussion}
\input discussion.tex
%
\section*{Acknowledgements}
We gratefully acknowledge all data contributors, i.e., the Authors and their Originating laboratories responsible for obtaining the specimens, and their Submitting laboratories for generating the genetic sequence and metadata and sharing via the GISAID Initiative, on which this research is based. This research was enabled in part by support provided by the Digital Research Alliance of Canada.
%
\bibliography{references}
%
\newpage
\appendix
    \newgeometry{
    textheight=9in,
    textwidth=6.5in,
    top=1in,
    headheight=12pt,
    headsep=25pt,
    footskip=30pt
    }
\input appendix.tex
\end{document}
\typeout{get arXiv to do 4 passes: Label(s) may have changed. Rerun}