%File: formatting-instructions-latex-2025.tex
%release 2025.0
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage{aaai25}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet}  % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\usepackage[T1]{fontenc}
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS


%File: anonymous-submission-latex-2025.tex
% \documentclass[letterpaper]{article} % DO NOT CHANGE THIS
% \usepackage[submission]{aaai25}  % DO NOT CHANGE THIS
% \usepackage{times}  % DO NOT CHANGE THIS
% \usepackage{helvet}  % DO NOT CHANGE THIS
% \usepackage{courier}  % DO NOT CHANGE THIS
% \usepackage[hyphens]{url}  % DO NOT CHANGE THIS
% \usepackage{graphicx} % DO NOT CHANGE THIS
% \urlstyle{rm} % DO NOT CHANGE THIS
% \def\UrlFont{\rm}  % DO NOT CHANGE THIS
% \usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
% \usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
% \frenchspacing  % DO NOT CHANGE THIS
% \setlength{\pdfpagewidth}{8.5in} % DO NOT CHANGE THIS
% \setlength{\pdfpageheight}{11in} % DO NOT CHANGE THIS
%
% These are recommended to typeset algorithms but not required. See the subsubsection on algorithms. Remove them if you don't have algorithms in your paper.
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath,amssymb}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{pdfpages}
%\usepackage{hyperref}
\makeatletter
\newcommand*{\rom}[1]{\romannumeral #1}
\makeatother

\newcommand{\framework}{Decentralized Cooperative Generative Agents}
\newcommand{\frameworkAbbr}{DCGA}
\newcommand{\hq}[1]{{\color{black} #1}}
\newcommand{\jd}[1]{{\color{black} #1}}

\usepackage{cite}



% These are are recommended to typeset listings but not required. See the subsubsection on listing. Remove this block if you don't have listings in your paper.
\usepackage{newfloat}
\usepackage{listings}
\DeclareCaptionStyle{ruled}{labelfont=normalfont,labelsep=colon,strut=off} % DO NOT CHANGE THIS
\lstset{%
	basicstyle={\footnotesize\ttfamily},% footnotesize acceptable for monospace
	numbers=left,numberstyle=\footnotesize,xleftmargin=2em,% show line numbers, remove this entire line if you don't want the numbers.
	aboveskip=0pt,belowskip=0pt,%
	showstringspaces=false,tabsize=2,breaklines=true}
\floatstyle{ruled}
\newfloat{listing}{tb}{lst}{}
\floatname{listing}{Listing}
%
% Keep the \pdfinfo as shown here. There's no need
% for you to add the /Title and /Author tags.
\pdfinfo{
/TemplateVersion (2025.1)
}

% DISALLOWED PACKAGES
% \usepackage{authblk} -- This package is specifically forbidden
% \usepackage{balance} -- This package is specifically forbidden
% \usepackage{color (if used in text)
% \usepackage{CJK} -- This package is specifically forbidden
% \usepackage{float} -- This package is specifically forbidden
% \usepackage{flushend} -- This package is specifically forbidden
% \usepackage{fontenc} -- This package is specifically forbidden
% \usepackage{fullpage} -- This package is specifically forbidden
% \usepackage{geometry} -- This package is specifically forbidden
% \usepackage{grffile} -- This package is specifically forbidden
% \usepackage{hyperref} -- This package is specifically forbidden
% \usepackage{navigator} -- This package is specifically forbidden
% (or any other package that embeds links such as navigator or hyperref)
% \indentfirst} -- This package is specifically forbidden
% \layout} -- This package is specifically forbidden
% \multicol} -- This package is specifically forbidden
% \nameref} -- This package is specifically forbidden
% \usepackage{savetrees} -- This package is specifically forbidden
% \usepackage{setspace} -- This package is specifically forbidden
% \usepackage{stfloats} -- This package is specifically forbidden
% \usepackage{tabu} -- This package is specifically forbidden
% \usepackage{titlesec} -- This package is specifically forbidden
% \usepackage{tocbibind} -- This package is specifically forbidden
% \usepackage{ulem} -- This package is specifically forbidden
% \usepackage{wrapfig} -- This package is specifically forbidden
% DISALLOWED COMMANDS
% \nocopyright -- Your paper will not be published if you use this command
% \addtolength -- This command may not be used
% \balance -- This command may not be used
% \baselinestretch -- Your paper will not be published if you use this command
% \clearpage -- No page breaks of any kind may be used for the final version of your paper
% \columnsep -- This command may not be used
% \newpage -- No page breaks of any kind may be used for the final version of your paper
% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
% \pagestyle -- This command may not be used
% \tiny -- This is not an acceptable font size.
% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference

\setcounter{secnumdepth}{2} %May be changed to 1 or 2 if section numbers are desired.

% The file aaai25.sty is the style file for AAAI Press
% proceedings, working notes, and technical reports.
%

% Titl

%Example, Single Author, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
\iffalse
\title{LLM-Powered Decentralized Generative Agents with Adaptive Hierarchical Knowledge Graph for Cooperative Planning}
\author {
    Author Name
}
\affiliations{
    Affiliation\\
    Affiliation Line 2\\
    name@example.com
}
\fi

%\iffalse
%Example, Multiple Authors, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it

\title{
\hq{
LLM-Powered Decentralized Generative Agents with Adaptive Hierarchical Knowledge Graph for Cooperative Planning
}
}
\author {
    % Authors
    Hanqing Yang\textsuperscript{\rm 1},
    Jingdi Chen\textsuperscript{\rm 1},
    Marie Siew\textsuperscript{\rm 2},
    Tania Lorido-Botran\textsuperscript{\rm 3 4},
    Carlee Joe-Wong\textsuperscript{\rm 1}
}
\affiliations {
    % Affiliations
    \textsuperscript{\rm 1}Carnegie Mellon University\\
    \textsuperscript{\rm 2}Singapore University of Technology and Design\\
    \textsuperscript{\rm 3}Roblox\\
    \textsuperscript{\rm 4}Northeastern University\\
    \{hanqing3, jingdic, cjoewong\}@andrew.cmu.edu, marie\_siew@sutd.edu.sg, t.loridobotran@northeastern.edu
}
%\fi

\renewcommand{\baselinestretch}{1}
\begin{document}

\maketitle

\begin{abstract}
Developing intelligent agents for long-term cooperation in dynamic open-world scenarios is a major challenge in multi-agent systems. Traditional Multi-agent Reinforcement Learning (MARL) frameworks like centralized training decentralized execution (CTDE) struggle with scalability and flexibility. They require centralized long-term planning, which is difficult without custom reward functions, and face challenges in processing multi-modal data. CTDE approaches also assume fixed cooperation strategies, making them impractical in dynamic environments where agents need to adapt and plan independently.
To address decentralized multi-agent cooperation, we propose Decentralized Adaptive Knowledge Graph Memory and Structured Communication System (\textbf{DAMCS}) in a novel Multi-agent Crafter environment. Our generative agents, powered by Large Language Models (LLMs), are more scalable than traditional MARL agents by leveraging external knowledge and language for long-term planning and reasoning. 
Instead of fully sharing information from all past experiences, DAMCS introduces a multi-modal memory system organized as a hierarchical knowledge graph and a structured communication protocol to optimize agent cooperation. This allows agents to reason from past interactions and share relevant information efficiently. Experiments on novel multi-agent open-world tasks show that DAMCS outperforms both MARL and LLM baselines in task efficiency and collaboration. Compared to single-agent scenarios, the two-agent scenario achieves the same goal with 63\% fewer steps, and the six-agent scenario with 74\% fewer steps, highlighting the importance of adaptive memory and structured communication in achieving long-term goals. \hq{We publicly release our project at: https://happyeureka.github.io/damcs}.
\end{abstract}

% Uncomment the following to link to your code, datasets, an extended version or similar.
%
% \begin{links}
%     \link{Code}{https://aaai.org/example/code}
%     \link{Datasets}{https://aaai.org/example/datasets}
%     \link{Extended version}{https://aaai.org/example/extended-version}
% \end{links}

\section{Introduction}
\input{AnonymousSubmission/LaTeX/1_Intro}



\section{Related Work}\label{sec:related}
\input{AnonymousSubmission/LaTeX/2_RelatedWork}



\section{Framework: DAMCS}\label{sec:method}
% \carlee{We should change the title to something specific to our method, e.g., \framework}

In this section, we give an overview of our framework. We first formally define how this framework interacts with our problem environment (Section~\ref{sec:setting}) and then describe the design of our multi-modal, adaptive memory system (Section~\ref{sec:memory}), structured LLM output for making agent decisions (Section~\ref{sec:output}) and communication protocol that enables agent cooperation (Section~\ref{sec:communication}).
\input{AnonymousSubmission/LaTeX/3_1_method_memory}
\input{AnonymousSubmission/LaTeX/3_2_method_communication}

\hq{

\section{Evaluation Challenges of LLM Agents}\label{sec:evaluation_challenges}
Evaluating LLM-powered multi-agent systems presents unique challenges. Unlike MARL-based agents, which are trained to optimize carefully crafted rewards, LLM agents rely on prompts and contextual information, making them highly adaptable but sensitive to the evaluation environment.

\textbf{Limitations of Existing Environments.} Existing multi-agent benchmarks are often too simple for meaningful collaboration~\cite{terry2021pettingzoo} or too complex~\cite{berner2019dota,vinyals2019grandmaster,fan2022minedojo}. Many focus on micro-level action management, whereas our work emphasizes macro-level planning, communication, and cooperation. Furthermore, MARL frameworks are known for scalability challenges, and existing environments are often not designed to support cooperative tasks that scale well with an increasing number of agents.

\textbf{Evaluation of Cooperation.} LLM-based collaboration is highly adaptable but difficult to quantify. Unlike RL agents that optimize reward signals, LLM-based collaboration relies on context and commonsense reasoning, making responses variable. No standardized metric exists for evaluating cooperation among LLM agents, and extensive modifications to benchmarks are often required. Testing with environment-specific prompts is also time-consuming.

\textbf{Quantifying LLM Agents' Capabilities.} Evaluating memory quality and adaptability in LLM agents is non-trivial. While our \textbf{A-KGMS} organizes past experiences, determining the quality of stored information and its impact on decision-making remains challenging. Adaptability is also difficult to measure, as LLM agents adjust dynamically rather than optimizing predefined objectives.

To address these challenges, we introduce Multi-Agent Crafter to evaluate strategic coordination, planning, and resource sharing in open-ended, scalable cooperative tasks.
}

\section{Multi-Agent Crafter: A Novel Testbed}\label{sec:crafter}
\input{AnonymousSubmission/LaTeX/4_multiagent_crafter}



\section{Evaluations}\label{sec:evaluation}
\input{AnonymousSubmission/LaTeX/5_results}

\section{Conclusion}\label{sec:conclusion}
\input{AnonymousSubmission/LaTeX/6_conclusion}



%\bibliographystyle{aaai25}
\bibliography{aaai25}

\input{AnonymousSubmission/LaTeX/7_appendix}

\end{document}
