%%%% ijcai25.tex

\typeout{IJCAI--25 Instructions for Authors}

% These are the instructions for authors for IJCAI-25.

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in

% The file ijcai25.sty is a copy from ijcai22.sty
% The file ijcai22.sty is NOT the same as previous years'
\usepackage{ijcai25}

% Use the postscript times font!
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
%\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage[switch]{lineno}



\usepackage{comment}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multirow}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}


\usepackage{tikz}
\usetikzlibrary{trees}
\usepackage{forest}
\usetikzlibrary{shapes,arrows,positioning}





% Comment out this line in the camera-ready submission
%\linenumbers

\urlstyle{same}

% the following package is optional:
%\usepackage{latexsym}

% See https://www.overleaf.com/learn/latex/theorems_and_proofs
% for a nice explanation of how to define new theorems, but keep
% in mind that the amsthm package is already included in this
% template and that you must *not* alter the styling.
\newtheorem{example}{Example}
\newtheorem{theorem}{Theorem}
\newtheorem{observation}{Observation}

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission.

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.

% added packages
\usepackage{xspace}

\newcommand{\ADDAND}{\ensuremath{\text{ADD}[\land]}\xspace}
\newcommand{\OBDDAND}{\ensuremath{\text{OBDD}[\land]}\xspace}

\makeatletter
\newcommand{\starfootnote}[1]{%
	\let\oldthefootnote\thefootnote% 保存原始脚注编号格式
	\renewcommand{\thefootnote}{\fnsymbol{footnote}}% 使用符号作为脚注编号
	\footnote{#1}% 添加脚注
	\let\thefootnote\oldthefootnote% 恢复原始脚注编号格式
}
\makeatother

\newcommand{\unnumberedfootnote}[1]{%
	\begingroup
	% 保存当前脚注计数器的值
	\edef\oldfootnote{\thefootnote}%
	% 将脚注计数器的值设置为一个很大的数，这样它不会影响后续正常脚注的编号
	\setcounter{footnote}{1000}%
	% 定义脚注编号为空，即不显示编号
	\renewcommand{\thefootnote}{}%
	% 插入脚注内容
	\footnote{#1}%
	% 恢复脚注计数器到之前保存的值
	\setcounter{footnote}{\oldfootnote}%
	\endgroup
}
% PDF Info Is REQUIRED.

% Please leave this \pdfinfo block untouched both for the submission and
% Camera Ready Copy. Do not include Title and Author information in the pdfinfo section
\pdfinfo{
	/TemplateVersion (IJCAI.2025.0)
}

\title{Scalable Precise Computation of Shannon
	Entropy}


% Single author syntax

%\affiliations
%    Affiliation
%\emails
%    email@example.com
\author{
	Yong Lai$^{1,2}$
	\and
	Haolong Tong$^{1,2}$
	\and
	Zhenghang Xu$^{1,2,3}$
	\and
	Minghao Yin$^{2,3}$\thanks{Corresponding author} % 使用 \thanks 标记通讯作者
	\affiliations
	$^1$College of Computer Science and Technology, Jilin University, Changchun 130012, China\\
	$^2$Key Laboratory of Symbolic Computation and Knowledge Engineering  Ministry of Education, Jilin University, Changchun 130012, China\\
	$^3$School of Computer Science and Information Technology, Northeast Normal University, Changchun 130017, China\\
	%\emails
	%\{laiyong, xuzhenghang\}@jlu.edu.cn,
	%yinminghao@nenu.edu.cn
}

% Multiple author syntax (remove the single-author syntax above and the \iffalse ... \fi here)
\iffalse
\author{
	First Author$^1$
	\and
	Second Author$^2$\and
	Third Author$^{2,3}$\And
	Fourth Author$^4$\\
	\affiliations
	$^1$First Affiliation\\
	$^2$Second Affiliation\\
	$^3$Third Affiliation\\
	$^4$Fourth Affiliation\\
	\emails
	\{first, second\}@example.com,
	third@other.example.com,
	fourth@example.com
}
\fi

\begin{document}
	
	\maketitle
	\begin{abstract}
		Quantitative information flow analyses (QIF) are a class of techniques for measuring the amount of confidential information leaked by a program to its public outputs. 
		Shannon entropy is an important method to quantify the amount of leakage in QIF.
		This paper focuses on the programs modeled in Boolean constraints and optimizes the two stages of the Shannon entropy computation to implement a scalable precise tool PSE.
		In the first stage, we design a knowledge compilation language called \ADDAND that combines Algebraic Decision Diagrams and conjunctive decomposition.
		\ADDAND avoids enumerating possible outputs of a program and supports tractable entropy computation. 
		In the second stage, we optimize the model counting queries that are used to compute the probabilities of outputs. 
		We compare PSE with the state-of-the-art probably approximately correct tool EntropyEstimation, which was shown to significantly outperform the existing precise tools.
		The experimental results demonstrate that PSE solved 55 more benchmarks compared to EntropyEstimation in a total of 441. For 98\% of the benchmarks that both PSE and EntropyEstimation solved, PSE is at least $10\times$ as efficient as EntropyEstimation.
		%\keywords{Shannon Entropy \and Quantitative Information Flow Analyses \and Knowledge Compilation.}
	\end{abstract}
	
	
	\unnumberedfootnote{Authors are listed alphabetically by last name.}
	
	\input{Sections/1.tex}
	
	\input{Sections/2.tex}
	
	\input{Sections/3.tex}
	
	\input{Sections/4.tex}
	
	\input{Sections/5.tex}
	
	\input{Sections/6.tex}
	
	\input{Sections/7.tex}
	
	
	%% The file named.bst is a bibliography style file for BibTeX 0.99c
	\bibliographystyle{named}
	\bibliography{main}
	
	\clearpage
	%\input{Sections/Appendix.tex}
	
\end{document}

