\pdfoutput=1
\documentclass{article}

\usepackage{PRIMEarxiv}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{xcolor}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{fancyhdr}       % header
\usepackage{graphicx}       % graphics
\graphicspath{{media/}}     % organize your images and other figures under media/ folder
\usepackage[thicklines]{cancel}
\usepackage{amsfonts,amssymb}
\usepackage{dsfont}
\usepackage{enumitem}
\usepackage{pifont}
\usepackage{subfigure}
\usepackage{wrapfig}
%Header
\pagestyle{fancy}
\thispagestyle{empty}
\rhead{ \textit{ }} 
\usepackage{amsmath}
\usepackage{amsthm}

\newtheorem{assumption}{Assumption}[section]
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}{Remark}[section]
\DeclareMathOperator*{\argminA}{arg\,min}
\DeclareMathOperator*{\argmaxA}{arg\,max}

% % Todonotes is useful during development; simply uncomment the next line
% %    and comment out the line below the next line to turn off comments
% % \usepackage[disable,textsize=tiny]{todonotes}
% \usepackage[textsize=tiny]{todonotes}


%Header
\pagestyle{fancy}
\thispagestyle{empty}
\rhead{ \textit{ }} 

% Update your Headers here
\fancyhead[LO]{Provably Efficient Multi-Objective Bandit Algorithms under Preference-Centric Customization}
% \fancyhead[RE]{Firstauthor and Secondauthor} % Firstauthor et al. if more than 2 - must use \documentclass[twoside]{article}



  
%% Title
\title{
Provably Efficient Multi-Objective Bandit Algorithms under Preference-Centric Customization
}

\author{
  Linfeng Cao \\
  Department of CSE \\
  The Ohio State University \\
  % City\\
  \texttt{cao.1378@osu.edu} \\
  %% examples of more authors
  \And
  Ming Shi \\
  Department of Electrical Engineering \\
  University at Buffalo \\
  % City\\
  \texttt{mshi24@buffalo.edu} \\
   \And
  Ness B. Shroff \\
  Department of ECE \\
  The Ohio State University \\
  % City\\
  \texttt{shroff.11@osu.edu} \\
}


\begin{document}
\maketitle


\begin{abstract}
Multi-objective multi-armed bandit (MO-MAB) problems traditionally aim to achieve Pareto optimality. However, real-world scenarios often involve users with varying preferences across objectives, resulting in a Pareto-optimal arm that may score high for one user but perform quite poorly for another. This highlights the need for \emph{customized learning}, a factor often overlooked in prior research.
To address this, we study a \emph{preference-aware} MO-MAB framework in the presence of explicit user preference.
% , where each user's overall-reward is modeled as the inner product of user preference and arm reward. 
It shifts the focus from achieving Pareto optimality to further optimizing within the Pareto front under preference-centric customization. To our knowledge, this is the first theoretical study of customized MO-MAB optimization with explicit user preferences.
Motivated by practical applications, we explore two scenarios: unknown preference and hidden preference, each presenting unique challenges for algorithm design and analysis. At the core of our algorithms are \emph{preference estimation} and \emph{preference-aware optimization} mechanisms to adapt to user preferences effectively. We further develop novel analytical techniques to establish near-optimal regret of the proposed algorithms. Strong empirical performance confirm the effectiveness of our approach.
\end{abstract}

\input{sections/introduction}
\input{sections/related_work}
\input{sections/porblem_definition}
\input{sections/case_unknown}
\input{sections/case_hidden}
\input{sections/conclusion}



\bibliography{reference}
\bibliographystyle{plain}

\newpage
\input{sections/appendix}

\end{document}