\section{Literature Review}
\subsection{2.1 Vertiport Location Selection}
Vertiports, which are essential infrastructure for UAM, should be installed in appropriate locations considering demand, economic feasibility, operational capability, and safety. Numerous studies employing various methodologies have been conducted to address these considerations.

The multi-criteria decision-making (MCDM) method involves making optimal choices by assessing multiple evaluation criteria across different options, employing techniques such as Analytic Hierarchy Process (AHP) and Focus Group Interviews (FGI). Fadhil (2018) analyzed the factors of UAM ground infrastructure placement and suitability using WLC and AHP techniques ____. Lee et al. (2023) proposed an optimal Vertiport and UAM network by calculating various topographic, population and social data of Seoul and its weights through the FGI technique for pilots ____. Recent studies have adequately considered whether the area is permissible for flight. 

From a data-driven perspective, recent studies on Vertiport location assessment have utilized data from four major categories—spatial, mobility, environmental, and socio-economic—for conducting multifaceted analyses. Straubinger et al. (2021) evaluated the UAM ecosystem in different urban spatial structures using population density data ____. Bulusu et al. (2021) analyzed feasible combinations of Vertiport locations using traffic data ____. Kotwicz Herniczek et al. (2022) evaluated the impact of airspace restrictions on the feasibility of Vertiports and UAM routes using airspace data ____. EASA (2021) analyzed the social acceptance of UAM in Europe using surveys and noise assessments ____. Other related literature is summarized in Table \ref{tab:Major Data}.


\subsection{2.2 Integration of UAM with Existing Transportation Systems}

Since the emergence of the UAM concept, researchers have proposed various approaches for integrating Vertiports into the existing transportation system. Fadhil (2018) utilized GIS to select locations for UAM among ground infrastructure locations ____. Rajendran et al. (2019) proposed Vertiport locations that enhance airport accessibility by analyzing taxi data in New York City ____. Wang et al. (2023) suggested utilizing UAM to transfer passengers from suburban areas to airports, thereby facilitating access to conventional air transportation through an on-demand transfer service ____. Those studies have focused on transfers between UAM and existing urban transportation networks, and the idea of linking highway infrastructure with UAM has not been proposed.


\begin{table}[H]
    \caption{List of Data Used for Vertiport Location Assessment}\label{tab:Major Data}
        \begin{center}
        \begin{tabular}{>{\raggedright\arraybackslash}m{0.2\linewidth}>{\raggedright\arraybackslash}m{0.22\linewidth}>{\raggedright\arraybackslash}m{0.5\linewidth}} \toprule[1.5pt]
            \multicolumn{1}{c}{\textbf{Category}} & \multicolumn{1}{c}{\textbf{Data}} & \multicolumn{1}{c}{\textbf{Application}} \\ 
            \midrule[1.5pt]
            \multirow{4}{*}[-0.5ex]{Spatial Data} & Population Density ____ & Potential demand evaluation of Vertiport location based on population distribution and density Information \\ \cline{2-3}
             & Employment Density ____ & Evaluation of access to key CBDs based on employment density information \\ \cline{2-3}
             & Land Use ____ & Selection of locations based on the status of land use, such as commercial districts, residential districts, and industrial districts \\ \cline{2-3}
             & Height of Building ____ & Evaluation of safety during take-off and landing through building data in the city center \\ \hline
            \multirow{4}{*}[-0.5ex]{Mobility Data} & Traffic ____ & Evaluation of accessibility by road traffic, bus and subway passengers \\ \cline{2-3}
             & Commuting Pattern ____ & Identification of key user groups by analyzing commuting flow over time \\ \cline{2-3}
             & Public Transport ____& Evaluation of accessibility by analyzing bus, subway line and stop locations \\ \cline{2-3}
             & Traffic Congestion ____ & Evaluation of traffic congestion level and to derive the optimal location \\ \hline
            \multirow{4}{*}[-0.5ex]{Environmental Data} & Air Space ____ & Selection of information-based locations related to aviation regulations, such as flight prohibitions/restrictions/control zones \\ \cline{2-3}
             & Nature Reserve ____ & Evaluation of environmental constraints such as Nature Reserve and Waterfront Areas \\ \cline{2-3}
             & Noise ____ & Evaluation of the noise level from flying UAM vehicles and Identification of areas where the noise levels is higher than criteria. \\ \cline{2-3}
             & Weather ____ & Analysis of historical wind data for UAM takeoff and landing direction assessment \\ \cline{2-3}
             & Cost of Construction ____ & Economic feasibility assessment based on cost information for the Vertiport construction \\ \hline
            \multirow{4}{*}[-0.5ex]{\raggedright \makecell{Socio-Economic \\ Data}} & Cost of Operation ____ & Long-term operational feasibility assessment based on cost information required to operate Vertiport \\ \cline{2-3}
             & Estimation Earnings ____ & Evaluation of economic feasibility by predicting revenue from Vertiport operations \\ \cline{2-3}
             & Income Level ____ & Economic feasibility assessment of UAM services through income distribution information \\ \cline{2-3}
             & Social Acceptance ____& Evaluation of social impact and acceptability based on survey or noise impact assessment \\ \cline{2-3}
             & Related Regulations ____ & Evaluation of regulations, policies, and restrictions for selecting Vertiport locations
             \\ 
             \bottomrule[1.5pt]
        \end{tabular}
    \end{center}
\end{table}