% \documentclass{article}
\documentclass[a4paper, UKenglish, cleveref, autoref, thm-restate, unicode, colorlinks=true, allcolors=blue]{lipics-v2021}

\nolinenumbers{}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
% \usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size
% \usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
% \usepackage[colorlinks=true, allcolors=blue]{hyperref}
% \usepackage[capitalize]{cleveref}
\usepackage{xspace}
\usepackage{xcolor}
\usepackage{thmtools}
\usepackage{thm-restate}
\usepackage{url}
\usepackage{svg}
\usepackage[normalem]{ulem}
\usepackage[T1]{fontenc}

\input{macros}

\title{U-index: A Universal Indexing Framework for Matching Long Patterns}%: Theory, Practice, and Applications}

\authorrunning{L. A. K. Ayad et al.}

\author{Lorraine A. K. Ayad}
       {Brunel University London, London, UK}
       {lorraine.ayad@brunel.ac.uk}
       {orcid.org/0000-0003-0846-2616}
       {}
\author{Gabriele Fici}
       {Dipartimento di Matematica e Informatica, Università di Palermo, Italy}
       {gabriele.fici@unipa.it}
       {https://orcid.org/0000-0002-3536-327X}
       {Supported by MIUR project PRIN 2022 APML – 20229BCXNW}
\author{Ragnar Groot Koerkamp}
       {ETH Zurich, Zurich, Switzerland}
       {ragnar.grootkoerkamp@inf.ethz.ch}
       {https://orcid.org/0000-0002-2091-1237}
       {ETH Research Grant ETH-1721-1 to Gunnar Rätsch.}
\author{Grigorios Loukides}
       {King's College London, London, UK}
       {grigorios.loukides@kcl.ac.uk}
       {https://orcid.org/0000-0003-0888-5061}
       {}
\author{Rob Patro}
       {University of Maryland, College Park, MD, USA}
       {rob@cs.umd.edu}
       {https://orcid.org/0000-0001-8463-1675}
       {NIH grant award number R01HG009937, NSF award CNS-1763680 and grants
         252586 and 2024342821 from the Chan Zuckerberg Initiative DAF, an
         advised fund of Silicon Valley Community Foundation.
         RP is a co-founder of Ocean Genomics, Inc.
       }
\author{Giulio Ermanno Pibiri}
       {Ca' Foscari University of Venice, Venice, Italy\\ISTI-CNR, Pisa, Italy}
       {giulioermanno.pibiri@unive.it}
       {https://orcid.org/0000-0003-0724-7092}
       {European Union’s Horizon Europe research and innovation programme (EFRA project, Grant Agreement Number 101093026). This work was also partially supported by DAIS – Ca’ Foscari University of Venice within the IRIDE program.}
\author{Solon P. Pissis}
       {CWI, Amsterdam, The Netherlands\\Vrije Universiteit, Amsterdam, The Netherlands}
       {solon.pissis@cwi.nl}
       {https://orcid.org/0000-0002-1445-1932}
       {Supported by the PANGAIA and ALPACA projects that have received funding from the European Union’s Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreements No 872539 and 956229, respectively.}

\ccsdesc{Theory of computation~Sketching and sampling}
\ccsdesc{Applied computing~Bioinformatics}

\supplementdetails[subcategory={Rust}, linktext={github.com/u-index/u-index-rs}]{Software}{
https://github.com/u-index/u-index-rs
}

\keywords{Sketching; Hashing; Minimizers; Text Indexing}

\hideLIPIcs

% \EventEditors{}
% \EventNoEds{}
% \EventLongTitle{}
% \EventShortTitle{}
% \EventAcronym{}
% \EventYear{}
% \EventDate{}
% \EventLocation{}
% \EventLogo{}
% \SeriesVolume{}
% \ArticleNo{}

\begin{document}
\maketitle

\input{abstract}

%% \newpage
\setcounter{page}{1}

\input{introduction}
\input{background}
\input{related-work}
\input{framework}
\input{experiments}
\input{applications}
\input{conclusions}

\bibliographystyle{plain}
\bibliography{bibliography}

\newpage
\appendix
\input{appendix}

%\newpage
%\input{moved}

\end{document}

\section{Implementation}

A few important remarks with respect to a C\texttt{++} implementation.

\begin{remark}
The simple construction algorithm (without $F$ and the tries) uses a single pass over $T$.
We never store $\Minimizers$. We compute the elements of $\Minimizers$ from left to right and construct $H,V,$ and $R$.
It may be the case that for long patterns, we need to implement vector $F$ (but not the tries).
\end{remark}

\begin{remark}
The value of $\tau$ can for example be set to $8$ so that the alphabet is $[0,256)$.
Thus if we have an implementation of \INDEX that works with \texttt{unsigned char}, it will work with $S$ over $[0,256)$.
\end{remark}

\begin{remark}\label{rem:V}
Vector $V$ is monotonically increasing and so it can be compressed using Elias-Fano coding~\cite{DBLP:journals/jacm/Elias74}; see \url{https://simongog.github.io/assets/data/sdsl-cheatsheet.pdf} (\texttt{enc\_vector}). Specifically, we can use \texttt{sdsl::enc\_vector<sdsl::coder::elias\_delta, 128> enc\_delta(V)}.
\end{remark}

\begin{remark}\label{rem:H}
Hash table $H$ can also be compressed as follows.
Sort the keys of $H$ in increasing order and store them in a vector $K$. 
We rename vector $R$ according to the following: $K[i]$ is the fingerprint for letter $i$.
Since $K$ is monotonically increasing we compress it using Elias-Fano coding.
For pattern matching queries, we binary search on $K$ to find the letter representing a fingerprint.
\end{remark}

\cref{rem:V} and \cref{rem:H} can be tested first to see if they give any improvements.
