\section{Conclusion}

Our experiments demonstrate that the safe exploration framework we developed successfully trains a goal-conditioned policy while preventing mistakes during learning. We showed that incorporating a notion of distance to terminal states is crucial for safety and examined the impact of thresholds on both safety and goal-space coverage. As zero error over the course of safe exploration is our objective, we analyzed failure modes to identify ways to reduce mistakes. The primary cause of failures lies in insufficient pretraining of the safety policy in certain regions of the state space. Improvements could be achieved by considering disagreement between critics as an intrinsic motivation during pretraining or incorporating it into the action selection mechanism. 
Additionally, fine-tuning the safety policy during the safe exploration phase could expand the set of reachable states, further enhancing exploration. Also, we will focus on improving the evaluation procedure of the safety policy. Our experiments reveal that the key is not only to have a safety policy we trust but also to have a clear understanding of its limitations.

%Through our experiments, we have shown that the safe exploration framework we developed 
%is able to train a goal-conditioned policy and prevent mistakes while learning. 
%We have also shown that taking into account the estimated distance to terminal states is crucial to achieve 
%safety, and studied the effects of thresholds both on safety during exploration and goal-space coverage.
%As absolute zero error is virtually impossible, we conducted a study to explain failure modes
%and identify perspectives to reduce the number of mistakes. The main cause of failures appears to 
%be the pretraining, where the safety policy has not been trained enough on specific regions of the state 
%space. Taking into account disagreement between critics may lead to improvements, 
%as an intrinsic motivation during pretraining and/or as a new component of the action selection mechanism.
%Also one could finetune the safety policy during the pretraining phase. {\color{blue} AND WE WILL DO THAT IN FUTURE WORK}
%\textcolor{blue}{long term}
