\section{Oblivious Transfer from $\DP$ Inner Product}\label{sec:main_theorems}

In this section, we state our main theorems. We first state and prove the information theoretic case in \cref{sec:DP TO OT IT}, and then prove the computational case in \cref{sec:DP TO OT Comp}.

\subsection{Information-Theoretic Case}\label{sec:DP TO OT IT}
The following is our main theorem (for the information theoretic case) which shows that given a sufficiently accurate and private $\DP$ channel, we can construct a statistically secure semi-honest oblivious transfer protocol.

\begin{theorem}\label{thm:DPIP-to-OT}
There exist constants $c_1,c_2>0$ and an oracle-aided \ppt protocol $\Pi$ such that the following holds for large enough $n \in \bbN$ and for 
$\eps \leq \log^{0.9} n$, $\delta \leq \frac1{3n}$, and $\ell \leq e^{-c_1  \eps}  c_2\cdot n^{1/6}$:
    Let $C = ((X,U),(Y,V))$ be an $(\eps,\delta)$-\DP channel with independent $X,Y \la \oo^n$ that is $(\ell,0.999)$-accurate for the inner-product functionality (i.e., $\ppr{C}{\size{O-\ip{X,Y}} \leq \ell} \geq 0.999$).
    Then $\Pi^C$ is a semi-honest statistically secure $\OT$ protocol.
\end{theorem}

To prove Theorem \ref{thm:DPIP-to-OT}, we will make use of the following technical Lemma.


\begin{lemma}\label{lemma:DPIP-to-AWEC}
        There exists a \ppt protocol $\Gamma=(\Ac,\Bc)$ such that for every $c_1,c_2,n,\eps,\delta,\ell$ and $C$ as in \cref{thm:DPIP-to-OT},
        the channel $\tilde{C}$ induced by the execution of $\Gamma^C$ is $(\ell, \alpha=0.001, p=0.001, q=0.001)$-\AWEC (\cref{def:AWEC}).
\remove{
	There exist constant $c_1,c_2 > 0$ and an oracle-aided \ppt protocol $\Gamma=(\Ac,\Bc)$ such that the following holds for large enough $n \in \bbN$ and for 
	$\eps \leq \log^{0.9} n$, $\delta \leq \frac1{3n}$, and $\ell \leq e^{-c_1  \eps}  c_2\cdot n^{1/6}$:
	Let $C = ((X,U),(Y,V))$ be an $(\eps,\delta)$-\DP channel with independent $X,Y \la \oo^n$ that is $(\ell,0.999)$-accurate for the inner-product functionality (i.e., $\ppr{C}{\size{O-\ip{X,Y}} \leq \ell} \geq 0.999$). Then the channel $\tilde{C}$ induced by the execution of $\Gamma^C$ is $(\ell, \alpha=0.001, p=0.001, q=0.001)$-\AWEC (\cref{def:AWEC}). 
}
    Furthermore, the proof is constructive in a black-box manner:
	\begin{enumerate}
		\item There exists an oracle-aided \ppt algorithm $\Act$ such that for every channel $C = ((X,U),(Y,V))$ and algorithm $\Ac$ violating the \AWEC secrecy property~\ref{AWEC:item:A} of $\tilde{C}$ (the channel of $\Gamma^C$), the following holds for $Y^*_i = \Act^{\Ac}(i,\: Y_{-i}, \: X, \: U)$:\label{item:privacy-of-Y}
		\begin{align*}
			\eex{i \la [n]}{\pr{Y^*_i = Y_i }} > e^{\eps} \cdot \eex{i \la [n]}{\pr{Y^*_i = -Y_i }} + \delta.
		\end{align*}
		
		%\Enote{The above two items suffice for breaking $(\eps,\delta)$-DP because if we let $p = \pr{Y^*_i = Y_i}$ and $q = \pr{Y^*_i = -Y_i}$ then $p \leq e^{\eps} q + \delta$, and if $q \geq e^{-\eps} \delta$ then we get that
			%$$\frac{p}{p+q} \leq \frac{e^{\eps} q + \delta}{e^{\eps} q + \delta + q} = \frac{1}{1 + \frac{1}{e^{\eps} + \delta/p}} \leq  \frac{1}{1 + \frac{1}{2\cdot e^{\eps} }} \leq \frac{2\cdot e^{\eps}}{2\cdot e^{\eps}+1}.$$}
		
		%\item \Enote{We should use Theorem 5.1 from our previous paper (or Theorem 6.2 directly).}
		
		
		\item There exists an oracle-aided \ppt algorithm $\Bct$ such that for every channel $C = ((X,U),(Y,V))$ and algorithm $\Bc$ violating the \AWEC secrecy property~\ref{AWEC:item:B} of $\tilde{C}$ (the channel of $\Gamma^C$),  the following holds for $X^*_i = \Bct^{\Bc, C}(i,\: X_{-i}, \: Y, \: V)$:\label{item:privacy-of-X}
		\begin{align*}
			\eex{i \la [n]}{\pr{X^*_i = X_i }} > e^{\eps} \cdot \eex{i \la [n]}{\pr{X^*_i = -X_i }} + \delta.
		\end{align*} 
	\end{enumerate}
\end{lemma}

The proof of \cref{lemma:DPIP-to-AWEC} is given in \cref{sec:DPIP_to_WAEC}. But first we use \cref{lemma:DPIP-to-AWEC} to prove \cref{thm:DPIP-to-OT}. 

\begin{proof}[Proof of \cref{thm:DPIP-to-OT}]
By \cref{lemma:DPIP-to-AWEC} there exists a \ppt protocol $\Gamma\coloneqq\Gamma^C$ such that channel $\tilde{C}\coloneqq C_{\Gamma(\pk)}$ is a $(\ell, \alpha=0.001, p=0.001, q=0.001)$-\AWEC. By \cref{lemma:AWEC-to-WEC} (and \cref{prop:hard-to-guess-inf}) there exists a \ppt protocol $\Lambda\coloneqq\Lambda^{\tilde{C}}$ such that the channel $\hhC=C_{\Lambda(\pk)}$ is a $(\alpha'=0.002,\: p' = 0.001 ,\:  q' = 1/2 + 0.022))$-\WEC. Since $44(\eps'+p')<1-q'$, by  \cref{thm:WEC TO OT IT} there exists a \ppt protocol $\Pi$, such that $\Pi^{\hhC}$ is a semi-honest statistically secure \OT, concluding the proof.
\end{proof}


\subsection{Computational Case}\label{sec:DP TO OT Comp}


In this section, we state and prove our results for the computational case. We show that a \CDP (computational differential private) protocol that estimates the inner product well implies a semi-honest computationally secure oblivious transfer protocol. 


\begin{theorem}[Restatement of \cref{thm:intro:main}]\label{thm:CDPIP-to-OT}
There exist constant $c_1,c_2 > 0$ and an oracle-aided \ppt protocol $\Pi$ such that the following holds for large enough $n \in \bbN$ and for 
$\eps \leq \log^{0.9} n$, $\delta \leq \frac1{3n}$, and $\ell \leq e^{-c_1  \eps}  c_2\cdot n^{1/6}$:
    Let $\Psi$ be an $(\eps,\delta)$-\CDP protocol that is $(\ell,0.999)$-accurate for the inner-product functionality. 
    Then $\Pi^\Psi$ is a semi-honest computationally secure oblivious transfer protocol.
\end{theorem}

By the result of \cite{GoldreichMW87}, in the computational setting, we can ``compile" any semi-honest computational oblivious transfer protocol into a protocol that is secure against any \ppt (malicious) adversary (assuming one-way functions). We state this formally in the following corollary.

\begin{corollary}\label{cor:mal OT}
   Let $\eps,\delta,\ell$ be as in \cref{thm:CDPIP-to-OT}. If there exists a protocol $\Psi$ that is $(\eps,\delta)$-\CDP  and is $(\ell,0.999)$-accurate for the inner-product functionality, then there exists a computationally secure oblivious transfer protocol. 
\end{corollary}
\begin{proof}[Proof of \cref{cor:mal OT}]
By \cref{thm:CDPIP-to-OT}, there exists a semi-honest computational secure oblivious transfer protocol $\Pi$. Note that by \cite{ImpagliazzoLu89}, $\Pi$ implies the existence of one-way functions and by \cite{GoldreichMW87}, using the one-way function, we can compile $\Pi$ into a (computational) \OT secure against arbitrary adversaries.
\end{proof}


\paragraph{Proof of \cref{thm:CDPIP-to-OT}.}

 In order to use \cref{lemma:DPIP-to-AWEC}, and similar to \cite{HaitnerMST22}, we first convert the protocol $\Psi$ into a (no input) protocol such that the \CDP-channel it induces, is uniform and accurately estimates the inner-product functionality. Such a transformation is simply the following protocol that invokes $\Psi$ over uniform inputs, and each party locally outputs its input. 
\begin{protocol}[$\hPsi^{\Psi} = (\hAc,\hBc)$]\label{prot:EDPtoSV}
	\item Common input: $1^\kappa$.
	%	\item Oracle: protocol $\Psi=(\Ac,\Bc)$.
	\item Operation:
	\begin{enumerate}
		
		\item $\hAc$ samples $x \gets \oo^{\pn(\kappa)}$ and $\hBc$ samples $y\gets \oo^{\pn(\kappa)}$. 
		
		\item The parties interact in \remove{a random execution protocol }$\Psi(1^\kappa)$, with $\hAc$ playing the role of $\Ac$ with private input $x$, and $\hBc$ playing the role of $\Bc$ with private input $y$.
		
		\item $\hAc$ locally outputs $x$ and $\hBc$ locally outputs $y$. 
	\end{enumerate}
\end{protocol}

Let $C$ be the channel ensemble induced by $\hPsi$, letting its designated output (the function $\out$) be the designated output of the embedded execution of $\Psi$. The following fact is immediate by definition.

\begin{proposition}\label{prop:EDP to SV}
	The channel ensemble $C$ is $(\eps,\delta)$-$\CDP$, and has the same accuracy for computing the inner product as protocol $\Psi$ has.
\end{proposition}


We first prove the computational version of \cref{lemma:DPIP-to-AWEC}.
\begin{claim}\label{claim:Comp DP to AWEC}
    There exist constant $c_1,c_2 > 0$ and an oracle-aided \ppt protocol $\Gamma=(\Ac,\Bc)$ such that the following holds for large enough $n \in \bbN$ and for 
	$\eps \leq \log^{0.9} n$, $\delta \leq \frac1{3n}$, and $\ell \leq e^{-c_1  \eps}  c_2\cdot n^{1/6}$:
    Let $\Psi$ be an $(\eps,\delta)$-\CDP protocol that is $(\ell,0.999)$-accurate for the inner-product functionality. Then the protocol $\Gamma^{\Psi}$ is an  $(\ell, \alpha=0.001, p=0.001, q=0.001)$-\CompAWEC protocol.
\end{claim}
\begin{proof}[Proof of \cref{claim:Comp DP to AWEC}]
Recall that $\Psi$ is an $(\eps,\delta)$-$\DP$ protocol that is $(\ell,0.999)$-acccurate for the inner-product functionality for 
	$\eps \leq \log^{0.9} n$, $\delta \leq \frac1{3n}$, and $\ell \leq e^{-c_1  \eps}  c_2\cdot n^{1/6}$. By \cref{prop:EDP to SV}, $\hPsi\coloneqq\hPsi^\Psi$ induces a channel ensemble $C=\set{C_\pk=((X_\pk,U_\pk),(Y_\pk,V_\pk))}_{\pk\in\N}$ that is $(\eps,\delta)$-\CDP and $(\ell,0.999)$-acccurate for the inner-product functionality. Let $\Gamma$ be the \ppt protocol guarrented by \cref{lemma:DPIP-to-AWEC}, we now claim that $\Gamma^{\Psi}(\pk)\coloneqq\Gamma^{C_\pk}(\pk)$ is an $(\ell, \alpha=0.001, p=0.001, q=0.001)$-\CompAWEC protocol. %That is, the ensemble $\tC=\set{\tC_\pk=(X,V)(}_{\pk\in\N}$
Assume towards contradiction that this does not hold, then by \cref{lemma:DPIP-to-AWEC}, there exists a \ppt $\Dc$, such that for infinity many $\pk\in\N$, $\Dc$ contradicts one of the two secrecy properties of $\CompAWEC$. Fix such $\pk\in\N$ and omit it from the notation when clear from the context, and without loss of generality, assume that $\Dc$ contradicts the first secrecy property of $\CompAWEC$ (the case where $D$ contradicts the second property is essentially identical). By the first item of \cref{lemma:DPIP-to-AWEC}, there exists a \ppt algorithm $\tilde{\Ac}$ such that 
$Y^*_i = \Act^{\Dc}(i,\: Y_{-i}, \: X, \: U)$  and it holds that:
\begin{align*}
			\eex{i \la [n]}{\pr{Y^*_i = Y_i }} > e^{\eps} \cdot \eex{i \la [n]}{\pr{Y^*_i = -Y_i }} + \delta.
\end{align*}
Thus, we get a contradiction since by \cref{prop:hard-to-guess-comp}, algorithm $\Act^{\Dc}$ breaks the \CDP property of $C$.  
\end{proof}


\begin{proof}[Proof of \cref{thm:CDPIP-to-OT}]
Recall that $\Psi$ is an $(\eps,\delta)$-$\DP$ protocol that is $(\ell,0.999)$-acccurate for the inner-product functionality. By \cref{claim:Comp DP to AWEC} there exists a \ppt protocol $\Gamma$ such that $\Gamma\coloneqq\Gamma^\Psi$ is an $(\ell, \alpha=0.001, p=0.001, q=0.001)$-\CompAWEC protocol. By \cref{cor:CompAWEC to CompWEC}, there exists a \ppt protocol $\Lambda$ such that $\Lambda\coloneqq \Lambda^\Gamma$ is $(\epsilon'=\epsilon+0.001,\: p' = p ,\:  q' = 1/2 + 2(q+0.01))$-\CompWEC. Finally, since $44(\eps'+p')<1-q'$ by  \cref{thm:WEC TO OT Comp} there exists a \ppt protocol $\Pi$, such that $\Pi^{\Lambda}$ is a semi-honest computationally secure \OT, as required.
\end{proof}