\section{Related Work}
This section briefly introduces the progress of another two-stage recommendation paradigm, the cold-start recommendation for new publish product. The cold-start item recommendation problem, can be challenging due to the limited interactions between users and items~\cite{wu2022adversarial}. Cold-start item recommendation is typically divided into two stages by the amount of user interactions. 

In the first stage, the new items have just entered into the system and thus received very limited exposure. At this moment, any existing recommendation models may fail because of the minimal user interactions. To avoid direct competition with existing items, industrial recommendation systems may offer a separate channel only for new items. In this case, high-quality new items only compete with each other, and they can warm up and enter the next stage after obtaining a certain level of interaction. After that, items can obtain simple real-time statistical information or incorporate a limited amount of user and item features for recommendation. The common approaches ~\cite{cold_cv,cold_context,cold_cross_domain,cold_graph,cold_poi} utilize generalization features or explore information across different domains and modalities. The most effective algorithms, of course, are usually the ones that involve introducing additional user-item interaction data from other domains~\cite{zhang2022keep,rec4ad}.

In the second stage, new items may receive a few interactions from the separate channel, And now, they compete directly with the existing items. 
ClassBalance~\cite{class_re} and LogQ~\cite{logq} attempted to enhance the importance of new items during the training process by adjusting the weights of the samples or their corresponding network parameters. 
DropoutNet~\cite{dropoutnet} and POSO~\cite{poso} enhanced the expressive ability and robustness of models for new items by designing specialized networks tailored to their characteristics. There are indeed many other research studies on this topic, we do not list them all here due to the content limit.