\section{Related Work}
\label{subsec:furtherwork}
The goal of this paper is to use approaches from the \ac{SW} to build the graph model, therefore the related work section focuses on  
\ac{SW} technologies are being explored in many areas for data modelling and describing semantic relationships between entities and attributes. Much research has also focused on how this technology can be connected to smart grid research areas.
%\todo{CGMS abgernzung einbauen für Produktiv Systeme verschiedenenn usecases im berreich Forschung aandere anforderung an Daten oft deutlich geruuinger manchmal anders -> perspektifich interoperabilität}
Several approaches exist to build ontologies for power systems using \ac{RDF}.
For instance, ____ focuses on multi-domain energy systems, while ____ concentrates on micro-grid communities and on the service side of the system.
Both use \ac{SW} technologies, but not with the purpose of bridging different domains and enable cyber resilience research.
The proposed models do not consider a holistic framework for describing smart grid use cases with all associated layers in both cases. In particular, they lack the connection to the \ac{SGAM}, which brings the concept in line with a widely used way to describe use cases, making the models much more accessible for researchers.
____ uses the \ac{SGAM} as basis for an ontology designed for threat modelling.
Although this work involves modelling cross-domain inter-dependencies of cyber-physical systems, it does not utilise SHACL, which allows validation through rules.
In contrast to the work presented, our paper proposes an approach that provides a holistic picture that presents a complete framework and shows how \ac{SW} technologies can be useful in all steps of definition, description, validation, and evaluation.
One of these additional aspects is a rule-based augmentation for transforming power systems in \ac{CPES}.
____ presents an approach of graph-based modelling for \ac{CPES} and emphasises the need for an automated augmentation process. However, the work focuses only on a specific \ac{SCADA} use case and does not utilise \ac{SW} technologies. 

\input{methodology}
\input{workflow}
\input{casestudy}
\input{conclusion}