\section*{Limitations}
\label{subsec:limitation} 

\textbf{Limited Availability of Comic Strips:} Our dataset contains a relatively small number of samples due to the scarcity of standalone short-story comic strips available online. Most comics are either serialized narratives or dialog-driven, making it challenging to collect a diverse set of independent stories.

\textbf{Narrow Focus on Comics:} While our dataset captures the narrative structure of comic strips, it does not encompass the full spectrum of sequential visual storytelling, such as photographic sequences, instructional diagrams, or movie storyboards. Future work could extend beyond comics to explore a broader range of visual sequences.

\textbf{Limited Training Data for Fine-Tuning:} Our findings indicate that fine-tuning significantly enhances model performance on the reordering task. However, the limited availability of training data constrains the model’s ability to fully develop temporal reasoning skills. Expanding the dataset or incorporating alternative sources, such as video sequences, could further improve performance.

\section*{Ethics Statement}
\label{subsec:ethics} 
The datasets used in our experiment are publicly released and labeled through interaction with humans in English. In this process, user privacy is protected, and no personal information is contained in the dataset. The scientific artifacts that we used are available for research with permissive licenses. And the use of these artifacts in this paper is consistent with their intended use. Therefore, we believe that our research work meets the ethics of ACL. 