\begin{abstract}
    Calibration is a fundamental concept that aims at ensuring the reliability of probabilistic predictions by aligning them with real-world outcomes. 
    There is a surge of studies on new calibration measures that are easier to optimize compared to the classical $\ell_1$-Calibration while still having strong implications for downstream applications.
    One recent such example is the work by~\citet{fishelsonfull} who show that it is possible to achieve $\tilde{\cO}(T^{1/3})$ pseudo $\ell_{2}$-Calibration error via minimizing pseudo swap regret of the squared loss, which in fact implies the same bound for all bounded proper losses with a smooth univariate form. In this work, we significantly generalize their result in the following ways: (a) in addition to smooth univariate forms, our algorithm also simultaneously achieves $\tilde{\cO}(T^{1/3})$ swap regret for any proper loss with a twice continuously differentiable univariate form (such as Tsallis entropy); (b) our bounds hold not only for pseudo swap regret that measures losses using the forecaster's distributions on predictions, but also hold for the actual swap regret that measures losses using the forecaster's actual realized predictions.
    
    We achieve so by introducing a new stronger notion of calibration called \textit{(pseudo) KL-Calibration}, which we show is equivalent to the (pseudo) swap regret with respect to  log loss. 
    We prove that there exists an algorithm that achieves $\tilde{\cO}(T^{1/3})$ KL-Calibration error and provide an explicit algorithm that achieves $\tilde{\cO}(T^{1/3})$ pseudo KL-Calibration error.
    Moreover, we show that the same algorithm achieves {${\cO}(T^{1/3} (\log T) ^ {-\frac{1}{3}}\log (T/{\delta}))$} swap regret with probability at least $1-\delta$ for any proper loss with a smooth univariate form, which implies $\tilde{\cO}(T^{1/3})$ $\ell_2$-Calibration error.
    A technical contribution of our work is a new randomized rounding procedure and a non-uniform discretization scheme to minimize the swap regret for log loss.     
\end{abstract}