\documentclass[twoside]{article}

\usepackage{amsmath, amsfonts, amsthm}
\usepackage{nicefrac}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{siunitx}
\usepackage{natbib}
\usepackage[margin=1in]{geometry}

\usepackage{etoolbox}
\newtoggle{icml}
\togglefalse{icml}

\usepackage{comment}

\usepackage{wrapfig}
\usepackage{tabularray}
\usepackage[shortlabels]{enumitem}

\usepackage{thmtools}
\usepackage{thm-restate}
\declaretheorem[name=Theorem,numberwithin=section]{thm}
\declaretheorem[name=Proposition,numberwithin=section]{prop}
\newtheorem{theorem}{Theorem}
\numberwithin{theorem}{section}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}
\usepackage{mathtools}

\input{z-macros}

\usepackage[dvipsnames]{xcolor}   

\usepackage{hyperref}
\hypersetup{
colorlinks=true,
linkcolor=DarkOrchid,
filecolor=magenta,
urlcolor=Orchid,
citecolor=CadetBlue
}

\usepackage[algo2e,ruled,noend]{algorithm2e}
\usepackage[font=small,tableposition=top]{caption}


\title{From Individual Experience to Collective Evidence: \\
A Reporting-Based Framework for Identifying Systemic Harms 
}
\author{Jessica Dai, Paula Gradu, Inioluwa Deborah Raji, Benjamin Recht\\{\textit{University of California, Berkeley}}
}
\date{February 2025}

\begin{document}

\maketitle

\thispagestyle{empty}
\begin{abstract}
When an individual reports  a negative interaction with some system, how can their personal experience be contextualized within broader patterns of system behavior?
We study the \textit{incident database} problem, where individual reports of adverse events arrive sequentially, and are aggregated over time.
In this work, our goal is to identify whether there are subgroups---defined by any combination of relevant features---that are disproportionately likely to experience harmful interactions with the system.
We formalize this problem as a sequential hypothesis test, and identify conditions on reporting behavior that are sufficient for making inferences about disparities in true rates of harm across subgroups.
We show that algorithms for sequential hypothesis tests can be applied to this problem with a standard multiple testing correction.
We then demonstrate our method on real-world datasets, including mortgage decisions and vaccine side effects; on each, our method (re-)identifies subgroups known to experience disproportionate harm using only a fraction of the data that was initially used to discover them.
\end{abstract}

\input{1-intro}
\input{2-model}
\input{3-preponderance}
\input{4-algorithm}
\input{5-experiments}
\input{6-discussion}

\newpage
\section*{Acknowledgements}
We are grateful to Ian Waudby-Smith, Kevin Jamieson, and Robert Nowak for helpful discussions in developing this work. 

JD is supported in part by the National Science Foundation Graduate Research Fellowship Program under Grant No. DGE 2146752. Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation. JD also thanks the AI Policy Hub at UC Berkeley for funding support in the 2023-2024 academic year. BR is generously supported in part by NSF CIF award 2326498, NSF IIS Award 2331881, and ONR Award N00014-24-1-2531. IDR is supported by the Mozilla Foundation and MacArthur Foundation. 

\bibliographystyle{plainnat}
\bibliography{z-biblio}

\newpage
\appendix
\input{a1-proofs}
\section{Further practical considerations}
\label{subsec:practical}
\input{a0-practical}
\end{document}
