\section{Related Work}

We discuss two lines of related work: the text-to-image prompt stealing attacks and the evolutionary algorithms in LLMs.

% \subsection{Text-to-Image Generation}
% % Diffusion Model \ VLLMs(MLLMs). DALL-E 3\ Midjourney
% % 相关工作-> 应用->prompt重要性

\subsection{Text-to-Image Prompt Stealing Attack}
Prompt stealing attacks, or prompt extraction attacks, aim to infer the input from a model's output. A successful attack infringes on intellectual property and poses significant risks to prompt trading platforms in the era of LLMs. However, such attacks are more challenging in text-to-image generation due to the greater uncertainty in image generation compared to text. CLIP Interrogator employs CLIP~\citep{radford2021learning} to extract the subject and then selects phrases matching the target image from predefined sets~\citep{udo2023image}. ~\citet{shen2024prompt} fine-tunes 2 models to extract image subjects and modifiers separately, combining them for the attack. Building on this,~\citet{naseh2024iteratively} employ GPT-4V to iteratively optimize the prompt, resulting in higher quality. Unlike these works, EvoStealer targets the extraction of a generalizable prompt template, offering greater practical value compared to stealing individual prompts.



\subsection{Evolutionary Algorithms in LLMs}
% 最近，联结进化算法和大语言模型的研究在众多任务上都表现出了强大而稳定的性能。一些工作借助LLMs具备丰富的领域知识和强大的文本分析的特性，使用LLMs作为优化器加速进化算法的搜索过程，如复杂推理，可解释性等。另一方面，也有些工作借助进化算法稳定的特性，使用LLM生成质量更高的提示。在本篇文章中，我们使用进化算法逐步产生和多个样例图片相似度更高的风格描述词，实现prompt template的窃取。

Recent researches combining evolutionary algorithms with LLMs have demonstrated strong and stable performance across various tasks~\citep{yang2023large, liu2023large}. Some studies leverage the rich domain knowledge and powerful text analysis capabilities of LLMs to accelerate the search process in evolutionary algorithms, particularly in tasks involving complex reasoning~\citep{meyerson2024language,liu2024large,lange2024large,brahmachary2024large} and interpretability~\citep{chiquier2025evolving}. Conversely, some studies capitalize on the stability of evolutionary algorithms to utilize LLMs for generating higher-quality prompt words~\citep{xu2022gps,prasad2022grips,guo2023connecting,fernando2023promptbreeder}. In this paper, we use evolutionary algorithms to progressively generate style descriptors that closely resemble multiple example images, thereby achieving prompt template stealing.