\section{Related Work}
\label{sec:related-work}

The increasing interest in fallacy detection has led to the creation of datasets with different characteristics in recent years (Table~\ref{tab:previous-work}). 
Some works focused on the educational domain, either by collecting data through gamification____ or from online quizzes____, whereas others studied transcripts of political debates____. Fallacies in news articles have been explored by____ and____ in climate change discourse, and by____ for analyzing COVID-19-related misinformation. 
Some works examined social media content using Reddit comments____ or tweets____, with a special interest on political discourse____.
We also analyze fallacies on social media but tackle previously unexplored topics over a four-year time frame and deal with the Italian language. 

Regarding annotation, current datasets mainly provide coarse-grained labels at the text (post, paragraph), snippet, or text pair level. The only exceptions are span-level datasets by____ and____ which, however, do not foresee overlapping annotations. Inspired by work on propaganda detection____ and persuasion techniques detection____, our dataset instead provides span-level annotations with overlaps for a better model transparency. Previous datasets include up to 3.6K \emph{human}-annotated instances____ across 14 fallacies____. Instead, ours provides 11K human-labeled spans across 20 fallacy types.

Recent work highlighted the importance of considering human label variation____, i.e., genuine disagreement____, subjectivity and perspectives____, and multiple plausible answers____ as signal rather than noise. Yet, in fallacy detection all the labels from multiple annotators are typically aggregated____, selectively chosen____, or adjudicated through discussion or by an expert____. 
More broadly, label variation has been mostly studied at the post or token level, with only few exceptions analyzing span-level disagreement in argument annotation____. 
In our work, we resolve annotation errors while keeping genuine disagreement and multiple plausible answers, propose the first fallacy detection dataset at the span level with parallel annotations, and use human label variation during evaluation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%