\documentclass[11pt]{article}
\usepackage[bottom=1in,left=1in, right=1in, top=1in]{geometry}
\geometry{letterpaper}

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{hyperref}
 
% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% For theorems and such
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{thmtools, thm-restate}
\usepackage{notation}
\usepackage{enumitem}
\newtheorem{proposition}{Proposition}
\newtheorem{claim}{Claim}

% if you use cleveref..
\usepackage[capitalize,noabbrev]{cleveref}

% Todonotes is useful during development; simply uncomment the next line
%    and comment out the line below the next line to turn off comments
%\usepackage[disable,textsize=tiny]{todonotes}
\usepackage[textsize=tiny]{todonotes}

\usepackage{multirow}

\usepackage{times}
\usepackage{fancyhdr}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{natbib}
\usepackage{eso-pic}
\usepackage{forloop}
\usepackage{url}
\usepackage{subfigure}

\usepackage{notation}

\newcommand{\yrcite}[1]{\citeyearpar{#1}}
\renewcommand{\cite}[1]{\citep{#1}}

\usepackage{authblk}

\title{Edge-Colored Clustering in Hypergraphs: Beyond Minimizing Unsatisfied Edges}
\date{}
\author[1]{Alex Crane}
\author[2]{Thomas Stanley}
\author[1]{Blair D. Sullivan}
\author[2]{Nate Veldt}

\affil[1]{University of Utah}
\affil[2]{Texas A\&M University}

\begin{document}

\maketitle

\begin{abstract}
    We consider a framework for clustering edge-colored hypergraphs, where the goal is to cluster (equivalently, to \emph{color}) objects based on the primary {type} of multiway interactions they participate in. One well-studied objective is to color nodes to minimize the number of \textit{unsatisfied} hyperedges -- those containing one or more nodes whose color does not match the hyperedge color. We motivate and present advances for several directions that extend beyond this minimization problem. We first provide new algorithms for maximizing \textit{satisfied} edges, which is the same at optimality but is much more challenging to approximate, with all prior work restricted to graphs. We develop the first approximation algorithm for hypergraphs, and then refine it to improve the best-known approximation factor for graphs. We then introduce new objective functions that incorporate notions of balance and fairness, and provide new hardness results, approximations, and fixed-parameter tractability results.
\end{abstract}


\input{sections/intro}

\input{sections/max-ecc}

\section{Alternative ECC Objectives}
We now turn our attention to more general~\ECC{} objectives, as well as several which incorporate notions of balance or fairness with respect to edge colors.
\input{sections/generalized-mean}

\input{sections/color-fair}

\input{sections/protected-class}

\input{sections/experiments}

\newpage
\section*{Acknowledgements}
This work was supported in part by the Gordon \& Betty Moore Foundation under award GBMF4560 to Blair D. Sullivan, by the National Science Foundation under award IIS-1956286 to Blair D. Sullivan, and by the Army Research Office under award W911NF‐24-1-0156 to Nate Veldt.

\bibliography{refs.bib}
\bibliographystyle{plainnat}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\newpage
\appendix
\onecolumn

\input{appendix/maxecc-proofs.tex}
% \input{appendix/generalized-mean.tex}
\input{appendix/color-fair.tex}
\input{appendix/protected-class.tex}

\end{document}

