
\documentclass[10pt]{article} % For LaTeX2e
\usepackage[table]{xcolor} 
% \usepackage{tmlr}
% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{tmlr}
% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
\usepackage[preprint]{tmlr}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
% \input{math_commands.tex}

\usepackage{hyperref}
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
% \usepackage[table]{xcolor}         % colors
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumitem}
\usepackage{wrapfig}
\input{utils/ju_math}
\input{utils/math_commands}
\definecolor{Gray}{gray}{0.92}
\usepackage{tabularx,arydshln}
\usepackage{tablefootnote}
\usepackage{sidecap}

\newcommand{\js}[1]{\textcolor{red}{\textbf{Ju: #1}}}
\newcommand{\hy}[1]{\textcolor{red}{\textbf{HY: #1}}}


\title{A Baseline Method for Removing Invisible Image Watermarks using Deep Image Prior}

% Authors must not appear in the submitted version. They should be hidden
% as long as the tmlr package is used without the [accepted] or [preprint] options.
% Non-anonymous submissions will be rejected without review.

\author{\name Hengyue Liang \email liang656@umn.edu \\
      \addr Electrical and Computer Engineering\\
      University of Minnesota, Twin Cities
      \AND
      \name Taihui Li \email lixx5027@umn.edu \\
      \addr Computer Science and Engineering\\
      University of Minnesota, Twin Cities
      \AND
      \name Ju Sun \email jusun@umn.edu\\
      \addr Computer Science and Engineering\\
      University of Minnesota, Twin Cities}

% The \author macro works with any number of authors. Use \AND 
% to separate the names and addresses of multiple authors.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\def\month{MM}  % Insert correct month for camera-ready version
\def\year{YYYY} % Insert correct year for camera-ready version
\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version


\begin{document}
\maketitle
\begin{abstract}
Image watermarks have been considered a promising technique to help detect AI-generated content, which can be used to protect copyright or prevent fake image abuse. In this work, we present a black-box method for removing \emph{invisible} image watermarks, without the need of any dataset of watermarked images or any knowledge about the watermark system. Our approach is simple to implement: given a \emph{single} watermarked image, we regress it by deep image prior (DIP). We show that from the intermediate steps of DIP one can reliably find an evasion image that can remove invisible watermarks while preserving high image quality. Due to its unique working mechanism and practical effectiveness, we advocate including DIP as a baseline invasion method for benchmarking the robustness of watermarking systems. Finally, by showing the limited ability of DIP and other existing black-box methods in evading training-based \emph{visible} watermarks, we discuss the positive implications on the practical use of training-based \emph{visible}  watermarks to prevent misinformation abuse. 
Our code is publicly available at \url{https://github.com/HengyueL/DIP_Watermark_Evasion}. 
\end{abstract}

\input{sections/intro}
\input{sections/related_works}
\input{sections/Experiments}
\input{sections/Summary}

% \subsubsection*{Author Contributions}
% If you'd like to, you may include a section for author contributions as is done
% in many journals. This is optional and at the discretion of the authors. Only add
% this information once your submission is accepted and deanonymized. 

\subsubsection*{Acknowledgments}
Liang H. is partially supported by NIH fund R01NS131314. Sun J. is partially supported by NIH funds R01NS131314 and R01CA287413. The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health. The authors acknowledge the Minnesota Supercomputing Institute
(MSI) at the University of Minnesota for providing resources that contributed to the research results reported in this article.

\bibliography{main}
\bibliographystyle{tmlr}

\appendix
\input{sections/Appendix}

\end{document}
