\begin{abstract}
  Modern Integrated Development Environments (IDEs) offer automated refactorings to aid programmers in developing and maintaining software.
  However, implementing sound automated refactorings is challenging, as refactorings may inadvertently introduce name-binding errors or cause references to resolve to incorrect declarations.
  To address these issues, previous work by Sch\"afer et al.\@ proposed replacing concrete references with \emph{locked references} to separate binding preservation from transformation.
  Locked references vacuously resolve to a specific declaration, and after transformation must be replaced with concrete references that also resolve to that declaration.
  Synthesizing these references requires a faithful inverse of the name lookup functions of the underlying language.

  Manually implementing such inverse lookup functions is challenging due to the complex name-binding features in modern programming languages.
  Instead, we propose to automatically derive this function from type system specifications written in the Statix meta-DSL.
  To guide the synthesis of qualified references we use \emph{scope graphs}, which represent the binding structure of a program, to infer their names and discover their syntactic structure.

  We evaluate our approach by synthesizing concrete references for locked references in \javaTotalTests{} Java, \chocopyTotalTests{} ChocoPy, and \fgjTotalTests{} Featherweight Generic Java test programs.
  Our approach yields a principled language-parametric method for synthesizing references.
\end{abstract}
