\section{Related Works}
Graph-based cooperative games and the properties of core imputations within them have been extensively studied. The seminal work of ____ established that in the assignment game, the core is precisely characterized by the set of optimal dual solutions. A natural extension of the assignment game is the $b$-matching game, where each vertex can be matched multiple times up to a predefined capacity.

There are two versions of the $b$-matching game. ____ distinguish them as simple and non-simple $b$-matching games. The simple $b$-matching game is the edge-constrained version where each edge is to be matched at most once. ____ demonstrated that verifying whether an imputation belongs to the core in this setting is co-NP-complete. Our work focuses on the non-simple $b$-matching game, the edge-unconstrained version, where edges can be used multiple times. These games are often referred to as \textit{transportation games} (see ____), and we establish a similar hardness result in this setting. Notably, ____) proved that the core of transportation games is always non-empty.

Beyond $b$-matching games, another well-studied class of cooperative games is the \textit{minimum-cost spanning tree (MST) games}. ____ introduced MST games and showed that their core is always non-empty. Several methods have been developed to compute core imputations, including Bird’s rule(____) and fairness-based approaches (____, ____, ____, ____). However, verifying whether an imputation belongs to the core of an MST game is co-NP-hard (____).

A closely related class of cooperative games is the max-flow game, introduced by ____, where the core is also always non-empty. Yet, similar to MST games, testing whether an imputation belongs to the core is co-NP-hard (____).

Transportation games share structural properties with these games. ____ demonstrated that any optimal dual solution of the linear programming (LP) formulation provides a core imputation. However, unlike in the assignment game, these dual-based imputations, collectively called the \textit{Owen set}, do not fully characterize the core of $b$-matching games (____).

The Owen set, first introduced by ____ for linear production games, was later formalized in ____. It consists of core imputations derived from dual solutions, where each dual variable represents an agent's shadow price. More recently, ____ extended the Owen set framework to MST and max-flow games, noting that in these settings, dual variables do not directly correspond to shadow prices, requiring careful interpretation of dual-optimal solutions.

Finally, fairness concepts such as minimax, maximin, leximin, and leximax have been widely explored in cooperative game theory(____). ____ developed an efficient method for computing leximin/leximax fair core imputations for the assignment game. Later, ____ established that computing such imputations for MST and max-flow games is NP-hard, although these imputations can be efficiently computed within the Owen set.


% \todo[inline]{RG: Should we add discussion about Nucleolus? Discuss with Vijay.}