\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{diagbox}
\usepackage{tikz}
\usepackage{float}
\usepackage{array}
\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\newcommand{\spaarc}{SAAP}
\begin{document}

    \title{SAAP: Spatial awareness and Association based Prefetching of Virtual Objects in Augmented Reality at the Edge}

    \author{\IEEEauthorblockN{Nikhil Sreekumar, Abhishek Chandra, Jon Weissman}
    \IEEEauthorblockA{\textit{University of Minnesota, Twin Cities} \\
    MN, USA \\
    \{sreek012, chandra, weiss039\}@umn.edu}}
    % \and
    % \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
    % \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
    % \textit{name of organization (of Aff.)}\\
    % City, Country \\
    % email address or ORCID}
    % \and
    % \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
    % \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
    % \textit{name of organization (of Aff.)}\\
    % City, Country \\
    % email address or ORCID}
    % \and
    % \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
    % \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
    % \textit{name of organization (of Aff.)}\\
    % City, Country \\
    % email address or ORCID}
    % \and
    % \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
    % \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
    % \textit{name of organization (of Aff.)}\\
    % City, Country \\
    % email address or ORCID}
    % \and
    % \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
    % \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
    % \textit{name of organization (of Aff.)}\\
    % City, Country \\
    % email address or ORCID}
    % }

    \maketitle

    \begin{abstract}
        Mobile Augmented Reality (MAR) applications face performance challenges due to their high computational demands and need for low-latency responses.  Traditional approaches like on-device storage or reactive data fetching from the cloud often result in limited AR experiences or unacceptable lag. Edge caching, which caches AR objects closer to the user, provides a promising solution. However, existing edge caching approaches do not consider AR-specific features such as AR object sizes, user interactions, and physical location. This paper investigates how to further optimize edge caching by employing AR-aware prefetching techniques. We present \spaarc, a Spatial Awareness and Association-based Prefetching policy specifically designed for MAR Caches.  \spaarc{} intelligently prioritizes the caching of virtual objects based on their association with other similar objects and the user's proximity to them. It also considers the recency of associations and uses a lazy fetching strategy to efficiently manage edge resources and maximize Quality of Experience (QoE).

        Through extensive evaluation using both synthetic and real-world workloads, we demonstrate that \spaarc{} significantly improves cache hit rates compared to standard caching algorithms, achieving gains ranging from 3\% to 40\% while reducing the need for on-demand data retrieval from the cloud. Further, we present an adaptive tuning algorithm that automatically tunes \spaarc{} parameters to achieve optimal performance. Our findings demonstrate the potential of \spaarc{} to substantially enhance the user experience in MAR applications by ensuring the timely availability of virtual objects.
    \end{abstract}

    \begin{IEEEkeywords}
    Augmented reality, caching, prefetching, association, proximity, support, confidence, edge computing
    \end{IEEEkeywords}
    \vspace{-3mm}
    \input{tex/intro.tex}
    \input{tex/motive.tex}
    \input{tex/spaarc.tex}
    % \input{tex/tuning.tex}
    \input{tex/eval.tex}
    % \input{tex/discuss.tex}
    \input{tex/related.tex}
    \input{tex/conclude.tex}
    % \section*{Acknowledgment}

% \section*{References}

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv, refer}

\end{document}
