\section{RECOVER A SOLUTION}\label{sec:recover_solution}

We state our results and proof of recovering a solution.
\begin{theorem}\label{thm:recover_solution}
Given a matrix $A$ and $W$, each entry can be written using $O(n^{\gamma})$ bits for $\gamma >0$.
Given $A \in \R^{n \times n}, W \in \R^{n \times n}, 1 \leq k \leq n$ and $ \epsilon \in (0,0.1)$. 
Assume $W$ has $r$ distinct columns and rows. 
Assume $A \circ W$ has at most $r \cdot p$ distinct columns and at most $r \cdot p$ distinct rows. 
Let $\OPT$ be defined as Definition~\ref{def:opt}. 
Then, with high probability,  
one can output two matrices $U,V \in \R^{n \times k}$ in time $n^{1+\gamma} \cdot 2^{O (k^2 r / \epsilon )} $ such that 
\begin{align*}
    \| (UV^\top- A) \circ W \|_F^2 \leq(1+\epsilon) \OPT .
\end{align*}

Further, if we choose (1) $k^2 r = O(\log n /\log\log n)$, (2) $\epsilon \in (0,0.1)$ to be a small constant, (3) $p = n^{o(1)}$ and (4) $\gamma = o(1)$, then the running time becomes $n^{1+o(1)}$.
\end{theorem}


\begin{proof}

Here, we show how to recover an approximate solution, not only the value of $\OPT$.

The idea is to recover the entries of $U$ and $V$ one by one and use the algorithm from the previous section for the corresponding decision problem. We initialize the semialgebraic set to be
\begin{align*}
    S= \{x \in \R^l \mid q(x) \neq 0, p(x) \leq \Lambda q(x) \}.
\end{align*}
We start by recovering the first entry of $U$. We perform the binary search to localize the entry, which takes $ \log  ( 2^{n^{\gamma}} )$ invocations of the decision algorithm. For each step of binary search, we use Theorem~\ref{thm:jpt13} to determine whether the following semi-algebraic set $S$ is empty or not,
\begin{align*}
S \cap \{U_{1,1}(x) \geq \wh{U}_{1,1}^{-}, U_{1,1}(x) \leq \wh{U}_{1,1}^{+} \}.
\end{align*}
After that, we declare the first entry of $U$ to be any point in this interval.  Then, we add an equality constraint that fixes the entry of $\wh{U}$ to this value and add a new constraint into $S$ permanently, e.g., $S  \leftarrow S \cap \{U_{1,1}(x)=\wh{U}_{1,1} \}$. Next, we repeat the same with the second entry of $U$ and so on.

This allows us to recover a solution of cost at most $(1+ \epsilon) \OPT$ in time
\begin{align*}
n^{1+\gamma} \cdot p \cdot 2^{O (k^{2} r / \epsilon )} .
\end{align*}
If we choose $\gamma = o(1)$, $\epsilon = \Theta(1)$, $p =n^{o(1)}$ and $k^2 r = O(\log n / \log\log n )$, then the running time becomes $n^{1+o(1)}$.
Thus, we complete the proof.
\end{proof}

