\section{Future Research Directions}
\label{sec:data_gen}

\url{https://ladderlogicworld.com/plc-architecture/}
\url{https://www.yokogawa.com/es/library/resources/white-papers/what-are-the-roles-of-dcs-and-scada-in-digital-transformation/}


### References for Recent PLC Attacks: PIPELIAR and Others

Recent years have seen various sophisticated attacks on Programmable Logic Controllers (PLCs) and critical infrastructure. Here are some notable examples and references:

### 1. **Stuxnet (2010)**
- **Description**: Stuxnet targeted Siemens PLCs at Iran’s Natanz uranium enrichment facility, causing centrifuges to malfunction while reporting normal operations.
- **Impact**: Damaged around 1,000 centrifuges, setting back Iran’s nuclear program.
- **References**:
  - [Wikipedia on Stuxnet](https://en.wikipedia.org/wiki/Stuxnet)
  - [Malwarebytes on Stuxnet](https://www.malwarebytes.com/stuxnet)
  - [9News on Stuxnet](https://www.9news.com.au/technology/stuxnet-worm-took-out-iranian-nuclear-facility-2010/)

### 2. **Industroyer/CrashOverride (2016)**
- **Description**: Industroyer targeted Ukraine’s power grid using ICS protocols like IEC 104, causing temporary outages.
- **Impact**: Demonstrated potential to disrupt critical infrastructure.
- **References**:
  - [Dragos Report](https://dragos.com/resource/industroyer-crashoverride/)
  - [Wired Article](https://www.wired.com/story/crashoverride-malware/)

### 3. **Triton/Trisis (2017)**
- **Description**: Triton targeted Schneider Electric’s Triconex safety systems, aiming to manipulate SIS controllers.
- **Impact**: Highlighted the vulnerability of safety systems in critical infrastructure.
- **References**:
  - [FireEye Report](https://www.fireeye.com/blog/threat-research/2017/12/triton-attack-safety-instrumented-systems.html)
  - [Dragos Report](https://dragos.com/resource/trisis/)

### 4. **EKANS Ransomware (2020)**
- **Description**: EKANS (Snake) ransomware targeted ICS environments, including PLCs, by encrypting files and halting processes.
- **Impact**: Showed the growing trend of ransomware in critical infrastructure.
- **References**:
  - [Dragos Blog](https://dragos.com/resource/ekans-ransomware/)
  - [Sophos on EKANS](https://news.sophos.com/en-us/2020/06/29/snake-ransomware-and-ics-environments/)

### 5. **INCONTROLLER/PIPEDREAM (2022)**
- **Description**: A toolkit targeting ICS devices, including PLCs, capable of manipulating configurations and executing commands.
- **Impact**: Demonstrated advanced capabilities for disrupting critical operations.
- **References**:
  - [Dragos Report](https://dragos.com/resource/incontroller-pipedream/)
  - [CISA Alert](https://www.cisa.gov/uscert/ncas/alerts/aa22-103a)

### 6. **PIPELIAR (2023)**
- **Description**: A sophisticated malware targeting oil and gas infrastructure, aimed at disrupting operations.
- **Impact**: Highlighted the risks faced by the energy sector from targeted cyberattacks.
- **References**:
  - [Canadian Centre for Cyber Security](https://www.cyber.gc.ca/en/guidance/cyber-threat-canadas-oil-and-gas-sector)
  - [OTORIO Report on Cyberattacks](https://www.otorio.com/resources/blog/cyberattacks-on-operational-environments/)
  - [The Register on OT Ransomware](https://www.theregister.com/2023/07/21/ot_ransomware_analysis/)

### General Trends and Mitigation
- **Unauthorized Access**: Often due to weak authentication mechanisms.
- **Firmware Manipulation**: Involves injecting malicious firmware.
- **Denial of Service (DoS)**: Overloading PLCs to disrupt operations.
- **Command Injection**: Sending unauthorized commands to PLCs.
- **Mitigation Strategies**:
  - **Strong Authentication**: Implement robust authentication mechanisms.
  - **Regular Patching**: Keep systems updated with the latest security patches.
  - **Network Segmentation**: Isolate ICS networks from other networks.
  - **Monitoring and Logging**: Continuously monitor for suspicious activity.

By understanding these threats and implementing robust security measures, organizations can better protect their PLCs and critical infrastructure from evolving cyber threats.
