\section{Conclusion}
This survey systematically reviews 448 studies and 46 public datasets to advance deep learning-driven analysis of EEG/iEEG signals across seven neurological diagnostic tasks: seizure detection, sleep staging and disorder, major depressive disorder, schizophrenia, Alzheimer’s disease, Parkinson’s disease, and ADHD.
Our work establishes three foundational contributions: First, we unify fragmented methodologies across neurological conditions by standardizing data processing, model architectures, and evaluation protocols.
Second, we identify self-supervised learning as the most promising paradigm for multi-task neurodiagnosis, providing a comprehensive overview of pre-trained SSL frameworks and their advancements. Third, we introduce \href{https://github.com/ZJU-BrainNet/BrainBenchmark}{BrainBenchmark} to enhance reproducibility by integrating neurological datasets and universal models under standardized evaluations.

Looking back, the pursuit of universal models capable of learning from diverse, multimodal data reflects the field's growing ambition. It lays the groundwork for a new era of intelligent and adaptable healthcare systems. Over the past decades, significant progress in traditional methods has established a strong foundation for neurological diagnostics based on electrical brain signals. Key contributions include advances in signal preprocessing techniques, curating large-scale, well-annotated datasets, and developing deep learning architectures for specific tasks.
Building on this foundation, the integration of self-supervised pretraining marks a paradigm shift, enabling models to extract rich and meaningful representations from vast amounts of unlabeled, heterogeneous data.

Looking forward, the ultimate goal is to develop genuinely universal and adaptable frameworks capable of transcending individual tasks and datasets to address a broader range of neurological disorders. These advancements will pave the way for intelligent diagnostic tools that deliver precise, efficient, and accessible healthcare solutions globally, driving transformative progress in biomedical research and clinical applications.