\begin{abstract}
% 
Leveraging the effective visual-text alignment and static generalizability from CLIP, recent video learners adopt CLIP initialization with further regularization or recombination for generalization in open-vocabulary action recognition in-context.
% while compromising their generalizability—especially when encountering novel, out-of-context, out-of-distribution (OOD) data. 
% 
However, due to the static bias of CLIP, such video learners tend to overfit on shortcut static features, thereby compromising their generalizability, especially to novel out-of-context actions.
% especially when encountering novel and out-of-context videos, \ie out-of-distribution (OOD) data.
% 
To address this issue, we introduce \textbf{Open-MeDe}, a novel \underline{Me}ta-optimization framework with static \underline{De}biasing for \underline{Open}-vocabulary action recognition.
% 
From a fresh perspective of generalization, Open-MeDe adopts a meta-learning approach to improve ``\textbf{known-to-open generalizing}'' and ``\textbf{image-to-video debiasing}'' in a cost-effective manner.
% 
Specifically, Open-MeDe introduces a cross-batch meta-optimization scheme that explicitly encourages video learners to quickly generalize to arbitrary subsequent data via virtual evaluation, steering a smoother optimization landscape.
% 
In effect, the free of CLIP regularization during optimization implicitly mitigates the inherent static bias of the video meta-learner.
% 
We further apply self-ensemble over the optimization trajectory to obtain generic optimal parameters that can achieve robust generalization to both in-context and out-of-context novel data.
% 
Extensive evaluations show that Open-MeDe not only surpasses state-of-the-art regularization methods tailored for in-context open-vocabulary action recognition but also substantially excels in out-of-context scenarios.
\end{abstract}

% Leveraging the effective visual-text alignment and static generalizability from CLIP, recent video learners adopt CLIP initialization with further regularization or recombination for generalization in open-vocabulary action recognition in-context. However, due to the static bias of CLIP, such video learners tend to overfit on shortcut static features, thereby compromising their generalizability, especially to novel out-of-context actions. To address this issue, we introduce \$textbf{Open-MeDe}$, a novel Meta-optimization framework with static Debiasing for Open-vocabulary action recognition. From a fresh perspective of generalization, Open-MeDe adopts a meta-learning approach to improve $\textbf{known-to-open generalizing}$ and $\textbf{image-to-video debiasing}$ in a cost-effective manner. Specifically, Open-MeDe introduces a cross-batch meta-optimization scheme that explicitly encourages video learners to quickly generalize to arbitrary subsequent data via virtual evaluation, steering a smoother optimization landscape. In effect, the free of CLIP regularization during optimization implicitly mitigates the inherent static bias of the video meta-learner. We further apply self-ensemble over the optimization trajectory to obtain generic optimal parameters that can achieve robust generalization to both in-context and out-of-context novel data. Extensive evaluations show that Open-MeDe not only surpasses state-of-the-art regularization methods tailored for in-context open-vocabulary action recognition but also substantially excels in out-of-context scenarios.