\section{Related Work}
\label{sec: related work}

\textbf{Fairness in Algorithmic Hiring} Studies both before and after the adoption of online hiring platforms have consistently found discrimination in hiring outcomes based on race, gender, and age____.  ____ study the screening stage of algorithmic hiring algorithms. These studies connect legal perspectives from hiring law with algorithmic approaches to mitigating disparate impact. In fact, mitigating biases has been described as crucial to ``ethicality of the AI tool design''____. As far as proposed solutions, ____ suggest ``augmentation-based" interventions where AI-assisted decision can best achieve equitable outcomes. 

\textbf{Strategic Classification}
Strategic Classification was initially introduced in the form of a Stackelberg game in ____ to address the impact of manipulative tactics on classification problems. Since then, several works have provided a modified Strategic Manipulation game that model disparities in manipulation abilities____, which we drew upon in our work. Furthermore, we utilize techniques from ____ to describe ``random'' classifiers in light of non-deterministic strategic manipulations. 

\textbf{Behavior and Risks of Generative Models} While guidance counselors and career coaches alike now recommend using generative AI tools to help with application materials____ despite a plethora of risks highlighted by recent research. For example, LLMs have been shown to hallucinate which may mislead employers____ or memorize text which can result in unintended plagiarism____. These behaviors are hard to anticipate a priori; thus the benefits of applying LLMs to application materials may be stochastic. 

% Hallucinations
% Memorization
% Stochastic 
% Responsible usage policies 
% Lack of diversity

% maybe 5-ish
%citiations to reference - noise in random classification aka randomized classifier, Lee cohen & FP objective, 
% appendix: paragraph on?
% don't have to justify every citation
% related work in strategic mainpulation - last sentence is the one that is most similar, and then maybe one sentence about what we've done.
%\textbf{Generative Models in the Hiring Ecosystem}

% as a last sentence
% what work do we add to that is the most similar --- does this apply / group / social cost?
% e.g. thus far, no works have focused on the LLM

% LLM and Societal Challenges