\documentclass{article}

\newif\ifarxiv
\arxivtrue

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} 
\usepackage[colorlinks,
linkcolor=red,
anchorcolor=blue,
citecolor=blue
]{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

\usepackage{fancyhdr}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}



\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}

\usepackage{comment}


\usepackage[capitalize,noabbrev]{cleveref}
\usepackage[textsize=tiny]{todonotes}


\usepackage{color-edits}
\addauthor{Jeffrey}{orange}
\addauthor{Konstantina}{teal}
\newcommand{\jointccp}{joint conditional conformal prediction }
\newcommand{\wcovdev}{\operatorname{wCD}}


\input{preamble}
\usepackage{fullpage}


% \captionsetup[subfigure]{singlelinecheck=off}

\title{Kandinsky Conformal Prediction: \\
Beyond Class- and Covariate-Conditional Coverage\thanks{All authors are listed in alphabetical order.}}
\author{Konstantina Bairaktari\thanks{Khoury College of Computer Sciences, Northeastern University. \texttt{bairaktari.k@northeastern.edu}. Supported by NSF awards CNS-2232692 and CCF-2311649.}
        \and
        Jiayun Wu\thanks{Department of Computer Science and Technology,
        Tsinghua University.
        \texttt{jiayun.wu.work@gmail.com}} 
        \and
        Zhiwei Steven Wu\thanks{School of Computer Science,
        Carnegie Mellon University.
        \texttt{zhiweiw@cs.cmu.edu}} 
}
\date{}




\begin{document}

\maketitle

\begin{abstract}
% Conformal prediction is a framework to produce prediction sets that cover the true label with certain probabilities. Since prediction sets with marginal coverage guarantees can perform worse for individuals, previous work has studied their coverage conditioned on covariates or labels. In this paper, we aim for a stronger conditional guarantee for events not measurable to the joint of covariates and labels. For example, the event can be defined by protected attributes that are restricted from algorithmic decision making. Extending the group conditional coverage guarantees considered by Mondrian conformal prediction, we propose Kandinsky conformal prediction for both overlapping groups and fractional group membership. We formulate the general group conditional coverage by weighted coverage deviation, which also implies valid coverage under subpopulation shifts. We study a simple quantile regression algorithm, which recovers (covariate-based) group conditional, class conditional, and Mondrian conformal prediction as special cases. We obtain a training conditional and group conditional coverage bound with a $\mathcal O(\sqrt{d\cdot \log n/n})$ rate of convergence to the target coverage. When groups are determined by covariates, our bound is near optimal up to a logarithmic factor. We perform experiments on real datasets for income prediction and toxic comment detection.
Conformal prediction is a powerful distribution-free framework for constructing prediction sets with coverage guarantees. Classical methods, such as split conformal prediction, provide \emph{marginal coverage}, ensuring that the prediction set contains the label of a random test point with a target probability. However, these guarantees may not hold uniformly across different subpopulations, leading to disparities in coverage. Prior work has explored coverage guarantees conditioned on events related to the covariates and label of the test point. 
We present \emph{Kandinsky conformal prediction}, a framework that significantly expands the scope of conditional coverage guarantees. In contrast to Mondrian conformal prediction, which restricts its coverage guarantees to disjoint groups—reminiscent of the rigid, structured grids of Piet Mondrian’s art—our framework flexibly handles overlapping and fractional group memberships defined jointly on covariates and labels, reflecting the layered, intersecting forms in Wassily Kandinsky’s compositions. Our algorithm unifies and extends existing methods, encompassing covariate-based group conditional, class conditional, and Mondrian conformal prediction as special cases, while achieving a minimax-optimal high-probability conditional coverage bound.  Finally, we demonstrate the practicality of our approach through empirical evaluation on real-world datasets.

% \Konstantinacomment{The submission instruction say that the abstract should be one paragraph.}


% In this work, we propose \emph{Kandinsky conformal prediction}, which extends the disjoint group conditional coverage framework of Mondrian conformal prediction to accommodate overlapping groups and fractional group memberships. We define this general group conditional coverage objective as a weighted coverage deviation and introduce a simple quantile regression algorithm that unifies and recovers existing methods, including covariate-based group conditional, class conditional, and Mondrian conformal prediction, as special cases. For this algorithm, we establish a high-probability conditional coverage bound with a convergence rate of $\mathcal{O}(\sqrt{d \cdot \log n / n})$, where $n$ is the number of calibration samples and $d$ is the number of groups. When groups are determined by covariates, this bound is optimal, up to a logarithmic factor. To demonstrate the practicality of our approach, we evaluate its performance on real-world datasets, highlighting its ability to achieve improved group conditional coverage.

%%OLD 
% Conformal prediction is a framework for constructing distribution-free prediction sets that cover the true label with a prespecified probability. Since prediction sets with marginal coverage guarantees can perform worse for some subpopulations, previous work has studied their coverage conditioned on events that depend on the covariates or the label of the test point. 
% \Jeffreydelete{In this paper, we aim for stronger conditional guarantees for events not measurable with respect to the joint of covariates and labels.}
% By extending the disjoint group conditional coverage guarantee considered in Mondrian conformal prediction, we propose Kandinsky conformal prediction that addresses both overlapping groups and fractional group membership. We formulate this general group conditional coverage objective as a weighted coverage deviation and study a simple quantile regression algorithm that recovers known algorithms for (covariate-based) group conditional, class conditional, and Mondrian conformal prediction as special cases. For this algorithm, we obtain a high-probability conditional coverage bound with a $\mathcal O(\sqrt{d\cdot \log n/n})$ convergence rate, where $n$ is the number of calibration data and $d$ is the number of groups. When the groups are determined by the covariates, our bound is nearly optimal up to a logarithmic factor. We evaluate our algorithm by performing experiments on real-world datasets.
\end{abstract}


\input{intro}

\input{prel}

\input{algo}

\input{experiments}

\clearpage

\ifarxiv
\else
\section*{Impact Statement}
In this paper, we introduce Kandinsky Conformal Prediction, which extends conformal prediction to provide coverage guarantees across a broader class of overlapping and fractional groups. This framework addresses fundamental challenges in fairness and distribution shift, ensuring robust, well-calibrated uncertainty quantification across diverse subpopulations. By maintaining computational efficiency, our method strengthens the technical foundation for reliable predictive modeling in real-world applications.
\fi

\bibliography{references}
\bibliographystyle{plainnat}

\newpage
\appendix

\input{appendix}

\end{document}
