% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
% \usepackage[review]{acl}
\usepackage[final]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}

%Including images in your LaTeX document requires adding
%additional package(s)
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables
% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2025} with \usepackage[nohyperref]{icml2025} above.
\usepackage{hyperref}
% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}
% For theorems and such
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{url}            % simple URL typesetting
\usepackage{bm}
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{adjustbox}
\usepackage{xcolor}         % colors
\usepackage{blindtext}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{pifont}
\usepackage{epsfig} % for postscript graphics files
\usepackage{times}
\usepackage{color}
% \usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{makecell}
\usepackage{algorithm}
\usepackage{rotating}
\usepackage{subfigure}
\usepackage[T1]{fontenc}
\usepackage{algorithm}
\usepackage{algorithmic}
% \usepackage[noend]{algpseudocode}
\usepackage{colortbl}
\usepackage{color, soul}
\usepackage{arydshln} % For dashline
\usepackage{tcolorbox}
\definecolor{babyblue}{rgb}{0.54, 0.81, 0.94}
\definecolor{bisque}{rgb}{1.0, 0.89, 0.77}
\definecolor{bshade}{rgb}{0.55,0.75,0.95}
\def\eg{{\it e.g.}\xspace}
\def\ie{{\it i.e.}\xspace}
\definecolor{mygray}{gray}{.6}
\definecolor{myblue}{RGB}{89,158,254}
\definecolor{mygreen1}{RGB}{81,150,111}
\definecolor{mygreen2}{RGB}{93,174,86}
\definecolor{myred}{RGB}{160,0,0}
\definecolor{myyellow}{RGB}{227,207,87}
% For pictures
\usepackage{caption}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage{threeparttable}
\usepackage{wrapfig}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\definecolor{babyblue}{rgb}{0.54, 0.81, 0.94}
\definecolor{bisque}{rgb}{1.0, 0.89, 0.77}
\definecolor{bshade}{rgb}{0.55,0.75,0.95}
\def\eg{{\it e.g.}\xspace}
\def\ie{{\it i.e.}\xspace}
\definecolor{mygray}{gray}{.6}
\definecolor{myblue}{RGB}{89,158,254}
\definecolor{mygreen1}{RGB}{81,150,111}
\definecolor{mygreen2}{RGB}{93,174,86}
\definecolor{myred}{RGB}{160,0,0}
\definecolor{myyellow}{RGB}{227,207,87}
\usepackage{bbding}
\newcommand{\cmark}{\ding{52}}%
\newcommand{\xmark}{\ding{55}}%
\usepackage{enumitem}
\let\oldding\ding% Store old \ding in \oldding
\renewcommand{\ding}[2][1]{\scalebox{#1}{\oldding{#2}}}
% Todonotes is useful during development; simply uncomment the next line
%    and comment out the line below the next line to turn off comments
%\usepackage[disable,textsize=tiny]{todonotes}
\usepackage[textsize=tiny]{todonotes}
\newcommand{\ceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\round}[1]{\left\lfloor #1 \right\rceil}
\newcommand{\sign}{\mathrm{sign}}
\newcommand{\tofloat}[1]{\mbox{float}\left(#1\right)}
\newcommand{\toint}[2]{\mbox{int}_{#1}\left(#2\right)}


\newcommand{\methodname}{GSQ-Tuning}
\newcommand{\dataname}{GSE}
\newcommand{\syz}[1]{\textcolor{teal}{SYZ:#1}}
% SamJ's commands
\input{text/command}


% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{GSQ-Tuning: \underline{G}roup-\underline{S}hared Exponents Integer in \\ Fully \underline{Q}uantized Training for LLMs On-Device Fine-tuning}


\author{
 \textbf{Sifan Zhou\textsuperscript{1,2$\dagger$$\ddagger$}},
 \textbf{Shuo Wang\textsuperscript{2$\dagger$}},
 \textbf{Zhihang Yuan\textsuperscript{2$\dagger$}},
 \textbf{Mingjia Shi\textsuperscript{2}},
 \textbf{Yuzhang Shang\textsuperscript{3}},
 \textbf{Dawei Yang\textsuperscript{2}}
\\
\\
\\
 \textsuperscript{1}Southeast University,
 \textsuperscript{2}Houmo AI,
 \textsuperscript{3}Illinois Institute of Technology
\\
 % \small{
   % \textbf{Correspondence:} \href{mailto:email@domain}{email@domain}
 % }
}

\begin{document}
\maketitle
\def\thefootnote{$\dagger$}\footnotetext{Equal contribution}
\def\thefootnote{$\ddagger$}\footnotetext{Work down as an intern at Houmo AI}
% \def\thefootnote{$\Diamond$}\footnotetext{Corresponding author}
\input{text/0_abs}
\input{text/1_intro}
\input{text/3_method}
\input{text/4_exp}
\input{text/2_related}
\input{text/5_con}

% \newpage
\section{Limitations and Future Work}

\paragraph{Limitations} While our GSQ-Tuning significantly advances on-device LLM adaptation through integer-focused optimization and parameter-efficient quantization, two key limitations warrant discussion:

\paragraph{Non-linear Operator Precision.} Our current implementation maintains non-linear operations (e.g., LayerNorm, Softmax) in 16-bit to preserve numerical stability. This introduces partial precision conversion overhead during computation. However, non-linear operations do not contain additional learnable parameters and thus do not consume memory. Moreover, these non-linear operations are generally computation-light, making their computational burden negligible. Future work could explore fully integer implementations for non-linear layers. \textbf{Bit-Width Range Constraints.}
The current framework operates effectively in 5-8bit configurations but didn't present the performance at extreme low bit ($\leq$ 4bit) precision. This stems from gradient direction distortion under extreme quantization—a challenge requiring new error compensation mechanisms. We plan to investigate two directions: (1) 4bit stochastic rounding with gradient-aware scaling, and (2) mixed-precision adapters allocating higher bits to critical gradient dimensions.

\paragraph{Future Work} Furthermore, future work could explore (1) full integer fine-tuning, (2) extreme low-bit quantized fine-tuning and (3) co-design with emerging integer-optimized AI accelerators. Our code will be publicly available to advance edge LLMs research.
% \section*{Acknowledgments}
\clearpage
\bibliography{custom}

\input{text/6_append}

\end{document}
