\section{Related Works}
\label{sec:related_works}
\textbf{Advanced Task Representation Learning:}Advanced representation learning techniques have been widely explored to improve task latents that effectively distinguish tasks. Recent meta-RL methods use contrastive learning ____ to enhance task differentiation through positive and negative pairs, improving task representation ____ and capturing task information in offline setups ____. The Bisimulation metric ____ is employed to capture behavioral similarities ____ and group similar tasks ____. Additionally, skill representation learning ____ addresses non-parametric meta-RL challenges ____, while task representation learning is increasingly applied in multi-task setups ____.

\noindent \textbf{Generalization for OOD Tasks:} Meta-RL techniques for improving policy generalization in OOD test environments have been actively studied ____. Model-based approaches ____, advanced representation learning with Gaussian Mixture Models ____, and Transformers ____ have been explored. Additionally, some studies tackle distributional shift challenges through robust learning ____.


\noindent \textbf{Model-based Sample Relabeling:} Model-based sample generation and relabeling techniques have gained attention in meta-RL ____, enabling the reuse of samples from other tasks using dynamics models ____. These methods address sparse rewards ____, mitigate distributional shifts in offline setups ____, and incorporate human preferences ____ or guided trajectory relabeling ____, expanding their applications.