\section{Related Works}
\label{sec:related_works}

\subsection{Weight-Space Learning}

While neural networks can learn effective representations for many traditional data modalities, effective representations for neural networks are still work in progress. As a first step, ____ proposed to observe simple statistics of weights, and  ____ on them. Others proposed to encode the weights by modeling the connections between the neurons ____. Recent methods ____ model a network as a graph where every neuron is a node, and train permutation-equivariant architectures ____ on these graphs.
Probing is an alternative paradigm that encodes the network by observing its outputs to a fixed set of inputs (probes)  ____. Differently from these approaches, we propose a probing-based method for zero-shot classification model search.

\subsection{Other Applications of Model Weights}

Learning on model weights has found many applications. Several approaches demonstrated advanced generation abilities using the weights ____, and others proposed to compress the weights to a smaller, more compact representation ____. A different line of research explored the relations between the weights for recovering the model graph ____ or for merging ____. Recently, a few works proposed to recover the exact black-boxed weights ____ by having access to their fine-tuned versions or to an API. Finally, some relevant works search for new adapters for generative models ____, however these approaches either rely on available metadata or tailored for generative models. Here, we propose an approach to search for new discriminative models which are capable of detecting a specific concept among other unrelated concepts seen in training time.