\begin{abstract}
    This paper introduces the Safe Protective and Assistive Robot Kit (\spark), a comprehensive benchmark designed to ensure safety in humanoid autonomy and teleoperation.
    Humanoid robots pose significant safety risks due to their physical capabilities of interacting with complex environments.
    % As such, robust safety measures are essential for both research and deployment in humanoid robots.
    The physical structures of humanoid robots further add complexity to the design of general safety solutions.
    To facilitate safe deployment of complex robot system, \spark can be used as a toolbox that comes with state-of-the-art safe control algorithms in a modular and composable robot control framework.
    Users can easily configure safety criteria and sensitivity levels to optimize the balance between safety and performance.
    To accelerate humanoid safety research and development, \spark provides simulation benchmark that compare safety approaches in a variety of environments, tasks, and robot models.
    % We establish benchmarks that can serve as a foundation for future research. 
    Furthermore, \spark allows quick deployment of synthesized safe controller on real robots.
    For hardware deployment, \spark supports Apple Vision Pro (AVP) or a Motion Capture System as external sensors, while offering interfaces for seamless integration with alternative hardware setups at the same time.
    This paper demonstrates \spark's capability with both simulation experiments and case studies with a Unitree G1 humanoid robot.
    Leveraging these advantages of \spark, users and researchers can significantly improve the safety of their humanoid systems as well as accelerate relevant research.
    The open source code is available at \href{https://github.com/intelligent-control-lab/spark}{{https://github.com/intelligent-control-lab/spark}}.
\end{abstract}