%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Template for USENIX papers.
%
% History:
%
% - TEMPLATE for Usenix papers, specifically to meet requirements of
%   USENIX '05. originally a template for producing IEEE-format
%   articles using LaTeX. written by Matthew Ward, CS Department,
%   Worcester Polytechnic Institute. adapted by David Beazley for his
%   excellent SWIG paper in Proceedings, Tcl 96. turned into a
%   smartass generic template by De Clarke, with thanks to both the
%   above pioneers. Use at your own risk. Complaints to /dev/null.
%   Make it two column with no page numbering, default is 10 point.
%
% - Munged by Fred Douglis <douglis@research.att.com> 10/97 to
%   separate the .sty file from the LaTeX source template, so that
%   people can more easily include the .sty file into an existing
%   document. Also changed to more closely follow the style guidelines
%   as represented by the Word sample file.
%
% - Note that since 2010, USENIX does not require endnotes. If you
%   want foot of page notes, don't include the endnotes package in the
%   usepackage command, below.
% - This version uses the latex2e styles, not the very ancient 2.09
%   stuff.
%
% - Updated July 2018: Text block size changed from 6.5" to 7"
%
% - Updated Dec 2018 for ATC'19:
%
%   * Revised text to pass HotCRP's auto-formatting check, with
%     hotcrp.settings.submission_form.body_font_size=10pt, and
%     hotcrp.settings.submission_form.line_height=12pt
%
%   * Switched from \endnote-s to \footnote-s to match Usenix's policy.
%
%   * \section* => \begin{abstract} ... \end{abstract}
%
%   * Make template self-contained in terms of bibtex entires, to allow
%     this file to be compiled. (And changing refs style to 'plain'.)
%
%   * Make template self-contained in terms of figures, to
%     allow this file to be compiled. 
%
%   * Added packages for hyperref, embedding fonts, and improving
%     appearance.
%   
%   * Removed outdated text.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Minor updates for SOUPS 2019 by Michelle Mazurek
%%% Minor updates for SOUPS 2022 by Rick Wash

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix2025_SOUPS}

% to be able to draw some self-contained figs
\usepackage{tikz}
\usepackage{amsmath}
\PassOptionsToPackage{table,xcdraw}{xcolor} 

\usepackage{textgreek}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{xcolor,colortbl}
\definecolor{gray}{rgb}{0.1,0.1,0.1}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage[flushleft]{threeparttable}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{longtable}
\usepackage{threeparttablex}

% inlined bib file
\usepackage{filecontents}

%-------------------------------------------------------------------------------
\begin{filecontents}{references.bib}
%-------------------------------------------------------------------------------
\end{filecontents}

\input{sections/header}


%-------------------------------------------------------------------------------
\begin{document}
%-------------------------------------------------------------------------------

%don't want date printed
\date{}

\title{\Large \bf Trust-Enabled Privacy:\\Social Media Designs to Support Adolescent User Boundary Regulation}

% if you leave this blank it will default to a possibly ugly attempt 
% to make the contents of the \author command below into a string
\def\plainauthor{Author name(s) for PDF metadata. Don't forget to anonymize for submission!}

%for single author (just remove % characters)
\author{
{\rm JaeWon Kim}\\
University of Washington
\and
{\rm Robert Wolfe}\\
University of Washington
\and
{\rm Ramya Bhagirathi Subramanian}\\
Smarten Spaces
\and
{\rm Mei-Hsuan Lee}\\
National Renewable Energy Laboratory
\and
{\rm Jessica Colnago}\\
{  }
\and
{\rm Alexis Hiniker}\\
University of Washington
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
} % end author

\maketitle
% \thecopyright

\begin{abstract}
Through a three-part co-design study involving 19 teens aged 13–18, we identify key barriers to effective boundary regulation on social media, including ambiguous audience expectations, social risks associated with oversharing, and the lack of design affordances that facilitate trust-building. Our findings reveal that while adolescents seek casual, frequent sharing to strengthen relationships, existing platform norms and designs often discourage such interactions, leading to withdrawal. To address these challenges, we introduce \textit{trust-enabled privacy} as a design framework that recognizes trust---whether building or eroding---as central to boundary regulation. When trust is supported, boundary regulation becomes more adaptive and empowering; when it erodes, users default to self-censorship or withdrawal. We propose concrete design affordances, including guided disclosure, contextual audience segmentation, intentional engagement signaling, and trust-centered norms, to help platforms foster a more dynamic and nuanced privacy experience for teen social media users. By reframing privacy as a trust-driven process rather than a rigid control-based trade-off, this work provides empirical insights and actionable guidelines for designing social media environments that empower teens to manage their online presence while fostering meaningful social connections.
\end{abstract}

\input{sections/1_introduction}
\input{sections/2_related-work}
\input{sections/3_method}
\input{sections/4-1_results1}
\input{sections/4-2_results2}
\input{sections/4-3_results3}
\input{sections/5_discussion}
\input{sections/6_conclusion}
\input{sections/7_appendix}

\section*{Acknowledgements}
The authors would like to acknowledge the CERES network, the University of Washington Global Innovation Funds (GIF), and the University of Washington Student Technology Funds (STF), which provided support for this work. We additionally thank the anonymous reviewers for their detailed feedback and the participants for sharing their thoughts. We truly appreciate all the help. Alexis Hiniker is a special government employee for the Federal Trade Commission. The content expressed in this manuscript does not reflect the views of the Commission or any of the Commissioners.

%-------------------------------------------------------------------------------
\bibliographystyle{plain}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%  LocalWords:  endnotes includegraphics fread ptr nobj noindent
%%  LocalWords:  pdflatex acks
