\section{Related Work}
Interactive drama driven by computer systems has been proposed for quite a long time ____.
More recently, ____ first discuss LLM-based interactive drama, defining its six primary elements. The director-actor architecture is first used in ____, while ____ model a single character with an ego and a superego to capture its development. In our work, we employ a director agent to program the plots periodically. ____ propose a curriculum learning method to fine-tune the role agents. Our work, in parallel, proposes a comprehensive framework suitable for any LLMs.

\paragraph{AI-Generated Stories}
It has been widely used in recent years to scale the training performance by using AI-generated data ____.
Our research focuses on the use of LLMs to assist humans in creating high-quality literature, such as poetry ____, music ____, stories ____. Our method generates a story given a short paragraph. This is more challenging than generating the drama script given a complete story ____.

\paragraph{Simulating Dramatic Characters with LLM Agents}
Recent years emerges many companion-oriented AI applications such as \texttt{character.ai}, where LLM agents are developed to simulate various characters ____. A new request arises, players are not satisfied with aimless conversations, but would like to go through a story, an adventure with AI, indicating the very need for our research, which connects NLP with research in psychology ____, art ____, and narratology ____.
Dramatic characters evolve over time, undergoing shifting inner motivations, which greatly distinguishes them from conversational agents.