\section{Related Work}
\paragraph{Existing Lattice-Based Hardness Assumptions}  
Lattice-based cryptography has gained widespread attention as a post-quantum alternative due to its worst-case hardness guarantees and efficiency in cryptographic applications. The Learning with Errors (LWE) problem____ was first introduced as a computationally hard assumption with worst-case reductions to GapSVP, forming the basis for lattice-based cryptographic schemes. However, LWE suffers from large ciphertext sizes and computational inefficiencies. To address these limitations, Ring-LWE (RLWE)____ was introduced, leveraging polynomial rings to improve efficiency while preserving security reductions to worst-case ideal lattice problems. Subsequent extensions such as Module-LWE (MLWE)____ and NTRU-based cryptosystems____ have further explored structured variants of LWE to balance security and performance. Recent standardization efforts, such as Kyber____ and Dilithium____,have demonstrated the practicality of RLWE-based schemes for key exchange and digital signatures. More recent studies____ have investigated the concrete security of LWE and its variants, refining worst-case reductions and assessing attack resistance under various parameter settings. A recent work by Peikert and Pepin____ introduces a unified framework that systematically connects various algebraically structured LWE variants, including Ring-LWE, Module-LWE, Polynomial-LWE, Order-LWE, and Middle-Product LWE. The proposed Variety-LWE (V-LWE) departs from the above unified paradigm by introducing a security assumption that does not solely depend on lattice structure: we will show that attacks relying on ideal-lattice reductions, such as hybrid Gröbner basis and lattice-reduction methods, do not trivially extend to V-LWE.

\paragraph{Multivariate Polynomial Rings in Cryptography}  
The use of multivariate polynomial rings in cryptography has primarily been explored in the context of multivariate public key cryptosystems (MPKCs)____. However, these approaches do not incorporate LWE-style noise distributions, which are crucial for worst-case to average-case reductions in lattice-based cryptography. Module-LWE____ extends RLWE by introducing module structures, but it remains within the single-variable polynomial framework and does not fully exploit the algebraic richness of multivariate polynomial rings. Newer works____ examine alternative structured lattice assumptions, providing additional perspectives on the interplay between multivariate polynomials and lattice-based security. Our work introduces \textit{algebraic varieties} as a component of an LWE-based cryptographic assumption, extending structured lattice problems into a multivariate setting.

\paragraph{Security Reductions and Structured Lattice Problems}  
The security of structured lattice problems relies on worst-case hardness assumptions. The Ideal Shortest Vector Problem (Ideal-SVP)____ has been extensively studied in the context of RLWE and Module-LWE, providing a foundation for their security reductions. Hybrid algebraic-lattice attacks, combining Gröbner basis techniques with lattice reduction algorithms, have been explored in cryptanalysis____, particularly in the context of structured ring-based schemes. While Gröbner basis methods have shown effectiveness in attacking certain algebraic cryptosystems____, their direct application to lattice-based cryptosystems remains constrained due to high computational costs____. Our work explores the impact of hybrid algebraic-lattice attacks on Variety-LWE, demonstrating that existing attack methodologies do not directly compromise its security.

\paragraph{Homomorphic Encryption and Vector Encryption}  
Fully homomorphic encryption (FHE) schemes, such as BFV____, BGV____, and CKKS____, enable secure computation on encrypted data but primarily operate within single-variable RLWE-based constructions. Batching methods using the Chinese Remainder Theorem (CRT) allow vectorized computations____, but they are constrained by modulus structure and require additional encoding layers. Our work introduces a direct multivariate polynomial structure that inherently supports vector encryption, providing an alternative approach to efficiently processing encrypted high-dimensional data. This approach aligns with research in privacy-preserving machine learning____, encrypted search____, and secure federated learning____. 


\paragraph{Attacks at Multivariate Polynomial LWE}
Bootland et al.____ proposed an attack on the multivariate Ring Learning with Errors (m-RLWE) problem in their 2020 paper ____. This attack exploits algebraic relationships between the roots of the defining polynomials in the multivariate ring to reduce the m-RLWE problem to multiple lower-dimensional RLWE problems. However, this attack is not applicable to the proposed Variety-LWE due to key distinctions between the two problems.
Firstly, Variety-LWE is defined over a multivariate polynomial ring constructed from an algebraic variety, which possesses a more complex structure than the multivariate rings considered in ____. Secondly, Variety-LWE enforces a ``no-mixed-terms'' constraint on the defining polynomials of the algebraic variety, ensuring that each polynomial depends on only one variable. This constraint guarantees coordinate-wise independence in Variety-LWE, preventing attackers from exploiting relationships between variables.