\section{Related Work}
\label{sec:related_work}







Researchers have proposed a variety of techniques to address issue management challenges and automate several tasks in the process____. For example, researchers have proposed automated techniques to better report issues____, assess issue quality____, 
predict the priority and severity of the issues____, categorize issue types____, assign developers to issues____, suggest potential duplicate issues____, reproduce buggy behavior____, localize buggy code files____,  and predict re-opened issues____. 

Researchers have studied issues for a variety of purposes: to understand  decision-making____ and 
the discourse used to describe issues____; extract decision information____; understand stakeholders' information needs____; characterize/predict different kinds of issues such as \textit{won't fix} issues____, fixed/resolved issues____, non-reproducible bugs____, and bug/issue types____; predict issue severity____; understand workarounds____ and visual content in issues____;  
questions____, and information types in issues____.










Researchers have used automated mining of issue data (\eg status changes) and version control/code review data to identify development processes and assess delays and inconsistencies____. 
They have utilized process mining techniques to integrate data from different sources (\eg VCS, issue trackers, and mail archives)____ and proposed process mining techniques____ to gain insight into development processes. Other work has studied the life cycle of issues by mining and analyzing issue state transitions____.
These works focused more broadly on issue management and identified transitions of issue states (\eg from ``Assigned" to ``In progress" to ``Closed"). 
However, issue states are often too broad to provide detailed insights into how stakeholders resolve issues in practice.




Unlike prior work, our research qualitatively analyzed issue reports to identify resolution stages, develop a process model, and uncover detailed patterns of issue resolution at Firefox. This in-depth analysis led to new insights into the issue resolution process. To our knowledge, we are the first to examine how the issue resolution process is actually implemented and discussed in practice, and how it differs from the theoretical models found in the literature.