\documentclass[conference]{IEEEtran}
%****************this is required for ICMLCN**********
\usepackage[letterpaper, left=0.65in, right=0.65in, bottom=1in, top=0.8in]{geometry}
\IEEEoverridecommandlockouts
\special{papersize=8.5in,11in}
%****************************************************
\input{files/acronym_list}
\input{files/pakages}
% \setlength{\textfloatsep}{0pt}
% \setlength{\parskip}{0pt} 
%****************************************************
\begin{document}
\title{A Low-Complexity Plug-and-Play Deep Learning Model for Massive MIMO Precoding Across Sites\\
\author{\IEEEauthorblockN{Ali~Hasanzadeh~Karkan$^{*}$, Ahmed~Ibrahim$^{\dagger}$, Jean-François~Frigon$^{*}$, and François~Leduc-Primeau$^{*}$\\
$^{*}$Department of Electrical Engineering, Polytechnique Montréal, Montréal, QC H3C 3A7, Canada\\
$^{*}$Emails: \{ali.hasanzadeh-karkan, j-f.frigon, francois.leduc-primeau\}@polymtl.ca. \\
$^{\dagger}$Ericsson Canada's R\&D,  Kanata, ON K2K 2V6, Canada, Email: ahmed.a.ibrahim@ericsson.com.
}
}
}

\maketitle
\IEEEpubidadjcol

\begin{abstract}
Massive multiple-input multiple-output (mMIMO) technology has transformed wireless communication by enhancing spectral efficiency and network capacity. This paper proposes a novel deep learning-based mMIMO precoder to tackle the complexity challenges of existing approaches, such as weighted minimum mean square error (WMMSE), while leveraging meta-learning domain generalization and a teacher-student architecture to improve generalization across diverse communication environments. 
When deployed to a previously unseen site, the proposed model
achieves excellent sum-rate performance while maintaining low computational complexity by avoiding matrix inversions and by using a simpler neural network structure.
The model is trained and tested on a custom ray-tracing dataset composed of several base station locations. The experimental results indicate that our method effectively balances computational efficiency with high sum-rate performance while showcasing strong generalization performance in unseen environments.
Furthermore, with fine-tuning, the proposed model outperforms WMMSE across all tested sites and SNR conditions while reducing complexity by at least 73$\times$.
\end{abstract}

% \begin{IEEEkeywords}
% \end{IEEEkeywords}
% \IEEEpeerreviewmaketitle

\section{Introduction} \label{Sec:Intro}
\input{1_intro}

\section{Problem Formulation} \label{Sec:Problem_Formulation}
\input{2_problem_formulation}

\section{Proposed Method} \label{Sec:Proposed}
\input{3_proposed_method}

\input{files/results_figure}
\section{Baselines and Complexity Analysis} \label{Sec:Complexity}
\input{complexity}


\section{Numerical Results} 
\label{Sec:Simulation}
\input{4_results}

\section{Conclusion} \label{Sec:conclusion}
In this paper, a novel \gls{DL}-based precoding method approach named PaPP is proposed for mMIMO systems to overcome the limitations of traditional methods in terms of complexity and generalizability. By combining WMMSE with DL techniques, we designed a computationally efficient precoder that leverages MLDG and a teacher-student training framework for improved adaptability. Our results demonstrate high sum-rate performance across diverse urban environments while achieving approximately 3.5$\times$ lower computational complexity than alternative generalizable DNN approaches (MAML-CNN).  Furthermore, when deployed in a new site without fine-tuning, PaPP achieves roughly the same performance as a model trained directly on the deployment site. With fine-tuning, PaPP outperforms all other methods. 

\section*{Acknowledgement}
This work was supported by Ericsson - Global Artificial Intelligence Accelerator AI-Hub Canada in Montr\'{e}al and jointly funded by NSERC Alliance Grant 566589-21 (Ericsson, ECCC, Innov\'{E}\'{E}).

\bibliographystyle{IEEEtran}
\bibliography{0_main}


\end{document}