% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\usepackage[preprint]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}
\usepackage{booktabs}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}

%Including images in your LaTeX document requires adding
%additional package(s)
\usepackage{graphicx}
\usepackage{float}
\usepackage{overpic}
\usepackage{wrapfig}
\usepackage{caption}
\usepackage{subfig}
\usepackage{multirow}

% math
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}

%symbol
\usepackage{bbding}
\usepackage{enumitem}

% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{MegaTTS 3: Sparse Alignment Enhanced Latent Diffusion Transformer \\ for Zero-Shot Speech Synthesis}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a separate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

% \author{First Author \\
%   Affiliation / Address line 1 \\
%   Affiliation / Address line 2 \\
%   Affiliation / Address line 3 \\
%   \texttt{email@domain} \\\And
%   Second Author \\
%   Affiliation / Address line 1 \\
%   Affiliation / Address line 2 \\
%   Affiliation / Address line 3 \\
%   \texttt{email@domain} \\}

\author{
 \textbf{Ziyue Jiang\textsuperscript{1,2}\thanks{
Intern at ByteDance.}},
 \textbf{Yi Ren\textsuperscript{2}},
 \textbf{Ruiqi Li\textsuperscript{1,2}},
 \textbf{Shengpeng Ji\textsuperscript{1}},
 \textbf{Boyang Zhang\textsuperscript{1}},
 \textbf{Zhenhui Ye\textsuperscript{1}},
 \textbf{Chen Zhang\textsuperscript{2}},
\\
 \textbf{Bai Jionghao\textsuperscript{1}},
 \textbf{Xiaoda Yang\textsuperscript{1}},
 \textbf{Jialong Zuo\textsuperscript{1}},
 \textbf{Yu Zhang\textsuperscript{1}},
 \textbf{Rui Liu\textsuperscript{3}},
 \textbf{Xiang Yin\textsuperscript{2}},
 \textbf{Zhou Zhao\textsuperscript{1}}
\\
 \textsuperscript{1}Zhejiang University,
 \textsuperscript{2}ByteDance,
 \textsuperscript{3}Inner Mongolia University
\\
 \small{
   \href{ziyuejiang341@gmail.com}{ziyuejiang341@gmail.com},
   \href{zhaozhou@zju.edu.cn}{zhaozhou@zju.edu.cn}
 }
}

\begin{document}
\maketitle
\begin{abstract}
While recent zero-shot text-to-speech (TTS) models have significantly improved speech quality and expressiveness, 
mainstream systems still suffer from issues related to speech-text alignment modeling: 1) models without explicit speech-text alignment modeling exhibit less robustness, especially for hard sentences in practical applications; 2) predefined alignment-based models suffer from naturalness constraints of forced alignments. This paper introduces \textit{MegaTTS 3}, a TTS system featuring an innovative sparse alignment algorithm that guides the latent diffusion transformer (DiT). Specifically, we provide sparse alignment boundaries to MegaTTS 3 to reduce the difficulty of alignment without limiting the search space, thereby achieving high naturalness. Moreover, we employ a multi-condition classifier-free guidance strategy for accent intensity adjustment and adopt the piecewise rectified flow technique to accelerate the generation process. Experiments demonstrate that MegaTTS 3 achieves state-of-the-art zero-shot TTS speech quality and supports highly flexible control over accent intensity. Notably, our system can generate high-quality one-minute speech with only 8 sampling steps. Audio samples are available at~\url{https://sditdemo.github.io/sditdemo/}.
\end{abstract}

% autoregressive large language models are inefficient and not robust in long-sentence inference; 
% The naturalness of predefined alignment-based models is also highly dependent on the performance of the duration model.

\input{sections/1_intro}
\input{sections/2_background}
\input{sections/3_method}
\input{sections/4_exp}
\input{sections/5_conclusion}


\section*{Limitations}
\label{app:limitation_and_future_work}
In this section, we discuss the limitations of the proposed method and outline potential strategies for addressing them in future research. 
\begin{itemize}
\item \textbf{Language Coverage.} Although our model currently supports both English and Chinese, there are far more languages in the world. We plan to incorporate additional training data from a wider range of languages and apply adaptation-based techniques, such as LoRA tuning~\citep{hu2021lora}, to enhance speech quality for low-resource languages.
\item \textbf{Function Coverage.} We can make MegaTTS 3 more user-friendly by enabling it to generate speech in various styles according to text descriptions through instruction-based fine-tuning. We can further fine-tune MegaTTS 3 on the paralinguistic corpus, allowing it to generate speech that is closer to a natural human style.
\end{itemize}


% Bibliography entries for the entire Anthology, followed by custom entries
%\bibliography{anthology,custom}
% Custom bibliography entries only
\bibliography{custom}

\input{sections/6_appendix}

\end{document}
