% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\usepackage[final]{acl}
\usepackage{fontawesome}
% Standard package includes
\usepackage{times}
\usepackage{latexsym}
\usepackage{xcolor}         % colors
\usepackage{commenting} 
% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets
\usepackage{subcaption}
% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}
\usepackage{float}
% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}
\usepackage{booktabs}
% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}


\usepackage{xcolor} % For colors
\usepackage{soul}   % For highlighting text
\usepackage{ulem}   % For underlining and strikethrough

% Define colors
\definecolor{emotion}{RGB}{200, 50, 50} % For emotional content
\definecolor{death}{RGB}{150, 0, 0}     % For implicit suicidal ideation
\definecolor{highlight}{RGB}{255, 223, 186} % For background highlighting

%Including images in your LaTeX document requires adding
%additional package(s)
\usepackage{graphicx}
\usepackage{amsmath} 
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{booktabs} 
\usepackage{array}    
\usepackage{multirow}
\usepackage{fancyvrb}
\newcommand{\di}[1]{\textcolor{red}{#1}}  
% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.
\newcommand{\ourbench}{\texttt{xxbench}~}
\newcommand{\ourdata}{\texttt{DeepSuiMind}~}

\definecolor{customblue}{HTML}{0f71e5}
\definecolor{customgreen}{HTML}{008a0d}

\title{Can Large Language Models Identify Implicit Suicidal Ideation? An Empirical Evaluation}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a separate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}


\author{
    \textbf{Tong Li\textsuperscript{1,2,3}}, 
    \textbf{Shu Yang\textsuperscript{1,2}}, 
    \textbf{Junchao Wu\textsuperscript{4}},  
    \textbf{Jiyao Wei\textsuperscript{5}}, \\
    \textbf{Lijie Hu\textsuperscript{1,2}}, 
    \textbf{Mengdi Li\textsuperscript{1,2}}, 
    \textbf{Derek F. Wong\textsuperscript{4}}, 
    \textbf{Joshua R. Oltmanns \textsuperscript{3}}
    \textbf{Di Wang\textsuperscript{1,2,†}}\\
  \textsuperscript{1}Provable Responsible AI and Data Analytics (PRADA) Lab, \\
  \textsuperscript{2}King Abdullah University of Science and Technology, \\
    \textsuperscript{3}Washington University in St.Louis,
  \textsuperscript{4}University of Macau, \\
    \textsuperscript{5}	Institute of Computing Technology, Chinese Academy of Sciences
}

%\author{
%  \textbf{First Author\textsuperscript{1}},
%  \textbf{Second Author\textsuperscript{1,2}},
%  \textbf{Third T. Author\textsuperscript{1}},
%  \textbf{Fourth Author\textsuperscript{1}},
%\\
%  \textbf{Fifth Author\textsuperscript{1,2}},
%  \textbf{Sixth Author\textsuperscript{1}},
%  \textbf{Seventh Author\textsuperscript{1}},
%  \textbf{Eighth Author \textsuperscript{1,2,3,4}},
%\\
%  \textbf{Ninth Author\textsuperscript{1}},
%  \textbf{Tenth Author\textsuperscript{1}},
%  \textbf{Eleventh E. Author\textsuperscript{1,2,3,4,5}},
%  \textbf{Twelfth Author\textsuperscript{1}},
%\\
%  \textbf{Thirteenth Author\textsuperscript{3}},
%  \textbf{Fourteenth F. Author\textsuperscript{2,4}},
%  \textbf{Fifteenth Author\textsuperscript{1}},
%  \textbf{Sixteenth Author\textsuperscript{1}},
%\\
%  \textbf{Seventeenth S. Author\textsuperscript{4,5}},
%  \textbf{Eighteenth Author\textsuperscript{3,4}},
%  \textbf{Nineteenth N. Author\textsuperscript{2,5}},
%  \textbf{Twentieth Author\textsuperscript{1}}
%\\
%\\
%  \textsuperscript{1}Affiliation 1,
%  \textsuperscript{2}Affiliation 2,
%  \textsuperscript{3}Affiliation 3,
%  \textsuperscript{4}Affiliation 4,
%  \textsuperscript{5}Affiliation 5
%\\
%  \small{
%    \textbf{Correspondence:} \href{mailto:email@domain}{email@domain}
%  }
%}

\begin{document}
\maketitle
\begin{abstract}
We present a comprehensive evaluation framework for assessing Large Language Models' (LLMs) capabilities in suicide prevention, focusing on two critical aspects: the Identification of Implicit Suicidal ideation (IIS) and the Provision of Appropriate Supportive responses (PAS). We introduce \ourdata, a novel dataset of 1,308 test cases built upon psychological frameworks including D/S-IAT and Negative Automatic Thinking, alongside real-world scenarios. Through extensive experiments with 8 widely used LLMs under different contextual settings, we find that current models struggle significantly with detecting implicit suicidal ideation and providing appropriate support, highlighting crucial limitations in applying LLMs to mental health contexts. Our findings underscore the need for more sophisticated approaches in developing and evaluating LLMs for sensitive psychological applications.
\textcolor{red}{\faExclamationTriangle~Content Warning: This paper contains examples of harmful language.}

\end{abstract}

\input{section/1_intro}
\input{section/2_relatedwork}
\input{section/3_0_theory}
\input{section/3_method}
\input{section/4_experiment}
\input{section/5_discussion}
\newpage
\section{Limitation}

Our study primarily focuses on English, which may limit its applicability across diverse linguistic and cultural contexts. Suicidal ideation can manifest differently across languages and cultures, and incorporating a broader range of languages would provide a more comprehensive assessment. Additionally, while our findings indicate that explicit warnings enhance models' sensitivity to implicit suicidal intent, further research is needed to explore which specific model architectures respond most effectively to different forms of explicit guidance.  Moreover, as large language models continue to evolve, their ability to detect and interpret implicit suicidal intent may vary significantly depending on model design, training data, and fine-tuning strategies. Future work should investigate how these factors influence model performance and explore more robust and generalizable approaches to suicide risk assessment.

\bibliography{custom}

\newpage
\appendix
\input{section/appendix}


\end{document}

