\section{Related works}

\noindent \textbf{Gender bias studies in MLMs:} 
%
Profession \cite{limisiewicz-marecek-2022-dont, bartl-etal-2020-unmasking} and behavioral concepts (e.g., intelligent) 
% (e.g., intelligent, enemy) 
% nadeem-etal-2021-stereoset,kaneko2022unmasking
\cite{guo2022auto,ahn-oh-2021-mitigating} are frequently explored targets in bias studies.
%
Less explored targets include physical appearances (e.g., beautiful) \cite{kaneko2022unmasking,nadeem-etal-2021-stereoset}, stimuli (e.g., career/family),
% , maths/science/arts
% kurita-etal-2019-measuring
and emotion \cite{bartl-etal-2020-unmasking}.
% , toxicity \cite{steed2022upstream,ousidhoum2021probing}.
%
Exploration of language model biases in human trait perceptions is novel. An exception is \citet{rao-etal-2023-chatgpt} exploring bias in personality perceptions in GPT-4 -- they find low gender bias. 
%
However, they do not compare with self-reported traits in psychology. The study of biases in \textit{personality} and \textit{character} perceptions is novel. 

\vspace{0.5em}
\noindent \textbf{Gender differences in traits from psychology:} Big Five personality traits  \cite{goldberg1992development}, have been studied extensively \cite{hartmann2023big,ock2020practical,russo2020gender,kusnierz2020examining,lippa2010gender}
%
using methods of self-reported questionnaires
% weisberg2011gender
%(e.g., \citet{{kumar2023big}}) 
and aggregated through meta-analyses \cite{lippa2010gender,feingold1994gender}.
%
These, in general, show that females score higher in \textit{agreeableness} and \textit{neuroticism} with small to little difference in the other personality traits.
%
Research on gender differences in character traits from a lexical approach is still lacking despite the framework being proposed in the early 2000s.

\vspace{0.5em}
\noindent \textbf{MLM bias detection with templates:}  
%
Prior works average bias scores across templates 
% delobelle2022measuring,qian2019reducing
\cite{limisiewicz-marecek-2022-dont,bartl-etal-2020-unmasking}, without considering template variability.
%
Some rely solely on magnitude of score differences while others \cite{steed2022upstream,bartl-etal-2020-unmasking,kurita-etal-2019-measuring} use significance test.
%
Bias quantification using effect size, common in psychology is overlooked.
