% !TEX root = main.tex

\section{SKE-CR-SKL with Key Testability}\label{sec:SKECRSKL-KT}
In this section, we show how to achieve SKE-CR-SKL introduced in~\cref{def:ske_cr_skl}.
\subsection{Construction}\label{sec:SKECRSKL-KT-construction}
We construct an SKE-CR-SKL scheme with key testability $\SKECRSKL=
\SKECRSKL.(\Setup,\qKG,\Enc,\qDec,\allowbreak\qVrfy)$ having the
additional algorithms $\CDec$ and $\KeyTest$, using the
following building blocks.

\begin{itemize}
\item BB84-based SKE-CD scheme (Definition \ref{def:bb84}) $\SKECD =
\SKECD.(\KG,\qEnc,\qDec,\allowbreak\qDel,\Vrfy)$ having the classical
decryption algorithm $\SKECD.\CDec$.

\item OWF $f:\bit^\secp\ra\bit^{p(\secp)}$ for some polynomial $p$.
\end{itemize}

Let $\cM \seteq \bit^{\msglen}$ be the plaintext space. The construction is as follows:

\begin{description}

\item[$\SKECRSKL.\Setup(1^\secp)$:] $ $
\begin{enumerate}
    \item Generate $r\gets\bit^{\msglen}$.
    \item Generate $\skecd.\sk\gets\SKECD.\KG(1^\secp)$.
%    \item Generate $(\sig.\vk,\sig.\sk)\gets\SIG.\KG(1^\secp)$.
    \item Output $\msk\seteq(\skecd.\sk, r)$.
\end{enumerate}

\item[$\SKECRSKL.\qKG(\msk)$:] $ $
\begin{enumerate}
    \item Parse $\msk=(\skecd.\sk, r)$.
    \item Generate
        $(\skecd.\qct,\skecd.\vk)\gets\SKECD.\qEnc(\skecd.\sk,r)$.
        $\skecd.\vk$ is of the form
        $(x,\theta)\in\bit^{\ctlen}\times\bit^{\ctlen}$, and
        $\skecd.\qct$ is of the form
        $\ket{\psi_1}_{\qreg{SKECD.CT_1}}\tensor\cdots\tensor\ket{\psi_{\ctlen}}_{\qreg{SKECD.CT_{\ctlen}}}$.

    \item Generate $s_{i,b}\la\bit^\secp$ and compute $t_{i,b}\la f(s_{i,b})$ for every $i\in[\ctlen]$ and $b\in\bit$. 
    Set $T\seteq
    t_{1,0}\|t_{1,1}\|\cdots\|t_{\ctlen,0}\|t_{\ctlen,1}$ and $S =
    \{s_{i,0} \xor s_{i, 1}\}_{i \in [\ctlen] \; : \;\theta[i] =
    1}$.
    \item Prepare a register $\qreg{S_i}$ that is initialized to
    $\ket{0^\secp}_{\qreg{S_i}}$ for every $i\in[\ctlen]$. 
    \item For every $i\in[\ctlen]$, apply the map
    \begin{align}
    \ket{u_i}_{\qreg{SKECD.CT_i}}\tensor\ket{v_i}_{\qreg{S_i}}
    \ra
    \ket{u_i}_{\qreg{SKECD.CT_i}}\tensor\ket{v_i\oplus s_{i,u_i}}_{\qreg{S_i}}
    \end{align}
    to the registers $\qreg{SKECD.CT_i}$ and $\qreg{S_i}$ and obtain the resulting state $\rho_i$.
    \item Output $\qdk = (\rho_i)_{i\in{[\ctlen]}}$,
    $\vk=(x,\theta,S)$, and $\tk=T$.
\end{enumerate}

\item[$\SKECRSKL.\Enc(\msk, \msg)$:] $ $
\begin{enumerate}
    \item Parse $\msk = (\skecd.\sk, r)$.
    \item Output $\ct\seteq(\skecd.\sk, r\oplus\msg)$.
    
\end{enumerate}

\item[$\SKECRSKL.\CDec(\dk, \ct)$:] $ $
\begin{enumerate}
\item Parse $\ct = (\skecd.\sk, z)$. Let
$\widetilde{\dk}$ be the sub-string of $\dk$ on register
$\qreg{SKECD.CT} = \qreg{SKECD.CT_1} \otimes \cdots \otimes \qreg{SKECD.CT_{\ctlen}}$.
\item Output $z \xor \SKECD.\CDec(\skecd.\sk, \widetilde{\dk})$.
\end{enumerate}

\item[$\SKECRSKL.\qDec(\qdk, \ct)$:] $ $
\begin{enumerate}
\item Parse $(\rho_i)_{i \in [\ctlen]}$. We denote the register
holding $\rho_i$ as $\qreg{SKECD.CT_i}\tensor\qreg{S_i}$ for
every $i\in[\ctlen]$.

\item Prepare a register $\qreg{MSG}$ of $\msglen$ qubits that is
initialized to $\ket{0\cdots0}_{\qreg{MSG}}$.

\item Apply the map
\begin{align}
&\ket{u}_{\bigotimes_{i\in[\ctlen]}\qreg{SKECD.
CT_i}} \tensor\ket{w}_{\qreg{MSG}} \ra\\
&\ket{u}_{\bigotimes_{i\in[\ctlen]}\qreg{SKECD.CT_i}}\tensor\ket{w\oplus
\SKECRSKL.\CDec(u, \ct)}_{\qreg{MSG}}
\end{align}

to the registers
$\bigotimes_{i\in[\ctlen]}\qreg{SKECD.CT_i}$ and
$\qreg{MSG}$.
\item Measure $\qreg{MSG}$ in the computational basis and output the result $\msg^\prime$.
\end{enumerate}

\item[$\SKECRSKL.\qVrfy(\vk,\widetilde{\qdk})$:] $ $
\begin{enumerate}
\item Parse $\vk = (x,\theta,S=\{s_{i,0} \xor
    s_{i,1}\}_{i\in[\ctlen]\; : \; \theta[i]=1})$ and
    $\qdk = (\rho_i)_{i\in[\ctlen]}$ where $\rho_i$ is a state on
    the registers $\qreg{SKECD.CT_i}$ and $\qreg{S_i}$.
\item For every $i \in [\ctlen]$, measure $\rho_i$ in the Hadamard
basis to get outcomes $c_i, d_i$ corresponding to 
the registers $\qreg{SKECD.CT_i}$ and $\qreg{S_i}$ respectively.

\item Output $\top$ if $x[i]=c_i \oplus d_i\cdot(s_{i,0}\oplus
        s_{i,1})$ holds for every $i\in[\ctlen]$ such that $\theta[i]=1$.
    Otherwise, output $\bot$.
\end{enumerate}

\item[$\SKECRSKL.\KeyTest(\tk,\dk)$:] $ $
\begin{enumerate}
\item Parse $\dk$ as a string over the registers
$\qreg{SKECD.CT} = \qreg{SKECD.CT_1} \otimes \cdots \otimes
\qreg{SKECD.CT_{\ctlen}}$ and $\qreg{S} = \qreg{S_1} \otimes
\cdots \otimes \qreg{S_{\ctlen}}$.
Let $u_i$ denote the value on
$\qreg{SKECD.CT_i}$ and $v_i$ the value on $\qreg{S_i}$. Parse $\tk$
as $T=t_{1,0}\|t_{1,1}\|\cdots \|t_{\ctlen,0}\|t_{\ctlen,1}$.


\item Let $\Check[t_{i,0},t_{i_1}](u_i,v_i)$ be the deterministic
algorithm that outputs $1$ if $f(v_i)=t_{i,u_i}$ holds and $0$
otherwise.

\item Output $\Check[t_{1,0},t_{1,1}](u_1,v_1) \land
\cdots \land
\Check[t_{\ctlen,0},t_{\ctlen,1}](u_{\ctlen},v_{\ctlen})$.
\end{enumerate}

\end{description}


\paragraph{Decryption correctness:} For a ciphertext $\ct =
(\skecd.\sk, r \xor \msg)$, the decryption algorithm $\qDec$ performs
the following computation:

\begin{align}
&\ket{u}_{\bigotimes_{i\in[\ctlen]}\qreg{SKECD.
CT_i}} \tensor\ket{w}_{\qreg{MSG}} \ra\\
&\ket{u}_{\bigotimes_{i\in[\ctlen]}\qreg{SKECD.CT_i}}\tensor\ket{w
\xor r \xor \msg \xor \SKECD.\CDec(\skecd.\sk, u)}_{\qreg{MSG}}
\end{align}

Recall that $\qreg{SKECD.CT} = \qreg{SKECD.CT_1} \otimes \cdots
\otimes \qreg{SKECD.CT_{\ctlen}}$ is a register corresponding to the
ciphertext of a BB84-based SKE-CD scheme $\SKECD$. Hence, from the Classical
Decryption property of $\SKECD$ (Definition \ref{def:bb84}), it must be that
$\SKECD.\CDec(\skecd.\sk, u) = r$ for every $u$ in the superposition
of $\skecd.\qct$. Consequently, $\msg$ is written onto the
$\qreg{MSG}$ register in each term of the superposition and decryption
correctness follows.


\paragraph{Verification correctness:}

Observe that for the Hadamard basis positions ($i \in [\ctlen]$ such that
$\theta[i] = 1$), $\rho_i$ is of the form:

$$\rho[i] = \ket{0}_{\qreg{SKECD.CT_i}}\ket{s_{i,0}}_{\qreg{S_i}}+
(-1)^{x[i]}\ket{1}_{\qreg{SKECD.CT_i}}\ket{s_{i,1}}_{\qreg{S_i}}$$

It is easy to see that measuring the state in the Hadamard basis gives
outcomes $c_i, d_i$ (on registers $\qreg{SKECD.CT_i}$ and
$\qreg{S_i}$ respectively) satisfying $x[i] = c_i \xor d_i \cdot
(s_{i,0} \xor s_{i,1})$. Hence, the verification correctness follows.

\begin{theorem} There exists an SKE-CR-SKL scheme satisfying
OT-IND-KLA security and Key-Testability, assuming the existence of a
BB84-based SKE-CD scheme and the existence of an OWF.
\end{theorem}

\ifnum\submission=1
We prove this theorem in~\cref{appsec:security_proofs_for_ske_cr_skl}.
\else
We prove this theorem in the subsequent sections.
\input{proof_SKE-CR-SKL}
\fi

