%
\def\llncs{0}
\def\fullpage{1}
\def\anonymous{0}
\def\authnote{0}
\def\notxfont{0}
\def\submission{0}
\def\reply{0}
\def\cameraready{0}
\def\noaux{1}
%%%%%%%%%%%% Define modes %%%%%%%%%%%%%
\ifnum\submission=1
\def\anonymous{1}
\def\llncs{1}
\else
\fi


\ifnum\cameraready=1
\def\llncs{1}
\def\anonymous{0}
\def\authnote{0}
\else
\fi

\ifnum\anonymous=1
%\def\llncs{1}
\def\authnote{0}
\else
\fi

\input{mac}

\ifnum\mac=0
\ifnum\llncs=1
	\documentclass[envcountsect,a4paper,runningheads]{llncs}
\else
	\documentclass[letterpaper,hmargin=1.05in,vmargin=1.05in]{article}
			\ifnum\fullpage=1
		\usepackage{fullpage}
		\fi
\fi
\else
\ifnum\llncs=1
	\documentclass[dvipdfmx,envcountsect,a4paper,runningheads]{llncs}
\else
	\documentclass[dvipdfmx,letterpaper,hmargin=1.05in,vmargin=1.05in]{article}
			\ifnum\fullpage=1
		\usepackage{fullpage}
		\fi
\fi
\fi

\ifnum\reply=1
\usepackage{ulem}
\renewcommand{\emph}{\textit}
\else
\fi


%%%%% Joc revision setting------------------------------------------
% \ifnum\reply=0  %%%%% Remove comments %%%%%
% \newcommand{\revise}[1]{#1}
% \newcommand{\strike}[1]{}
% \else
% \newcommand{\revise}[1]{{\color{red}{#1}}}
% \newcommand{\strike}[1]{\sout{#1}}
% \fi


%=========  Preamble Part ===========
\input{preamble_usepackages.tex}
\input{macros_basic.tex}
\input{macros_adhoc.tex}

\newcommand{\authornote}[3]{\textcolor{#3}{[\textsc{#1:} {#2}]}}
\ifnum\authnote=1
\newcommand{\fuyuki}[1]{\authornote{Fuyuki}{#1}{chocolate}}
\newcommand{\ryo}[1]{\authornote{Ryo}{#1}{darkblue}}
\newcommand{\nikhil}[1]{\authornote{Nikhil}{#1}{red}}
\else
\newcommand{\fuyuki}[1]{}
\newcommand{\ryo}[1]{}
\newcommand{\nikhil}[1]{}
\fi


%-------- START: LNCS branch ----------
\ifnum\llncs=1
%  See the following url for the meaning of the commands.   https://tex.stackexchange.com/questions/160053/vec-conflict-between-llncs-cls-and-mnsymbol
\let\oldvec\vec% Store \vec in \oldvec
%\usepackage{llncsdoc}
\let\vec\oldvec% Restore \vec from \oldvec
%
% for page number llncs
%\pagestyle{headings} 
%
% make a proper TOC despite llncs
\setcounter{tocdepth}{2}
\makeatletter
\renewcommand*\l@author[2]{}
\renewcommand*\l@title[2]{}
\makeatletter
\fi    
%-------- END: LNCS branch ----------

\theoremstyle{remark}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{experiment}{Experiment}
\newtheorem{construction}[theorem]{Construction}
\newtheorem{counterexample}[theorem]{Counterexample}




%================================
%=========  MAIN BODY ============    
%================================

\ifnum\submission=1
\title{
\textbf{PKE and ABE with \\ Collusion-Resistant Secure Key Leasing}%\thanks{{\color{red}{\emph{We attached the full version of this paper as a supplementary material}}}.}
}
\else
\title{
\textbf{PKE and ABE with \\ Collusion-Resistant Secure Key Leasing}
}
\fi


%\date{}
\begin{document}
%\author{}
%\institute{}


\ifnum\anonymous=1 
\ifnum\llncs=1
% for anonymized LNCS version
\author{\empty}\institute{\empty}
\else
% for anonymized full version
\author{}
\fi
\else
%
%  For camera ready version.
%
\ifnum\llncs=1
\author{
	Fuyuki Kitagawa\inst{1,2} \and Ryo Nishimaki\inst{1,2} \and Nikhil Pappu\inst{3}
}
\institute{
	NTT Social Informatics Laboratories, Tokyo, Japan \and NTT Research Center for Theoretical Quantum Information, Atsugi, Japan \and Portland State University, USA
}
\else
%
%   For full/eprint version, etc.
%
\author[$\dagger$ $\diamondsuit$]{\hskip 1em Fuyuki Kitagawa}
\author[$\dagger$ $\diamondsuit$]{\hskip 1em Ryo Nishimaki}
\author[$\star$]{\hskip 1em Nikhil Pappu\thanks{Part of this work was done while visiting NTT Social Informatics Laboratories as an internship.}
\thanks{Supported by the US National Science Foundation (NSF) via Fang Song's Career Award (CCF-2054758). }}
\affil[$\dagger$]{{\small NTT Social Informatics Laboratories, Tokyo, Japan}\authorcr{\small \{fuyuki.kitagawa,ryo.nishimaki\}@ntt.com}}
\affil[$\diamondsuit$]{{\small NTT Research Center for Theoretical Quantum Information, Atsugi, Japan}}
\affil[$\star$]{{\small Portland State University, USA}\authorcr{\small nikpappu@pdx.edu}}
\renewcommand\Authands{, }
\fi %%%%% END OF LNCS branch
\fi

\ifnum\llncs=1
\date{}
\else
\ifnum\anonymous=0
\date{\today}
\else
\date{}
\fi
\fi


\maketitle

\begin{abstract}
Secure key leasing (SKL) is an advanced encryption functionality that allows a secret key holder to generate a quantum decryption key and securely lease it to a user.
Once the user returns the quantum decryption key (or provides a classical certificate confirming its deletion), they lose their decryption capability.
Previous works on public key encryption with SKL (PKE-SKL) have only considered the single-key security model, where the adversary receives at most one quantum decryption key. However, this model does not accurately reflect real-world applications of PKE-SKL.
To address this limitation, we introduce \emph{collusion-resistant security} for PKE-SKL (denoted as PKE-CR-SKL). In this model, the adversary can adaptively obtain multiple quantum decryption keys and access a verification oracle which validates the correctness of queried quantum decryption keys. Importantly, the size of the public key and ciphertexts must remain independent of the total number of generated quantum decryption keys.
We present the following constructions:
\begin{itemize}
	\item A PKE-CR-SKL scheme based on the learning with errors (LWE) assumption.
	\item An attribute-based encryption scheme with collusion-resistant SKL (ABE-CR-SKL), also based on the LWE assumption.
	\item An ABE-CR-SKL scheme with classical certificates, relying on multi-input ABE with polynomial arity.
\end{itemize}
\end{abstract}



\ifnum\llncs=1
\else
\newpage
\setcounter{tocdepth}{2}
\tableofcontents

\newpage
\fi

\input{intro}
%\input{todo}
\input{overview2}
%\input{overview}
\input{preliminaries}
\input{def-CR-SKL}
%\input{def-SKL-RCD}
\input{const-CDSKE-CDec-NM}
\input{const-PKE-CR-SKL}
\ifnum\submission=0
\input{const-ABE-SKL}
\input{const-PKE-CR3-SKL}
\else
\fi

\ifnum\anonymous=1
\else
% \section*{Acknowledgement}

\fi

	\ifnum\llncs=1
%\bibliographystyle{bib/extreme_alpha}
\bibliographystyle{splncs04}
\bibliography{bib/abbrev3,bib/crypto,bib/siamcomp_jacm,bib/other}
	% \input{compressed_bib}
	\else
\bibliographystyle{alpha} 
\bibliography{bib/abbrev3,bib/crypto,bib/siamcomp_jacm,bib/other}
	\fi



\ifnum\cameraready=0
	\ifnum\llncs=0
	%%%%%% Full version region %%%%%%
	\appendix
	
	\input{quantum-secure-ABE}
	\input{SKECD_BB84}
	\input{const-SKFESKL}
	
%%% appendix files here
\else
%%%%% LNCS-style submission version region %%%%%
	\newpage
	 	\appendix
	 	\setcounter{page}{1}
 	{
	\noindent
 	\begin{center}
	{\Large SUPPLEMENTAL MATERIALS}
	\end{center}
 	}
	\setcounter{tocdepth}{2}
	\ifnum\noaux=1
 	\else
% {\color{red}{We attached the full version of this paper as a separated file (auxiliary supplemental material) for readability. It is available from the program committee members.}}
	\fi
\input{more_related_work}
\input{omitted_prelim}
\input{proof_O2HwOracle}
\input{quantum-secure-ABE}
\input{SKECD_BB84}
\input{proof_SKE-CR-SKL}
\input{const-ABE-SKL}
\input{const-SKFESKL}
\input{const-PKE-CR3-SKL}

	



	
	

	\setcounter{tocdepth}{2}
	\tableofcontents

\fi
\else
	%%%%%%% Camera-ready region (no appendix) %%%%%%
\fi


%
%
%

\end{document}
