\documentclass{article}
\usepackage{hyperref}       % hyperlinks
\usepackage{url}      
% simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb}
\usepackage{nicefrac}    
\usepackage[numbers,sort]{natbib}
% compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}      
\input{commands}
\usepackage[a4paper, margin=1in]{geometry}

\usepackage[ruled]{algorithm2e} % For algorithms


%% ADDED BY KRISHNA for subcaption
\let\subfigure\relax
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\newcommand{\norm}[1]{\left\| #1 \right\|}

%% END ADDED BY KRISHNA 
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}



% FOR COMMENTS

\newcommand{\juba}[1]{\textcolor{red}{[Juba: #1]}}
\newcommand{\vidya}[1]{\textcolor{red}{[Vidya: #1]}}
\newcommand{\gua}[1]{\textcolor{green}{[Guanghui: #1]}}
\newcommand{\kri}[1]{\textcolor{blue}{[Krishna: #1]}}

\renewcommand{\juba}[1]{}
 \renewcommand{\vidya}[1]{}
 \renewcommand{\gua}[1]{}
 \renewcommand{\kri}[1]{}
% FOR COMMENTS


\begin{document}


\begin{center}

{\bf{\Large{Multi-Agent Performative Prediction Beyond the Insensitivity Assumption: A Case Study for Mortgage Competition}}}

\vspace*{.2in}



{\large{
\begin{tabular}{cccc}
 Guanghui Wang$^1$, Krishna Acharya$^2$, Lokranjan Lakshmikanthan$^2$\\
 Vidya Muthukumar$^{2,3}$, Juba Ziani$^{2}$  \\
\end{tabular}}}

\vspace*{.05in}

\begin{tabular}{c}
$^1$College of Computing, Georgia Institute of Technology\\
$^2$School of Industrial and Systems Engineering, Georgia Institute of Technology\\
$^3$School of Electrical and Computer Engineering, Georgia Institute of Technology\\
  \texttt{\{gwang369,krishna.acharya,llakshmi3,vmuthukumar8,jziani3\}@gatech.edu}
\end{tabular}

\vspace*{.2in}
\date{}

\end{center}


% FOR COMMENTS

\begin{abstract}
Performative prediction models account for feedback loops in decision-making processes where predictions influence future data distributions. While existing work largely assumes insensitivity of data distributions to small strategy changes, this assumption usually fails in real-world competitive (i.e. multi-agent) settings. For example, in Bertrand-type competitions, a small reduction in one firm's price can lead that firm to capture the entire demand, while all others sharply lose all of their customers. 

We study a representative setting of multi-agent performative prediction in which insensitivity assumptions do not hold, and investigate the convergence of natural dynamics. To do so, we focus on a specific game that we call the ``Bank Game'', where two lenders compete over interest rates and credit score thresholds. Consumers act similarly as to in a Bertrand Competition, with each consumer selecting the firm with the lowest interest rate \emph{that they are eligible for based on the firms' credit thresholds}. Our analysis characterizes the equilibria of this game and demonstrates that when both firms use a common and natural no-regret learning dynamic---exponential weights---with proper initialization, the dynamics \emph{always} converge to stable outcomes despite the general-sum structure. Notably, our setting admits multiple stable equilibria, with convergence dependent on initial conditions. We also provide theoretical convergence results in the stochastic case when the utility matrix is not fully known, but each learner can observe sufficiently many samples of consumers at each time step to estimate it, showing robustness to slight mis-specifications. Finally, we provide experimental results that validate our theoretical findings.
\end{abstract}

\section{Introduction}
%\textcolor{red}{draft that lays out the mian story but needs some rewriting and a pass for clarity for a non-expert reader}
\input{intro.tex}

\section{Model}\label{sec1}
\input{model.tex}

\section{Equilibria of the One-Shot Game under Two Discrete Actions}
\label{sec:understand NE}
In this section, we introduce the equilibrium concepts that we use for solving the static, one-shot version of our game. We then fully characterize the equilibria of our problem. 

\input{equilibria.tex}

\section{Convergence of Natural Dynamics to Equilibria}
\label{sec:learntoconvege}
\input{convergence.tex}


\section{Experimental Results}
\label{sec: exp-results}
\input{experiments-intro}


\subsection{Experimental Setup} 
\input{experiments-setup-mainbody}

\subsection{Dynamics for Algorithm \ref{alg:Hedge}}
\label{sec:exp-dyna}
\input{experiments-dynamics-mainbody}

\subsection{Distance to Nash Equilibrium}
\label{sec:exp-dist}
\input{experiments-distance-mainbody}


\section{Conclusion and The Future Work}
In this work, we investigate the multi-agent performative prediction problem in settings where self-selection effects play a significant role and insensitivity assumptions do not hold. We introduce the "Bank Game", in which two mortgage companies compete to set minimum credit score requirements and interest rates. For this simple two-player game, we fully characterize the equilibria and demonstrate that when each player employs Exponential Weights, the dynamics always converge to equilibrium in the last-iterate sense. Finally, we present experimental results that validate our theoretical findings and extend our analysis to scenarios not covered under our assumptions. 

Our work also motivates various future directions to explore. A very natural open question is how our results extend to other scenarios in performative prediction where the insensitivity assumption does not hold. More specifically: i) Can we identify additional compelling applications where the insensitivity assumption is violated, finding new use cases for our framework? ii) Can we establish general-purpose results under relaxed insensitivity assumptions? For instance, one promising approach is to analyze cases where this assumption breaks down at specific points in the decision space (e.g., when both banks offer the same rate in our framework, leading to symmetry between players) but holds across most of the decision space. This could yield deeper theoretical insights and broaden the applicability of performative prediction models in multi-agent systems.




% \juba{ Future work is weak. A very natural open question we should ask is how our results expand to other situations in performative prediction where the insensitivity assumption is not satisfied. i) can we find further compelling applications where this is the case and ii) can we prove general-purpose results when relaxing the insensitivity assumption? A direction for ii) is to look at when this assumption does not hold at specific points (e.g. both banks have the same rate in our case, which suggests points at which players are symmetric) but holds on most of the space. Etc.} 
% \gua{I added more directions}


% \section{All plots placeholder section}
% \input{EC/all_plots}

\bibliographystyle{unsrt}
\bibliography{sample-bibliography}



% Appendix
\appendix
\input{equilibria-proof.tex}
\input{convergence-proof.tex}

\input{appendix-distNE}

\input{three-gamma.tex}
\end{document}
