\section{Introduction}

Satisfiability Modulo Theories (SMT) is the problem of deciding the satisfiability of a logical formula w.r.t.\ some background theory, such as linear and nonlinear arithmetic, bit-vectors, arrays, or uninterpreted functions~\cite{barrettSatisfiabilityModuloTheories2021}.
%
Many SMT-encodable problems also require the capability of finding models that are optimal w.r.t.\ some objective functions. These problems are grouped under the term Optimization Modulo Theories
(OMT)~\cite{nieuwenhuisSATModuloTheories2006,sebastianiOptimizationSMTLAQ2012,bjornerNZOptimizingSMT2015}.
%
% making it a powerful tool for solving complex constraint optimization problems in various domains.
OMT has been successfully applied to a wide range of problems, such as
verification of timed and hybrid
systems~\cite{sebastianiOptimizationSMTLAQ2012,henryHowComputeWorstcase2014},
numeric~\cite{leofanteOptimalPlanningModulo2021} and temporal
planning~\cite{panjkovicExpressiveOptimalTemporal2023,panjkovicAbstractActionScheduling2024},
optimal scheduling~\cite{bofillEfficientSMTApproach2017},
constrained goal modelling~\cite{nguyenMultiobjectiveReasoningConstrained2018}, 
hybrid machine learning~\cite{tesoStructuredLearningModulo2017}, GAS optimization for smart
contracts~\cite{albertGASOLGasAnalysis2020}, and optimum
encodings for quantum annealing~\cite{bianSolvingSATMaxSAT2020,dingEffectivePrimeFactorization2024}, 
establishing OMT solvers as powerful tools for solving complex constraint optimization problems in various domains.

% Several other contexts have seen the successful application of OMT techniques, e.g., verification of parametric systems~\cite{liSymbolicOptimizationSMT2014},quantum annealing~\cite{bianSolvingSATMaxSAT2020}, constrained goal
% modelling~\cite{nguyenRequirementsEvolutionEvolution2016,nguyenModelingReasoningRequirements2017,nguyenMultiobjectiveReasoningConstrained2018}, and hybrid machine learning~\cite{structured}.

% OMT techniques have been developed for 
% \larat~\cite{bjornerNZOptimizingSMT2015,sebastianiOptimizationSMTLAQ2012,sebastianiOptimizationModuloTheories2015}, 
% \laratint~\cite{bjornerNZOptimizingSMT2015,sebastianiPushingEnvelopeOptimization2015},
% \nlaratint~\cite{bigarellaOptimizationModuloNonlinear2021},
% \mem~\cite{nadelBitVectorOptimization2016,trentinOptimizationModuloTheories2021}, and 
% \fl~\cite{trentinOptimizationModuloTheories2021}.


\input{01a-related-work.tex}

\paragraph{Contributions.}
In this paper, we study the applicability of enumeration-based techniques to OMT solving, and, in particular, the usage of partial truth assignment reduction to improve the effectiveness and efficiency of OMT solving.
First, we notice that OMT solvers typically invoke the \T-optimizer on total truth assignments, and we spotlight how this can be suboptimal in many cases.
Second, we propose some ways to exploit partial truth assignments in OMT solving, tailoring existing techniques to the OMT context.
We show through an empirical evaluation over \omlarat{} benchmarks that these strategies can improve both the efficiency of OMT optimal solving and the quality of obtained solutions for anytime solving. %, with a small overhead.

\paragraph{Organization.}
The rest of the paper is organized as follows.
In \sref{sec:background}, we provide the necessary background on SMT and OMT solving.
In \sref{sec:analysis}, we analyze the role of total and partial truth assignments in OMT solving.
In \sref{sec:approach}, we propose two strategies to exploit partial truth assignments in OMT solving.
In \sref{sec:experiments}, we present an experimental evaluation of the proposed strategies over \omlarat{} benchmarks.
Finally, in \sref{sec:conclusions}, we conclude the paper and discuss future work.
