\section{Related Work}
\label{sec:related}

\textbf{Open-Ended Discovery with Foundation Models.}
The field of open-endedness____ aims to continually discover diverse and novel artifacts forever.
Recent methods leverage the generative capabilities and vast prior knowledge of foundation models (FMs) to accelerate this process____ by harnessing a foundation model's intrinsic notion of interestingness____ to construct the next proposal, analogous to human innovation.
Notable examples include ELM____ which evolves novel robot morphologies; OMNI-EPIC____, which automatically designs novel environments for reinforcement learning (RL) agents; DiscoPOP which discovers new loss functions for preference optimization algorithms____; ADAS____, which evolves novel designs for LLM-based agentic systems; and The AI Scientist____, which seeks to automate the entire scientific process by proposing novel ideas, conducting experiments, and writing a scientific paper summarizing the results.

\textbf{Automated Evaluation of Foundation Models.} Recent research also investigates automated evaluation of FMs, moving beyond static, human-designed test suites.
Rainbow Teaming____ applies Quality-Diversity algorithms____ to find novel adversarial attacks that stress-test FMs for safety.
Similarly, works like____ automate the red teaming (probing a system for weaknesses) process.
These works expand the comprehensiveness of existing safety checks but do not have the ability to create new tasks that might reveal completely unknown FM capabilities.
Other techniques generate new debate topics and evaluate FMs through multi-round debate between them____, discover open-ended programming challenges____, or devise visual recognition and reasoning tasks from a collection of visual assets____.
Meanwhile, ____ produces challenging math problems from existing datasets and human-in-the-loop supervision.
However, the generated tasks in these works tend to focus on a restricted domain, which fails to provide an overview of a model's abilities across a wide array of skills and limits the discovery of surprising capabilities of FMs.