% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;

\documentclass[runningheads]{llncs}
\usepackage{framed}
\usepackage{mdframed}
\usepackage{xcolor}
\usepackage{svg}
\usepackage{booktabs}
\usepackage{enumitem}
%\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{url}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{balance}
\usepackage{makecell}
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\usepackage{marvosym}
\usepackage[colorlinks,
            linkcolor=blue,
            anchorcolor=black,
            citecolor=blue]{hyperref}

\definecolor{shadecolor}{gray}{0.9}

\newcommand{\sq}[1]{\textcolor{cyan}{\textbf{Sihang: #1}}}

\begin{document}
%
\title{AutoS$^2$earch: Unlocking the Reasoning Potential of Large Models for Web-based Source Search}
%
\titlerunning{Large Models for Web-based Source Search}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
% \author{Anonymous Author}
\author{Zhengqiu Zhu\inst{1} \and
Yatai Ji\inst{1} \and
Jiaheng Huang\inst{1} \and
Yong Zhao\inst{1} \and
Sihang Qiu\inst{1} \and 
Rusheng Ju\inst{1}
}
%
\authorrunning{Z. ZHU et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{National University of Defense Technology, Changsha, China \email{zhuzhengqiu@nudt.edu.cn,jiyatai\_1209@nudt.edu.cn,12254745966@qq.com,\ zhaoyong15@nudt.edu.cn,sihangq@acm.org,jrscy@sina.com}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}

Web-based management systems have been widely used in risk control and industrial safety. However, effectively integrating source search capabilities into these systems, to enable decision-makers to locate and address the hazard (e.g., gas leak detection) remains a challenge. While prior efforts have explored using web crowdsourcing and AI algorithms for source search decision support, these approaches suffer from overheads in recruiting human participants and slow response times in time-sensitive situations.
%Web crowdsourcing has emerged as a transformative paradigm for tackling critical source search issues in complex environments, spanning applications from industrial safety (e.g., gas leak detection) to ecological monitoring (e.g., biological signal tracking). By integrating human intelligence, it effectively overcomes purely algorithmic limitations such as local optima and deadlocks. However, this approach also entails substantial costs and places considerable burdens on human workers. 
To address this, we introduce AutoS$^2$earch, a novel framework leveraging large models for zero-shot source search in web applications. AutoS$^2$earch operates on a simplified visual environment projected through a web-based display, utilizing a chain-of-thought prompt designed to emulate human reasoning. The multi-modal large language model (MLLMs) dynamically converts visual observations into language descriptions, enabling the LLM to perform linguistic reasoning on four directional choices. Extensive experiments demonstrate that AutoS$^2$earch achieves performance nearly equivalent to human-AI collaborative source search while eliminating dependency on crowdsourced labor. Our work offers valuable insights in using web engineering to design such autonomous systems in other industrial applications.






\keywords{Web Crowdsourcing \and Source Search \and Multi-modal Large Language Model \and Human-AI Collaboration.}
\end{abstract}


\input{1introduction}
\input{2background}
\input{3design}
\input{4method}
\input{5evaluation}
\input{6discussion}
\input{7conclusion}

% \begin{credits}
% \subsubsection{\ackname} This study is funded by Youth Independent Innovation Foundation of NUDT (ZK-2023-21) and National Natural Science Foundation of China (no. 62202477, 62173337).

% % \subsubsection{\discintname}
% % The authors have no competing interests to declare that are relevant to the content of this article.
% \end{credits}


% ---- Bibliography ----

\bibliographystyle{splncs04}
\bibliography{ref}

\end{document}
