
\subsection{Demographic Information of Participants}
\label{Demographic Information of Participants}
\setlength{\aboverulesep}{0pt}
\setlength{\belowrulesep}{0pt}
\begin{longtable}{|p{1cm}|p{1.5cm}|p{0.8cm}|p{4cm}|p{4cm}|}
\caption{Demographic details of Participants (N=47)} \\ 
\toprule
\textbf{Number} & \textbf{Gender} & \textbf{Age} & \textbf{Education Level} & \textbf{Profession} \\
\midrule

\endfirsthead
\caption[]{Demographic details of Participants (continued)} \\
\toprule
\textbf{Number} & \textbf{Gender} & \textbf{Age} & \textbf{Education Level} & \textbf{Profession} \\ 
\midrule
\endhead

\hline
\endfoot

\bottomrule
\endlastfoot

N1  & Female & 21 & Bachelor's degree & Industrial Design \\ \hline
N2  & Female & 21 & Bachelor's degree & Energy and Power Engineering \\ \hline
N3  & Female & 24 & Bachelor's degree & Safety Engineering \\ \hline
N4  & Female & 32 & Bachelor's degree & Public Relations \& Advertising Professional \\ \hline
N5  & Female & 31 & Bachelor's degree & Computer science \\ \hline
N6  & Male   & 35 & Bachelor's degree & Illustration \\ \hline
N7  & Female & 29 & Master's degree   & Media and communication \\ \hline
N8  & Male   & 36 & Associate degree  & Unity \\ \hline
N9  & Female & 27 & Master's degree & Culture Industry \\ \hline
N10 & Male   & 29 & Master's degree   & Nuclear Science and Technology \\ \hline
N11 & Male   & 24 & Master's degree   & Game Design \\ \hline
N12 & Female & 24 & Master's degree   & Art \\ \hline
N13 & Female & 20 & Bachelor's degree & Visual Communication Design \\ \hline
N14 & Female & 20 & Bachelor's degree & Design \\ \hline
N15 & Male   & 28 & Master's degree   & Computer Science \\ \hline
N16 & Female & 21 & Bachelor's degree & Finance \\ \hline
N17 & Female & 24 & Bachelor's degree & Art and Design \\ \hline
N18 & Female & 22 & Bachelor's degree & Art and Science \& Technology \\ \hline
N19 & Male   & 26 & Master's degree   & Software Development \\ \hline
N20 & Female   & 26 & Bachelor’s degree  & Business English \\ \hline
N21 & Male & 25 & Bachelor's degree & The Internet of Things Engineering \\ \hline
N22 & Male   & 26 & Associate degree   & Law,Psychology,Finance \\ \hline
N23 & Female & 23 & Master's degree & Design \\ \hline
N24 & Male & 25 & Master's degree   & Computer science \\ \hline
N25 & Female & 21 & Bachelor's degree & Electronic and Information Science and Technology \\ \hline
N26 & Female & 28 & Master’s degree & Art and Design \\ \hline
N27 & Female & 29 & Bachelor's degree & Marketing and Planning \\ \hline
N28 & Female & 28 & Master’s degree & Linguistics \\ \hline
N29 & Female & 22 & Bachelor's degree & Journalism and Communication \\ \hline
N30 & Male   & 22 & Bachelor's degree & Mechatronic Engineering \\ \hline
N31 & Female & 23 & Bachelor's degree & Film/Cinema/Media Studies\\ \hline
N32 & Male   & 38 & Bachelor’s degree & IT \\ \hline
N33 & Female   & 20 & Bachelor's degree & Psychology \\ \hline
N34 & Male   & 21 & Bachelor’s degree   & Computer science \\ \hline
N35 & Male   & 21 & Bachelor's degree & New Energy Vehicle Engineering \\ \hline
N36 & Female   & 22 & Bachelor's degree & Accounting \\ \hline
N37 & Female & 57 & Master's degree   & Mathematics and Computer Science \\ \hline
N38 & Prefer not to say & 23 & Bachelor's degree & Media \\ \hline
N39 & Female & 25 & Master’s degree & HCI \\ \hline
N40 & Male & 26 & Master’s degree & Visualization and Visual Analytics and Big Data \\ \hline
N41 & Male   & 21 & Bachelor’s degree   & Computer science \\ \hline
N42 & Female   & 26 & Bachelor’s degree   & Artificial Intelligence \\ \hline
N43 & Male & 26 & Associate degree    & Other \\ \hline
N44 & Female & 28 & Bachelor’s degree   & Internet of Things Engineering \\ \hline
N45 & Female & 24 & Bachelor’s degree   & Human-computer interaction \\ \hline
N46 & Male   & 23 & Master's degree   & Human-Computer Interaction \\ \hline
N47 & Male & 23 & Master's degree   & Design \\ 

\end{longtable}

\subsection{In-game News}
Below, we present the "News" in four rounds. 

\textbf{Round 1}
Dr. Amy claims that the herbal ingredients in R medicine have antiviral, anti-inflammatory, and immune-regulating properties that could potentially treat the Zinc virus. Dr. Amy has a medical background from a top university, has worked at the largest hospital in Southland City for 20 years, and is a member of the Academy of Medicine in South Country. She recently joined a newly established traditional medicine research institution and has started advocating for traditional medicine. She has a large online following, and many people trust her words. Her findings have gained support from those worried about the virus. Eternal Wellness claims that 20\% of  R medicine sales will go toward Zinc virus research and prevention, which has boosted public goodwill. The product, being affordable, has sold out in many places. This newspaper, known for its anti-corruption stance, has further enhanced Dr. Amy’s positive image.

\textbf{Round 2} More patients are dying in the city, with some cases attributed to severe Zinc virus infections. Among those who have died, a few had used the new medicine Max, raising concerns among the public. One widow shared a story on social media about how her husband contracted the Zinc virus, his time in the hospital, and the treatment he received. She suspects that Max was ineffective and believes it may have caused renal impairment, eventually leading to his death. She claims, “He was given a medication we demanded he NOT receive, and his health quickly went downhill,” ultimately resulting in him being “on a ventilator working most of the time at 100\%.” This story has gained significant attention on the largest social media platform, though it remains unclear how widespread these experiences are or whether there is any scientific basis for her claims.

\textbf{Round 3} A journalist investigating the case has discovered that the traditional medicine institution where Dr. Amy works received significant funding from billionaire Jack. Although this institution is associated with a prestigious research institute, there are concerns that it might be leveraging the institute’s name without a true affiliation. Adding to the complexity, billionaire Jack's ex-wife owns a company that produces and promotes traditional medicine products. These connections have led to suspicions that the recent promotion of traditional medicine might be part of a strategic collaboration between billionaire Jack and his ex-wife, raising questions about possible financial motivations behind this push.

\textbf{Round 4} A video circulating on a popular short video platform claims that a doctor practicing alternative medicine and R was targeted and killed to protect the financial interests of "Big Pharma." In the video, his friend, an herbalist with over 14,000 followers, states, "They are cracking down because Big Pharma is losing money." She adds, "My friend was healthy and not suicidal, but we need to stand up because this, this is going to get us all killed." The video has gained significant traction, with 100K views and shared over 30K times. As a result, more people are rallying in favour of traditional medicine, believing it has been effective throughout history and is a proud part of their culture. They advocate for the use of traditional medicine and oppose new drugs, seeing resistance to traditional medicine as an attack on their cultural heritage.
