\documentclass[a4paper, UKenglish, cleveref, autoref, thm-restate]{lipics-v2021}

\usepackage{cite}
%\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

\usepackage{xspace}
\usepackage{bbm}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{thmtools,thm-restate}
\usepackage[textsize=small]{todonotes}
\usepackage{tikz}
\usetikzlibrary{arrows,arrows.meta,automata,positioning} 
\usepackage{stmaryrd}
\usepackage{mathdots}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{thmtools}
\usepackage{lineno}
\usepackage{enumerate}
\usepackage{apxproof}
\usepackage{hyperref}
\usepackage{url}
%\usepackage[bookmarks,bookmarksopen,bookmarksdepth=2]{hyperref}
%\usepackage[appendix=strip]{apxproof}
\renewcommand{\appendixprelim}{\newpage}

\addtolength{\textwidth}{1cm}
\addtolength{\oddsidemargin}{-0.5cm}
\addtolength{\evensidemargin}{-0.5cm}

\addtolength{\textheight}{0.5cm}
\addtolength{\topmargin}{0.5cm}


\title{Reachability in 3-VASS is Elementary}

% anonymised


\author{Wojciech Czerwi\'nski}
{University of Warsaw}
{wczerwin@mimuw.edu.pl}
{https://orcid.org/0000-0002-6169-868X}
{Supported by the ERC grant INFSYS, agreement no. 950398.}

\author{Isma{\"e}l Jecker}
{FEMTO-ST, CNRS, Univ. Franche-Comté, France}
{ismael.jecker@gmail.com}
{}
{}

\author{Sławomir Lasota}
{University of Warsaw}
{s.lasota@uw.edu.pl}
{https://orcid.org/0000-0001-8674-4470}
{Supported by the ERC grant INFSYS, agreement no. 950398 and by the NCN grant 2021/41/B/ST6/00535.}

\author{Łukasz Orlikowski}
{University of Warsaw}
{l.orlikowski@.mimuw.edu.pl}
{https://orcid.org/0009-0001-4727-2068}
{Supported by the ERC grant INFSYS, agreement no. 950398.}


\authorrunning{W. Czerwi{\'n}ski, I. Jecker, S. Lasota, and {\L}. Orlikowski} 
\Copyright{Wojciech Czerwi{\'n}ski, Isma{\"e}l Jecker, S{\l}awomir Lasota, and {\L}ukasz Orlikowski}

\ccsdesc[500]{Theory of computation~Parallel computing models}
\keywords{vector addition systems, Petri nets, reachability problem, dimension three, doubly exponential space, length of shortest path} 

\pagestyle{plain}
\pagenumbering{arabic}

\input{macros}
\input{macros-new}

\nolinenumbers

\begin{document}

\maketitle

\begin{abstract}
The reachability problem in 3-dimensional vector addition systems with states (3-VASS) is known to
be PSpace-hard, and to belong to Tower.
We significantly narrow down the complexity gap by proving the problem to be solvable 
in doubly-exponential space.
The result follows from a new upper bound on the length of the shortest path: 
if there is a path between two configurations of a 3-VASS then there is also
one of at most triply-exponential length.
We show it by introducing a novel technique of approximating the reachability sets of 2-VASS by small semi-linear sets.
\end{abstract}


\maketitle

\input{intro}
\input{defs}
\input{overview}
\input{1comp}
\input{tools}
\input{mainproof}
%\input{sandwich}
\input{future}

%\end{document}

% anonymised
%
\paragraph*{Acknowledgements}
We thank Radosław Piórkowski for many inspiring discussions about cones in VASS a few years ago
and Filip Mazowiecki for the discussions about the reachability problem for \tvass.
We thank also Henry Sinclair-Banks for helpful discussions about \cite[Thm 4.16]{DBLP:conf/focs/0001CMOSW24}.
We thank Qizhe Yang and Yangluo Zheng for the discussions on $3$-VASS, which are geometrically $2$-dimensional.

\newpage

\bibliographystyle{plain}
\bibliography{citat}

\newpage


\appendix

%\input{ATTIC/1comp-copy}

\end{document}
