% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
\usepackage{caption}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%

\usepackage{xcolor}
\usepackage{soul}
\usepackage{fancyhdr}


\begin{document}
%

\newcommand{\acronym}{GNN-XAR}
\title{\acronym: A Graph Neural Network for
Explainable Activity Recognition in Smart Homes}
%
\titlerunning{Explainable GNN for Smart Home HAR}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{
Michele Fiori \and
Davide Mor\and
Gabriele Civitarese\and 
Claudio Bettini}
%
\authorrunning{Fiori et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{University of Milan, Milan, Italy \\
\email{\{michele.fiori, gabriele.civitarese, claudio.bettini\}@unimi.it}\\ \email{d.mor1@campus.unimib.it}}
%
\maketitle              % typeset the header of the contribution
%

\begin{abstract}
Sensor-based Human Activity Recognition (HAR) in smart home environments is crucial for several applications, especially in the healthcare domain. The majority of the existing approaches leverage deep learning models. While these approaches are effective, the rationale behind their outputs is opaque. Recently, eXplainable Artificial Intelligence (XAI) approaches emerged to provide intuitive explanations to the output of HAR models. To the best of our knowledge, these approaches leverage classic deep models like CNNs or RNNs. Recently, Graph Neural Networks (GNNs) proved to be effective for sensor-based HAR. However, existing approaches are not designed with explainability in mind. In this work, we propose the first explainable Graph Neural Network explicitly designed for smart home HAR.
Our results on two public datasets show that this approach provides better explanations than state-of-the-art methods while also slightly improving the recognition rate.

\keywords{Human Activity Recognition  \and Graph Neural Networks \and Smart Homes \and eXplainable AI.}
\end{abstract}

\textit{This is a preprint. Paper accepted for publication at the 21st EAI International Conference on Mobile and Ubiquitous Systems: Computing, Networking and Services (Mobiquitous).}

\input{sections/1-introduction}

\input{sections/2-related}

\input{sections/3-method}

\input{sections/4-experiments}

\input{sections/5-conclusion}


\begin{credits}
\subsubsection{\ackname} 

This work was supported in part by MUSA, SERICS, and FAIR projects under the NRRP MUR program funded by the EU-NGEU. Views and opinions expressed are those of the authors only and do not necessarily reflect those of the European Union or the Italian MUR. Neither the European Union nor the Italian MUR can be held responsible for them.

\subsubsection{\discintname}
The authors have no competing interests to declare that are
relevant to the content of this article.
%It is now necessary to declare any competing interests or to specifically
%state that the authors have no competing interests. Please place the
%statement with a bold run-in heading in small font size beneath the
%(optional) acknowledgments\footnote{If EquinOCS, our proceedings submission
%system, is used, then the disclaimer can be provided directly in the system.},
%for example: The authors have no competing interests to declare that are
%relevant to the content of this article. Or: Author A has received research
%grants from Company W. Author B has received a speaker honorarium from
%Company X and owns stock in Company Y. Author C is a member of committee Z.
\end{credits}

\bibliographystyle{splncs04}
\bibliography{references}

\end{document}