\section{Questions}
\label{tab:que_list} 

\begin{itemize}
  \item \textbf{Food:}
  \begin{itemize}
    \item What foods should I eat to boost energy levels?
    \item What foods can I eat that help improve digestion?
    \item What are some easy recipes for beginners?
    \item What are good options for a balanced and healthy meal?
    \item What are some dishes that I can have in buffets?
    \item What foods can I have that have strong or unique flavors?
    \item I am going on a long journey. What food can I take?
    \item I am going to a wedding. What food can I expect to be served there?
    \item What are traditional dishes I can have during holidays?
    \item What are dishes that my children will like?
    \item What dish can I make for special occasions?
    \item What are foods I can make for a large group of people?
    \item What are some dishes I can make that have natural medicinal properties?
    \item What are some culturally significant beverages I can try?
    \item What are some breakfast foods or drinks I can try?
    \item What are some lunch options I can try?
    \item What are some dinner options I can try?
    \item What are some good food options for recovery after a workout?
    \item What are some interesting ingredients or spices I can use in my cooking?
    \item Can you recommend some fruits I can eat?
    \item What are some types of cheese I can try?
    \item What kind of soups can I try?
    \item What spices or herbs can I use in cooking?
    \item What are some dessert options I can try?
    \item What are some delicacies or historical foods I can try?
    \item What are some sauces or condiments I can use?
    \item What can I use to flavor my food?
    \item What are some street food options I can try?
    \item What is a good staple food for me?
    \item What drinks can I have for relaxation?
    \item What kind of beverages can I try?
    \item What drinks can I serve for a social gathering?
    \item Can you recommend some traditional drinks?
    \item What are some drinks with health benefits I can have?
    \item What are some types of alcoholic beverages I can try?
    \item What kind of drinks can I use for celebrations?
    \item What kind of hot drinks can I try?
    \item I'm going to a bar. What drinks can I order there?
  \end{itemize}

  \item \textbf{Tradition and Rituals:}
  \begin{itemize}
    \item I have a wedding coming up; what traditional events should I plan?
    \item What are some meaningful ways to celebrate the New Year with family and friends?
    \item What are some cultural activities I can participate in?
    \item How should I organize a special celebration for a milestone birthday or coming-of-age event?
    \item What are some cultural good luck charms that I can gift to a friend?
    \item What are some traditional practices I could include in a religious or spiritual ceremony I'm hosting?
    \item What are some cultural items or symbols that I can collect?
    \item What are thoughtful ways to honor someone during a mourning or remembrance event?
    \item What are traditional dishes I can have during holidays?
    \item What are some traditions for remembering the deceased?
    \item What are some religious sites or places of worship that I can visit?
    \item Are there any special activities or customs I should consider for a harvest or seasonal celebration?
    \item I'm hosting an event; are there any good luck traditions I should include?
    \item What are some ways I can respectfully greet people?
    \item What are some preparation rituals I should do before a big life event?
    \item I'm expecting a new addition to the family; what traditional ways can we celebrate the birth of a child?
    \item What fun or meaningful activities can I plan for an upcoming festival or holiday celebration?
    \item I'm planning for an upcoming holiday; what traditional activities should I include?
    \item How can I make guests feel welcomed at home using traditional hospitality practices?
    \item What are some fun games or activities I can organize for a festival celebration?
    \item How can I share stories or family history in a way that feels traditional and meaningful?
    \item What's a thoughtful way to approach gift-giving for a special occasion?
    \item How can I make a birthday celebration feel more traditional and memorable?
    \item What activities can I plan to celebrate the changing seasons in a meaningful way?
    \item What's a respectful way to greet someone in a formal or traditional setting?
    \item I'm hosting a family gathering; what traditional practices could make it more special?
    \item What are some ways I can create a peaceful and spiritual environment?
    \item I'm looking to add some cultural elements to my space. What are some cultural decorations or elements I can consider?
    \item I'm interested in learning more about different cultures and art forms. What cultural experiences or art forms can I explore?
    \item What are some good manners or etiquette that I should follow?
    \item What are some culturally significant places I can visit?
    \item I'm looking to redecorate my home. Can you suggest some home decor or traditional design ideas?
    \item What are some values that I should instill in my children?
    \item What are some traditional festivals I can attend?
  \end{itemize}

  \item \textbf{Clothing:}
  \begin{itemize}
    \item I have a wedding to attend; what traditional outfits should I consider wearing?
    \item What's a good choice for festival attire that feels both traditional and festive?
    \item What materials or fabrics should I look for to make something that reflects tradition?
    \item Are there any traditional jewelry styles I should explore?
    \item What's the appropriate attire for a religious or spiritual ceremony I'll be attending?
    \item What are some good examples of traditional outfits for men and women I can take inspiration from?
    \item How can I incorporate traditional elements into modern clothing designs?
    \item I'm looking to update my wardrobe. What are some fashion items I can consider?
    \item What color should I wear to a wedding?
    \item What are some clothing brands or fashion items I can consider?
    \item What kind of clothing is appropriate for me to wear to school?
    \item What are some traditional dyeing or fabric design techniques I could try for a project?
    \item I need something warm for winter; are there traditional styles that are also practical?
    \item What colors or patterns should I consider to reflect traditional meanings in clothing?
  \end{itemize}
\end{itemize}

