\section{Related Work}
% \todo{Weixin: Consider enhancing Figure 7 by opening with a clear topic sentence that captures the key finding. his would help readers quickly grasp the main relationship you're demonstrating between xxx and yyy.}
% \subsection{DSL for LLM}
% DSL has been used to define the interface between users and LLM agents. ____ designs a mapper DSL as an interface, enabling an existing prompt optimizer to optimize a parallel computing system problem. Differently, their DSL has their own syntax, while ours is embedded in existing language.
\subsection{Self-improvement learning for LLMs}
Self-improvement learning for LLMs typically involves two stages: scoring generated samples (trajectories) and incorporating those samples to enhance the model. Scoring can be achieved through human labeling____ or through automated methods such as verifiers and heuristics____. Our method stands out in this context by utilizing AST analysis, offering a more interpretable approach to scoring. When it comes to incorporating samples, models may rely on retrieval____, reflection____, or reward feedback____. Our method introduces a new mechanism in this stage by adaptively extending and prioritizing high-scoring samples. 

Our method shares similar reinforcement principles with self-improvement learning at the post-training stage____ but is rewarded at a task-level granularity instead of token-level. Specifically, each action is a program implementation instead of token prediction and the state is defined by earned experiences rather than generated sequences.

\subsection{Agentic system organization for specialized tasks}
Task-specific organization has proven effective in enhancing the performance of agentic systems across diverse coding tasks____. We adopt an agentic system organization specifically for ML library development using an ASPL. Such domain knowledge can be further augmented with automatic agentic system design tools____. Furthermore, well-designed interfaces between agents, tools, and other agents have been shown to improve performance____, and a structural IR enables these interfaces to be highly task-aligned in our system.

% LLM agents have been developed for text-to-code____, lifting____, task completion____ using low-resource programming languages. The solutions include: using external tools to correct, guided with formal methods, and embedding into common languages.