\section{Related Work}
To the best of our knowledge, this work is the first to systematically survey and benchmark open-source ZKP frameworks spanning \textit{all} constructions for practical settings and realization. ____ have considered a very limited amount of frameworks, but their industry-led work is largely comparing different constructions to each other (zk-SNARK vs. zk-STARK), rather than comparing frameworks of the same construction to each other (zk-SNARK vs. zk-SNARK). While there are very interesting insights made, we believe that our work is much more objective, systematic and extensive, while also adding the element of usability and accessibility analysis.
Another survey on ZKP frameworks has been conducted ____, however this work only focuses on zk-SNARKs and does not look into the usability, accessibility, or performance of the chosen frameworks. Also, the work is largely focused on the application of zk-SNARKs in the blockchain. Similarly, ____ conducts a survey on non-interactive ZK applications in the blockchain. The main focus of this work is the analysis of privacy-protection schemes for smart contracts. This work approaches ZK scheme analysis from a theoretical standpoint. Rather than focusing on usability and accessibility, this work primarily focuses on analyzing the asymptotics of the available schemes. Unfortunately, this is not fully representative of the performance of these schemes in practice, due to some schemes having high constants in their asymptotic complexities. While we believe that surveying zk-SNARKs is very important, we note that many zk-SNARK schemes are not post-quantum secure. As post-quantum security becomes a rapidly growing concern, our work purposefully inspects every available ZKP construction to provide insights into post-quantum secure frameworks, alongside more established zk-SNARK frameworks. We believe that limiting our work to zk-SNARKs would not be fully representative of the ZK landscape.

We model our paper after the seminal surveys in privacy-preserving technology centered around MPC ____ and FHE ____. 
% As these techniques have been more extensively studied, the accompanying tools are more mature and their insights are more conclusive.
Like these works, we aim to provide as detailed of a description as we can surrounding the usability, accessibility, and performance of our chosen frameworks, while providing a digestible guide for developers choosing a tool for their ZK-based applications.