\section{Conclusion and Discussion} \label{sec:discussion}

We explore the question of privately answering linear queries with both relative and additive errors. Perhaps surprisingly, we give an approximate-DP algorithm which, if a constant relative error is allowed, can achieve the additive error that is polylogarithmic in $n, |\cX|, |\cF|, 1/\delta$. In comparison, without any relative error, the additive errors have to be polynomial in one of these parameters. We also show a nearly-matching lower bound for a large regime of parameters.

Perhaps the most obvious open question from our work is the pure-DP case. Our algorithm's additive error still has a dependency of $\sqrt{n}$. Is it possible to, similar to approximate-DP, achieve an additive error that is poly-logarithmic in $n, |\cX|, |\cF|, 1/\delta$ (for any small constant relative error $\zeta$)? We remark that the main challenge in obtaining this through our PREM framework lies in the SVT algorithm. Namely, to discover many queries above/below thresholds with pure-DP, the privacy budget grows with the number of discovered queries. 
If such a task can be accomplished without the increased budget, then applying PREM would result in the desired error bound. We note that, in a recent work, \cite{GhaziKRMS24} devises such a procedure, but only for \emph{above} threshold queries; extending their technique to \emph{below} threshold queries might answer this question.

Another research direction is to consider specific set of queries, such as $k$-marginal or graph cut queries. In fact, this direction was undertaken already by aforementioned previous work, e.g.,~\cite{EpastoMMMVZ23,Ghazi:2023} who considered range queries and hierarchical queries, respectively. Focusing on a specific query family may allow for better error guarantees, and more efficient algorithms. For the latter, we note once again that our algorithm runs in time polynomial in $n, |\cF|, |\cX|$. 
One might hope for an algorithm that runs in time polynomial in $n, |\cF|, \log|\cX|$; however, under cryptographic assumptions, this is known to be impossible for general family $\cF$ (and even for $k$-marginal queries)~\citep{UllmanV20}. However, such an efficient algorithm may exist for other specific query families, and this remains an intriguing direction for future research.

