\documentclass[
    sigconf,
    nonacm
]{acmart}

\newcommand\vldbdoi{1}
\newcommand\vldbpages{1 - 1}
\newcommand\vldbvolume{1}
\newcommand\vldbissue{1}
\newcommand\vldbyear{1}
\newcommand\vldbauthors{\authors}
\newcommand\vldbtitle{\shorttitle} 
\newcommand\vldbavailabilityurl{https://github.com/eth-easl/mixtera}

\newcommand\vldbpagestyle{plain} 

\newcommand{\ana}[1]{{\textcolor{orange}{[~ANA:~#1~]}}}
\newcommand{\xzyao}[1]{{\textcolor{blue}{[#1]}}}

\input{meta/packages}
\input{meta/commands}


\begin{document}

\title{\mixtera: A Data Plane for Foundation Model Training}

\settopmatter{authorsperrow=3}

\author{Maximilian Böther}
\orcid{0000-0003-4093-4361}
\authornote{Correspondence to \href{mailto:mboether@ethz.ch}
{mboether@ethz.ch}.}
\affiliation{%
  \institution{ETH Zurich, Switzerland}
}

\author{Xiaozhe Yao}
\orcid{0000-0002-4661-533X}
\affiliation{%
  \institution{ETH Zurich, Switzerland}
}

\author{Tolga Kerimoglu}
\orcid{0009-0004-1175-338X}
\affiliation{%
  \institution{ETH Zurich, Switzerland}
}

\author{Dan Graur}
\orcid{0009-0001-0682-2422}
\affiliation{%
  \institution{ETH Zurich, Switzerland}
}


\author{Viktor Gsteiger}
\orcid{0000-0002-6750-5500}
\affiliation{%
  \institution{ETH Zurich, Switzerland}
}


\author{Ana Klimovic}
\orcid{0000-0001-8559-0529}
\affiliation{%
  \institution{ETH Zurich, Switzerland}
}

\begin{abstract}
    \input{content/1_abstract}
\end{abstract}

\maketitle

\pagestyle{\vldbpagestyle}
\begingroup\small\noindent\raggedright\textbf{PVLDB Reference Format:}\\
\vldbauthors. \vldbtitle. PVLDB, \vldbvolume(\vldbissue): \vldbpages, \vldbyear.
\href{https://doi.org/\vldbdoi}{doi:\vldbdoi}
\endgroup
\begingroup
\renewcommand\thefootnote{}\footnote{\noindent
This work is licensed under the Creative Commons BY-NC-ND 4.0 International License. Visit \url{https://creativecommons.org/licenses/by-nc-nd/4.0/} to view a copy of this license. For any use beyond those covered by this license, obtain permission by emailing \href{mailto:info@vldb.org}{info@vldb.org}. Copyright is held by the owner/author(s). Publication rights licensed to the VLDB Endowment. \\
\raggedright Proceedings of the VLDB Endowment, Vol. \vldbvolume, No. \vldbissue\ %
ISSN 2150-8097. \\
\href{https://doi.org/\vldbdoi}{doi:\vldbdoi} \\
}\addtocounter{footnote}{-1}\endgroup

\ifdefempty{\vldbavailabilityurl}{}{
\vspace{.3cm}
\begingroup\small\noindent\raggedright\textbf{PVLDB Artifact Availability:}\\
The source code, data, and/or other artifacts have been made available at \url{\vldbavailabilityurl}.
\endgroup
}

\input{content/2_intro}
\input{content/3_background}
\input{content/4_challenges}

\input{content/5_design}
\input{content/6_impl}
\input{content/7_eval}
\input{content/8_conclusion}


\begin{acks}
We thank Beste Güney for her contributions to \mixtera's codebase.
We thank  Antoni-Joan Solergibert i Llaquet, Imanol Schlag, Steven Hand, Martin Jaggi, Antoine Bosselut, Alexander Hägele, Loubna Ben Allal, Quentin de Laroussilhe, Paul Barham, Yiding Jiang, Theodoros Rekatsinas, and Foteini Strati for helpful discussions.
This work was supported as part of the Swiss AI Initiative by a grant from the Swiss National Supercomputing Centre (CSCS) under project ID a09 on Alps.
Maximilian Böther is supported by the Swiss National Science Foundation (project number 200021\_204620).

\end{acks}

\pagebreak
\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\end{document}
