\section{Related Work}
\label{sec:related}
\paratitle{LLMs in Spatial-temporal Tasks}
Considerable efforts have been dedicated to using LLMs to improve the performance of spatial-temporal tasks~\cite{yu2024bigcity}. For instance, GeoGPT~\cite{geogpt} introduced an LLM-based tool capable of automating the processing of geographic data, but it does not delve into extracting detailed information about locations. GEOLLM~\cite{geollm} designed prompts that include coordinates, address, and surrounding buildings, but it can only address simple questions in a Q\&A format and are unable to handle complex tasks like POI recommendation. Besides, they fail to fully extract the semantic information of POIs.
Some researchers have used LLMs as backbones to tackle complex real-world tasks. For example, GATGPT~\cite{gatgpt} input spatial-temporal features into a frozen LLM to predict traffic speeds, while ST-LLM~\cite{stllm} used a partially frozen LLM to forecast traffic flow. However, these methods are designed for specific individual problems and cannot be applied across multiple tasks.
To solve these limitations, we designed three types of special prompts to extract the semantic information of POIs from LLMs effectively.

\paratitle{POI Representation with Semantic Information}
POI representation aims to turn each POI into a vector that can be utilized in various downstream tasks like traffic forecasting tasks~\cite{MLPST, PDFormer,stssl} and trajectory tasks~\cite{controltraj, START, gan}. Most existing methods like ~\cite{ppr}, leverage textual features typically using one-hot code to encode POI categories and then concatenate them with the embedding vectors. For data types like check-in content,~\cite{ge} model the similarity between POIs by constructing a POI-Word relationship graph, while ~\cite{cape} draws inspiration from Word2Vec method, simultaneously training word vectors and POI vectors. However, these methods often fall short in preserving semantic information and achieving a more comprehensive integration during the fusion process.
To address this issue, we designed three modules in the Embedding Enhancement to improve the preservation and integration of semantic information in the POI embedding.

%%%6666666666_conclusion_66666666666%%%%