\begin{thebibliography}{}

\bibitem[Aiyagari, 1994]{aiyagari1994uninsured}
Aiyagari, S.~R. (1994).
\newblock Uninsured idiosyncratic risk and aggregate saving.
\newblock {\em The Quarterly Journal of Economics}, 109(3):659--684.

\bibitem[Bilbiie, 2008]{bilbiie2008limited}
Bilbiie, F.~O. (2008).
\newblock Limited asset markets participation, monetary policy and (inverted) aggregate demand logic.
\newblock {\em Journal of Economic Theory}, 140(1):162--196.

\bibitem[Brock and Hommes, 1998]{brock1998heterogeneous}
Brock, W.~A. and Hommes, C.~H. (1998).
\newblock Heterogeneous beliefs and routes to chaos in a simple asset pricing model.
\newblock {\em Journal of Economic dynamics and Control}, 22(8-9):1235--1274.

\bibitem[Chen et~al., 2023]{chen2023}
Chen, Y., Liu, T.~X., Shan, Y., and Zhong, S. (2023).
\newblock The emergence of economic rationality of gpt.
\newblock {\em Proceedings of the National Academy of Sciences}, 120(51):e2316205120.

\bibitem[Gal{\'\i} et~al., 2004]{gali2004rule}
Gal{\'\i}, J., L{\'o}pez-Salido, J.~D., and Vall{\'e}s, J. (2004).
\newblock Rule-of-thumb consumers and the design of interest rate rules.
\newblock {\em Journal of Money, Credit and Banking}, pages 739--763.

\bibitem[Horton, 2023]{horton2023}
Horton, J.~J. (2023).
\newblock Large language models as simulated economic agents: What can we learn from homo silicus?
\newblock {\em arXiv preprint arXiv:2301.07543}.

\bibitem[Kaplan et~al., 2018]{kaplan2018monetary}
Kaplan, G., Moll, B., and Violante, G.~L. (2018).
\newblock Monetary policy according to hank.
\newblock {\em American Economic Review}, 108(3):697--743.

\bibitem[Krusell and Smith, 1998]{krusell1998income}
Krusell, P. and Smith, Jr, A.~A. (1998).
\newblock Income and wealth heterogeneity in the macroeconomy.
\newblock {\em Journal of political Economy}, 106(5):867--896.

\bibitem[Li et~al., 2024]{li2024}
Li, N., Gao, C., Li, Y., and Liao, Q. (2024).
\newblock Large language model-empowered agents for simulating macroeconomic activities.
\newblock Available at SSRN: https://ssrn.com/abstract=4606937.

\bibitem[Ma, 2024]{ma2024}
Ma, J. (2024).
\newblock Can machines think like humans? a behavioral evaluation of llm-agents in dictator games.
\newblock {\em arXiv preprint arXiv:2410.21359}.

\bibitem[Mei et~al., 2024]{mei2024}
Mei, Q., Xie, Y., Yuan, W., and Jackson, M.~O. (2024).
\newblock A turing test of whether ai chatbots are behaviorally similar to humans.
\newblock {\em Proceedings of the National Academy of Sciences}, 121(9):e2313925121.

\bibitem[Pang, 2024]{pang2024}
Pang, X. (2024).
\newblock 人工智能赋能社会科学研究探析——生成式行动者、复杂因果分析与人机科研协同.
\newblock {\em 世界经济与政治}, (7).

\bibitem[Tesfatsion, 2006]{tesfatsion2006handbook}
Tesfatsion, L. (2006).
\newblock Agent-based computational economics: A constructive approach to economic theory.
\newblock In Tesfatsion, L. and Judd, K.~L., editors, {\em Handbook of Computational Economics}, volume~2, chapter~16, pages 831--880. Elsevier.

\bibitem[Wu et~al., 2024]{wu2024}
Wu, X., Whittington, D., Chen, Y., and Zuo, R. (2024).
\newblock The role of generative ai in navigating trade-offs in policy research design: balancing validity, rigour, and innovation.
\newblock {\em Journal of Asian Public Policy}, pages 1--27.

\bibitem[Xie et~al., 2024]{xie2024}
Xie, Y., Liu, Y., Ma, Z., Shi, L., Wang, X., Yuan, W., Jackson, M.~O., and Mei, Q. (2024).
\newblock How different ai chatbots behave? benchmarking large language models in behavioral economics games.
\newblock {\em arXiv preprint arXiv:2412.12362}.

\end{thebibliography}
