\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
% \usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{algpseudocode}
\usepackage{color}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[inline]{enumitem}
\usepackage{ulem}
\usepackage{makecell}
\newcommand{\jy}[1]{\textcolor{blue}{#1}}
\newcommand{\gq}[1]{\textcolor{green}{#1}}
\newcommand{\yl}[1]{\textcolor{brown}{#1}}
\newcommand{\st}[1]{\textcolor{red}{#1}}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Transfer Risk Map: Mitigating Pixel-level Negative Transfer in Medical Segmentation \\
% {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used}
\thanks{\IEEEauthorrefmark{2} Corresponding author. 
% This work is supported in part by the Natural Science Foundation of China (Grant 62371270) and Shenzhen Key Laboratory of Ubiquitous Data Enabling (No.ZDSYS20220527171406015).
}
}

% \author{\IEEEauthorblockN{Shutong Duan$^{1}$, Jingyun Yang$^1$, Yang Tan$^{1}$, Guoqing Zhang$^{1}$, Yang Li$^{1,}$\IEEEauthorrefmark{2}, Xiao-Ping Zhang$^{1,}$}
%  \IEEEauthorblockA{\textit{1 Shenzhen Key Laboratory of Ubiquitous Data Enabling, Tsinghua Shenzhen International Graduate School, Tsinghua University}  \\
%   yangli@sz.tsinghua.edu.cn, xpzhang@ieee.org
%   }
%   % \IEEEauthorblockA{\IEEEauthorrefmark{2} Corresponding author. E-mail: yangli@sz.tsinghua.edu.cn}
%   }

\author{\IEEEauthorblockN{Shutong Duan, Jingyun Yang, Yang Tan, Guoqing Zhang, Yang Li\IEEEauthorrefmark{2}, Xiao-Ping Zhang}
 \IEEEauthorblockA{\textit{Shenzhen Key Laboratory of Ubiquitous Data Enabling}  \\
\textit{Tsinghua Shenzhen International Graduate School, Tsinghua University}  \\
  yangli@sz.tsinghua.edu.cn
  }
  % \IEEEauthorblockA{\IEEEauthorrefmark{2} Corresponding author. E-mail: yangli@sz.tsinghua.edu.cn}
  }
  


\maketitle

\begin{abstract}
How to mitigate negative transfer in transfer learning is a long-standing and challenging issue, especially in the application of medical image segmentation. Existing methods for reducing negative transfer focuses on classification or regression tasks, ignoring the non-uniform  negative transfer risk in different image regions.  
%risk of negative transfer of a given model does not uniformly distribute over the whole image, which necessitates the need for a fine-grained transfer learning approach that could adapt to variable negative transfer risk at different image regions.
%Previous works have incorporated transferability into fine-tuning to address this issue, but their performance on medical segmentation tasks is limited due to fewer semantic categories, tiny segmentation foreground, and severed class imbalance of medical images, limiting their practicality.
%{However, the risk of negative transfer does not uniformly distribute over the image. How to accurately evaluate the local negative transfer risk remains a challenging problem. }
In this work, we propose a simple yet effective weighted fine-tuning method that directs the model’s attention towards regions with significant transfer risk for medical semantic segmentation. Specifically, we compute a transferability-guided transfer risk map to quantify the transfer hardness for each pixel and the potential risks of negative transfer. During the fine-tuning phase, we introduce a map-weighted loss function, normalized with image foreground size to counter class imbalance. 
Extensive experiments on brain segmentation datasets show our method significantly improves the target task performance, with gains of 4.37\% on FeTS2021 and 1.81\% on iSeg-2019, avoiding negative transfer across modalities and tasks. Meanwhile, a 2.9\% gain under a few-shot scenario validates the robustness of our approach.
% Extensive experiments on brain tumor and brain matter segmentation datasets demonstrate that our proposed method achieves significantly enhanced performance when transferring knowledge between distinct modalities and tasks, with 4.37\% gain on FeTS2021 and 1.81\% gain on iSeg-2019, indicating that it indeed avoids negative transfer from diverse modalities and tasks. 
%{we propose a transferability based pixel-level transfer risk map that quantitatively represents the potential risks of negative transfer for each pixel. Furthermore, we propose a novel weighted fine-tuning approach guided by the transfer risk map. The proposed method is designed to direct the model's attention towards regions with significant transfer hardness, thus mitigating the potential risk of local negative transfer and enhancing the overall effectiveness of the fine-tuning process.}
% Extensive experiments on brain segmentation datasets demonstrate that our proposed weighted fine-tuning consistently outperforms the vanilla fine-tuning in all transfer experiments, with 4.37\% gain on FeTS 2021 and 1.81\% gain on iSeg-2019. A 2.9\% gain in Dice score under a few-shot scenario validates the robustness of our approach, highlighting its capacity to facilitate model adaptation and refinement during the fine-tuning phase by mitigating the impact of localized negative transfer.

\end{abstract}

\begin{IEEEkeywords}
Medical Image Segmentation, Transfer Learning, Transferability Estimation, Negative Transfer, Model Fine-tuning
\end{IEEEkeywords}
\input{sections/introduction}
\input{sections/method}
\input{sections/experiments}
\input{sections/conclusion}

\section*{Acknowledgment}
This work was supported in part by the Natural Science Foundation of China (Grant 62371270), Shenzhen Key Laboratory of Ubiquitous Data Enabling (No.ZDSYS20220527171406015), and Tsinghua Shenzhen International Graduate School-Shenzhen Pengrui Endowed Professorship Scheme of Shenzhen Pengrui Foundation.

\bibliographystyle{IEEEtran}
\normalem
\bibliography{main}

\vspace{12pt}

\end{document}
