\documentclass[12pt]{article}
\usepackage[margin=1.25in]{geometry}
\usepackage{times}

\usepackage[round]{natbib}
\usepackage{hyperref,url,booktabs}
\usepackage{nicefrac,microtype,xcolor}
\usepackage{graphicx,tikz}
\usepackage{subfigure}
\usepackage{amsmath,amsthm,amssymb,amsfonts}
\usepackage{mathtools,mathrsfs,bm}
\usepackage{algorithm,algorithmic}

\usepackage{enumitem}

\mathtoolsset{showonlyrefs}

\usepackage{caption}
\usepackage{authblk}
\usepackage{etoc}
\usepackage{wrapfig} 

\renewcommand{\ttfamily}{\fontencoding{T1}\fontfamily{lmtt}\selectfont}


\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{backgrounds}
\usetikzlibrary{calc}
\tikzset{>={angle 60}}

\usepackage{accents}
\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}

\allowdisplaybreaks

\input{junos_math}
\hypersetup{
  colorlinks=true,
  linkcolor={red!60!black},
  citecolor={blue!60!black},
  urlcolor={magenta!60!black}
}  
\setlist[itemize,enumerate]{}
\setenumerate[1]{label={\upshape(\arabic*)}}


\newtheorem{manualtheoreminner}{Assumption}
\newenvironment{manualtheorem}[1]{%
  \IfBlankTF{#1}
    {\renewcommand{\themanualtheoreminner}{\unskip}}
    {\renewcommand\themanualtheoreminner{#1}}%
  \manualtheoreminner
}{\endmanualtheoreminner}

\newcommand{\DW}[1]{\textcolor{violet}{\bf (DW: #1)}}

\usepackage{comment}

\title{
\makebox[\textwidth][c]{Metastable Dynamics of Chain-of-Thought Reasoning:}\\
Provable Benefits of Search, RL and Distillation} 


\author{
Juno Kim\thanks{University of Tokyo and RIKEN AIP. \texttt{junokim@g.ecc.u-tokyo.ac.jp}.} ,\,\,
Denny Wu\thanks{New York University and Flatiron Institute. \texttt{dennywu@nyu.edu}.} ,\,\,
Jason D.~Lee\thanks{Princeton University. \texttt{jasonlee@princeton.edu}.} ,\,\,
Taiji Suzuki\thanks{University of Tokyo and RIKEN AIP. \texttt{taiji@mist.i.u-tokyo.ac.jp}.}
 \vspace{-2.5mm}
}


%\let\usebigfont\undefined
%\def\usebigfont{1}
\ifdefined\usebigfont
\renewcommand{\mddefault}{bx}
\renewcommand\seriesdefault{b}
\renewcommand\bfdefault{bx} % just to make sure
\usepackage{times}
\usepackage[fontsize=13pt]{scrextend}
\makeatletter
\@ifpackageloaded{geometry}{\AtBeginDocument{\newgeometry{letterpaper,left=1.56in,right=1.56in,top=1.71in,bottom=1.77in}}}{\usepackage[letterpaper,left=1.56in,right=1.56in,top=1.71in,bottom=1.77in]{geometry}}
\AtBeginDocument{\newgeometry{letterpaper,left=1.56in,right=1.56in,top=1.71in,bottom=1.77in}}
\linespread{1}
\usepackage{hyperref} % watch out for error: ! Illegal parameter number in definition of \Hy@tempa.

\else
\fi

\begin{document}
\etocdepthtag.toc{mtchapter}
\etocsettagdepth{mtchapter}{subsection}
\etocsettagdepth{mtappendix}{none}

\maketitle 

\input{abstract}
\input{intro}
\input{result}
\input{conclusion}

\bigskip

\subsection*{Acknowledgments} 

The authors thank Jimmy Ba and Yuexiang Zhai for helpful discussions. JK was partially supported by JST CREST (JPMJCR2015). JDL acknowledges support of the NSF CCF 2002272, NSF IIS 2107304, and NSF CAREER Award 2144994. TS was partially supported by JSPS KAKENHI (24K02905, 20H00576) and JST CREST (JPMJCR2115). This research is unrelated to DW's work at xAI. 

\bigskip

{
\begin{small} 

\bibliography{ref}
\bibliographystyle{plainnat} 
\end{small} 
}

\newpage
{
\small 
\hypersetup{linkcolor=black}
\renewcommand{\contentsname}{Table of Contents}
\tableofcontents
}

\newpage
\appendix
\input{appendix}


\end{document}