\label{dataset} We conduct our experiments on various datasets related to different graph learning tasks, as described below.
% [labelindent=0pt,labelwidth=\widthof{\ref{last-item}},itemindent=1em,leftmargin=!]


\begin{enumerate}
\item \label{data:iot}\textbf{IoT} \cite{guo2019deep}: This dataset is used for malware confinement tasks. It includes three subsets with graph orders of 20, 40, and 60, and average node degrees of 7.42, 11.22, and 9.51, respectively. Each subset contains 343 graph pairs with contextual parameters: infection rate, recovery rate, and decay rate. Node attributes indicate infection status, and edge attributes represent the distance between nodes.

\item \label{data:chem}\textbf{Chemistry Reaction} \cite{guo2019deep, lowe2014patent}: This dataset is used to forecast the outcome of a chemical reaction based on the reactants. Molecular graphs have atoms as nodes and bonds as edges. There are 7,180 atom-mapped reactant-product pairs in the dataset with an average node degree of 1.75. Atom characteristics consist of elemental identity, connectivity, hydrogen count, valence, and aromaticity. Moreover, bond characteristics encompass bond type and connectivity.


% \item \label{data:iot}\textbf{IoT} \cite{guo2019deep}: This dataset is used for malware confinement task. There are three subsets with graph orders 20, 40, and 60, having average node degree of 7.42, 11.22, 9.51, respectively. 343 graph pairs are available in each subset. Each graph pair has three contextual parameters - infection rate, recovery rate and decay rate. Node attribute is infection status (i.e., whether the device corresponding to a node is compromised or not) and edge attribute is distance between two nodes.

% \item \label{data:chem}\textbf{Chemistry Reaction} \cite{guo2019deep}\cite{lowe2014patent}: This dataset is used for predicting the product (target graph) of chemical reaction given the reactant (input graph). Each molecular graph consists of atoms as nodes and bonds as edges. The reactions used for training are atom-mapped so that each atom in the product graph has a unique corresponding atom in the reactants. 
% The dataset contains 7,180 reactions (reactant-product graph pairs) with average node degree of 1.75. Atom (node) features include its elemental identity, degree of connectivity, number of attached hydrogen atoms, implicit valence, and aromaticity. Bond (edge) features include bond type (single, double, triple, or aromatic), and whether it is connected.

\item \label{data:socailnet}\textbf{Social Networks}: We have selected two datasets used for directed link prediction task.
\begin{enumerate}
\item \label{data:epi}\textbf{Epinions}
\cite{wang2020edge2vec}\cite{Tang-etal12c}: It is a large social network with 22,166 nodes and 355,217 directed edges. Here, nodes are users and edges represent trust between users. Users provided ratings (1-5) of different products under 27 categories and each rating has a helpfulness score (1-5). 
\item \label{data:slash}\textbf{Slashdot}
\cite{wang2020edge2vec}\cite{leskovec2009community}: It is another large social network graph with 82,168 nodes and 870,161 directed edges. Here, nodes are users and edges represent friendship between users.
\end{enumerate}

\item \label{data:community}\textbf{Community Detection}: In this task, label of only one node from each community is known during training, making it a semi-supervised task. Here, edges represent interactions between nodes.
\begin{enumerate}
\item \label{data:zach}\textbf{Zachary’s karate club} \cite{zachary1977information}\cite{lu2018community}\cite{DBLP:journals/debu/HamiltonYL17}\cite{cai2019community}: It is a small bidirectional graph with 34 nodes and 156 edges. Some researchers (on the basis of their notion of community) have reported the optimal number of communities to be two, whereas others have reported it to be four.
\item \label{data:acf}\textbf{American College football} \cite{girvan2002community}\cite{lu2018community}\cite{cai2019community}: It is another small bidirectional graph with 115 nodes, 1,226 edges and 12 communities. The community sizes are non-uniform.
\item \label{data:books}\textbf{Books about US politics} \cite{krebs2004books}\cite{cai2019community}: It is also a small bidirectional graph with 105 nodes, 882 edges and 3 communities (i.e., liberal, conservative and neutral). Collection size under neutral category is much less than the other two categories.
\end{enumerate}

\item \label{data:biomed}\textbf{Biomedical}:
\begin{enumerate}
\item \label{data:hcp}\textbf{Human Connectome Project (HCP\footnote{https://compneuro.neuromatch.io/projects/fMRI/README.html})} \cite{van2013wu}: The dataset comprises 339 samples with structural (SC) and functional (FC) brain connectivity data across five cognitive tasks (rest, gambling, language, motor, emotion). Each network consists of 360 nodes that correspond to distinct brain regions, featuring adjacency matrices based on correlation. The objective is to forecast functional connectivity (FC) utilizing structural connectivity (SC).

\item \label{data:ddi}\textbf{Drug Drug Interaction (BioSNAP-DDI\footnote{http://snap.stanford.edu/biodata/datasets/10001/10001-ChCh-Miner.html})} \cite{zitnikbiosnap}: This dataset contains interactions between 1,514 FDA-approved drugs, with 48,514 bidirectional edges. The task is to predict drug interactions (edges).

\item \label{data:dti}\textbf{Drug Target Interaction (BioSNAP-DTI\footnote{https://snap.stanford.edu/biodata/datasets/10002/10002-ChG-Miner.html})} \cite{zitnikbiosnap}: This dataset includes interactions between 5,017 drugs and 2,324 target proteins, with 15,138 bidirectional edges. The task is to predict drug-target interactions.

\item \label{data:ppi}\textbf{Protein Protein Interaction (HuRI-PPI\footnote{http://www.interactome-atlas.org/data/HuRI.tsv})} \cite{luck2020reference}: This dataset contains interactions between 8,245 proteins with 52,068 bidirectional edges. The task is to predict protein-protein interactions.
\end{enumerate}





% \begin{enumerate}
% \item \label{data:hcp}\textbf{Human Connectome Project (HCP\footnote{https://compneuro.neuromatch.io/projects/fMRI/README.html}) }\cite{van2013wu}: It is a subset of the full HCP dataset of 1,200 young adults. Here, 339 subjects are present with their structural connectivity (SC) and functional connectivity (FC) of brain while performing five different tasks (rest, gambling, language, motor, and emotion). 360 regions of the brain are represented by the nodes in both the graphs (SC, FC). All graphs are fully connected. A correlation based connectedness score ([-1, 1]) is available between all possible pair of regions. This correlation based matrices are the adjacency matrices of the graphs. The task is to predict the FC given SC for different tasks.

% \item \label{data:ddi}\textbf{Drug Drug Interaction (BioSNAP-DDI\footnote{http://snap.stanford.edu/biodata/datasets/10001/10001-ChCh-Miner.html})} \cite{zitnikbiosnap}: This is a network of interactions betweeen drugs, which are approved by the U.S. Food and Drug Administration. The dataset consists of 1,514 nodes representing drugs and 48,514 bidirectional edges between them representing drug interactions. The task is to predict presence of edges (i.e., interactions) between two nodes (i.e., drugs).

% \item \label{data:dti}\textbf{Drug Target Interaction (BioSNAP-DTI\footnote{https://snap.stanford.edu/biodata/datasets/10002/10002-ChG-Miner.html})} \cite{zitnikbiosnap}: This is a drug-target interaction network that contains information on which genes (i.e., proteins encoded by genes) are targeted by which drugs that are on the U.S. market. The drugs on average have 5-10 unique target proteins. The DTI dataset consists of 5,017 nodes that represent drugs, 2,324 nodes that represent targets and 15,138 bidirectional edges that represent drug target interactions. The task is to predict presence of edges i.e., interactions between a drug-node and a target-node.

% \item \label{data:ppi}\textbf{Protein Protein Interaction (HuRI-PPI\footnote{http://www.interactome-atlas.org/data/HuRI.tsv})} \cite{luck2020reference}: To detect protein-protein interactions (PPI), pairwise combinations of human protein-coding genes are tested systematically using high throughput yeast two-hybrid screens. The quality of these interactions is further validated in multiple orthogonal assays. The dataset consists of 8,245 nodes representing proteins and 52,068 bidirectional edges between them representing PPI. Given two proteins as input, the task is to predict whether or not any interaction will happen between them.
% \end{enumerate}

\begin{comment}
\item \label{data:movie}\textbf{MovieLens-100k} \cite{miller2003movielens}: This is a popular recommendation system dataset, a medium sized bipartite graph with 943 user nodes and 1682 movie nodes. There are total 100,000 ratings (1-5) from user nodes to the movie nodes. User attributes are age (7-73), gender (male, female), occupation (21 types) and movie features are release date (year 1922-1998) and genre (19 types).
\end{comment}
%\item 

\item \label{data:yago}\textbf{YAGO} \cite{suchanek2007yago}: YAGO\footnote{https://yago-knowledge.org/} is a Knowledge Graph that adds common knowledge facts from Wikipedia to WordNet, transforming it into a common knowledge base. It is a large directed knowledge graph where nodes represent entities and directed edges describe facts between them. YAGO includes 37 relation types and each entity has at least 10 relations. The knowledge graph contains 123,131 entities and 1,089,040 relations between them.
% \item \textbf{The Penn Treebank} \cite{marcus-etal-1993-building}: The task of dependency parsing is projected as a graph translation problem. The source is word-level co-occurrence graph of the sentence to be parsed and the target is the dependency tree of the sentence. Total 45 different dependency relations (edges) are present in the dataset. \hl{Order of edge-set is very less because} source graph is similar to line-graph and target graph is a tree.
% \item \textbf{SemEval-2010 Task 8} \cite{hendrickx2010semeval}: In the SemEval-2010 Task 8 on ``Multi-way classification of mutually exclusive semantic relations between pairs of nominals'', given a sentence and two tagged nominals, the task is to predict the directed relation between those nominals. There are 10 different types of relations present in the dataset among which 9 relations are directional and one relation is bidirecional. Total 10,717 annotated examples (sentences along with tagged entity-pairs) are present in the dataset.

\end{enumerate}



%Fig. \ref{fig:datasets} pictorially presents the different datasets belonging to different tasks in different domains on which we carried out experiments with our model.

%\input{Latex/FigDef/model_datasets}