\documentclass[11pt,fullpage,letterpaper]{article} % For LaTeX2e

\usepackage{natbib}

\bibliographystyle{plainnat}
% modification to natbib citations
\setcitestyle{authoryear,round,citesep={;},aysep={,},yysep={;}}
\usepackage[english]{babel}

\oddsidemargin 0in
\evensidemargin 0in
\textwidth 6.5in
\topmargin -0.5in
\textheight 9.0in
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[C]{Non-Monetary Mechanism Design without Distributional Information}
\fancyhead[L]{}
\fancyhead[R]{}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts

\usepackage{titletoc}
\usepackage[toc, page, header]{appendix} %%% MAKE SURE TO PUT THIS BEFORE hyperref PACKAGE

\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue,urlcolor=black]{hyperref}
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

\usepackage{indentfirst}

\title{Non-Monetary Mechanism Design without Distributional Information: Using Scarce Audits Wisely}

\author{%
    Yan Dai~\thanks{Massachusetts Institute of Technology. Email: \texttt{yandai20@mit.edu}.}\and Mo\"ise Blanchard~\thanks{Columbia University. Email: \texttt{mb5414@columbia.edu}.}\and Patrick Jaillet~\thanks{Massachusetts Institute of Technology. Email: \texttt{jaillet@mit.edu}.}
}
\date{}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}

\usepackage{cleveref} % ===== BE SURE TO PUT THIS BEFORE \newtheorem's
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\mathsc}[1]{{\normalfont\textsc{#1}}}
\newcommand{\E}{\operatornamewithlimits{\mathbb{E}}}
\newcommand{\argmax}{\operatornamewithlimits{\mathrm{argmax}}}
\newcommand{\argmin}{\operatornamewithlimits{\mathrm{argmin}}}
\newcommand{\tr}{\operatorname{\mathrm{Tr}}}
\renewcommand{\O}{\operatorname{\mathcal O}}
\newcommand{\Otil}{\operatorname{\tilde{\mathcal O}}}
\renewcommand{\P}{\operatorname{\mathbb P}}
\newcommand{\D}{\operatorname{\mathbb{D}}}
\newcommand{\T}{\operatorname{\mathcal{T}}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\mS}{\mathcal{S}}
\newcommand{\mA}{\mathcal{A}}
\newcommand{\mF}{\mathcal{F}}
\newcommand{\mW}{\mathcal{W}}
\newcommand{\mT}{\mathcal{T}}
\newcommand{\mH}{\mathcal{H}}
\newcommand{\mE}{\mathcal{E}}
\newcommand{\mD}{\mathcal{D}}
\newcommand{\mB}{\mathcal{B}}
\newcommand{\mI}{\mathcal{I}}
\newcommand{\mV}{\mathcal{U}}
\newcommand{\Gap}{\mathsc{Gap}}
\newcommand{\Unif}{\mathrm{Unif}}
\newcommand{\poly}{\mathrm{poly}}
\newcommand{\Cov}{\mathrm{Cov}}

\renewcommand{\tilde}{\widetilde}
\renewcommand{\hat}{\widehat}
\renewcommand{\bar}{\overline}
\newcommand{\tappa}{\mathfrak t}

\newcommand{\truth}{\textbf{truth}}

\newcommand{\paren}[1]{\left( #1 \right)}
\newcommand{\sqb}[1]{\left[ #1 \right]}
\newcommand{\set}[1]{\left\{ #1 \right\}}
\newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
\newcommand{\ceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\1}{\mathbbm{1}}
\newcommand{\comment}[1]{}

\usepackage{nicefrac}
\usepackage{multirow}
\usepackage{footnote}
\usepackage{enumitem}
\setitemize{leftmargin=*, nosep, topsep=2pt, itemsep=2pt}
\setenumerate{leftmargin=*, nosep, topsep=2pt, itemsep=2pt}
\usepackage{comment}
\usepackage{bbm}
\usepackage{bm}

\allowdisplaybreaks

\usepackage{tikz,graphicx}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{patterns}

%%%%% BE SURE TO REMOVE algorithm2e %%%%%
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\newcommand{\alglinelabel}{%
\addtocounter{ALG@line}{-1}% Reduce line counter by 1
\refstepcounter{ALG@line}% Increment line counter with reference capability
\label% Regular \label
}
\usepackage{algorithm}
\usepackage[noEnd,commentColor=blue!70!white]{algpseudocodex}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\Crefname{ALG@line}{Line}{Lines}
%%%%% END OF algpseudocodex %%%%%

\Crefname{assumption}{Assumption}{Assumptions}
\Crefformat{equation}{Eq. #2(#1)#3}
\Crefrangeformat{equation}{Eqs. #3(#1)#4 to #5(#2)#6}
\Crefmultiformat{equation}{Eqs. #2(#1)#3}{ and #2(#1)#3}{, #2(#1)#3}{ and #2(#1)#3}
\Crefrangemultiformat{equation}{Eqs. #3(#1)#4 to #5(#2)#6}{ and #3(#1)#4 to #5(#2)#6}{, #3(#1)#4 to #5(#2)#6}{ and #3(#1)#4 to #5(#2)#6}

\Crefname{enumi}{Restriction}{Restrictions}
\Crefname{footnote}{Footnote}{Footnotes}

\newcommand\scalemath[2]{\scalebox{#1}{\mbox{\ensuremath{\displaystyle #2}}}}
\let\originalmiddle=\middle
\def\middle#1{\mathrel{}\originalmiddle#1\mathrel{}}

\usepackage{xspace}
\newcommand{\mechname}{\texttt{AdaAudit}\xspace}

\renewcommand{\paragraph}[1]{\vspace{2pt}\noindent\textbf{#1}}

\begin{document}
\maketitle


\begin{abstract}
We study a repeated resource allocation problem with strategic agents where monetary transfers are disallowed and the central planner has no prior information on agents' utility distributions. In light of Arrow's impossibility theorem, acquiring information about agent preferences through some form of feedback is necessary. We assume that the central planner can request powerful but expensive audits on the winner in any round, revealing the true utility of the winner in that round. We design a mechanism achieving $T$-independent $\mathcal O(K^2)$ regret in social welfare while requesting $\mathcal O(K^3 \log T)$ audits in expectation, where $K$ is the number of agents and $T$ is the number of rounds. We also show an $\Omega(K)$ lower bound on the regret and an $\Omega(1)$ lower bound on the number of audits when having low regret.
Algorithmically, we show that incentive-compatibility can be mostly enforced with an accurate estimation of the winning probability of each agent under truthful reporting. To do so, we impose future punishments and introduce a \emph{flagging} component, allowing agents to flag any biased estimate (we show that doing so aligns with individual incentives). On the technical side, without monetary transfers and distributional information, the central planner cannot ensure that truthful reporting is exactly an equilibrium. Instead, we characterize the equilibrium via a reduction to a simpler \emph{auxiliary game}, in which agents cannot strategize until late in the $T$ rounds of the allocation problem. The tools developed therein may be of independent interest for other mechanism design problems in which the revelation principle cannot be readily applied. 
\end{abstract}


\input{introduction}
\input{technical-overview}
\input{setup}
\input{main_results}
\input{proof_sketch}

\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\onecolumn
\newpage
\crefalias{section}{appendix}
\crefalias{subsection}{appendix}
\appendix

\startcontents[section]
\printcontents[section]{l}{1}{\setcounter{tocdepth}{2}}

\input{appendix-varying-p-regret}
\input{appendix-varying-p-audits}

\input{appendix-fixed-p}
\input{appendix-lower-bound}

\end{document}
