\section{Related Work}
\label{relatedwork}

\textbf{Self-Referential Self-Improving Systems.} Numerous frameworks ____ have been proposed to address the design of multi-agent scaffolding. EvoAgent ____ extends single expert agents to multi-agent scaffolds via evolutionary algorithms, whilst \textsc{AgentBreeder} evolves the entire system as a unit. EvoMAC ____ evolves agents and their connections during test time to improve code generation, whereas \textsc{AgentBreeder} is domain agnostic and can explore the entire search space of scaffolds. ADAS ____, ComfyAgent ____ and Gödel Agent ____ search in the space of code for novel scaffolds, but unlike \textsc{AgentBreeder} they do not incorporate a quality-diversity mechanism for exploring agent design space. FunSearch ____ is an evolutionary method to search the function space for high-performing computer programs but not necessarily scaffolds. PromptBreeder ____ is an evolutionary self-improving framework that evolves prompts for a given domain, but does not focus on the scaffold as a whole.

\textbf{Multi-Agent Safety Research.} ____ evaluate the safety of multi-agent scaffolds from a psychological perspective by injecting agents with malicious traits, and provide mitigation strategies such as performing psychological assessments and therapy for agents. Polaris ____ introduces a safety-focused scaffold for real-time patient healthcare conversations.
____ explore the resilience of multi-agent scaffolds when injected with malicious or error-prone agents. ____ provide a more thorough discussion of the safety risks associated with scaffolded LLMs.