\section{Related Work}
\label{sec:related}

\paragraph{Multi-Agent Frameworks}\;
LLM agents are employed in advanced tasks for their ability to reason over diverse inputs and generate responses in complex scenarios~\cite{wei2022chain, caoyun2024canllm, liu2024agentbench, shinn2024reflexion}. Past studies have explored different evaluation frameworks~\cite{debenedetti2024agentdojo, ye2024justice, asgari2024mmlupro}, emphasizing the need for rigorous practices in developing productized LLM agents. While these agents have shown remarkable success, their experiments are often limited by a lack of dynamic interactions.

Multi-agent systems explore LLMs' capabilities in complex environments~\cite{wu2024autogen, huang2024far, zhao2024competeai, wu-etal-2024-shall}. For instance, \citet{gu2024agent} examined simulations where agents interact and collaborate within group chat settings, while \citet{zhang2024llmmastermindsurveystrategic} explored social simulations and board games, finding that even advanced models struggle to fully utilize their reasoning potential. \citet{huang2024far} introduced an evaluation metric to assess LLMs' gaming abilities in multi-agent settings through game-theoretic experiments. \citet{shinn2024reflexion} proposed a method to enhance LLMs through linguistic feedback, and \citet{huang2024far} developed methodologies to evaluate their decision-making capabilities.

Our work extends AucArena~\cite{chen2023put} by exploring persona dynamics in a bidding framework that enables multiple agents to compete for maximum profit in an auction. Our framework aims for a deeper investigation of a single agent's ability to profile other agents. Moreover, we explore how theory of mind influences an agent's performance in a competitive multi-agent setting.

\vspace{-0.05in}
\paragraph{Persona in LLM Agents}\;
Personas function as identities assigned to LLM agents, enhancing their ability to generate personalized or specialized outputs~\cite{xu2024character, chen2024from, sun2024building, li2025helloagainllmpoweredpersonalized}. Prior research has extensively examined the influence of personas across various roles~\cite{hu-collier-2024-quantifying, samuel2024personagymevaluatingpersonaagents, kim2024persona, dong-etal-2024-llm}. Leveraging their role-playing nature, some studies have applied personas to social simulations by assigning diverse identities to entire agent populations~\cite{lee2024language, tseng-etal-2024-two, hu2024quantifying}. \citet{yang2024} find that persona prompting can introduce shortcut learning, causing LLM agents to deviate from rational objectives.  
Building on this work, we incorporate persona into a multi-agent auction to examine its effects in competitive settings.

\vspace{-0.05in}
\paragraph{Theory of Mind (ToM)}\;\;
Theory of Mind, the ability to understand and infer one's own and others' mental states, is fundamental to human social interaction and a crucial capability for LLMs to achieve human-like reasoning~\cite{leslie2004core, sap-etal-2022-neural, van-duijn-etal-2023-theory, cross2024hypothetical, chan-etal-2024-negotiationtom}. An accurate ToM in others' intentions and actions provides significant advantages~\cite{street2024llm, amirizaniani2024llms}. Past studies have applied ToM to simulate social behaviors. \citet{de2017negotiating} explored its role in negotiation tasks, while \citet{wang-etal-2024-sotopia} proposed an interactive learning environment to train LLMs in social interactions. 

Previous studies have examined collaborative multi-agent settings. For example, \citet{li2023camel} and \citet{Li_2023} explored LLMs using ToM for task collaboration, while others demonstrated its effectiveness in structured games with rules and rewards. \citet{light2023avalonbench} investigated how LLMs infer players' motives in the Avalon board game, highlighting ToM capabilities within a constrained and predefined context. Similarly, \citet{yim2024evaluating} evaluated ToM in cooperative scenarios through a poker game called Guandan. In contrast, our work investigates whether ToM can enhance agent performance in competing multi-agent environments with auction as the testbed.