\section{Literature Review}
%The modern science of networks has witnessed significant advancements in our comprehension of complex systems, with community structure emerging as a crucial aspect in the analysis of graphs representing real-world systems. Community structure, characterized by the organization of vertices into clusters with dense internal connections and sparse inter-cluster connections, mirrors the compartmentalization observed in various systems, resembling tissues or organs in biological organisms.
Modern-day network research has made great strides towards understanding complex systems, and community structure has become an important tool for analyzing graphs that model real-world systems. The organizing of vertices into clusters with sparse connections between clusters and dense interior connections reflects the compartmentalization seen in many systems, such as biological organisms' tissues or organs. Detecting communities holds substantial importance across disciplines such as sociology, biology, and computer science, where networks serve as fundamental representations of systems.

% The paper "Community Detection in Graphs" by Fortunato \cite{cd_in_graphs_Fortunato} provides a comprehensive overview of the complexities inherent in community detection,
Fortunato's paper "Community Detection in Graphs" \cite{cd_in_graphs_Fortunato} offers a thorough summary of the challenges involved in community detection, emphasizing the need to delineate key elements, explore methodologies, and address critical issues such as clustering significance and method evaluation. With graph clustering lacking precise definitions, the literature presents a plethora of algorithms and approaches, spanning graph partitioning, modularity-based methods, spectral algorithms, dynamic techniques, and more. Modularity-based methods like Newman-Girvan modularity and spectral algorithms have garnered attention for their efficiency, while dynamic algorithms and statistical inference methods offer promising avenues. Despite progress, challenges persist in algorithm evaluation, understanding clustering in real-world contexts, and exploring dynamic community structures. The evolving field continues to refine existing methods, explore new approaches, and illuminate the properties and applications of communities in real-world networks.

"Community detection in networks: A user guide," \cite{cd_in_networks_Fortunato} offers an in-depth exploration of community detection tailored for practitioners and accessible to those with basic network science knowledge. Rather than striving for exhaustive coverage, the paper focuses on elucidating fundamental aspects of community detection, organized into three main sections. The first section explores conceptual foundations, tracing the evolution of the community concept and laying the groundwork for understanding diverse interpretations. The second section tackles validation challenges, emphasizing the importance of validation and discussing various validation methods. The final section delves into algorithmic approaches, critically examining popular clustering algorithms and methodological considerations. Additionally, the paper provides guidance on accessing software tools and concludes with key findings and suggestions for future research. Overall, it serves as a valuable resource for navigating community detection complexities, addressing both theoretical foundations and practical considerations.

The paper "Community Detection in Social Networks" by Bedi et al. \cite{cd_in_social_networks_Bedi} underscores the importance of community detection within the expansive realm of social networking. As individuals increasingly engage in virtual communities within social networking sites, understanding and identifying these clusters become paramount for various purposes, from collaborative research to targeted marketing strategies. Through a comprehensive survey of existing algorithms, categorized by their methodologies and applications across domains, the paper navigates through the fundamental concepts of social networks and community structure. By highlighting the potential applications and inclusivity of datasets utilized by these algorithms, the authors emphasize the versatility and significance of community detection, offering insights into its effective utilization across commercial, educational, and developmental landscapes.

The paper "Community detection algorithms: A comparative analysis" by Lancichinetti and Fortunato \cite{cda_comp_analysis_Fortunato} addresses the complexity of community detection in networks and the necessity for rigorous evaluation of algorithmic performance. Recognizing the limitations of previous evaluations, which often relied on small or simplified networks, the authors conduct comprehensive tests using diverse benchmarks to assess algorithmic accuracy and computational efficiency. Among the tested algorithms, Rosvall and Bergstrom's Infomap method emerges as particularly effective, especially on challenging benchmarks. However, the paper acknowledges ongoing challenges, such as the need for algorithms capable of handling hierarchical structures and multipartite graphs, highlighting avenues for further research in the field.