% Human motion generation 이걸 넣고 Recurrency? 관련은 안넣는게 맞는 거 같기도 하고
Human motion generation is one of the most interested field in computer vision studies. Due to the high level of interest, it has witnessed remarkable advances through many researches. Despite this significant growth, human motion generation remains a challenging problem, often plagued by issues such as foot-sliding and a lack of consistency or smoothness due to abrupt motion changes.

Roughly human motion generation is based on either regression models or generative models. In recent years, generative models have become the dominant approach. For instance, Kinetic-GAN \cite{degardin2022generative} utilizes a GAN combined with GCN. It leverages latent space disentanglement to separate different aspects of motion, facilitating easier manipulation and diverse generation. Furthermore, stochastic variations help produce a wider range of realistic motions.
% 여기부터는 그냥 대충 뭐 diffusion 쓰는 모델 소개

MDM \cite{tevet2022human} and MLD \cite{chen2023executing}, for example, utilize diffusion model to generate desired motions. \cite{ahn2018text2action, ahuja2019language2pose, ghosh2021synthesis, petrovich2022temos, athanasiou2022teach} are among the studies to generate human motion sequences from text descriptions. Since above methods can face challenges in zero-shot generation, \cite{tevet2022motionclip, hong2022avatarclip} employs CLIP, a pre-trained vision-language model, for better zero-shot generation performance.
% 여기에 이제 좀 recently 뭐하는지 쓴거긴한데 MDM이랑 MLD는 겹침
% 안 띄워도 됨! 이해 안가는 부분 같은거 주석으로 달아주셔도 됩니다 보고 의견 달아드릴게요
% MDM 그냥 모델임  diffusion 기반 motion 생성, for example을 달려고 했었는데 위에랑 말이 좀 안맞을까봐

VQ-VAE is another popular method for generating human motion. For instance, T2MT \cite{guo2022tm2t} uses VQ-VAE for training text-to-motion and motion-to-text tasks. T2M-GPT \cite{zhang2023t2m} leverages a GPT-like transformer architecture and VQ-VAE combined with an EMA. %SOTA에용? if sota: [FLAME, MoFusion, MDM, MLD] diffusion-based text-to-motion models represent the current state-of-the-art in human motion generation.

