


\subsection{Additional NLOS Applications \& Extensions}\label{sec:discussion} 

In addition to the applications described above, \name\ also paves the way for future work in multiple directions:
\begin{itemize}
    \item \textit{NLOS Object Completion:} In some cases, the mmWave images do not capture the entire object due to specularity (See supplementary for more details).  It would be interesting future work to create an object completion network that learns the missing parts of an object \& creates more photo-realistic NLOS images.
    \item \textit{NLOS 6DoF Estimation:} Another interesting problem is 6DoF estimation from mmWave images. This could be used to ensure proper liquid packaging to avoid spills, or enable efficient grasp planning in NLOS (e.g., grasping an item under cloth or within a box with packing peanuts).
    \item \textit{NLOS Robotic Manipulation: }Future work can leverage many of the above NLOS perception tasks to enable new tasks in robotic manipulation. For example, accurate mmWave object segmentation and classification can significantly increase the efficiency of mechanical search and retrieval (i.e., the problem of searching for a specific item in clutter).  Further, mmWave images can be used to analyze pile stability to improve push and grasp planners. NLOS perception can also help autonomous robots determine which packages contain fragile materials to be handled with special care. More generally, NLOS perception can help autonomous robots to operate in more realistic, cluttered environments.
\end{itemize}
