\section{NOTE}
OUTPUT CURL  
\\\textbf{$->$ }certs Output the certificate chain with details. Supported only by the OpenSSL, GnuTLS, Schannel, NSS, GSKit and Secure Transport backends (Added in 7.88.0)
\\\textbf{$->$ }content\_type The Content-Type of the requested document, if there was any.
\\\textbf{$->$ }errormsg The error message. (Added in 7.75.0)
\\\textbf{$->$ }exitcode The numerical exitcode of the transfer. (Added in 7.75.0)
\\\textbf{$->$ }filename\_effective The ultimate filename that curl writes out to. This is only meaningful if curl is told to write to a file with the -O, --remote-name or -o, --output option. It's most useful in combination with the -J, --remote-header-name option.
\\\textbf{$->$ }ftp\_entry\_path The initial path curl ended up in when logging on to the remote FTP server.
\\\textbf{$->$ }header\_json A JSON object with all HTTP response headers from the recent transfer. Values are provided as arrays, since in the case of multiple headers there can be multiple values.
\\\textbf{$->$ }The header names provided in lowercase, listed in order of appearance over the wire. Except for duplicated headers. They are grouped on the first occurrence of that header, each value is presented in the JSON array.
\\\textbf{$->$ }http\_code The numerical response code that was found in the last retrieved HTTP(S) or FTP(s) transfer.
\\\textbf{$->$ }http\_connect The numerical code that was found in the last response (from a proxy) to a curl CONNECT request.
\\\textbf{$->$ }http\_version The http version that was effectively used. (Added in 7.50.0)
\\\textbf{$->$ }json A JSON object with all available keys.
\\\textbf{$->$ }local\_ip The IP address of the local end of the most recently done connection \\\textbf{$->$ }can be either IPv4 or IPv6.
\\\textbf{$->$ }local\_port The local port number of the most recently done connection.
\\\textbf{$->$ }method The http method used in the most recent HTTP request. (Added in 7.72.0)
\\\textbf{$->$ }num\_certs Number of server certificates received in the TLS handshake. Supported only by the OpenSSL, GnuTLS, Schannel, NSS, GSKit and Secure Transport backends (Added in 7.88.0)
\\\textbf{$->$ }num\_connects Number of new connects made in the recent transfer.
\\\textbf{$->$ }num\_headers The number of response headers in the most recent request (restarted at each redirect). Note that the status line IS NOT a header. (Added in 7.73.0)
\\\textbf{$->$ }num\_redirects Number of redirects that were followed in the request.
\\\textbf{$->$ }onerror The rest of the output is only shown if the transfer returned a non-zero error (Added in 7.75.0)
\\\textbf{$->$ }proxy\_ssl\_verify\_result The result of the HTTPS proxy's SSL peer certificate verification that was requested. 0 means the verification was successful. (Added in 7.52.0)
\\\textbf{$->$ }redirect\_url When an HTTP request was made without -L, --location to follow redirects (or when --max-redirs is met), this variable will show the actual URL a redirect would have gone to.
\\\textbf{$->$ }referer The Referer: header, if there was any. (Added in 7.76.0)
\\\textbf{$->$ }remote\_ip The remote IP address of the most recently done connection \\\textbf{$->$ }can be either IPv4 or IPv6.
\\\textbf{$->$ }remote\_port The remote port number of the most recently done connection.
\\\textbf{$->$ }response\_code The numerical response code that was found in the last transfer (formerly known as "http\_code").
\\\textbf{$->$ }scheme The URL scheme (sometimes called protocol) that was effectively used. (Added in 7.52.0)
\\\textbf{$->$ }size\_download The total amount of bytes that were downloaded. This is the size of the body/data that was transferred, excluding headers.
\\\textbf{$->$ }size\_header The total amount of bytes of the downloaded headers.
\\\textbf{$->$ }size\_request The total amount of bytes that were sent in the HTTP request.
\\\textbf{$->$ }size\_upload The total amount of bytes that were uploaded. This is the size of the body/data that was transferred, excluding headers.
\\\textbf{$->$ }speed\_download The average download speed that curl measured for the complete download. Bytes per second.
\\\textbf{$->$ }speed\_upload The average upload speed that curl measured for the complete upload. Bytes per second.
\\\textbf{$->$ }ssl\_verify\_result The result of the SSL peer certificate verification that was requested. 0 means the verification was successful.
\\\textbf{$->$ }stderr From this point on, the -w, --write-out output will be written to standard error. (Added in 7.63.0)
\\\textbf{$->$ }stdout From this point on, the -w, --write-out output will be written to standard output. This is the default, but can be used to switch back after switching to stderr. (Added in 7.63.0)
\\\textbf{$->$ }time\_appconnect The time, in seconds, it took from the start until the SSL/SSH/etc connect/handshake to the remote host was completed.
\\\textbf{$->$ }time\_connect The time, in seconds, it took from the start until the TCP connect to the remote host (or proxy) was completed.
\\\textbf{$->$ }time\_namelookup The time, in seconds, it took from the start until the name resolving was completed.
\\\textbf{$->$ }time\_pretransfer The time, in seconds, it took from the start until the file transfer was just about to begin. This includes all pre-transfer commands and negotiations that are specific to the particular protocol(s) involved.
\\\textbf{$->$ }time\_redirect The time, in seconds, it took for all redirection steps including name lookup, connect, pretransfer and transfer before the final transaction was started. time\_redirect shows the complete execution time for multiple redirections.
\\\textbf{$->$ }time\_starttransfer The time, in seconds, it took from the start until the first byte was just about to be transferred. This includes time\_pretransfer and also the time the server needed to calculate the result.
\\\textbf{$->$ }time\_total The total time, in seconds, that the full operation lasted.
\\\textbf{$->$ }url The URL that was fetched. (Added in 7.75.0)
\\\textbf{$->$ }urlnum The URL index number of this transfer, 0-indexed. De-globbed URLs share the same index number as the origin globbed URL. (Added in 7.75.0)
\\\textbf{$->$ }url\_effective The URL that was fetched last. This is most meaningful if you have told curl to follow location: headers.
