\begin{abstract}
The login functionality, being the gateway to app usage, plays a critical role in both user experience and application security. As Android apps increasingly incorporate login functionalities, they support a variety of authentication methods with complicated login processes, catering to personalized user experiences.
However, the complexities in managing different operations in login processes make it difficult for developers to handle them correctly.
In this paper, we present the first empirical study of login issues in Android apps. We analyze 361 issues from 44 popular open-source Android repositories, examining the root causes, symptoms, and trigger conditions of these issues. Our findings indicate that the vast majority of the login issues are induced by the improper handling of complex state transitions during the login process, which can prevent users from logging in or misdirect them to incorrect subsequent actions. Additionally, we observed that issues related to this cause typically require the convergence of multiple trigger conditions to manifest. These findings can help developers to model the login processes which can help them to identify the causes of issues and design targeted test cases and precise test oracles. 
Our dataset has been made openly available to facilitate future research in this area.
% \lili{Highlight one key take-away message here. Why are the findings important?}\zixu{Add two sentences}
% \lili{THe previous two sentences needs to be revised again. They are still simply saying that they are useful. Can we be a bit more concrete?}\zixu{I add three concrete points}


\end{abstract}

\begin{IEEEkeywords}
Login, Empirical Study, Android, Issue
\end{IEEEkeywords}
