% \documentclass[conference]{IEEEtran}
\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
\usepackage{hyperref}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
% \usepackage{cite}
% \usepackage{amsmath,amssymb,amsfonts}
% \usepackage{algorithmic}
% \usepackage{graphicx}
% \usepackage{textcomp}
% \usepackage{xcolor}
% \usepackage{enumitem}
% \usepackage{booktabs}
% \usepackage{cellspace}
% \usepackage[breaklinks,colorlinks]{hyperref}
% \usepackage{subcaption}
% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
%     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\overrideIEEEmargins                                      % Needed to meet printer requirements.
\usepackage{multirow}

\title{\LARGE \bf
Range and Bird's Eye View Fused Cross-Modal Visual Place Recognition
}

\author{Jianyi Peng, Fan Lu, Bin Li, Yuan Huang, Sanqing Qu, Guang Chen$^{*}$% <-this % stops a space
%\thanks{$^{*}$equal contribution}% <-this % stops a space
\thanks{* corresponding author}
\thanks{Jianyi Peng, Fan Lu, Bin Li, Sanqing Qu and Guang Chen are with Tongji University, Shanghai 201804, China. E-mail: 2233589@tongji.edu.cn, lufan@tongji.edu.cn, libin2021@tongji.edu.cn, sanqingqu@tongji.edu.cn, guangchen@tongji.edu.cn.}
\thanks{Yuan Huang is with Automation, Beijing Institute of Control Engineering. E-mail: hy0103053428@gmail.com.}
\thanks{This work was supported by the National Key Research and Development Program of China (No. 2024YFE0211000), in part by the National Natural Science Foundation of China (No. 62372329), in part by Shanghai Scientific Innovation Foundation (No. 23DZ1203400), in part by Tongji-Qomolo Autonomous Driving Commercial Vehicle Joint Lab Project, and in part by Xiaomi Young Talents Program.}
}

\usepackage{graphicx}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{amsmath}
\usepackage{stfloats}
\usepackage{amssymb}
\usepackage[dvipsnames]{xcolor}
\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\todo}[1]{{\color{red}#1}}
\newcommand{\TODO}[1]{\textbf{\color{red}[TODO: #1]}}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\input{0_abstract}
\input{1_intro}
\input{2_related_work}
\input{3_method}
\input{4_experiment}
\input{5_conclusion}


% \addtolength{\textheight}{-5cm} 

\bibliographystyle{IEEEtran}
\bibliography{IEEEfull}

\clearpage
\section*{Appendix}
% \appendices
\input{6_appendix}
\end{document}