\section{Related Work}
\textbf{RAG Models.} 
RAG models have shown superior performance in QA tasks____. 
These models typically employ the \textit{retriever-reader} architecture, which consists of a retriever____ % for retrieving relevant documents 
and a reader____. 
Efforts to improves RAG models generally follows three main directions: 
(1) enhance the retriever for better retrieval performance____; 
(2) enhance the reader for better comprehension and answer generation____; 
(3) introduce additional modules to bridge the retriever and the reader____. 

\vspace{0.5em} \noindent \textbf{Iterative RAG Models for Multi-Hop QA.}
Iterative RAG models____ address multi-hop QA by performing multiple steps of retrieval and reasoning. 
For instance, IRCoT____ use LLM-generated chain-of-thoughts for retrieval, while  DRAGIN____ dynamically decides when and what to retrieve based on the LLM's information needs. 
However, these models all rely on LLM-generated thoughts, making them prone to hallucination. In contrast, \OURS{} employs knowledge triples and a trained retriever to actively identify and retrieve missing information, enabling a more reliable and accurate retrieval for multi-hop QA. 

\vspace{0.5em} \noindent \textbf{KG-Enhanced RAG Models.} 
Recently, KGs have been integrated into RAG models____. Some studies leverage information from existing KGs____ for additional context____, while others generate KGs from documents to improve knowledge organisation____ or enhance reader comprehension____. 
These models primarily follow the standard RAG pipeline, whereas our work focuses on the iRAG pipelines. Moreover, while they rely on single-step retrieval with pre-existing retrievers, \OURS{} employs a trained retriever tailored for iterative retrieval, allowing it to dynamically adapt to the evolving information needs in multi-step reasoning.