\def\isarxiv{1}

\documentclass[11pt]{article}
\pagestyle{plain}

\usepackage[margin=1in]{geometry}
\usepackage{latexsym, amscd, amsfonts, mathrsfs, amsmath, amssymb, amsthm}
\usepackage{bm}
\usepackage{color}
\usepackage[noend]{algpseudocode}
\usepackage[colorlinks,citecolor=blue,linkcolor=blue,urlcolor=red,pagebackref]{hyperref}
\usepackage[capitalise]{cleveref}

\usepackage{stmaryrd, tikz-cd, enumitem, mathrsfs, bbm, algorithm, url, esint, listings, moreverb, makecell, thm-restate, soul}


\def\bC{\mathbb{C}}
\def\bE{\mathbb{E}}
\def\bN{\mathbb{N}}
\def\bO{\mathbb{O}}
\def\bP{\mathbb{P}}
\def\bQ{\mathbb{Q}}
\def\bR{\mathbb{R}}
\def\bZ{\mathbb{Z}}

\def\bfI{\mathbf{I}}
\def\bfV{\mathbf{V}}

\def\cA{\mathcal{A}}
\def\cB{\mathcal{B}}
\def\cC{\mathcal{C}}
\def\cD{\mathcal{D}}
\def\cE{\mathcal{E}}
\def\cI{\mathcal{I}}
\def\cN{\mathcal{N}}
\def\cM{\mathcal{M}}
\def\cP{\mathcal{P}}
\def\cS{\mathcal{S}}
\def\cX{\mathcal{X}}
\def\cY{\mathcal{Y}}

\def\fX{\mathfrak{X}}

\def\contig{\vartriangleleft}

\def\wt{\widetilde}
\DeclareMathOperator\BEC{\mathrm{BEC}}
\DeclareMathOperator\Ber{\mathrm{Ber}}
\DeclareMathOperator\Bin{\mathrm{Bin}}
\DeclareMathOperator\BSC{\mathrm{BSC}}
\DeclareMathOperator\dom{\mathrm{dom}}
\DeclareMathOperator\dist{\mathrm{dist}}
\DeclareMathOperator\Ent{\mathrm{Ent}}
\DeclareMathOperator\KL{\mathrm{KL}}
\DeclareMathOperator\sgn{\mathrm{sgn}}
\DeclareMathOperator\supp{\mathrm{supp}}
\DeclareMathOperator\Tr{\mathrm{Tr}}
\DeclareMathOperator\TV{\mathrm{TV}}
\DeclareMathOperator\Pois{\mathrm{Pois}}
\DeclareMathOperator\Unif{\mathrm{Unif}}
\DeclareMathOperator\Var{\mathrm{Var}}
\DeclareMathOperator\Inf{\mathsf{Inf}}
\DeclareMathOperator\I{\mathsf{I}}
\DeclareMathOperator\bs{\mathsf{bs}}
\DeclareMathOperator\s{\mathsf{s}}
\DeclareMathOperator\R{\mathsf{R}}


\renewcommand{\setminus}{\backslash}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\newcommand{\todo}[1]{{\color{red}[TODO: #1]}}
\newcommand{\yuzhou}[1]{{\color{blue}[Yuzhou: #1]}}
\newcommand{\yinzhan}[1]{{\color{purple}[Yinzhan: #1]}}
\newcommand{\xin}[1]{{\color{magenta}[Xin: #1]}}


\newcommand{\rom}[1]{\textup{\uppercase\expandafter{\romannumeral#1}}}

\pagestyle{plain}


\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{observation}[theorem]{Observation}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{open}{Open Problem}

\newcommand{\DKLs}[0]{D_{\KL}}
\newcommand{\DKL}[0]{\DKLs(p \parallel 1-p)}
\newcommand{\Threshold}[2]{\mathrm{TH}_{#2}^{#1}}

\begin{document}
\title{Tight Bounds for Noisy Computation of High-Influence Functions, Connectivity, and Threshold}
\ifdefined\isarxiv
\author{Yuzhou Gu\thanks{\texttt{yuzhougu@nyu.edu}. New York University. Part of the work was done while supported by the National Science Foundation under Grant No.~DMS-1926686.}
\and
Xin Li\thanks{\texttt{lixints@cs.jhu.edu}. Johns Hopkins University. Supported by NSF CAREER Award CCF-1845349 and NSF Award CCF-2127575.}
\and
Yinzhan Xu\thanks{\texttt{xyzhan@ucsd.edu}. University of California, San Diego. Supported by NSF Grant CCF-2330048, HDR TRIPODS Phase II grant 2217058 and a Simons Investigator Award.}}
\else
\author{Anonymous Author(s)}
\fi
\date{}

\maketitle

\pagenumbering{gobble}

\begin{abstract}
  The noisy query model (a.k.a.~noisy decision tree model) was formally introduced by [Feige, Raghavan, Peleg and  Upfal, SICOMP 1994].
  In this model, the (binary) return value of every query (possibly repeated) is independently flipped with some fixed probability $p \in (0, 1/2)$.
  The noisy query complexity (a.k.a.~noisy decision tree complexity) is the smallest number of queries an algorithm needs to make to achieve error probability $\le \frac 13$.

  Previous works often focus on specific problems, and it is of great interest to have a characterization of noisy query complexity for general functions. We show that any Boolean function with total influence $\Omega(n)$ has noisy query complexity $\Omega(n\log n)$, matching the trivial upper bound of $O(n \log n)$. Our result is the first noisy query complexity lower bound of this generality, beyond what was known for random Boolean functions [Reischuk and Schmeltz, FOCS 1991].

  Our second contribution is to determine the asymptotic noisy query complexity of the Graph Connectivity problem. In this problem, the goal is to determine whether an undirected graph is connected, where each query asks for the existence of an edge in the graph.
  A simple algorithm can solve the problem with error probability $o(1)$ using $O(n^2 \log n)$ noisy queries, but no non-trivial lower bounds were known prior to this work.
  We show that any algorithm that solves the problem with error probability $\le \frac 13$ uses $\Omega(n^2\log n)$ noisy queries.

  For the proofs of the above lower bounds, we develop a three-phase method, which is a refinement of the two-phase method of [Feige, Raghavan,  Peleg and  Upfal, SICOMP 1994]. Our three-phase method adds an extra step where carefully designed ``free'' information is disclosed to the algorithm. This way, we can precisely control the posterior distribution of the input given the query results, empowering a more refined analysis.

  Last but not least, we determine the exact number of noisy queries (up to lower order terms) needed to solve the $k$-Threshold problem and the Counting problem. The $k$-Threshold problem asks to decide whether there are at least $k$ ones among $n$ bits, given noisy query access to the bits.
  We prove that $(1\pm o(1)) \frac{n\log (\min\{k,n-k+1\}/\delta)}{(1-2p)\log \frac{1-p}p}$ queries are both sufficient and necessary to achieve error probability $\delta = o(1)$. Previously, such a result was only known when $\min\{k,n-k+1\}=o(n)$ [Wang, Ghaddar, Zhu and Wang, arXiv 2024]. We also show a similar $(1\pm o(1)) \frac{n\log (\min\{k+1,n-k+1\}/\delta)}{(1-2p)\log \frac{1-p}p}$ bound for the Counting problem, where one needs to count the number of ones among $n$ bits given noisy query access and $k$ denotes the answer.
\end{abstract}

\newpage
\pagenumbering{arabic}


\section{Introduction}
\input{intro}

\section{Technical overview}
\input{overview}

\section{Preliminaries}
\input{prelim}

\section{High-influence functions}
\input{influence}

\section{Graph Connectivity}
\input{connectivity}

\section{Threshold and Counting}
\input{threshold}




\ifdefined\isarxiv
\section*{Acknowledgments}
We thank Ziao Wang for pointing out an error in the statement of \cref{open:k-selection} in a previous version of the paper.
\else
\fi

\bibliographystyle{alpha}
\bibliography{ref}

\appendix
\section{A simpler proof of hardness of Threshold for small \texorpdfstring{$k$}{k}}
\label{sec:th-small}
\input{threshold-small-k}


\end{document}
