\section{Related Work}
\vspace{-0.05in}
\noindent\textbf{AI-powered Chatbots for Seniors: }
Social interaction plays a crucial role in the cognitive health of seniors ____.
%
% 
AI-powered chatbots have emerged as a promising alternative ____.
%
Recent efforts span commercial products ____ and research prototypes focused on emotional support and audio assistance ____. Unlike prior work, our approach prioritizes user conversational \textit{engagement}, which has been empirically shown to enhance both cognitive and emotional status.
%


\noindent\textbf{Conversational Strategies: }
%
____ curated a dataset with annotated strategies, demonstrating the effectiveness of Helping Skills Theory____ in providing emotional support. 
%
____ examined the causal relationships between dialogue acts (DAs) and participants’ emotional states in a clinical trial ____, emphasizing the impact of strategic interventions in tele-mediated dialogues.
%
____ identified key strategies for improving child patient-provider communication through semi-structured interviews. 
%
{Few works have systematically integrated these strategies into the reasoning flow of AI chatbots. Our approach fills this gap by embedding structured conversational strategies for automatically enhancing user engagement.}

 
%
%
\noindent\textbf{Multi-turn chatbot exploration: } Recent advances in AI dialogue systems predominantly focused on short-turn interactions ____. While some pioneering research has explored multi-turn optimization through Reinforcement Learning (RL) approaches for LLMs ____, these methods were not specifically designed for supporting senior dialogue engagement. Our approach is  tuning-free yet compatible with RL-driven methods, enabling future extensions through hierarchical RL to further enhance multi-turn conversation engagement.

%
%

\vspace{-0.05in}