%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
\IEEEoverridecommandlockouts    % This command is only needed if you want to use the \thanks command
\overrideIEEEmargins                                      % Needed to meet printer requirements.

%In case you encounter the following error:
%Error 1010 The PDF file may be corrupt (unable to open PDF file) OR
%Error 1000 An error occurred while parsing a contents stream. Unable to analyze the PDF file.
%This is a known problem with pdfLaTeX conversion filter. The file cannot be opened with acrobat reader
%Please use one of the alternatives below to circumvent this error by uncommenting one or the other
%\pdfobjcompresslevel=0
%\pdfminorversion=4

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

% my additions
\usepackage{hyperref}
\usepackage{hhline}
\usepackage{soul}
\usepackage{xcolor}
\usepackage{cite}
\usepackage{import}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amsfonts} 
\usepackage{bm}
\usepackage[capitalise]{cleveref}
%\graphicspath{./figures/}
\usepackage{booktabs}
\usepackage{authblk}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\newcommand{\etal}{\textit{et al}. }
\newcommand{\ie}{\textit{i}.\textit{e}., }
\newcommand{\eg}{\textit{e}.\textit{g}. }
\newcommand{\secref}[1]{Sec.~\ref{#1}}
% \newcommand{\eqtref}[1]{Eq.~\ref{#1}}
% \newcommand{\figref}[1]{Fig.~\ref{#1}}
% \newcommand{\tabref}[1]{Table~\ref{#1}}
% \newcommand{\appref}[1]{App.~\ref{#1}}

\title{\LARGE \bf
Dynamic object goal pushing with mobile manipulators through model-free constrained reinforcement learning
}

%
\author{Ioannis Dadiotis$^{1,2}$, Mayank Mittal$^{3,4}$, Nikos Tsagarakis$^{1}$, Marco Hutter$^{3}$

\thanks{\hspace{-12pt} This project has received funding from the European Union’s Horizon Europe Framework Programme under grant agreement No 101070596 (euROBIN) and an ETH RobotX research grant funded through the ETH Zurich Foundation. Contact: {\tt\small ioannis.dadiotis@iit.it}}
\thanks{$^{1}$HHCM lab, IIT, Genoa 16163, Italy}
\thanks{$^{2}$DIBRIS, University of Genoa, Genoa 16145, Italy.}
\thanks{$^{3}$RSL, ETH Z\"{u}rich, Z\"{u}rich 8092, Switzerland. $^{4}$NVIDIA.}
}

% *** IEEE Copyright notice with TikZ ***

\usepackage{tikz}
\usepackage{textcomp}
\usepackage{lipsum}
\usepackage{outlines}

\usepackage{multirow}   % table with subcolumns

\newcommand\copyrighttext{%
  \scriptsize \textcopyright 2025 IEEE. Personal use of this material is permitted.
  Permission from IEEE must be obtained for all other uses, in any current or future
  media, including reprinting/republishing this material for advertising or promotional
  purposes, creating new collective works, for resale or redistribution to servers or
  lists, or reuse of any copyrighted component of this work in other works.}
\newcommand\copyrightnotice{%
\begin{tikzpicture}[remember picture,overlay]
\node[anchor=south,yshift=10pt] at (current page.south) {\parbox{\dimexpr\textwidth-\fboxsep-\fboxrule\relax}{\copyrighttext}};
\end{tikzpicture}%
}

\usepackage{array}
\usepackage{color, colortbl}

\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}

\begin{document}
\bstctlcite{IEEEexample:BSTcontrol}

\maketitle
%\copyrightnotice

\thispagestyle{empty}
\pagestyle{empty}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Non-prehensile pushing to move and reorient objects to a goal is a versatile loco-manipulation skill. In the real world, the object's physical properties and friction with the floor contain significant uncertainties, which makes the task challenging for a mobile manipulator. In this paper, we develop a learning-based controller for a mobile manipulator to move an unknown object to a desired position and yaw orientation through a sequence of pushing actions. The proposed controller for the robotic arm and the mobile base motion is trained using a constrained Reinforcement Learning (RL) formulation. We demonstrate its capability in experiments with a quadrupedal robot equipped with an arm. The learned policy achieves a success rate of 91.35\% in simulation and at least 80\% on hardware in challenging scenarios. Through our extensive hardware experiments, we show that the approach demonstrates high robustness against unknown objects of different masses, materials, sizes, and shapes. It reactively discovers the pushing location and direction, thus achieving contact-rich behavior while observing only the pose of the object. Additionally, we demonstrate the adaptive behavior of the learned policy towards preventing the object from toppling.
\end{abstract}

\input{Introduction}

\input{RelatedWork}

% \input{Background}

\input{Method}

\input{Results}

\input{Conclusion}

% \section*{APPENDIX}
% Appendixes should appear before the acknowledgment.

% \section*{ACKNOWLEDGMENT}
% The authors would like to thank...

\input{references}

\end{document}
