\section{Pandora's Cabinets} \label{sec:cabinets}

We next consider one more intermediate problem that will enable our general results, and is also interesting in its own right.
The problem captures e.g. Pandora's Box with Nonobligatory Inspection as a special case (see Appendix \ref{app:hardness}).

The \emph{Pandora's Cabinets$(\F)$} problem is the same as Cabinets-Prophets, but each arriving cabinet $i \in [n]$ now contains in each drawer $j \in [m_i]$ a bandit process $\M_i^j$.
Any algorithm can open at most one drawer per cabinet.
An offline algorithm $\OPT$ in general may interact adaptively with the cabinets, e.g. opening $j^*(i)$ on $i$, advancing the bandit $\M_i^{j^*(i)}$, opening some other drawer $j^*(i')$ on another cabinet, then coming back to $i$ to continue advancing $\M_i^{j^*(i)}$.
An online algorithm $\ALG$ must select a drawer $j(i)$ and employ a policy denoted $\pi_i^{j(i)}$ on the bandit $\M_i^{j(i)}$ until it halts, including with a decision of whether to claim $i$ before continuing to $i+1$.

The \emph{welfare} of $\ALG$ is the sum of all rewards claimed minus all search costs incurred, and can be written
  \[ \Welfof{\ALG} = \sum_{i=1}^n \Perfofon{\pi_i^{j(i)}}{\M_i^{j(i)}}~~, \]
where we recall from Section \ref{subsec:MSP} that the performance of a policy on an MSP is the value if claimed minus the sum of all search costs.
$\Welfof{\OPT}$ is similarly the sum of values claimed minus search costs incurred.

Recall from Section \ref{subsec:multistage} that each bandit process $\M_i^j$ has an associated capped value, which we denote $\kappa_i^j$, a random variable that represents an amortization of the net value attainable from interacting with $\M_i^j$.
Given a cabinet $i$, the randomness of the processes $(\M_i^j : j \in [m_i])$ in its drawers may be arbitrarily coupled and correlated.
This will be useful for Section \ref{sec:dags}.



We define \SAUP{} for the Pandora Cabinets setting.
\begin{definition}[\SAUP{} for Pandora's Cabinets] \label{def:pcsaup}
	  The \emph{Pandora's Cabinets Single-Agent Utility Problem (\SAUP$(i,\tau)$)}, given a cabinet $i$ with bandits $(\M_i^j : j \in [m_i])$ and a threshold $\tau$, is to compute a drawer $j$ and a policy $\pi$ for $\M_i^{j}$ in order to solve
	  \begin{align}
		    \max_{j,\pi} \E \left[ \Perfofon{\pi}{\M_i^j} - \Bof{\pi} \tau \right] , \label{eq:pcsaup}
		  \end{align}
	  where the expectation is over randomness in cabinet $i$ and the policy.
\end{definition}


Our main result in this setting is another reduction.
\begin{theorem}[Cabinet-Prophet to Pandora-Cabinets] \label{thm:cp-pc}
  For any downward-closed $\F$, if there exists an ex-ante prophet inequality $\alpha$ for Cabinets-Prophets$(\F)$, then there exists an ex-ante prophet inequality $\alpha$ for Pandora's Cabinets$(\F)$.
  If the former is efficient (respectively, \SAUP-based) the latter is efficient (respectively, \SAUP-based).
\end{theorem}
We prove this theorem by reducing the Pandora-Cabinets problem to a related Cabinet-Prophet instance, as follows.
Given an instance $\calI$ of Pandora's Cabinets$(\F)$ specified by $(\M_i^j : i \in [n], j \in [m_i])$, construct the instance $\calI'$ of Cabinets-Prophets$(\F)$ with random variables $(\kappa_i^j : i \in [n], j \in [m_i])$.
We relate the welfare of the two instances as follows.
Let $\OPT$ (respectively, $\OPT'$) be the optimal ex-ante feasible algorithm for Pandora's Cabinets$(\F)$ (respectively, Cabinets-Prophets$(\F)$).

The proof of Theorem \ref{thm:cp-pc} follows directly from the following lemmas, each a relatively direct consequence of non-exposure and Lemma \ref{lemma:magic-multistage}.
Proofs appear in Appendix \ref{app:cabinets}.
\begin{lemma} \label{lemma:cabinets-opt-upper}
	For any instance $\calI$, $\E[ \Welfofon{\OPT}{\calI}] \leq \E[ \Welfofon{\OPT'}{\calI'} ]$.
\end{lemma}

\begin{lemma} \label{lemma:cabinets-alg-lower}
	Given an online algorithm $\ALG'$ for Cabinets-Prophets$(\F)$, there is an online algorithm $\ALG$ for Pandora's Cabinets$(\F)$ such that for all instances $\mathcal{I}$, $\E[ \Welf_{\ALG}(\mathcal{I}) ] \geq \E[ \Welf_{\ALG'}(\mathcal{I'}) ]$.
	If $\ALG'$ is efficient (respectively, \SAUP-based), then $\ALG$ is efficient (respectively, \SAUP-based).
\end{lemma}

As a corollary of Theorems \ref{thm:cp-pc} and \ref{thm:classic-cp-saup}, we obtain an ex ante prophet inequality for matroid constraints.
\begin{corollary}\label{cor:cp-saup}
	For any matroid constraint $\F$, there is a \SAUP{}-based, efficient ex ante prophet inequality $\tfrac{1}{2}$ for Pandora's Cabinets$(\F)$.
\end{corollary}





