
\section{Conclusion and Future Work}\label{sec:6}

In this paper, we designed a limiting covariance matrix estimator for sketched stochastic Newton methods.~Our estimator is fully online and constructed entirely from the Newton iterates.~We~\mbox{established}~the consistency and convergence rate of the estimator. Compared to plug-in estimators for \mbox{second-order} methods, our estimator is asymptotically consistent and more computationally efficient, \mbox{requiring} no matrix factorization. Compared to batch-means estimators for first-order methods, our \mbox{estimator} is batch-free and exhibits faster convergence. Based on our study, we can then construct asymptotically valid confidence intervals/regions for the model parameters using sketched Newton methods. We~also discussed the generalization of our estimator to constrained stochastic problems. Extensive~experiments on regression problems demonstrate the superior performance of our estimator.



For future research, it would be of interest to explore the lower bound for the online \mbox{covariance} estimation problem, as it provides insights into the statistical efficiency of our weighted sample covariance. Additionally, constructing difference test statistics with different asymptotic distributions based on (sketched) Newton iterates could be promising. In particular, although~the~\mbox{normality}~achieved~by~Newton methods is asymptotically minimax optimal \citep{Na2022Statistical}, recent studies have observed that confidence intervals constructed using other test statistics, such as $t$-statistics \citep{Zhu2024High} and their variants \citep{Lee2022Fast, Luo2022Covariance, Chen2024Online}, may exhibit~\mbox{better}~coverage rates in some problems due to the absence of further covariance estimation.~Lastly, \mbox{performing}~inference based on Newton methods in non-asymptotic and \mbox{high-dimensional}~settings, where the~problem dimension grows with the sample size, would also be an interesting direction.


