\documentclass[sigconf]{acmart}

\usepackage{soul}
\usepackage{listings}

\lstset{
    basicstyle=\ttfamily\small,  
    breaklines=true,            
    frame=single,               
    backgroundcolor=\color{gray!10},  
    rulecolor=\color{black},      
    xleftmargin=0pt,            
    framexleftmargin=0pt,       
    showstringspaces=false,     
    fontadjust=true  % <- Ensures it follows document-wide font settings
}


\usepackage{multirow}
\usepackage{subfigure}
\usepackage{xspace}
\usepackage{rotating}
\usepackage{algorithm}



\usepackage{algorithmic}
\usepackage{mathtools}


%math
\usepackage{amsmath}% http://ctan.org/pkg/amsmath

\newcommand{\eg}{e.g.\@\xspace}
\newcommand{\ie}{i.e.\@\xspace}
\newcommand{\etal}{et~al.\@\xspace}
\newcommand{\etc}{{\em etc}\xspace}
\newcommand{\TK}{{\bf TK}\xspace}
\newcommand{\tk}{\TK}

\newcommand{\financialcnt}{12\xspace}


\newcommand{\platform}{\textit{TermLens}\xspace}
\newcommand{\termname}{unfavorable financial\xspace}
\newcommand{\TermName}{Unfavorable Financial\xspace}
\newcommand{\Termname}{Unfavorable financial\xspace}
\newcommand{\websitecnt}{8,979\xspace}
\newcommand{\termcnt}{1.9 million\xspace}
\newcommand{\websitepct}{42.06\%\xspace}
\newcommand{\termpct}{0.5\%\xspace}
\newcommand{\termtypecnt}{22\xspace}
\newcommand{\termcatcnt}{4\xspace}
\newcommand{\myparagraph}[1]{\textbf{\textit{#1}:}\hspace{3pt}}


\usepackage{pifont} % Approved package
\newcommand{\filledCircle}{\ding{108}} % ● Full circle
\newcommand{\halfFilledCircle}{\ding{109}} % ◐ Half-filled circle



\copyrightyear{2025}
\acmYear{2025}
\setcopyright{cc}
\setcctype{by}
\acmConference[WWW '25]{Proceedings of the ACM Web Conference 2025}{April 28-May 2, 2025}{Sydney, NSW, Australia}
\acmBooktitle{Proceedings of the ACM Web Conference 2025 (WWW '25), April 28-May 2, 2025, Sydney, NSW, Australia}
\acmDOI{10.1145/3696410.3714573}
\acmISBN{979-8-4007-1274-6/25/04}

\settopmatter{printacmref=true}


%\showthe\font



\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title[Harmful Terms and Where to Find Them]{Harmful Terms and Where to Find Them: Measuring and Modeling Unfavorable Financial Terms and Conditions in Shopping Websites at Scale}


\author{Elisa Tsai}
\affiliation{%
  \institution{University of Michigan}
  \city{Ann Arbor}
  \state{Michigan}
  \country{USA}
}
\email{eltsai@umich.edu}


\author{Neal Mangaokar}
\affiliation{
  \institution{University of Michigan}
  \city{Ann Arbor}
  \state{Michigan}
  \country{USA}
}
\email{nealmgkr@umich.edu}


\author{Boyuan Zheng}
\affiliation{
  \institution{University of Michigan}
  \city{Ann Arbor}
  \state{Michigan}
  \country{USA}
}
\email{boyuann@umich.edu}

\author{Haizhong Zheng}
\affiliation{
  \institution{University of Michigan}
  \city{Ann Arbor}
  \state{Michigan}
  \country{USA}
}
\email{hzzheng@umich.edu}

\author{Atul Prakash}
\affiliation{
  \institution{University of Michigan}
  \city{Ann Arbor}
  \state{Michigan}
  \country{USA}
}
\email{aprakash@umich.edu}



%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Terms and conditions for online shopping websites often contain terms that can have significant financial consequences for customers. 
Despite their impact, there is currently no comprehensive understanding of the types and potential risks associated with unfavorable financial terms. Furthermore, there are no publicly available detection systems or datasets to systematically identify or mitigate these terms.
In this paper, we take the first steps toward solving this problem with three key contributions.

\textit{First}, we introduce \textit{TermMiner}, an automated data collection and topic modeling pipeline to understand the landscape of unfavorable financial terms.
\textit{Second}, we create \textit{ShopTC-100K}, a dataset of terms and conditions from shopping websites in the Tranco top 100K list, comprising 1.8 million terms from 8,251 websites. Consequently, we develop a taxonomy of 22 types from 4 categories of unfavorable financial terms---spanning purchase, post-purchase, account termination, and legal aspects.
\textit{Third}, we build \textit{TermLens}, an automated detector that uses Large Language Models (LLMs) to identify unfavorable financial terms. 

Fine-tuned on an annotated dataset, \textit{TermLens} achieves an F1 score of 94.6\% and a false positive rate of 2.3\% using GPT-4o. 
When applied to shopping websites from the Tranco top 100K, we find that 42.06\% of these sites contain at least one unfavorable financial term, with such terms being more prevalent on less popular websites. Case studies further highlight the financial risks and customer dissatisfaction associated with unfavorable financial terms, as well as the limitations of existing ecosystem defenses.

\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10002951.10003260.10003277</concept_id>
       <concept_desc>Information systems~Web mining</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002978.10002997.10003000</concept_id>
       <concept_desc>Security and privacy~Social engineering attacks</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10003456.10003462.10003544.10011709</concept_id>
       <concept_desc>Social and professional topics~Consumer products policy</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Web mining}
\ccsdesc[300]{Security and privacy~Social engineering attacks}
\ccsdesc[500]{Social and professional topics~Consumer products policy}



%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Topic modeling; Unfavorable financial terms; Consumer protection; Terms and conditions dataset; Deceptive content}

%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle



%\section{CCS Concepts and User-Defined Keywords}

%Two elements of the ``acmart'' document class provide powerful
%taxonomic tools for you to help readers find your work in an online search.

%The ACM Computing Classification System ---
%\url{https://www.acm.org/publications/class-2012} --- is a set of
%classifiers and concepts that describe the computing
%discipline. Authors can select entries from this classification
%system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
%commands to be included in the \LaTeX\ source.

%User-defined keywords are a comma-separated list of words and phrases
%of the authors' choosing, providing a more flexible way of describing
%the research being presented.

%CCS concepts and user-defined keywords are required for for all
%articles over two pages in length, and are optional for one- and
%two-page articles (or abstracts).



%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{01introduction}
\input{02related_work}
\input{03data}
\input{04system}
\input{05evaluation}
\input{06limitation}
\input{07conclusion}

\bibliographystyle{ACM-Reference-Format}
% \bibliography{ref}
\input{00main.bbl}


%%
%% If your work has an appendix, this is the place to put it.
\input{appendix}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.
