\section{Related Work}\label{sec:lit}
% Our literature review revolves around studies in sarcasm detection and sarcasm-related generation tasks.
\input{tex/dataset}
% \paragraph{Sarcasm Detection:} \label{sec:lit-review-sarcasm-detection}
Sarcasm detection is a task that involves detecting whether sarcasm is present in a sample or not. The reason behind what makes a sample sarcastic is not in the scope of this task. In the beginning, researchers focused only on the textual modality for this task. For instance, \citet{Bouazizi-etal-sarcasm-detection-2016} and \citet{felbo-etal-2017-using} explored using hand-crafted features such as emojis, certain punctuation marks, and lexicons to detect whether an utterance is sarcastic or not. Neural network-based architectures have been used here as well (\citealp{tay-etal-2018-reasoning}; \citealp{babanejad-etal-2020-affective}). 

However, recognizing the multimodal nature of online content, \citet{schifanella-etal-2016} proposed the task of multimodal sarcasm detection. They proposed a system that uses Convolutional Neural Networks \citep{ma-etal-cnn-2015} to fuse textual and visual information and detect the presence of sarcasm. Having identified some limitations with this work, \citet{Qiao_Jing_Song_Chen_Zhu_Nie_2023}, \citet{kumar-etal-2022-become} and \citet{chakrabarty-etal-2020-r} explored ways to resolve them using Graph Convolutional Networks \citep{kipf2017semisupervised}. \citet{castro-etal-2019-towards} extended this task to conversational dialog systems and even proposed a new dataset, MUStARD, for the same. \citet{bedi-etal-code-mixed-2023} explored sarcasm detection in Hindi-English code-mixed conversations.

% \paragraph{Sarcasm Generation} \label{sec:lit-review-sarcasm-gen}
Researchers have also explored sarcasm analysis from a generative perspective, using only textual inputs in the beginning. For example, \citet{peled-reichart-2017-sarcasm} and \citet{dubey-2019-sarcasm-generation} explored the conversion of sarcastic texts into their non-sarcastic interpretations using machine translation techniques. On the contrary, \citet{mishra-etal-2019-modular} have worked on using sentences that express a negative sentiment to generate corresponding sarcastic text. 

\citet{Desai_Chakraborty_Akhtar_2022} proposed the task of multimodal sarcasm explanation (MuSE) and also proposed the MORE dataset for the same. Additionally, they proposed a model to benchmark MORE on this task, called ExMORE, which was built on a BART \citep{lewis-etal-2020-bart} backbone. 
However, ExMORE suffered from some limitations which were addressed by \citet{jing-etal-2023-multi} in their model, TEAM, which exceeded the performance of all existing baselines by large margins. Despite this, TEAM suffers from some limitations discussed in Section~\ref{sec:intro}. 