\vspace{-8pt}
\section{Conclusion}
In this study, we introduce \method{}, a novel framework designed to advance drug discovery by integrating structured biomedical knowledge and unstructured textual data. By leveraging the Drug Repurposing Knowledge Graph (DRKG) and enriching it with relevant contextual information from PubMed Central and Clinical Trials, we curated a comprehensive dataset that enhances the accuracy of drug recommendation and the explainability of reasoning processes.

Our experimental results underscore the efficacy of \method{} across both drug selection and reasoning generation tasks. For drug selection, \method{} consistently outperformed baseline models, achieving the highest F1 scores on both the xDrugRecommendation and MIMIC-III datasets. In reasoning generation, \method{} demonstrated superior ROUGE scores, highlighting its ability to produce high-quality, interpretable rationales for drug recommendations. Importantly, the inclusion of enriched background information significantly boosted the performance of all evaluated models, further validating the value of integrating domain-specific knowledge into language modeling frameworks.

This work emphasizes the potential of combining graph-based structured knowledge with text-based contextual data to enhance precision and interpretability in drug discovery. By bridging advanced language modeling techniques with biomedical data, \method{} provides a robust foundation for more effective and explainable AI-driven drug discovery pipelines.

In future work, we aim to extend the scalability of \method{} to larger and more diverse datasets, exploring its potential applications in other critical biomedical tasks, such as drug repurposing and adverse effect prediction. These advancements will further contribute to the integration of AI in accelerating therapeutic innovation and improving patient outcomes.