% ----------------------- cvpr ----------------------------------
% CVPR 2025 Paper Template; see https://github.com/cvpr-org/author-kit
\documentclass[10pt,twocolumn,letterpaper]{article}

%%%%%%%%% PAPER TYPE  - PLEASE UPDATE FOR FINAL VERSION
\usepackage{cvpr}              % To produce the CAMERA-READY version
% \usepackage[review]{cvpr}      % To produce the REVIEW version
% \usepackage[pagenumbers]{cvpr} % To force page numbers, e.g. for an arXiv version

\usepackage[accsupp]{axessibility}
\usepackage{pgf}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
\usepackage{pgfplotstable}
\usepackage{tikz}
\usepackage{csvsimple}
\usepackage{graphicx}
\usepackage{xifthen}
\usepackage{arydshln}
\usepackage{adjustbox}
\usepackage{balance}
\usepackage{tcolorbox}
\usepackage{etoolbox}
\usetikzlibrary{shapes,calc,positioning, decorations.text, arrows.meta, calc, shadows.blur, shadings, fit}

\definecolor{cvprblue}{rgb}{0.21,0.49,0.74}
\usepackage[pagebackref,breaklinks,colorlinks,allcolors=cvprblue]{hyperref}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\paperID{10138} % *** Enter the Paper ID here
\def\confName{CVPR}
\def\confYear{2025}

%%%%%%%%% TITLE - PLEASE UPDATE
\title{ILIAS: Instance-Level Image retrieval At Scale}

%%%%%%%%% AUTHORS - PLEASE UPDATE
\author{Giorgos Kordopatis-Zilos \quad Vladan Stojni\'{c} \quad Anna Manko \quad Pavel \v{S}uma  \\ Nikolaos-Antonios Ypsilantis \quad Nikos Efthymiadis \quad Zakaria Laskar \\ Ji\v{r}\'{i} Matas \quad Ond\v{r}ej Chum \quad Giorgos Tolias \\ \\
VRG, FEE, Czech Technical University in Prague \vspace{-7pt}
}

\begin{document}

\maketitle

\input{tex/abbrev}
\input{tex/plots}

\input{tex/abstract}

\input{tex/intro}
\input{tex/related}
\input{tex/dataset}
\input{tex/methods}
\input{tex/experiments}
\input{tex/conclusions}

\newpage

\begin{center}
\Large \textbf{Supplementary materials}
\vspace{25pt}
\end{center}

\renewcommand\thesection{\Alph{section}}

\setcounter{section}{0}
\setcounter{table}{0}
\setcounter{figure}{0}

\input{tex/supplementary_arxiv}

\section*{Acknowledgments}

\noindent\textbf{Funding acknowledgement.} This work was supported by the Junior Star GACR (grant no. GM 21-28830M), Horizon MSCA-PF (grant no. 101154126), Programme Johannes Amos Comenius (grant no. CZ.02.01.01/00/22\_010/0003405), CTU in Prague (grant no. SGS23/173/OHK3/3T/13), the CTU institutional support (Future fund). JM was supported by Ministry of the Interior of the Czech Republic (project no. VJ02010041).
We acknowledge VSB – Technical University of Ostrava, IT4Innovations National Supercomputing Center, Czech Republic, for awarding this project access to the LUMI supercomputer, owned by the EuroHPC Joint Undertaking, hosted by CSC (Finland) and the LUMI consortium through the Ministry of Education, Youth and Sports of the Czech Republic through the e-INFRA CZ (grant ID: 90254).

\medskip

\noindent\textbf{Contributors acknowledgments.} We want to thank Larysa Ivashechkina for her work on the annotation of object bounding boxes and masks and the initial filtering of text queries. We also want to thank all contributors for the collection of \ours dataset. Additionally to all authors, the list of external contributor in alphabetical order: Aggeliki Tserota, Anna Nisyraiou, Celeste Abreu, Charalambos Tzamos, Christina Tserota, Dimitris Karageorgiou, Dmytro Mishkin, Eleni Karantali, Eleni Papadopoulou, Eva Tsiliakou, Kelly Kordopati-Zilou, Markos Zampoglou, Noa Garcia, Panagiotis Tassis, Paraskevas Kordopatis, Pavlos Alexantonakis, Ruslan Rozumnyi, Sne\v{z}ana \'{C}urguz, Tereza Nejedl\'{a}, Tom\'{a}\v{s} Jel\'{i}nek, Yannis Kalantidis, Vasilis Alexiadis, Yankun Wu.

{   
    \small
    \bibliographystyle{ieeenat_fullname}
    \bibliography{main}
}

\end{document}
