\section{Limitations}
\label{sec:limitation}

Both p-SSA and r-SSA demonstrated promising results when mitigating infeasible safe control optimizations for humanoids by relaxing the constraints with a minimal amount.
However, as long as relaxation is needed, the safety violation cannot be bounded, preventing any safety guarantee to hold.
To enhance our work with certain guarantees, it may be possible to adapt the original control constraints based on the solved minimal relaxation.
Then, we might be able to temporarily restore safety guarantees, enhancing post-infeasible-QP safety.

\section{Discussions and Conclusions}
\label{sec:conclusion}

In this paper, we proposes the Relaxed Safe Set Algorithm (r-SSA) and Projected Safe Set Algorithm (p-SSA) to mitigate infeasible safe control optimizations in the problem of dexterous safety in cluttered environments.
We showed that there are multiple sources of infeasibility, making it extremely hard to synthesize permanently feasible safe controllers for multi-constraint cases.
Through both simulation and real experiments on a Unitree G1 humanoid, we validated both r-SSA and p-SSA as practical approaches to computing safe control in challenging collision avoidance tasks.
We also showed that p-SSA achieved the optimal performance-safety trade-off across various tasks with zero parameter tuning.
For future work, we are interested in investigating optimal ways to balance between different safety constraints for p-SSA and improve the safety index design to account for higher order dynamics.






