\chapter{Contact-free non-planar haptics with a spherical electromagnet}
\label{ch:shared:contact}


\contribution{
We investigated a shared variable interface using a contact-free volumetric haptic feedback device. This device employs a symmetric electromagnet in combination with a dipole magnet model and a simple control law to deliver dynamically adjustable forces onto a handheld tool. The tool requires only an embedded permanent magnet, allowing it to remain completely untethered. Despite being contact-free, the force is grounded via the spherical electromagnet, enabling the user to feel relatively large forces (up to 1N at contact). The device can render both attracting and repulsive forces within a thin shell around the electromagnet. We conducted a user experiment with six participants to characterize the force delivery aspects and perceived precision of our system. Our findings indicate that users can discern at least 25 locations for repulsive forces. This research marks the first step towards understanding shared variable interfaces. By eliciting forces on the tool manipulated by the user, we enable haptic feedback in the same modality as the user's exertion, resulting in natural and intuitive interactions.
}

\input{\dir/contact_free/sections/01_introduction_v2.tex}

\input{\dir/contact_free/sections/02_related_work.tex}

\input{\dir/contact_free/sections/03_system_description.tex}

\input{\dir/contact_free/sections/04_system_evaluation.tex}

\input{\dir/contact_free/sections/05_user_evaluation.tex}

\input{\dir/contact_free/sections/06_discussion_conclusion.tex}
