
\section{Discussion and Conclusion}
In this work, we investigated the phenomenon of hallucinations occurring with high certainty, even when the model possesses the knowledge required to generate correct answers (\chk). While hallucinations are often correlated with uncertainty, we demonstrated that some hallucinations defy this trend, occurring with both high certainty and sufficient knowledge. Furthermore, we showed that large models also produce high-certainty hallucinations and that instruction-tuned models exhibit poorer calibration between uncertainty and hallucinations, amplifying this issue. Importantly, we provided evidence that these hallucinations are not mere noise but represent a distinct and robust subset of model behavior. Finally, we demonstrated that existing mitigation methods based on uncertainty fail to address these high-certainty hallucinations (\chk). 

The phenomenon of \chk has significant implications for understanding the reliability and trustworthiness of LLMs. These examples reveal a critical vulnerability: Models may produce factually incorrect responses while
 maintaining a high degree of certainty, even when they possess the capability to provide the correct answer.

This presents a significant challenge for the deployment of LLMs in high-stakes domains, such as healthcare, legal applications, and scientific research, where erroneous outputs can have severe consequences.



\section{Limitations}

This work demonstrates the existence of \chk, which poses a challenge for uncertainty-based methods in fully detecting and mitigating hallucinations. However, this study does not propose a solution to this problem.
 Additionally, this work solely demonstrated the existence of \emph{certain hallucinations despite knowledge} and did not consider other types of hallucinations. Finally, we employed a single method for calculating the threshold for certainty hallucinations. While this approach provides a reasonable definition, alternative methods for defining the threshold may yield varying results.
