\section{Related Work}
\label{sec:2}

\subsection{Language Models for Scientific Literature}%1311

The introduction of the transformer architecture ____ revolutionized natural language processing, marking a new era in the field, with pretrained language models like BERT ____ significantly advancing performance across a wide range of tasks. This progress quickly extended to the scientific domain, leading to the development of domain-specific models such as SciBERT ____, which set new benchmarks on various scientific NLP tasks. SciBERT and similar models demonstrate clear advantages over general-purpose models ____, and have therefore been applied to a variety of tasks within the scientific domain, including literature search and similarity assessment ____, classification ____, and summarization ____, with similar pretrained models having been trained for the general biomedical domain ____ as well as for the biodiversity domain ____.

More recently, the improved performance of autoregressive language models ____ has driven a shift toward leveraging these models for a wide range of tasks. Openly available models, such as Llama-2 ____, alongside proprietary systems like ChatGPT, have established new state-of-the-art results in various scientific document processing tasks, including structured information extraction ____, term extraction ____, text classification, named entity recognition and and question answering ____.

A range of benchmarks has been developed specifically for information extraction from scientific full texts, often accompanied by proposed models. These benchmarks target various tasks, including dataset mention detection ____, entity and relation extraction ____, general information extraction ____, and summarization ____.

\subsection{Language Models for Biodiversity Science}

In the specific domain of biodiversity science, transformer encoder architectures have been employed to tackle tasks such as hypothesis classification ____, biodiversity analysis ____, named entity recognition and relation extraction ____, as well as hypothesis evidence localization ____. Additionally, autoregressive models have been applied to tasks such as literature review, question answering ____, and structured information extraction ____, with further potential applications continuing to emerge ____.