\section{Related Works}
\label{sec:related}
%
This research primarily pertains to the field of deraining studies.
%
Many methods have been proposed to develop state-of-the-art deraining networks.
%
These works include deep network designs ____, residual networks ____, recurrent networks ____, multi-task ____ and multi-scale designs ____, sparsity-based image modeling ____, low-rank prior ____, model-driven solutions ____, attention mechanisms ____, Transformer-based networks ____, adversarial learning ____, representation learning ____, semi-supervised ____ and unsupervised learning ____.
%
Deep learning methods are data-hungry but collecting rain streaks and background image pairs is challenging.
%
A lot of works have been proposed to synthesize rain streaks with better results.
%
Garg \etal ____ first propose a physically-based photo-realistic rendering method for synthesizing rain streaks.
%
Zhang \etal ____ and Fu \etal ____ use Photoshop software to manually add rain effects to images to build the synthetic paired data.
%
Due to the poor generalization performance of existing methods, models trained on synthetic images were found to be ineffective in real-world scenarios.
%
Some works ____ that have contributed to real collected deraining datasets.
%
However, acquiring these datasets is still expensive and cannot solve the problem of poor generalization.

There are also works that mention the generalization issue of the deraining models.
%
Xiao \etal ____ and Zhou \etal ____ attempt to improve the generalization ability of deraining networks by accumulating knowledge from multiple synthetic rain datasets, as most existing methods can only learn the mapping on a single dataset for the deraining task.
%
However, this attempt does not allow the network to generalize beyond the training set.
%
In addition, semi-supervised methods ____ have also been used to improve the deraining effect on real images, and we also include the representative method Syn2Real ____.
%
There are some semi-supervised deraining methods ____ that are proposed to improve the performance of deraining models in real-world scenarios.
%
When obtaining some real images similar to the test images, these works can indeed achieve some improvement. 
%
However, these improvements are not brought about by improving the generalization ability.
%
Their solution is to include real test images in the training set, even if we don't have corresponding clean images. 
%
These methods are effective when we can determine the characteristics of the test image. 
%
However, this does not solve the generalization problem. 
%
Because these methods manage to convert ``rain outside the training set'' to ``rain inside the training set''. 
%
Since data collection is extremely difficult, this method still faces great challenges in practice.
%
Unlike the majority of existing deraining research, we do not propose new network structures, loss functions, or datasets.
%
Our objective is to analyze and understand the generalization problem within the context of the deraining task.
%
We will proceed to review previous works focusing on interpretability and understanding generalization in low-level vision.


Deep learning interpretability research aims to understand the mechanism of deep learning methods and to obtain clues about the success or failure of these methods.
%
Without a deep understanding of these working mechanisms, we are not convinced to move forward in the right direction.
%
The research on deep learning interpretability follows a long line of works, most focusing on the classification task ____.
%
Low-level vision tasks have embraced great success with powerful deep learning techniques.
%
There are also works on interpretability for these deep low-level networks ____.
%
Gu and Dong ____ bring the first interpretability tool for image super-resolution (SR) networks.
%
Xie \etal ____ find the most discriminative filters for each specific degradation in a blind SR network, whose weights, positions, and connections are important for the specific function in blind SR.
%
Magid \etal ____ use a texture classifier to assign patches with semantic labels, to identify global and local sources of SR errors.
%
Shi \etal ____ show that Transformers can directly utilize multi-frame information from unaligned frames, and adopting alignment methods is sometimes harmful to Transformers in video super-resolution.
%
Hu \etal ____ propose the causal effect map to interpret low-level vision models by focusing on causation rather than mere correlation.
%
The closest work to this paper is the deep degradation representation proposed by ____.
%
They argue that SR networks tend to overfit to degradations and show degradation ``semantics'' inside the network.
%
The presence of these representations often means a decrease in generalization ability.
%
The utilization of this knowledge can guide us to analyze the generalization performance of SR methods ____.

The generalization problem in low-level vision often arises when the testing degradation does not match the training degradation, \eg, different downsampling kernel ____ and noise distribution ____.
%
The existing works either develop blind methods to include more degradation possibilities in the training process or make the training data closer to real-world applications.
%
Only a little work has been proposed to study the reasons for this lack of generalization performance ____.
%
No research has attempted to investigate the interpretation of the training process of low-level vision networks, especially from the perspective of the generalization problem.