\documentclass{article}

% --------------------- Packages & Macros ---------------------
\usepackage{takeuchi_mcr}  % (Local macro file, if desired)
\usepackage{bm}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}    % professional-quality tables
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{booktabs}
\usepackage[numbers]{natbib}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{slantsc}
\usepackage[textsize=tiny]{todonotes}
\usepackage{authblk}
\usepackage{rotating}

% -------------------------------------------------------------
% Remove the line: \usepackage{icml2025}
% and any ICML macros (e.g., \icmltitle, \icmlauthorlist, etc.)
% Replace them with standard LaTeX title/author/affiliation.
% -------------------------------------------------------------

% THEOREM ENVIRONMENTS
% -------------------------------------------------------------
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% -------------------------------------------------------------
% Standard LaTeX Title and Author
% -------------------------------------------------------------
\title{Generalized Kernel Inducing Points by Duality Gap for Dataset Distillation}

\author[1]{Tatsuya Aoyama\thanks{Equal contribution.}}
\author[1]{Hanting Yang\thanks{Equal contribution.}}
\author[2]{Hiroyuki Hanada}
\author[1]{Satoshi Akahane}
\author[1]{Tomonari Tanaka}
\author[1]{Yoshito Okura}
\author[3]{Yu Inatsu}
\author[2]{Noriaki Hashimoto}
\author[4]{Taro Murayama}
\author[4]{Hanju Lee}
\author[4]{Shinya Kojima}
\author[1,2]{Ichiro Takeuchi}

\affil[1]{Nagoya University, Nagoya, Japan}
\affil[2]{RIKEN, Nagoya, Japan}
\affil[3]{Nagoya Institute of Technology, Nagoya, Japan}
\affil[4]{DENSO CORPORATION, Kariya, Japan}
\affil[]{{\tt takeuchi.ichiro.n6@f.mail.nagoya-u.ac.jp}}

% -------------------------------------------------------------
\begin{document}
\maketitle

\begin{abstract}
We propose \textit{Duality Gap KIP} (DGKIP), an extension of the Kernel Inducing Points (KIP) method for dataset distillation. While existing dataset distillation methods often rely on bi-level optimization, DGKIP eliminates the need for such optimization by leveraging duality theory in convex programming. The KIP method has been introduced as a way to avoid bi-level optimization; however, it is limited to the squared loss and does not support other loss functions (e.g., cross-entropy or hinge loss) that are more suitable for classification tasks. DGKIP addresses this limitation by exploiting an upper bound on parameter changes after dataset distillation using the duality gap, enabling its application to a wider range of loss functions. We also characterize theoretical properties of DGKIP by providing upper bounds on the test error and prediction consistency after dataset distillation. Experimental results on standard benchmarks such as MNIST and CIFAR-10 demonstrate that DGKIP retains the efficiency of KIP while offering broader applicability and robust performance.
\end{abstract}

% ------------------------ Main Text ---------------------------

\input{Sections/Introduction}
\input{Sections/Preliminary}
\input{Sections/Methods}
\input{Sections/Experiments}
\input{Sections/Conclusion}
\input{Sections/Impact_Statement}
% -------------------------------------------------------------
% Acknowledgments
% -------------------------------------------------------------
\section*{Acknowledgments}
This work was partially supported by MEXT KAKENHI (JP20H00601, JP23K16943, JP24K15080), JST CREST (JPMJCR21D3 including AIP challenge program, JPMJCR22N2), JST Moonshot R\&D (JPMJMS2033-05), JST AIP Acceleration Research (JPMJCR21U2), JST ACT-X Grant Number JPMJAX24C3, NEDO (JPNP20006) and RIKEN Center for Advanced Intelligence Project.

% -------------------------------------------------------------
% Impact Statement (Optional or could be removed for arXiv)
% -------------------------------------------------------------



% -------------------------------------------------------------
% Bibliography
% -------------------------------------------------------------
\bibliographystyle{plain}
\bibliography{ref}

% -------------------------------------------------------------
% Appendix
% -------------------------------------------------------------
\appendix
\input{Sections/Appendix}

\end{document}
