\subsection{Comparative Statics: Money Redistribution for Welfare Analysis}
\label{subsec:comparative_money}

% In this section, we assume that $f''_i(k^*_i) > 0$ for model regularity. 

In this section, we study comparative statics, \ie, how the players' utilities will change if the model parameters are modified by an infinitesimal amount. 
We characterize the infinitesimal modification by money redistribution, \ie, replace $\{f_i(k_i)\}_\iinn$ by $\{f_i(k_i + \delta_i t)\}_\iinn$, where $\bmdd = (\delta_1,\dots, \delta_n)\in\bbR^n$ is called the direction of money redistribution and $t\in \bbR$ is called the change magnitude. 
Overall, there is a $\bmdd t$ shift in the gain level of players.
The goal of infinitesimal change drives us to study the case $t\to 0$.

% Let $\bmdd$ be an infinitesimal change of money redistribution, \ie, the direction of money change. We model the change magnitude by $t > 0$.
In this way, the utility of player $i$ becomes
\begin{eqnarray*}
  u_i(\bmx;t) = f_i(k_i + \delta_i t) - c_i(x_i).   
\end{eqnarray*}

Denote $\bmx^*(t)$ as the NE when the change magnitude is $t$. We do not assume the uniqueness of NE anymore, and $\bmx^*(t)$ might be not unique. However, we assume the first-order differentiability of $\bmx^*(t)$ with respect to $t$, as well as that $\bmx^*(t)$ is an inner point of $X$. 
These assumptions are quite natural. 
For the first assumption, if the game changes with an infinitesimal magnitude and players always achieve the rational outcome, \ie, NE, then it is imaginable and intuitive that the outcome of players should also change minimally.
The second assumption is only technical.
We denote $u_i(t) = u_i(\bmx^*(t);t)$ for a little abuse of notation when the context is clear. We are mainly concerned about $u'_i(0)$, which means that what the marginal change of $\delta$ would affect the players' utilities. Thus, we have the following result.

%We build this result as follows.

\input{Contents/Theorems/thm_comparative_money}

We also demonstrate some examples to illustrate the implications of the result in \cref{thm:comparative:money}.
\input{Contents/Examples/comparative_money}

% \subsection{Prosociality}

% In this section, we study the effect of prosociality on the players' utilities. We denote $V = \{v_{ij}\}_\ijinn$ as the infinitesimal change of prosociality, \ie, the direction of prosociality change. $v_{ij},\ \inej$ represnets the altruistic degree of player $i$ on player $j$. We normalize that $v_{ii} = 0$ and we admit the case that $v_{ij}<0$. We also model the change magnitude by $t > 0$. In this way, the utility of player $i$ becomes,
% \begin{align*}
%     u_i(\bmx;t) = f_i(k_i) - c_i(x_i) + \sum_\jnei (v_{ij} t) f_j(k_j) ,
% \end{align*}

% The $u_i(t)$ and $\bmx^*(t)$ is similarly defined. We have following results,

% \begin{theorem}
% \label{thm:comparative:prosociality}
% \begin{align*}
%     \bmu'(0) = \myx{TBA}
% \end{align*}
% \end{theorem}
