\documentclass[10pt,conference,letterpaper]{IEEEtran}
\IEEEoverridecommandlockouts

\listfiles

\input{packages.tex}
\input{commands.tex}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{End-to-End Reliability in Wireless IEEE~802.1Qbv Time-Sensitive Networks}

\author{\IEEEauthorblockN{Simon Egger$^\ast$, James Gross$^\dagger$, Joachim Sachs$^\mathsection$, Gourav P. Sharma$^\dagger$, Christian Becker$^\ast$, Frank Dürr$^\ast$}
\IEEEauthorblockA{{}$^\ast$University of Stuttgart, {}$^\dagger$KTH Royal Institute of Technology, Stockholm, {}$^\mathsection$Ericsson Research, Aachen\\
\{simon.egger,$\,$christian.becker,$\,$frank.duerr\}@ipvs.uni-stuttgart.de, \{jamesgr,$\,$gpsharma\}@kth.se, joachim.sachs@ericsson.com
}}

\maketitle

\begin{abstract}
  Industrial cyber-physical systems require dependable network communication with formal end-to-end reliability guarantees.
  Striving towards this goal, recent efforts aim to advance the integration of 5G into Time-Sensitive Networking~(TSN).
  However, we show that IEEE~802.1Qbv TSN schedulers that are unattuned to 5G packet delay variations may jeopardize any reliability guarantees provided by the 5G system.
  We demonstrate this on a case where a $\qty{99.99}{\percent}$ reliability in the inner 5G network diminishes to below $\qty{10}{\percent}$ when looking at end-to-end communication in TSN.
  In this paper, we overcome this shortcoming by introducing Full Interleaving Packet Scheduling~(FIPS) as a wireless-friendly IEEE~802.1Qbv scheduler.
  To the best of our knowledge, FIPS is the first to provide formal end-to-end QoS guarantees in wireless TSN.
  FIPS allows a controlled batching of TSN streams, which improves schedulability in terms of the number of wireless TSN streams by a factor of up to~$\times 45$.
  Even in failure cases, FIPS isolates the otherwise cascading QoS violations to the affected streams and protects all other streams.  
  With formal end-to-end reliability, improved schedulability, and fault isolation, FIPS makes a substantial advance towards dependability in wireless TSN. 
\end{abstract}

\input{01_introduction.tex}
\input{02_related_work.tex}
\input{03_background.tex}
\input{04_problem_statement.tex}
\input{05_robustness.tex}
\input{06_fips.tex}
\input{07_evaluation.tex}
\input{08_conclusion.tex}

\section*{Acknowledgment}
This work was supported by the European Commission through the H2020 project DETERMINISTIC6G (Grant Agreement no. 101096504).
We thank Melanie Heck and the DETERMINISTIC6G consortium for their valuable feedback.

\bibliographystyle{IEEEtran.bst}
\bibliography{IEEEabrv,bibliography.bib}

\clearpage
\onecolumn

\appendix
\input{09_appendix.tex}

\end{document}

\typeout{get arXiv to do 4 passes: Label(s) may have changed. Rerun}
