%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

%In case you encounter the following error:
%Error 1010 The PDF file may be corrupt (unable to open PDF file) OR
%Error 1000 An error occurred while parsing a contents stream. Unable to analyze the PDF file.
%This is a known problem with pdfLaTeX conversion filter. The file cannot be opened with acrobat reader
%Please use one of the alternatives below to circumvent this error by uncommenting one or the other
%\pdfobjcompresslevel=0
%\pdfminorversion=4

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{graphicx}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{bm}
\usepackage{subfig}
\usepackage{url}
\newcommand\figsize{0.46}

\title{\LARGE \bf Real-Time Fast Marching Tree for Mobile Robot Motion Planning in Dynamic Environments*
\thanks{*This research was supported in part by the Natural Science and Engineering Research Council of Canada (NSERC) under grant number DNDPJ 533392-18, General Dynamics Land Systems (Canada), and Defence R\&D Canada (DRDC).}
}

\author{Jefferson~Silveira$^{1}$\thanks{$^{1}$ J.\ Silveira is with the Department of Electrical \& Computer Engineering and the Ingenuity Labs Research Institute, Queen's University, Kingston, ON K7L 3N6 Canada \texttt{jefferson.silveira@queensu.ca}}, %~\IEEEmembership{Member,~IEEE,}
        Kleber~Cabral$^{2}$\thanks{$^{2}$K.\ Cabral is with the School of Computing, Queen's University, Kingston, ON K7L 3N6 Canada \texttt{kleber.cabral@queensu.ca}}, %~\IEEEmembership{Fellow,~OSA,}
        Sidney~Givigi$^{3}$\thanks{$^{3}$S.\ Givigi is with the School of Computing and the Ingenuity Labs Research Institute, Queen's University, Kingston, ON K7L 3N6 Canada \texttt{sidney.givigi@queensu.ca}}
        and~Joshua~A.~Marshall$^{4}$\thanks{$^{4}$J.\ Marshall is with the Department of Electrical \& Computer Engineering and the Ingenuity Labs Research Institute, Queen's University, Kingston, ON K7L 3N6 Canada \texttt{joshua.marshall@queensu.ca}}%,~\IEEEmembership{Senior Member,~IEEE}% <-this % stops a space
        }
        
% \author{Albert Author$^{1}$ and Bernard D. Researcher$^{2}$% <-this % stops a space
% \thanks{*This work was not supported by any organization}% <-this % stops a space
% \thanks{$^{1}$Albert Author is with Faculty of Electrical Engineering, Mathematics and Computer Science,
%         University of Twente, 7500 AE Enschede, The Netherlands
%         {\tt\small albert.author@papercept.net}}%
% \thanks{$^{2}$Bernard D. Researcheris with the Department of Electrical Engineering, Wright State University,
%         Dayton, OH 45435, USA
%         {\tt\small b.d.researcher@ieee.org}}%
% }

\urlstyle{same}
\begin{document}



\maketitle
\thispagestyle{empty}
\pagestyle{empty}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

This paper proposes the Real-Time Fast Marching Tree (RT-FMT), a real-time planning algorithm that features local and global path generation, multiple-query planning, and dynamic obstacle avoidance. During the search, RT-FMT quickly looks for the global solution and, in the meantime, generates local paths that can be used by the robot to start execution faster. In addition, our algorithm constantly rewires the tree to keep branches from forming inside the dynamic obstacles and to maintain the tree root near the robot, which allows the tree to be reused multiple times for different goals. Our algorithm is based on the planners Fast Marching Tree (FMT*) and Real-time Rapidly-Exploring Random Tree (RT-RRT*). We show via simulations that RT-FMT outperforms RT-RRT* in both execution cost and arrival time, in most cases. Moreover, we also demonstrate via simulation that it is worthwhile taking the local path before the global path is available in order to reduce arrival time, even though there is a small possibility of taking an inferior path. 

\end{abstract}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\input{1_introduction}

\input{2_related}

%\section{Problem Formulation and Notations}

\input{4_RTFMT_algorithm}

\input{5_methodology}

\input{6_results}

\input{7_conclusions}






%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\addtolength{\textheight}{-15cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Generated by IEEEtran.bst, version: 1.14 (2015/08/26)
\begin{thebibliography}{10}
\providecommand{\url}[1]{#1}
\csname url@samestyle\endcsname
\providecommand{\newblock}{\relax}
\providecommand{\bibinfo}[2]{#2}
\providecommand{\BIBentrySTDinterwordspacing}{\spaceskip=0pt\relax}
\providecommand{\BIBentryALTinterwordstretchfactor}{4}
\providecommand{\BIBentryALTinterwordspacing}{\spaceskip=\fontdimen2\font plus
\BIBentryALTinterwordstretchfactor\fontdimen3\font minus
  \fontdimen4\font\relax}
\providecommand{\BIBforeignlanguage}[2]{{%
\expandafter\ifx\csname l@#1\endcsname\relax
\typeout{** WARNING: IEEEtran.bst: No hyphenation pattern has been}%
\typeout{** loaded for the language `#1'. Using the pattern for}%
\typeout{** the default language instead.}%
\else
\language=\csname l@#1\endcsname
\fi
#2}}
\providecommand{\BIBdecl}{\relax}
\BIBdecl

\bibitem{tian2021trajectory}
F.~Tian, R.~Zhou, Z.~Li, L.~Li, Y.~Gao, D.~Cao, and L.~Chen, ``Trajectory
  planning for autonomous mining trucks considering terrain constraints,''
  \emph{IEEE Transactions on Intelligent Vehicles}, vol.~6, no.~4, pp.
  772--786, 2021.

\bibitem{hayat2020multi}
S.~Hayat, E.~Yanmaz, C.~Bettstetter, and T.~X. Brown, ``Multi-objective drone
  path planning for search and rescue with quality-of-service requirements,''
  \emph{Autonomous Robots}, vol.~44, no.~7, pp. 1183--1198, 2020.

\bibitem{karaman2011anytime}
S.~Karaman, M.~R. Walter, A.~Perez, E.~Frazzoli, and S.~Teller, ``Anytime
  motion planning using the rrt,'' in \emph{Proceedings of the 2011 IEEE
  International Conference on Robotics and Automation (ICRA)}.\hskip 1em plus
  0.5em minus 0.4em\relax IEEE, 2011, pp. 1478--1483.

\bibitem{xu2020informed}
J.~Xu, K.~Song, D.~Zhang, H.~Dong, Y.~Yan, and Q.~Meng, ``Informed anytime fast
  marching tree for asymptotically optimal motion planning,'' \emph{IEEE
  Transactions on Industrial Electronics}, vol.~68, no.~6, pp. 5068--5077,
  2020.

\bibitem{naderi2015rt}
K.~Naderi, J.~Rajam{\"a}ki, and P.~H{\"a}m{\"a}l{\"a}inen, ``{RT-RRT*:} {a}
  real-time path planning algorithm based on {RRT},'' in \emph{Proceedings of
  the 8th ACM SIGGRAPH Conference on Motion in Games}, 2015, pp. 113--118.

\bibitem{cserna2016anytime}
B.~Cserna, M.~Bogochow, S.~Chambers, M.~Tremblay, S.~Katt, and W.~Ruml,
  ``Anytime versus real-time heuristic search for on-line planning,'' in
  \emph{Proceedings of the International Symposium on Combinatorial Search},
  vol.~7, no.~1, 2016.

\bibitem{karaman2011sampling}
S.~Karaman and E.~Frazzoli, ``Sampling-based algorithms for optimal motion
  planning,'' \emph{The International Journal of Robotics Research}, vol.~30,
  no.~7, pp. 846--894, 2011.

\bibitem{fmt}
L.~Janson, E.~Schmerling, A.~Clark, and M.~Pavone, ``Fast marching tree: A fast
  marching sampling-based method for optimal motion planning in many
  dimensions,'' \emph{The International Journal of Robotics Research}, vol.~34,
  no.~7, pp. 883--921, 2015.

\bibitem{grothe2022st}
F.~Grothe, V.~N. Hartmann, A.~Orthey, and M.~Toussaint, ``St-rrt*:
  Asymptotically-optimal bidirectional motion planning through space-time,'' in
  \emph{Proceedings of the 2022 International Conference on Robotics and
  Automation (ICRA)}, 2022, pp. 3314--3320.

\bibitem{luders2010bounds}
B.~D. Luders, S.~Karaman, E.~Frazzoli, and J.~P. How, ``Bounds on tracking
  error using closed-loop rapidly-exploring random trees,'' in
  \emph{Proceedings of the 2010 American Control Conference}.\hskip 1em plus
  0.5em minus 0.4em\relax IEEE, 2010, pp. 5406--5412.

\bibitem{chandler2017online}
B.~Chandler and M.~A. Goodrich, ``Online rrt and online fmt: Rapid replanning
  with dynamic cost,'' in \emph{Proceedings of the 2017 IEEE/RSJ International
  Conference on Intelligent Robots and Systems (IROS)}, 2017, pp. 6313--6318.

\bibitem{tong2019rrt}
B.~Tong, Q.~Liu, and C.~Dai, ``A rrt* fn based path replanning algorithm,'' in
  \emph{Proceedings of the 2019 IEEE 4th Advanced Information Technology,
  Electronic and Automation Control Conference (IAEAC)}, vol.~1.\hskip 1em plus
  0.5em minus 0.4em\relax IEEE, 2019, pp. 1435--1445.

\bibitem{sucan2012the-open-motion-planning-library}
I.~A. Sucan, M.~Moll, and L.~E. Kavraki, ``The {O}pen {M}otion {P}lanning
  {L}ibrary,'' \emph{{IEEE} Robotics \& Automation Magazine}, vol.~19, no.~4,
  pp. 72--82, December 2012.

\end{thebibliography}

\end{document}
