\documentclass[letterpaper, 11pt]{article}
\usepackage{amsmath,amsthm,graphicx,url}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs} % For formal tables
% \usepackage[ruled,noline]{algorithm2e} % For algorithms
%  \renewcommand{\algorithmcfname}{ALGORITHM}
%  \SetKwBlock{Initialize}{Initialize:}{} \SetKwFor{From}{from}{do}{}
% \SetKwFor{ForAllInParallel}{for all}{in parallel do}{}
% \SetAlFnt{\small}
% \SetAlCapFnt{\small}
% \SetAlCapNameFnt{\small}
% \SetAlCapHSkip{0pt}
%\IncMargin{-\parindent}
\usepackage{authblk}
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
\usepackage[numbers]{natbib}
% \newtheorem{definition}[theorem]{Definition}%[section]

% The packages, macros and environments are in the settings folder
\input{settings/packages.tex}
\input{settings/macros.tex}



\title{Multi-Platform Autobidding with and without Predictions}
\author[a]{Gagan Aggarwal}
\author[b]{Anupam Gupta}
\author[c]{Xizhi Tan \thanks{This work was done while the author was visiting Google as a student researcher.}}
\author[a]{Mingfei Zhao}

\affil[a]{Google Research: \texttt{{\{gagana, mingfei\}@google.com}}}
\affil[b]{New York University: \texttt{anupam.g@nyu.edu}}
\affil[c]{Drexel University: \texttt{xizhi@drexel.edu}}

\date{}

\begin{document}
\begin{titlepage}

\maketitle


\begin{abstract}
We study the problem of finding the optimal bidding strategy for an advertiser in a multi-platform auction setting. The competition on a platform is captured by a value and a cost function, mapping bidding strategies to value and cost respectively. We assume a diminishing returns property, whereby the marginal cost is increasing in value. The advertiser uses an autobidder that selects a bidding strategy for each platform, aiming to maximize total value subject to budget and return-on-spend constraint. The advertiser
has no prior information and learns about the value and cost functions by querying a platform with a specific bidding strategy. Our goal is to design algorithms that find the optimal bidding strategy with a small number of queries.

We first present an algorithm that requires \(O(m \log (mn) \log n)\) queries, where $m$ is the number of platforms and $n$ is 
the number of possible bidding strategies in each platform. 
Moreover, we adopt the learning-augmented framework and propose an algorithm that utilizes a (possibly erroneous) prediction of the optimal bidding strategy. We provide a $O(m \log (m\eta) \log \eta)$ query-complexity bound on our algorithm as a function of the prediction error $\eta$. This guarantee gracefully degrades to \(O(m \log (mn) \log n)\). This achieves a ``best-of-both-worlds'' scenario: \(O(m)\) queries when given a correct prediction, and \(O(m \log (mn) \log n)\) even for an arbitrary incorrect prediction.

% Lastly, we employ the classic centroid method to provide an algorithm that finds a (randomized) bidding strategy that is \(\epsilon\)-close to the optimal strategy using \(O(m^2 \log \frac{Gn}{\epsilon})\) queries, where \(G\) is an upper bound on the derivative of the value functions. \mznote{Remove this paragraph.}
\end{abstract}
\thispagestyle{empty} 
\end{titlepage}

\input{sections/10intro}

\input{sections/20prelims}

\input{sections/30optsolution}

\input{sections/40medianofmedian}

\input{sections/45lowerbounds}
\input{sections/50-withpredictions}
% \input{sections/42centroidmethod}

\input{sections/60-conclusion}


\bibliographystyle{abbrvnat}
\bibliography{biblio}



\end{document}