\section{Related Works}
\vspace{-0.5em}

\textbf{BO for black-box games.} 
In the BO community, the game-theoretic approach has been researched as the \emph{application} to multi-objective BO (MOBO; ____). Direct consideration of the multi-agentic scenario is limited, which typically assumes the specific aggregation rule (Nash equilibrium; ____, Kalai-Smorodinsky solution; ____) or Chebyshev scalarisation function ____, which requires discrete domain to ensure the existence of solution. Ours is the \emph{first-of-its-kind} principled work that addresses the social-influence issues on continuous domain.
%\textbf{Fair BO.} 
%____ worked on fair budget allocation in batch BO settings, using welfare functions as a fairness criterion. While they treat fairness as a constraint in batch BO, ours considers aggregated utility as the main objective in sequential BO.

\textbf{Preferential BO.} 
Preferential BO is a single-agentic preference maximisation algorithms ____, extended to diverse scenarios; choice data ____, top-$k$ ranking ____, preference over objectives on MOBO ____, human-AI collaboration ____. Our work is the first to study the multi-agentic social influence, and is orthogonal to these works.

\textbf{Other BO.} 
Multitask BO ____ addresses scenarios where cheap but lower-fidelity information is available. However, ____ showed that this approach can converge significantly more slowly than standard BO if the low-fidelity information is unreliable, as is the case in our setting. While cost-aware BO ____ accounts for location-dependent cost functions, our public votes do not exhibit such dependency.
%%%%%
\vspace{-0.5em}