% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encodings may result in incorrect characters.
%

\usepackage{graphicx}

\usepackage{booktabs}

\usepackage{multirow}  % added
\usepackage{tabularx}
\usepackage[hidelinks]{hyperref}
%\hypersetup{colorlinks=False}
%\urlstyle{same}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
%\title{Detecting student intent in LLM-powered tutoring applications}
\title{Detecting Student Intent for Chat-Based Intelligent Tutoring Systems}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Ella Cutler\inst{1}\orcidID{0009-0001-2661-142X} \and
Zachary Levonian\inst{2}\orcidID{0000-0002-8932-1489} \and
S. Thomas Christie\inst{2}\orcidID{0009-0006-2900-4939}}
%\author{Author 1\inst{1}\orcidID{0000-0000-0000-0000} \and
%Author 2\inst{2}\orcidID{0000-0000-0000-0000} \and
%Author 3\inst{2}\orcidID{0000-0000-0000-0000}}
%Owen Henkel\inst{2}\orcidID{0009-0001-8850-067X}
%
\authorrunning{E. Cutler et al.}
%\authorrunning{Anonymous et al.}

% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Math Academy\\
\and
Digital Harbor Foundation\\
}
%\institute{Institution 1\\
%\and
%Institution 2\\
%}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Chat interfaces for intelligent tutoring systems (ITSs) enable interactivity and flexibility.
However, when students interact with chat interfaces, they expect dialogue-driven navigation from the system and can express frustration and disinterest if this is not provided.
Intent detection systems help students navigate within an ITS, but detecting students' intent during open-ended dialogue is challenging.
We designed an intent detection system in a chatbot ITS, classifying a student's intent between continuing the current lesson or switching to a new lesson.
We explore the utility of four machine learning approaches for this task---including both conventional classification approaches and fine-tuned large language models---finding that using an intent classifier introduces trade-offs around implementation cost, accuracy, and prediction time.
We argue that implementing intent detection in chat interfaces can reduce frustration and support student learning.


\keywords{classification \and intelligent tutoring systems \and large language models \and intent detection.}
\end{abstract}
%
%
%

\input{body}

\begin{credits}
\subsubsection{\ackname}
We would like to thank Owen Henkel, Millie-Ellen Postle, Maria Dyshel, Greg Thompson, Bill Roberts, and the staff of Rising Academies.
This work was supported by the Learning Engineering Virtual Institute (LEVI) and by the Digital Harbor Foundation.
%Omitted for anonymous review.

\subsubsection{\discintname}
The authors have no competing interests to declare that are relevant to the content of this article. No Generative AI technologies were used in the drafting of this article.
\end{credits}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{levi_enghub}

\end{document}
