\section{Geometric method for dataset generation}
\label{sec:Overview}
In this section, we introduce a rapid geometric approach to generate datasets with labels for inaccessible and occlusion regions.

\input{figures/fig-collision-detection}

\subsection{Voronoi-based accessibility analysis}
\label{sec:acc}
We use the subtractive collision detection method from \cite{zhong2023vasco} to gather accessibility training data, as it's efficient. 
We introduced a slight modification to their method by incorporating a detection box for pre-detection, enabling faster calculation.
Details of the method are outlined below.

\paragraph{Voronoi-based sampling.} The inset figure demonstrates the use of \input{figures/fig-voronoi} Lloyd's Voronoi relaxation for uniform sampling on the surface of $M$, with each Voronoi cell represented by its site $s_i\in S$, where $S = \{ s_1, s_2, \dots, s_n \}$. The surface of $M$ can be simplified as $M = \cup_{i=1}^n s_i$.

\paragraph{Ball-end cutter} 
In finishing machining of CNC, a ball-end cutter is commonly employed to finish the surface, modeled as a hemisphere combined with two differently sized cylinders.
\autoref{fig:collision-detection}(a) illustrates the cross-section of the cutter, which can be characterized using four parameters: two radii ($CR$ and $FR$) and two heights ($CH$ and $FH$). To ensure successful manufacturing, it is essential to guarantee the cutter $\mathcal{C}$ does not collide with $M$ in any direction $d_k$:
\begin{equation} 
\forall s_i \in S, \forall d_k \in D, s_i \cap \mathcal{C}(d_k) = \varnothing
\end{equation}

\paragraph{Inaccessible points.} 
For each $s_i \in S$, collision detection is performed with other sites, $\forall s_j\in S$ $(j \neq i)$, using the method from \cite{zhong2023vasco}. We first uniformly sample cutter directions $D = \{d_1, d_2, \dots, d_m\}$ using the Fibonacci Sphere sampling method~\cite{vorobiev2002fibonacci} on the upper Gaussian hemisphere and then rotate $M$ contrarily along each direction. To accelerate computation, a cylinder with a radius of $FR + \sigma$ is added as a detecting box before collision detection, allowing only points within the cylinder to undergo finer detection, as shown in \autoref{fig:collision-detection}(b).
Next, as illustrated in \autoref{fig:collision-detection}(c), each $s_j$ is rapidly evaluated for collision by calculating its horizontal distance from the center of $\mathcal{C}$. If the Z-coordinate of $s_j$ exceeds $CR + CH + FH$, it is immediately classified as colliding with the infinitely large shaft space (called global collision). 
After the traversal, if $s_i$ collides with at least one $s_j$ in all cutter directions, it is classified as an inaccessible point:
\begin{equation} 
s_i \leftarrow l_I \iff \forall d_k \in D, |S \cap \mathcal{C}(d_k)| > 0
\end{equation}
Compared to the triangle-facet-based approach~\cite{dhaliwal2003algorithms}, which involves collision detection between the cutter’s cylindrical surface and the triangular mesh, the proposed method is based on discrete sampling points (sites of Voronoi cells), significantly improving computational efficiency.
Even so, this method still has a worst-case complexity of $O(mn^2)$. Besides, it is crucial to ensure that the shortest edge length of the smallest Voronoi cell is greater than $2*CR$ to prevent the cutter from passing through the cell without detection, as illustrated in \autoref{fig:collision-detection}(d).

\input{figures/fig-Occlusion-analyze}

\subsection{Occlusion analysis}
To further assist designers in modifying the "culprit" causing inaccessible points, we compute the "occlusion factor" $\beta_i$ for each $s_i \in S$ to quantify the severity of its occlusion for the inaccessible points:
\begin{equation} 
\beta_i = \sum_{s_j \in S}\sum_{d_k \in D} \left(
\begin{cases} 
1  \text{, if } s_j \text{ is inaccessible and } s_i \text{ occludes } \text{$s_j$ in} \; d_k \\
0  \text{, otherwise}
\end{cases} \right)
\end{equation}
The top 10$\%$ of $s_i$ with the highest $\beta_i$ values are defined as "occlusion points" ($s_i \leftarrow l_O$). As shown in \autoref{fig:Occlusion-analyze}, the points in the upper region of the 2D shape are marked as "occlusion points."



