\documentclass[12pt]{article}
\usepackage{amsmath,dsfont}
\usepackage{graphicx}
\usepackage{enumerate}

\usepackage{natbib} 

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%


%---------------------------------------------------
%
% Personalization: Self-defined macros
%
\input{macros/packages.tex}
\input{macros/self-defined.tex}
\input{macros/box-theorems.tex}

\RequirePackage{algorithm}
\RequirePackage{algorithmic}

%
%
%---------------------------------------------------


\begin{document}
\pagenumbering{arabic}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%---------------------------------------------------
%
% Titile page
%
% https://www.math.uh.edu/~torok/math_6298/latex/MANUALS/NASA_Hypertext-Help-with-LaTeX/latex/ltx-407.html
%

\def\TITLE{Transition Transfer $Q$-Learning for Composite Markov Decision Processes}


\if0\blind
{
  \title{\bf \TITLE}
  \author{
  	Jinhang Chai$^\flat$ \hspace{5ex}
    Elynn Chen$^\sharp$ \hspace{5ex}
    Lin Yang$^\dag$ \thanks{Correspondence to {\em Elynn Chen} \textless{}\url{elynn.chen@stern.nyu.edu}\textgreater{} and {\em Lin Yang} \textless{}\url{linyang@ee.ucla.edu}\textgreater{} } \hspace{2ex} 
    \\ \normalsize
    $^{\flat\sharp}$Princeton University \hspace{3ex}
    $^{\flat\sharp}$New York University \hspace{3ex} \\ \normalsize
    $^{\dag }$ University of California, Los Angeles
    }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Title}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
\spacingset{1.08}
To bridge the gap between empirical success and theoretical understanding in transfer reinforcement learning (RL), we study a principled approach with provable performance guarantees.
We introduce a novel composite MDP framework where high-dimensional transition dynamics are modeled as the sum of a low-rank component representing shared structure and a sparse component capturing task-specific variations. This relaxes the common assumption of purely low-rank transition models, allowing for more realistic scenarios where tasks share core dynamics but maintain individual variations.
We introduce UCB-TQL (Upper Confidence Bound Transfer Q-Learning), designed for transfer RL scenarios where multiple tasks share core linear MDP dynamics but diverge along sparse dimensions. When applying UCB-TQL to a target task after training on a source task with sufficient trajectories, we achieve a regret bound of $\tilde{\mathcal{O}}(\sqrt{eH^5N})$ that scales independently of the ambient dimension. Here, $N$ represents the number of trajectories in the target task, while $e$ quantifies the sparse differences between tasks. This result demonstrates substantial improvement over single task RL by effectively leveraging their structural similarities. Our theoretical analysis provides rigorous guarantees for how UCB-TQL simultaneously exploits shared dynamics while adapting to task-specific variations.
\end{abstract}

\noindent%
{\it Keywords:}  Transfer Learning; Reinforcement Learning; Online $Q$-Learning; Upper Confidence Bound Algorithms; Low-Rank plus Sparse Structure. 
\vfill

%%---------------------------------------------------
%%
%% Main Text
%5

\newpage
\spacingset{1.9} % DON'T change the spacing!

\addtolength{\textheight}{.1in}%
\input{core.tex}

%%---------------------------------------------------
%%
%% Bibliography
%%
\spacingset{1.18}
\bibliographystyle{agsm}
\bibliography{main}

%%---------------------------------------------------
%%
%% Appendix in a seperate file
%%
\newpage
\begin{appendices}

\begin{center}
\input{suppl.tex}
\end{center}

%\section{Proofs} \label{sec-proof}
%\listoffigures
%\listoftables

\end{appendices}


\end{document}
