\section{Diary Study Questions}
\label{ref:diary}

\textbf{PID}: What is your participant ID?

\subsubsection*{Branch}
Which of the following best describes your most recent interaction with social media?

\noindent\textbf{Options:}
\begin{itemize}
    \item I shared something on social media.
    \item I thought about sharing something on social media but didn't/couldn't.
    \item I did not think about sharing anything on social media and did not share anything.
\end{itemize}

\subsubsection*{Shared}

\begin{enumerate}
    \item What did you share on social media today? Please describe it in as much detail as possible, but feel free not to share any parts of it that you don’t feel comfortable sharing with us. (If you posted multiple posts, you can choose to share about all of them in one survey or choose one that is most different from what you shared with us so far.)

    \item Who did you share the content with? What are they called on the platform? (e.g., Followers, Friends, Close Friends) How would you describe them? (e.g., my best friends, people that share my interest in knitting) Please be as detailed as possible.

    \item Where (e.g., Story, regular post, Snap, 1:1 messaging), and on which platform did you share the content? If Instagram, please specify if it is a Rinsta (real Instagram account) or a Finsta (fake Instagram account just for close friends).

    \item How comfortable did you feel sharing the content on the platform?
    \begin{itemize}
        \item 1: very uncomfortable
        \item 2: somewhat uncomfortable
        \item 3: neither comfortable nor uncomfortable
        \item 4: somewhat comfortable
        \item 5: very comfortable
    \end{itemize}

    \item How did you choose to post on that specific platform? Are there any features that helped you feel comfortable sharing in particular? Why do those features help? Please share as much detail as possible.

    \item What risks or concerns do you worry about, if any, about sharing this content in this way?

    \item What may be some benefits, if any, of sharing the content on social media?

    \item Is there anything else you want to share with us?
\end{enumerate}

\subsubsection*{Not Shared}

\begin{enumerate}
    \item What did you want to share on social media today that you did not share? Please describe it in as much detail as possible, but feel free not to share any parts of it that you don’t feel comfortable sharing with us.

    \item Who did you intend/want to share with? What are they called on the platform? (e.g., Followers, Friends, Close Friends) How would you describe them? (e.g., my best friends, people that share my interest in knitting) Please be as detailed as possible.

    \item What might be the benefits of sharing those with your friends? Please be specific.

    \item What stopped you from sharing? Please be specific.

    \item Do you expect any risks to sharing such content online? Please be specific.

    \item What are some features that could have possibly helped you share the content you could/did not share more (comfortably)? How would they help? Please explain in as much detail as possible.

    \item Is there anything else you want to share with us?
\end{enumerate}

\subsubsection*{No Thought}

\begin{enumerate}
    \item Is there anything that you think would have been nice to share with your close friends or casual friends/acquaintances on social media? If you can't think of something, what's something that you feel comfortable sharing in offline settings but not online on social media?

    \item Who might you want to share the content with? What are they called on the platform? (e.g., Followers, Friends, Close Friends) How would you describe them? (e.g., my best friends, people that share my interest in knitting) Please be as detailed as possible.

    \item What might be the benefits of sharing those with your friends? Please be specific.

    \item What stops you from sharing those? Please be specific.

    \item Do you expect any risks to sharing such content online? Please be specific.

    \item What are some ways that could possibly help you share the content more (comfortably)? How would they help? Please explain in as much detail as possible.

    \item Is there anything else you want to share with us?
\end{enumerate}