\pdfoutput=1

\documentclass[twoside,11pt]{article}
\usepackage{jmlr2e}



\usepackage{xcolor}
\usepackage[utf8]{inputenc} %
\usepackage[T1]{fontenc}    %
\usepackage{hyperref}       %
\usepackage{url}            %
\usepackage{booktabs}       %
\usepackage{amsfonts}       %

\usepackage{booktabs} %
\usepackage{longtable}

\usepackage{hyperref}

\newcommand{\theHalgorithm}{\arabic{algorithm}}



\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%

\usepackage{amsmath,amssymb}%
\usepackage{graphicx}
\usepackage{mathrsfs}%

\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%

\usepackage{enumitem}

 \usepackage[english]{babel}
\usepackage{mathtools}
\usepackage{thmtools}
\usepackage{thm-restate}
\usepackage{algorithm}[l]
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{chngpage}
\usepackage{xfrac}

\usepackage{mathrsfs}
\usepackage{xspace}
\usepackage{bm}
\usepackage{upgreek}
\usepackage{bbm}


\usepackage{subfig}

\include{definitions}


\looseness=-1





\ShortHeadings{VI Approach to Independent Learning in Static MFGs}{VI Approach to Independent Learning in Static MFGs}
\title{
A Variational Inequality Approach to Independent Learning in Static Mean-Field Games
}


\author{\name Batuhan Yardim \email alibatuhan.yardim@inf.ethz.ch \\
     \addr Department of Computer Science\\
          ETH Z\"urich\\
          Z\"urich, Switzerland\\
     \AND
     \name Semih Cayci   \email cayci@mathc.rwth-aachen.de \\
     \addr Department of Mathematics\\
          RWTH Aachen University \\
          Aachen, Germany\\
     \AND
     \name Niao He   \email niao.he@inf.ethz.ch \\
     \addr Department of Computer Science\\
          ETH Z\"urich\\
          Z\"urich, Switzerland
}



\begin{document}

\maketitle

\begin{abstract}%
Competitive games involving thousands or even millions of players are prevalent in real-world contexts, such as transportation, communications, and computer networks. However, learning in these large-scale multi-agent environments presents a grand challenge, often referred to as the "curse of many agents". In this paper, we formalize and analyze the Static Mean-Field Game (SMFG) under both full and bandit feedback, offering a generic framework for modeling large population interactions while enabling independent learning. 


We first establish close connections between SMFG and variational inequality (VI),  showing that SMFG can be framed as a VI problem in the infinite agent limit. Building on the VI perspective, we propose independent learning and exploration algorithms that efficiently converge to approximate Nash equilibria, when dealing with a finite number of agents.   Theoretically, we provide explicit finite sample complexity guarantees for independent learning across various feedback models in repeated play scenarios, assuming (strongly-)monotone payoffs. Numerically, we validate our results through both simulations and real-world applications in city traffic and network access management. 

\end{abstract}

\begin{keywords}
variational inequality, independent learning, mean-field games, multi-agent systems, bandit feedback
\end{keywords}


\input{intro}
\input{related_work}
\input{formulation}

\input{approximation}
\input{fullfeedback}
\input{banditfeedback}

\input{experiments}

\input{discussion}

\section*{Acknowledgements}

We thank the editor and reviewers for their detailed comments and insightful feedback. This project was supported by Swiss National Science Foundation (SNSF) under the framework of NCCR Automation and SNSF Starting Grant.
S. Cayci's research was funded under the Excellence Strategy of the Federal Government and the L{\"a}nder.

\bibliography{sn-bibliography}

\appendix

\input{appendix/notationtable}
\input{appendix/comparison_setting}
\input{appendix/formalization_alg}
\input{appendix/basic_inequalities}
\input{appendix/technical_results}
\input{appendix/experiment_details}


\end{document}
