\documentclass[conference]{IEEEtran}
%% IEEE CNS addition:
\makeatletter
\def\ps@headings{%
\def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
\def\@evenhead{\scriptsize\thepage \hfil \leftmark\mbox{}}%
\def\@oddfoot{}%
\def\@evenfoot{}}
\makeatother
\pagestyle{empty}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{cite}
\usepackage{adjustbox}
\usepackage{xcolor}
\usepackage{color,soul}
\usepackage{multirow}
\usepackage{enumerate}
\usepackage[
  separate-uncertainty = true,
  multi-part-units = repeat
]{siunitx}
\usepackage{algorithm}
\usepackage{graphicx}
\usepackage{textcomp}
%\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{authblk}
\usepackage{caption}
\usepackage{amsfonts}
\usepackage{algpseudocode}
\usepackage{url}
\usepackage[justification=centering]{caption}
\newcommand{\R}{\mathbb{R}}

\definecolor{colorEntityBack}{rgb}{0.01, 0.01, 0.4}
\definecolor{colorPolicyBackDarkDark}{rgb}{0.81, 0.81, 0.86}
\definecolor{ashgrey}{rgb}{0.7, 0.75, 0.71}
\definecolor{beaublue}{rgb}{0.74, 0.83, 0.9}
\definecolor{hotpink}{rgb}{1.0, 0.41, 0.71}
\definecolor{darkgreen}{rgb}{0.09, 0.45, 0.27}
\definecolor{cerulean}{rgb}{0.105, 0.672, 0.836}
\definecolor{candyapplered}{rgb}{1.0, 0.03, 0.0}
\definecolor{cadetgrey}{rgb}{0.57, 0.64, 0.69}
\definecolor{trolleygrey}{rgb}{0.5, 0.5, 0.5}

\newcommand{\og}[1]{\textcolor{red}{\textsf{Onat:~~#1}}}
\newcommand{\sean}[1]{\textcolor{blue}{\textsf{Sean:~~#1}}}
\newcommand{\changes}[1]{\textcolor{darkgreen}{#1}}
\begin{document}

\date{18 February 2024}

\title{CND-IDS: Continual Novelty Detection for Intrusion Detection Systems}

%\author{}
\author[1]{Sean Fuhrman}
\author[2]{Onat Gungor}
\author[2]{Tajana Rosing}

\affil[1]{Department of Electrical and Computer Engineering, University of California, San Diego}
\affil[2]{Department of Computer Science and Engineering, University of California, San Diego}
\affil[ ]{\textit{\{stfuhrma, ogungor, tajana\}@ucsd.edu}}
%\IEEEoverridecommandlockouts
%\IEEEpubid{\makebox[\columnwidth]{978-1-7281-6127-3/21/\$31.00~\copyright2022 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}

\maketitle
\pagestyle{plain}
\pagenumbering{gobble}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\Design}[0]{CND-IDS}

\begin{abstract}
%The Industrial Internet of Things (IIoT) security has become a significant challenge due to increased inter-connectivity and large scale networks. 
Intrusion detection systems (IDS) play a crucial role in IoT and network security by monitoring system data and alerting to suspicious activities. Machine learning (ML) has emerged as a promising solution for IDS, offering highly accurate intrusion detection. However, ML-IDS solutions often overlook two critical aspects needed to build reliable systems: continually changing data streams and a lack of attack labels. Streaming network traffic and associated cyber attacks are continually changing, which can degrade the performance of deployed  ML models. Labeling attack data, such as zero-day attacks, in real-world intrusion scenarios may not be feasible, making the use of ML solutions that do not rely on attack labels necessary. To address both these challenges, we propose \Design{}, a continual novelty detection IDS framework which consists of (i) a  learning-based feature extractor that continuously updates new feature representations of the system data, and (ii) a novelty detector that identifies new cyber attacks by leveraging principal component analysis (PCA) reconstruction. Our results on realistic intrusion datasets show that \Design{} achieves up to 6.1$\times$ F-score improvement, and up to 6.5$\times$ improved forward transfer over the SOTA unsupervised continual learning algorithm. Our code will be released upon acceptance. 
%\Design{} also outperforms all other state-of-the-art novelty detection algorithms consistently. Novelty detection (ND) algorithms offer a solution to detect attacks without labels.

\end{abstract}

%\begin{IEEEkeywords}
%Industrial internet of things (I-IoT), predictive maintenance, intelligent fault diagnosis, hyper-dimensional computing, adversarial attacks, resilient machine learning 
%\end{IEEEkeywords}

\section{Introduction}
\input{Intro.tex}

\section{Related Work}
\input{RelatedWork.tex}

\section{\Design{} Framework}
\label{framework}
\input{Framework.tex}

\section{Experimental Analysis}
\label{experimental}
\input{Experiment.tex}

\section{Conclusion}
\label{conclusion}
\input{Conclusion.tex}

\section*{Acknowledgments}
This work has been funded in part by NSF, with award numbers \#1826967, \#1911095, \#2003279, \#2052809, \#2100237, \#2112167, \#2112665, and in part by PRISM and CoCoSys, centers in JUMP 2.0, an SRC program sponsored by DARPA.

\newpage
\bibliographystyle{ieeetr}
\bibliography{bibfile}


\end{document}