\documentclass[final,5p,times,twocolumn]{elsarticle}

\input{arXiv_macros.tex}

\journal{Journal of \LaTeX\ Templates}
\bibliographystyle{model5}\biboptions{authoryear}

\begin{document}
\begin{frontmatter}
\title{A Stochastic Fundamental Lemma with Reduced Disturbance Data Requirements}

\author{Ruchuan Ou}\ead{ruchuan.ou@tuhh.de}
\author{Guanru Pan}\ead{guanru.pan@tuhh.de}  

\author{Timm Faulwasser\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{timm.faulwasser@ieee.org}

\address{Institute of Control Systems, Hamburg University of Technology, 21079 Hamburg, Germany}                                        

\begin{keyword}                            
Fundamental lemma, stochastic systems, polynomial chaos, reduced disturbance data, non-Gaussian uncertainties, data-driven control
\end{keyword}                           


\begin{abstract}
Recently, the fundamental lemma by Willems et. al has been extended towards stochastic LTI systems subject to process disturbances. Using this lemma requires previously recorded data of inputs, outputs, and disturbances. In this paper, we exploit causality concepts of stochastic control to propose a variant of the stochastic fundamental lemma that does not require past disturbance data in the Hankel matrices. Our developments rely on polynomial chaos expansions and on the knowledge of the disturbance distribution. Similar to our previous results, the proposed variant of the fundamental lemma allows to predict future input-output trajectories of stochastic LTI systems. We draw upon a numerical example to illustrate the proposed variant in data-driven control context.
\end{abstract}
\end{frontmatter}

\input{sources/arXiv_Sec1_Introduction.tex}
\input{sources/arXiv_Sec2_Preliminaries.tex}
\input{sources/arXiv_Sec3_StochLemma.tex}
\input{sources/arXiv_Sec4_Simluation.tex}
\input{sources/arXiv_Sec5_Conclusion.tex}

\section*{Acknowledgements}
The authors would like to thank Jonas Schie{\ss}l, Michael Heinrich Baumann, and Lars Gr{\"u}ne at Mathematical Institute, University of Bayreuth, Germany for their valuable feedback. The authors also acknowledge funding by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) - project number 499435839.
                         
\bibliography{arXivRef}
\end{document}