\section{Limitations and Future Work}\label{sec:limitations}

\subsection{Limitations}
The distributed network we relied on provides a challenge for financial payment. We believe that meaningful compensation is an important standard in data collection that relies on community-based knowledge is important. However, we acknowledge a stalemate, in that to the best of our knowledge, normative standards around payment mechanisms and logistics for our novel decentralised approach have yet to be established. In preparing this work, we have---and continue to, extensively---engaged with other practitioners and community members to understand best practices. At this moment, we rely on a process of transparency and informed consent rooted in direct attempts to empower participants such that the practice is less extractive. We maintain a constant communication channel with all Contributors (who are rightfully acknowledged in the appendix). 

\subsection{Future work}

Finally, we conclude with suggestions for how future work can engage in infrastructure to support data work. 

The work presented here is not meant to be one size fits all, but rather as a starting point from which to change the way we approach ML datasets and position their importance in terms of the \textit{human processes} behind them. Data collection that relies on community input results in high-quality data, but the process is slow, iterative, and very hard to scale by any other means than what we propose here. We hope that other researchers use this as a foundation for future work.

As mentioned above, simply creating an access point for data submission does not mean that data will easily flow in. Complete infrastructure in the future should involve both accessible technical infrastructure as well as direct engagement with stakeholder communities.

 Data has value, and promises of ``increased representation'' should be critically assessed to make sure the promises made to Contributors are sincere. Improvements can come from increasing infrastructure for payments in a decentralised and distributed network that allows for \textit{meaningful} compensation for data labour, and we continue to advocate for Contributors to be the owners of their data, as well as maintaining~\cite{tonja2024inkubalm} transparency in the process.  
Additional efforts should support increased accessibility that supports submission in a Contributor's local language to increase accessibility.

Acknowledging the burden of data collection, even with the distributed workload we propose with the stakeholder distinction between Community Ambassador and Contributor, sustainable processes for all stakeholders need to be considered. 
 