Based on our study, we conclude that VFL algorithms are not yet ready for widespread deployment in real-world applications. We identify three critical gaps that must be addressed to enable successful implementation.

\paragraph{Fuzzy and Latent VFL.}  
Many potential VFL applications involve fuzzy or latent data scenarios; however, research in this area remains limited. The primary challenge is to design algorithms capable of handling noise in shared features and uncovering latent relationships within the data. Developing efficient and robust algorithms for these prevalent scenarios would significantly broaden the applicability of VFL.

\paragraph{VFL for Heterogeneous Data.}  
Most existing VFL algorithms are designed for homogeneous data, whereas real-world applications often involve highly heterogeneous data. The key challenge is to create algorithms that can effectively handle data with diverse structures and distributions. Conducting experiments on imbalanced datasets would provide strong evidence of VFL algorithms' effectiveness in practical applications.

\paragraph{Trustworthiness in VFL.}  
Ensuring trustworthiness remains a critical challenge for VFL systems. Our observations indicate that several privacy risks are still unaddressed. Strict privacy and security regulations often impede VFL deployment, as even a single unresolved risk can prevent adoption. Future research must focus on developing comprehensive security mechanisms to enhance the privacy and trustworthiness of VFL systems.  
