
\noindent
The main material of this thesis appeared in the following publications:

{\footnotesize
\setstretch{1.05}
\begin{itemize}
    \item \fullcite{fernandez2022sslwatermarking}
    \item \fullcite{fernandez2022active}
    \item \fullcite{fernandez2023stable}
    \item \fullcite{fernandez2023three}
    \item \fullcite{fernandez2024functional}
    \item \fullcite{san2024proactive}
    \item \fullcite{fernandez2024what}
    \item \fullcite{sander2024watermarking}
\end{itemize}}
\vspace*{1em}

\noindent
Some works were adapted to French venues for the national community: 

{\footnotesize
\setstretch{1.05}
\begin{itemize}
    \item \fullcite{fernandez2022tatouage}
    \item \fullcite{fernandez2023tatouage}
\end{itemize}}
\vspace*{1em}

\noindent
Contributions were also made to the following publications, not included in this thesis, but relevant for the field:

{\footnotesize
\setstretch{1.05}
\begin{itemize}
    \item \fullcite{sander2025watermark} 
    \begin{spacing}{1.0}
        \scriptsize Redefines watermarking as a segmentation task, introducing a method to embed and extract watermarks localized to specific areas of an image. \\
        \scriptsize Contributions: co-first author with Tom Sander.
    \end{spacing}
    \item \fullcite{san2024latent} 
    \begin{spacing}{1.0}
        \scriptsize Introduces a method to watermark audio generative models at the latent space level, through watermarking training data in a way that the watermark is preserved by the audio tokenizer. \\
        \scriptsize Contributions: advised on the watermarking part, and wrote the paper with Robin San Roman.
    \end{spacing}
    \item \fullcite{kinakh2024evaluation} 
    \begin{spacing}{1.0}
        \scriptsize Introduces a copy attack on the scheme presented in Chap.~\ref{chapter:ssl-watermarking}. \\
        \scriptsize Contributions: advised at the beginning of the project, and contributed to the writing.
    \end{spacing}
    \item \fullcite{labiad2024log} 
    \begin{spacing}{1.0}
        \scriptsize Introduces new black-box attacks against fake detection systems, and evaluates if they can be detected. \\
        \scriptsize Contributions: regularly advised Ismail on the watermarking and fake detection parts, and contributed to the writing.
    \end{spacing}
\end{itemize}
}


\noindent
Contributions were also made to the bigger projects at FAIR:

{\footnotesize
\setstretch{1.05}
\begin{itemize}
    \item \fullcite{oquab2023dinov2}
    \begin{spacing}{1.0}
        \scriptsize Contributions: first experiments with the Kozachenko-Leonenko regularizer, classification fine-tuning on ImageNet, zero-shot and fine-tuning evaluations on video benchmarks.
    \end{spacing}
    \item \fullcite{balestriero2023cookbook}
    \begin{spacing}{1.0}
        \scriptsize Contributions: wrote parts on vision transformers and video.
    \end{spacing}
    \item \fullcite{seamless2023} 
    \begin{spacing}{1.0}
        \scriptsize Contributions: wrote part on watermarking, jointly with Hady Elsahar and Robin San Roman.
    \end{spacing}
\end{itemize}
}
