\section{Related work}
\label{sec:related-work}

The preordering problem is formulated and studied from the perspective of computational complexity by ____ who establishes \textsc{NP}-hardness of the problem itself and many specializations.
The geometry of the preorder polytope is discussed briefly by ____.
The problem has received less attention so far than more constrained clustering and ordering problems:

Preordering is related to correlation clustering ____, more specifically, to maximum value weighted correlation clustering for complete graphs, which is also known as clique partitioning ____, where values $c_{ij}$ are associated with unordered pairs $\{i,j\}$, and the task is to find an equivalence relation $\sim$ on $V$ so as to maximize the sum of the values of those pairs $\{i,j\}$ for which $i \sim j$.
This problem is more specific than preordering in that equivalence relations are precisely the \emph{symmetric} preorders. 
The complexity and hardness of approximation of correlation clustering have been studied extensively ____.
In particular, it is known that clique partitioning cannot be approximated to within $n^{1-\epsilon}$ unless \textsc{P}$=$\textsc{NP} for $n$ the number of elements and any $\epsilon > 0$ ____.
Much work has been devoted also to the study of the clique partitioning polytope
____.
Correlation clustering has many applications, including community detection in social networks ____ and image analysis ____.
Variants that emphasize local errors are studied by ____.

Preordering is related also to partial ordering ____ where the task is to find a partial order $\leq$ on $V$ so as to maximize the sum of the values of those pairs $ij$ for which $i \leq j$.
Partial ordering is more specific than preordering in that partial orders are precisely the \emph{antisymmetric} preorders.
The polyhedral geometry of partial orders is studied by ____.
Also this problem has received less attention so far than even more constrained problems, like the linear ordering problem ____, the closely related rank aggregation problem ____ and the feedback arc set problem on tournament graphs ____.

In summary, preordering is a joint relaxation of correlation clustering, from which the symmetry constraint is dropped, and partial ordering, from which the antisymmetry constraint is dropped.
A joint relaxation of correlation clustering and \emph{linear} ordering is the bucket ordering problem, sometimes called the weak ordering problem, that asks for both a clustering and a \emph{linear} order on the clusters ____.
This problem has been applied to the tasks of dating archeological sites ____, finding consensus among voters ____ and constructing user recommendations ____.