 \documentclass[snmathphys,Numbered]{sn-jnl}% Math and Physical Sciences Reference Style
% \documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
% \documentclass[sn-vancouver,Numbered]{sn-jnl}% Vancouver Reference Style
%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style 
% \documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
% \documentclass[default]{sn-jnl}% Default
% \documentclass[default,iicol]{sn-jnl}% Default with double column layout

%%%% Standard Packages
%%<additional latex packages if required can be included here>
\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{bbding}
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%
\usepackage[normalem]{ulem} 

\usepackage{boldline}
\usepackage{makecell}
\usepackage{colortbl}
\usepackage{epsfig}
\usepackage{dsfont}
\usepackage{siunitx}
\usepackage{soul}
\usepackage{cite}
\usepackage{cases}
\usepackage{url}
\usepackage{epstopdf}
\usepackage[T1]{fontenc}
\newcommand{\gr}{\rowcolor[gray]{.95}}

% \usepackage{algorithmic}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{threeparttable}
\usepackage{hyperref}
\usepackage{adjustbox}
\usepackage{pifont}
\newcommand{\xmark}{\ding{55}} % 定义 \xmark 为叉号

% \usepackage[colorlinks]{hyperref}

% \usepackage{cite}%
%%%%

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published 
%%%%  by Springer Nature. The guidance has been prepared in partnership with 
%%%%  production teams to conform to Springer Nature technical requirements. 
%%%%  Editorial and presentation requirements differ among journal portfolios and 
%%%%  research disciplines. You may find sections in this template are irrelevant 
%%%%  to your work and are empowered to omit any such section if allowed by the 
%%%%  journal you intend to submit to. The submission guidelines and policies 
%%%%  of the journal take precedence. A detailed User Manual is available in the 
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

%\jyear{2021}%
% \hypersetup{hidelinks,
% 	colorlinks=true,
% 	allcolors=black,
% 	pdfstartview=Fit,
% 	breaklinks=true}
 
%% as per the requirement new theorem styles can be included as shown below
\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}% 
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%

\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}%

\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads

\begin{document}

\title[Article Title]{H3DE-Net: Efficient and Accurate 3D Landmark Detection in Medical Imaging}

%\authorrunning{Z. Huang et al.}
%\titlerunning{PELE scores}

%%=============================================================%%
%% Prefix	-> \pfx{Dr}
%% GivenName	-> \fnm{Joergen W.}
%% Particle	-> \spfx{van der} -> surname prefix
%% FamilyName	-> \sur{Ploeg}
%% Suffix	-> \sfx{IV}
%% NatureName	-> \tanm{Poet Laureate} -> Title after name
%% Degrees	-> \dgr{MSc, PhD}
%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate} 
%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
%%=============================================================%%

\author[1,2]{\fnm{Zhen} \sur{Huang}}

\author[3,4]{\fnm{Ronghao} \sur{Xu}} 

\author[3,4]{\fnm{Xiaoqian} \sur{Zhou}}

\author[1,5]{\fnm{Yangbo} \sur{Wei}}

\author[6]{\fnm{Suhua} \sur{Wang}}

\author[7]{\fnm{Xiaoxin} \sur{Sun}}

\author[8]{\fnm{Han} \sur{Li}}

\author[9]{\fnm{Qingsong} \sur{Yao}}

\affil[1]{\orgdiv{School of Computer Science and Technology}, \orgname{University of Science and Technology of China (USTC)}, \orgaddress{\city{Hefei}, \postcode{230026}, \country{China}}}

\affil[2]{\orgdiv{School of Information Science and Technology}, \orgname{Eastern Institute of Technology (EIT)}, \orgaddress{\city{Ningbo}, \postcode{315200}, \country{China}}}

\affil[3]{\orgdiv{School of Biomedical Engineering, Division of Life Sciences and Medicine}, \orgname{USTC}, \orgaddress{\city{Suzhou}, \postcode{215123}, \country{China}}}

\affil[4]{\orgname{Suzhou Institute for Advanced Research},\orgname{USTC}, \orgaddress{\city{Suzhou}, \postcode{215123}, \country{China}}}

\affil[5]{\orgdiv{Shanghai Jiao Tong University}, \orgaddress{\city{Shanghai}, \postcode{200030}, \country{China}}}

\affil[6]{\orgname{Computer Department},\orgname{Changchun Humanities and Sciences College}, \orgaddress{\city{Changchun}, \postcode{130117}, \country{China}}}

\affil[7]{\orgname{School of Information Science and Technology},\orgname{Northeast Normal University}, \orgaddress{\city{Changchun}, \postcode{130117}, \country{China}}}

\affil[8]{\orgname{Computer Aided Medical Procedures (CAMP), School of Computation, Information and Technology},\orgname{Technische Universitaet Muenchen (TUM)}, \orgaddress{\country{Germany}}}

\affil[9]{\orgdiv{Stanford University}, \orgaddress{\city{California}, \postcode{94305}, \country{United States}}}



%%==================================%%
%% sample for unstructured abstract %%
%%==================================%%

% \abstract{The pelvis, the lower part of the trunk, supports and balances the trunk. Landmark detection from a pelvic X-ray (PXR) facilitates downstream analysis and computer-assisted diagnosis and treatment of pelvic diseases. Although PXRs have the advantages of low radiation and reduced cost compared to computed tomography (CT) images, their 2D pelvis-tissue superposition of 3D structures confuses clinical decision-making. In this paper, we propose a PELvis Extraction (PELE) module that utilizes 3D prior anatomical knowledge in CT to guide and well isolate the pelvis from PXRs, thereby eliminating the influence of soft tissue. We conduct an extensive evaluation based on two public datasets and one private dataset, totaling 850 PXRs. The experimental results show that the proposed PELE module significantly improves the accuracy of PXRs landmark detection and achieves state-of-the-art performances in several benchmark metrics, thus better serving downstream tasks.}

%%================================%%
%% Sample for structured abstract %%
%%================================%%

\abstract{3D landmark detection is a critical task in medical image analysis, and accurately detecting anatomical landmarks is essential for subsequent medical imaging tasks. However, mainstream deep learning methods in this field struggle to simultaneously capture fine-grained local features and model global spatial relationships, while maintaining a balance between accuracy and computational efficiency. Local feature extraction requires capturing fine-grained anatomical details, while global modeling requires understanding the spatial relationships within complex anatomical structures. The high-dimensional nature of 3D volume further exacerbates these challenges, as landmarks are sparsely distributed, leading to significant computational costs. Therefore, achieving efficient and precise 3D landmark detection remains a pressing challenge in medical image analysis.

In this work, We propose a \textbf{H}ybrid \textbf{3}D \textbf{DE}tection \textbf{Net}(H3DE-Net), a novel framework that combines CNNs for local feature extraction with a lightweight attention mechanism designed to efficiently capture global dependencies in 3D volumetric data. This mechanism employs a hierarchical routing strategy to reduce computational cost while maintaining global context modeling. To our knowledge, H3DE-Net is the first 3D landmark detection model that integrates such a lightweight attention mechanism with CNNs. Additionally, integrating multi-scale feature fusion further enhances detection accuracy and robustness.

Experimental results on a public CT dataset demonstrate that H3DE-Net achieves state-of-the-art(SOTA) performance, significantly improving accuracy and robustness, particularly in scenarios with missing landmarks or complex anatomical variations. We aready open-source our project, including code, data and model weights on \href{https://github.com/ECNUACRush/H3DE-Net}{https://github.com/ECNUACRush/H3DE-Net}.
}

\keywords{3D Landmark Detection,
Medical Image Analysis,
Transformer-CNN Hybrid Model.}
%%\pacs[JEL Classification]{D8, H51}

%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}

\maketitle



% \input{sections/abstract.tex}
\input{introduction.tex}
\input{method.tex}
\input{experiment.tex}
\input{discussion.tex}
\input{conclusion.tex}
\input{statement.tex}

% \section{Introduction}\label{sec1}

% The Introduction section, of referenced text \cite{bib1} expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.

% Springer Nature does not impose a strict layout as standard however authors are advised to check the individual requirements for the journal they are planning to submit to as there may be journal-level preferences. When preparing your text please also be aware that some stylistic choices are not supported in full text XML (publication version), including coloured font. These will not be replicated in the typeset article if it is accepted. 
%\newpage

% \bibliographystyle{snmathphys}
\bibliographystyle{splncs04}
\bibliography{sn_bibliography}% common bib file
%\input{appendix.tex}
\section*{Statements and Declarations}

\subsection*{Funding}
The authors declare that no funds, grants, or other support were received during the preparation of this manuscript.

\subsection*{Competing Interests}
The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

% \subsection*{Author Contributions}
% All authors contributed to the study conception and design. Material preparation, data collection, and analysis were performed by Zhen Huang and Yangbo Wei. Experiments were conducted by Zhen Huang, Xiaoqian Zhou, and He Xu. The first draft of the manuscript was written by Zhen Huang, Xiaoxin Sun, and Wenkai Yang. The paper was reviewed and guided by Suhua Wang and Han Li. All authors commented on previous versions of the manuscript. All authors read and approved the final manuscript.

\subsection*{Data Availability Statement}
Data and code are already open-source.

\subsection*{Ethical approval} All procedures performed in studies involving human participants were in accordance with the ethical standards of the institutional and national research committee and with the 1964 Helsinki declaration and its later amendments or comparable ethical standards. The IRB approval is obtained.

\end{document}
