\section{Conclusion and Future Works}
Our work advances the integration of ethical reasoning into AI, particularly in reinforcement learning frameworks, by leveraging ethically aligned LLM agents. While effective for low-stakes applications like personal or kitchen assistant robots, our approach is best viewed as complementary to human oversight, especially in high-stakes scenarios requiring rigorous testing. The framework excels in bootstrapping learning algorithms where human feedback is scarce, with the potential to replace LLM feedback with evaluations from multiple human evaluators, safeguarding against malicious manipulation.

Applications include smart home assistants, where AMULED promotes ethical and sustainable choices, and social media platforms, where it fosters constructive, inclusive interactions. However, categorizing ethical theories into clusters may oversimplify moral complexities, and operationalizing these theories involves abstracting philosophical principles. Our framework assumes that the selected ethical clusters and decision factors are representative and computationally translatable, though cultural and contextual variations pose challenges.

Future research should focus on hybrid ethical models, real-time adaptation, and interpretability to enhance trust and relevance. Addressing biases, exploring multi-agent scenarios, and establishing evaluation benchmarks are critical for equitable and robust outcomes. By integrating Belief Probability Assignment (BPA) into reinforcement learning, AMULED enables nuanced ethical decision-making, balancing competing moral frameworks in dynamic environments. Leveraging the Dempster-Shafer Theory for robust belief aggregation and the Jensen-Shannon Belief Divergence for precise information fusion, AMULED effectively synthesizes diverse ethical perspectives while quantifying uncertainty and resolving conflicts. These advanced techniques ensure that the system can adaptively weigh and harmonize competing moral values, such as utilitarianism and deontology, with a mathematically grounded approach. This paves the way for responsible, adaptable AI systems that are not only aligned with societal values but also capable of navigating the complexities of real-world ethical dilemmas with confidence and transparency.