\section{Related Work}
The development of robust Optical Character Recognition (OCR) systems has been extensively studied across document layout analysis ____, table detection ____, and document understanding ____. While English OCR benefits from rich datasets like PubLayNet ____, DocBank ____, M6Doc ____, and DocLayNet ____, Arabic lacks standardized benchmarks for diverse fonts and layouts. Recent efforts like MIDAD ____ curates extensive training data for Arabic OCR and handwriting recognition, while Peacock ____ introduces culturally-aware Arabic multimodal models. Existing resources such as CAMEL-Bench ____, LAraBench ____, MADAR ____, OSACT ____, and Tashkeela ____ focus on language modeling or specific tasks rather than full-page OCR evaluation. Handwriting datasets including HistoryAr ____, IFN/ENIT ____, KHATT ____, APTI ____, and Muharaf ____ emphasize word/line recognition over document structure analysis.

\input{tables/distribution}

\noindent Arabic table recognition faces challenges from merged cells and RTL formatting ____. While methods like GTE ____, GFTE ____, CascadeTabNet ____, TableNet ____, and TableFormer ____ advance Latin table detection, their effectiveness on Arabic documents remains unproven. Document conversion pipelines (CCS ____, Tesseract ____, Docling ____, Surya ____, Marker ____, MinerU ____, PaddleOCR ____) lack Arabic-specific optimizations for segmentation and diacritic handling ____. This highlights the critical need for comprehensive Arabic OCR benchmarks addressing text recognition, table detection, and layout parsing.