\begin{abstract}
\label{sec:abstract}
\vspace{-0.2cm}

We present \textbf{Step-Video-T2V}, a state-of-the-art text-to-video pre-trained model with 30B parameters and the ability to generate videos up to 204 frames in length. 
A deep compression Variational Autoencoder, \textbf{Video-VAE}, is designed for video generation tasks, achieving 16x16 spatial and 8x temporal compression ratios, while maintaining exceptional video reconstruction quality.
User prompts are encoded using two bilingual text encoders to handle both English and Chinese. A DiT with 3D full attention is trained using Flow Matching and is employed to denoise input noise into latent frames. A video-based DPO approach, \textbf{Video-DPO}, is applied to reduce artifacts and improve the visual quality of the generated videos.
We also detail our training strategies and share key observations and insights. Step-Video-T2V's performance is evaluated on a novel video generation benchmark, \textbf{Step-Video-T2V-Eval}, demonstrating its state-of-the-art text-to-video quality when compared with both open-source and commercial engines. Additionally, we discuss the limitations of current diffusion-based model paradigm and outline future directions for video foundation models. We make both Step-Video-T2V and Step-Video-T2V-Eval available at \hyperlink{https://github.com/stepfun-ai/Step-Video-T2V}{https://github.com/stepfun-ai/Step-Video-T2V}. The online version can be accessed from \hyperlink{https://yuewen.cn/videos}{https://yuewen.cn/videos} as well. Our goal is to accelerate the innovation of video foundation models and empower video content creators. 

\end{abstract}