\pdfoutput=1

\documentclass[11pt]{article}

\usepackage[preprint]{acl}

\usepackage{times}
\usepackage{latexsym}

\usepackage[T1]{fontenc}

\usepackage[utf8]{inputenc}

\usepackage{microtype}

\usepackage{inconsolata}

\usepackage{graphicx}


\usepackage{arydshln}
\usepackage{float}
\usepackage{pifont}%
\usepackage{wrapfig}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{soul}
\usepackage[most]{tcolorbox}
\usepackage{geometry}
\usepackage{svg}
\usepackage{amsmath}
\usepackage{soul}
\usepackage{makecell}
\usepackage{subfigure}
\usepackage[subfigure]{tocloft}
\usepackage{multirow}
\usepackage{bbold}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{placeins}
\usepackage{makecell}
\usepackage{svg}
\usepackage{authblk}
\usepackage{annotation}
\usepackage{hyperref}
\tcbuselibrary{breakable}

\newcommand{\judge}{\textbf{DRIFTJudge}}
\newcommand{\policy}{\textbf{DRIFTPolicy}}
\newcommand{\metric}{\textbf{FOCUS}}
\newcommand{\dataset}{\textbf{DRIFTEval}}

\title{Stay Focused: Problem Drift in Multi-Agent Debate}


\author[1]{\bf Jonas Becker}
\author[1]{\bf Lars Benedikt Kaesberg}
\author[2]{\bf Andreas Stephan}
\author[1]{\bf Jan Philip Wahle}
\author[1]{\\\bf Terry Ruas}
\author[1]{\bf Bela Gipp}
\affil[1]{University of Göttingen}
\affil[ ]{\texttt{\{jonas.becker, larsbenedikt.kaesberg, wahle, ruas, gipp\}@uni-goettingen.de}}
\affil[2]{University of Vienna}
\affil[ ]{\texttt{\{andreas.stephan\}@univie.ac.at}}


\begin{document}
\maketitle
\AddAnnotationRef
\begin{abstract}
Multi-agent debate --- multiple instances of large language models discussing problems in turn-based interaction --- has shown promise for solving knowledge and reasoning tasks.
However, these methods show limitations, particularly when scaling them to longer reasoning chains.
In this study, we unveil a new issue of multi-agent debate: discussions drift away from the initial problem over multiple turns.
We define this phenomenon as \textbf{problem drift} and quantify its presence across ten tasks (i.e., three generative, three knowledge, three reasoning, and one instruction-following task).
To identify the reasons for this issue, we perform a human study with eight experts on discussions suffering from problem drift, who find the most common issues are a lack of progress (35\% of cases), low-quality feedback (26\% of cases), and a lack of clarity (25\% of cases).
To systematically address the issue of problem drift, we propose \judge, a method based on LLM-as-a-judge, to detect problem drift at test-time.
We further propose \policy, a method to mitigate 31\% of problem drift cases.
Our study can be seen as a first step to understanding a key limitation of multi-agent debate, highlighting pathways for improving their effectiveness in the future.
\end{abstract}

\input{1_introduction}
\input{2_relatedWork}
\input{3_definition}
\input{4_methodlogy}
\input{5_experiments}
\input{6_epilogue}

\bibliography{25_ProblemDrift}

\appendix

\input{7_appendix}

\clearpage
\onecolumn
\input{annotation.tex}
\end{document}
