\section{Related Work}
\paragraph{Trajectory dataset.} Data-driven approaches for control like imitation learning ____ and offline reinforcement learning ____ have promoted the public availability of trajectory datasets. However, these datasets are rarely utilized as unified big data for foundation models, likely due to their isolated characteristics, such as differences in policy levels, observation spaces, and action spaces. In fact, the largest robotics dataset, Open X-Embodiment ____, is typically used for imitation learning with homogeneous visual observations and end-effector actions ____. Gato ____ collects a large-scale dataset across diverse environments for a generalist agent, but it is not publicly available. In contrast, we curate public heterogeneous datasets, targeting a more capable trajectory world model.

\paragraph{Cross-environment architecture.} Zero-padding to fit a maximum length ____ or using separate neural network heads ____ hinders knowledge transfer between heterogeneous environments with mismatched or differently sized state and action spaces. Previous work has resorted to flexible architectures like graph neural networks ____ and Transformers ____ for policy learning. Our method leverages a similar architecture for world modeling ____, but with a novel two-dimensional attention design to enhance cross-environment transfer.

\paragraph{World model pre-training.} The homogeneity of videos across diverse tasks, environments, and even embodiments has driven rapid advancements in large-scale video pre-training for world models ____. However, heterogeneity across different sets of sensors and actuators poses significant challenges to developing general world models based on low-dimensional sensor information.

Our work is particularly relevant to ____, which trains a generalist transformer dynamics model from 80 heterogeneous environments. Still, they only observe positive transfer when adapting to a simple cart-pole environment and fail for a more complex walker environment. In contrast, our work, for the first time, validates the positive transfer benefits across such more complex environments.