
\documentclass{article} % For LaTeX2e
\usepackage{iclr2025_conference,times}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
\input{math_commands.tex}

\usepackage{hyperref}
\usepackage{url}

%% Extra packages - START - %%
\usepackage{booktabs}       % professional-quality tables
\usepackage{graphicx}
\usepackage{derivative}
\usepackage{amsmath}
\usepackage{multirow, multicol}
\usepackage{makecell}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}
\usepackage{svg}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{comment}
\usepackage[ruled]{algorithm2e} % For algorithms

\usepackage{colortbl} 
\definecolor{firstcolor}{RGB}{252, 235, 193}%{236 168 169}
\definecolor{secondcolor}{RGB}{241, 233, 223} %{211 226 183}
\definecolor{thirdcolor}{RGB}{219, 200, 189} %{116 174 212}
\newcommand{\markfirst}[1]{\cellcolor{firstcolor}{\textbf{#1}}}
\newcommand{\marksecond}[1]{\cellcolor{secondcolor}{#1}}
\newcommand{\markthird}[1]{\cellcolor{thirdcolor}{#1}}
\newcommand{\colorfirsttext}[1]{\colorbox{firstcolor}{\textbf{#1}}}
\newcommand{\colorsecondtext}[1]{\colorbox{secondcolor}{#1}}
\newcommand{\colorthirdtext}[1]{\colorbox{thirdcolor}{#1}}

\definecolor{GammaColor}{rgb}{0.5,0,0.7}
\newcommand{\yajie}[1]{{\color{GammaColor} Yajie: #1 }}
% \newcommand{\edit}[1]{\textcolor{magenta}{#1}}
\newcommand{\edit}[1]{\textcolor{black}{#1}}

%% Extra packages - END - %%


\title{SC-OmniGS: Self-Calibrating \\Omnidirectional Gaussian Splatting}

% Authors must not appear in the submitted version. They should be hidden
% as long as the \iclrfinalcopy macro remains commented out below.
% Non-anonymous submissions will be rejected without review.

\author{Huajian Huang$^1$\thanks{ Equal contribution. {$\dagger$ Corresponding author: huajian@ust.hk}} $\,^\dagger$ \quad Yingshu Chen$^{1 *}$ \quad Longwei Li$^2$ \quad Hui Cheng$^2$ \quad Tristan Braud$^1$ \\ 
\textbf{Yajie Zhao$^3$ \quad Sai-Kit Yeung$^1$} \\
$^1$ The Hong Kong University of Science and Technology \\
$^2$ Sun Yat-sen University \quad
$^3$ ICT, University of Southern California \\
% \And
% Yingshu Chen
% \And
% Longwei Li
% \And
% Hui Cheng
% \And
% Tristan Braud
% \And
% Yajie Zhao
% \And
% Sai-Kit Yeung\\
% \thanks{Use fo.} \\
% $^1$Department of Computer Science\\
% \thanks{ Use footnote}\\
% Cranberry-Lemon University\\
% Pittsburgh, PA 15213, USA \\
% \texttt{\{hhuangbg, susan.yingshu.chen\}@connect.ust.hk} \\
% \And
% Ji Q. Ren \& Yevgeny LeNet \\
% Department of Computational Neuroscience \\
% University of the Witwatersrand \\
% Joburg, South Africa \\
% \texttt{\{robot,net\}@wits.ac.za} \\
% \AND
% Coauthor \\
% Affiliation \\
% Address \\
% \texttt{email}
% \And
% Coauthor \\
% Affiliation \\
% Address \\
% \texttt{email}
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\iclrfinalcopy % Uncomment for camera-ready version, but NOT for submission.
\begin{document}


\maketitle

\begin{abstract}
360-degree cameras streamline data collection for radiance field 3D reconstruction by capturing comprehensive scene data. However, traditional radiance field methods do not address the specific challenges inherent to 360-degree images.
We present SC-OmniGS, a novel self-calibrating omnidirectional Gaussian splatting system for fast and accurate omnidirectional radiance field reconstruction using 360-degree images. Rather than converting 360-degree images to cube maps and performing perspective image calibration, we treat 360-degree images as a whole sphere and derive a mathematical framework that enables direct omnidirectional camera pose calibration accompanied by 3D Gaussians optimization.  Furthermore, we introduce a differentiable omnidirectional camera model in order to rectify the distortion of real-world data for performance enhancement. 
Overall, the omnidirectional camera intrinsic model, extrinsic poses, and 3D Gaussians are jointly optimized by minimizing weighted spherical photometric loss. 
Extensive experiments have demonstrated that our proposed SC-OmniGS is able to recover a high-quality radiance field from noisy camera poses or even no pose prior in challenging scenarios characterized by wide baselines and non-object-centric configurations. 
The noticeable performance gain in the real-world dataset captured by consumer-grade omnidirectional cameras verifies the effectiveness of our general omnidirectional camera model in reducing the distortion of 360-degree images. 
\end{abstract}

\input{journals}

\end{document}
