%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf, nonacm]{acmart}
%\documentclass[manuscript]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.


%\copyrightyear{2025}
%\acmYear{2025}
%\setcopyright{cc}
%\setcctype{by-nc}
%\acmConference[CHI '25]{CHI Conference on Human Factors in Computing
%Systems}{April 26-May 1, 2025}{Yokohama, Japan}
%\acmBooktitle{CHI Conference on Human Factors in Computing Systems (CHI
%'25), April 26-May 1, 2025, Yokohama,
%Japan}
%\acmDOI{10.1145/3706598.3713362}
%\acmISBN{979-8-4007-1394-1/25/04}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%\citestyle{acmauthoryear}


%%
%% end of the preamble, start of the body of the document source.

\usepackage{booktabs}  % For professional-looking tables
\usepackage{array}  % For custom column widths
\usepackage{multirow}  % For multirow cells
\usepackage{tikz}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions}
\usepackage{tikz,tikz-3dplot}
\usepackage{caption}
\captionsetup {labelfont=bf}
\usepackage{float}
\restylefloat{table}
\usepackage{placeins}
\usepackage{graphicx}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{filecontents}
\usepackage{soul}


\begin{document}


%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title[Integrating Cultural Sensitivity in an LLM Chatbot for Reproductive Health]{``\textit{Kya} family planning after marriage \textit{hoti hai}?'': Integrating Cultural Sensitivity in an LLM Chatbot for Reproductive Health
}


%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.


\author{Roshini Deva}
\affiliation{%
 \institution{Emory University}
 \city{Atlanta}
 \state{Georgia}
 \country{United States}}
 \email{deva.roshini@emory.edu}

\author{Dhruv Ramani}
\affiliation{%
  \institution{Georgetown University}
  \city{Washington}
  \state{District of Columbia}
  \country{United States}}
  \email{dr973@georgetown.edu}
  
  \author{Tanvi Divate}
\affiliation{%
  \institution{Myna Mahila Foundation}
  \city{Mumbai}
  \state{Maharastra}
  \country{India}}
  \email{tanvi@mynafoundation.com}
  
  \author{Suhani Jalota}
\affiliation{%
  \institution{Stanford University}
  \city{Stanford}
  \state{California}
  \country{United States}}
  \email{suhani@stanford.edu}

\author{Azra Ismail}
\affiliation{%
  \institution{Emory University}
  \city{Atlanta}
  \state{Georgia}
  \country{United States}}
\email{azra.ismail@emory.edu}


\renewcommand{\shortauthors}{Deva et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
 Access to sexual and reproductive health information remains a challenge in many communities globally, due to cultural taboos and limited availability of healthcare providers. Public health organizations are increasingly turning to Large Language Models (LLMs) to improve access to timely and personalized information. However, recent HCI scholarship indicates that significant challenges remain in incorporating context awareness and mitigating bias in LLMs. In this paper, we study the development of a culturally-appropriate LLM-based chatbot for reproductive health with underserved women in urban India. Through user interactions, focus groups, and interviews with multiple stakeholders, we examine the chatbot’s response to sensitive and highly contextual queries on reproductive health. Our findings reveal strengths and limitations of the system in capturing local context, and complexities around what constitutes ``culture''. Finally, we discuss how local context might be better integrated, and present a framework to inform the design of culturally-sensitive chatbots for community health.

\end{abstract}


\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003120.10003121.10011748</concept_id>
<concept_desc>Human-centered computing~Empirical studies in HCI</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Human-centered computing~Empirical studies in HCI}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{LLM, chatbot, reproductive health, HCI4D}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.


% \received{20 February 2007}
% \received[revised]{12 March 2009}
% \received[accepted]{5 June 2009}


%\newcommand{\AI}[1]{\textcolor{magenta}{[#1 - Azra]}}
%\newcommand{\RD}[1]{\textcolor{red}{#1}}
%\newcommand{\DR}[1]{\st{#1}}
\newcommand{\DR}[1]{}
\newcommand{\RD}[1]{\textcolor{black}{#1}}

%\newcommand{\X}[1]{\textcolor{magenta}{[#1 - X]}}
%\newcommand{\chiadd}[1]{\textcolor{red}{#1}}
%\newcommand{\chirm}[1]{\st{#1}}
 \newcommand{\chiadd}[1]{\textcolor{black}{#1}}
 \newcommand{\chirm}[1]{}



\maketitle
\renewcommand{\thefootnote}{} 
\footnotetext{This is a preprint of the paper accepted at CHI 2025. The final version will be available in the ACM Digital Library.}
\renewcommand{\thefootnote}{\arabic{footnote}}
\input{introduction.tex}
\input{litreview.tex}
\input{methods.tex}
\input{design.tex}
\input{findings.tex}
\input{discussion.tex}

\section{CONCLUSION}
% Our study shows that while the chatbot provides useful health information, it struggles with integrating local dialects, cultural nuances, and behavioral patterns specific to the user's context. This tells the need for more culturally sensitive tools that can effectively bridge the gap between global health technology and local realities. The study emphasizes the importance of incorporating cultural context into the design and evaluation of healthcare chatbots to ensure they meet the needs of diverse populations.
% Future development should be focused on enhancing the chatbot’s ability to recognize local linguistic and cultural variations. Also, feedback from users, healthcare professionals, and developers is essential to refine the chatbot’s responses and improve its cultural sensitivity. By addressing these challenges, we can develop more effective and inclusive public health tools that provide accurate, relevant, and respectful information to all users, eventually improving health outcomes in varied cultural contexts.

This study examined the development of an LLM-based chatbot for providing sexual and reproductive health information to underserved women in urban India. By analyzing user interactions, focus groups, and interviews with multiple stakeholders, we identified both the strengths and limitations of the system in addressing highly contextual and sensitive health queries. Our research highlighted the complexities of delivering culturally sensitive information, and the need for a nuanced understanding of local beliefs and practices. While the chatbot showed promise in delivering relevant health information, challenges remain in ensuring cultural appropriateness and adapting to diverse and evolving community needs. Our paper thus contributes to HCI literature by presenting a framework for designing culturally sensitive LLM-based chatbots. We thus aim to inform future efforts to implement LLMs and chatbots in healthcare settings, particularly with underserved populations.

\section{ACKNOWLEDGEMENTS}
\RD{We thank our participants for generously sharing their time and experiences, which were invaluable to this study. We are also grateful to the external reviewers for their insightful feedback that helped strengthen our work. We want to extend our gratitude to Manvi S and Sumon Kanti Dey from CARE Lab at Emory University, Meet Shah and Harsh Vitra from the technical team, Zeel Mehta from the program team, and Dr. Shraddha Kale Kapile, Associate Director of the Health Vertical at Myna Mahila Foundation for their support throughout the study. This research was made possible with the support of the Google Award for Inclusion Research, and The Agency Fund through Myna Mahila Foundation.}

\bibliographystyle{plain}
\bibliography{sample-base}
%\bibliographystyle{ACM-Reference-Format}

\input{appendix.tex}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.
