Despite the effectiveness of the proposed \textbf{RIDE} method in enhancing LLM alignment, several limitations and potential risks should be acknowledged.

\paragraph{Limited Scope of ICL Demonstrations.} One key limitation of this study is the restricted selection of ICL demonstrations. The candidate ICL demos were drawn from a subset of a large dataset, which may limit their diversity and generalizability. Given that alignment performance is highly dependent on the variety of training examples, a more extensive and diverse selection of candidate ICL exemplars could potentially yield stronger results. Future work should explore the impact of expanding the candidate pool by incorporating demonstrations from multiple datasets across different domains.

\paragraph{Dependency on LLM-as-a-Judge for Evaluation.} The evaluation methodology relies on using a strong LLM-as-a-judge (ChatGPT or Claude-3.5 Sonnet) to assess the effectiveness of restyled demonstrations. While this provides a cost-effective alternative to human evaluation, it introduces potential biases. LLMs used for scoring may favor responses that align with their own training data and reward certain styles over others in a way that may not fully reflect human preferences. Future work should incorporate human evaluations to validate the robustness of the results.

\paragraph{Potential for Misuse and Ethical Considerations.} Although \textbf{RIDE} aims to enhance LLM alignment, there exists a risk of its misuse. If adversarial actors manipulate ICL demonstrations using the same restyling approach, they could attempt to bypass safety constraints or generate misleading outputs. Additionally, optimizing for alignment does not eliminate the potential for biases present in the base LLMs, which may still surface despite restyling efforts. Ensuring continuous auditing and ethical oversight in deploying such methods is essential.

\paragraph{Future Directions.} To address these limitations, future research should: (i) Expand the candidate ICL demo pool to improve generalization across diverse datasets.
(ii) Reduce dependency on LLM-as-a-judge by integrating human assessments and alternative evaluation methods.
(iii) Establish safeguards against potential adversarial uses of restyled ICL demonstrations.
