\section{Related Work}
% \paragraph{World Modeling with LLMs.}
% Neural World Modeling
% https://arxiv.org/pdf/2412.03572
% merler2024generating 基于mcts生成code world model
% wong2023word
% li2022emergent leverage transformer model to predict next legal step on a board game...
% tang2024worldcoder construct a code world model via interacting with the environments
% direct application of LLMs for state prediction during planning____, where benchmarks like ByteSized32-State-Prediction____ have been developed to evaluate LLMs' predictive capabilities. However, this evaluation approach faces significant challenges: exhaustively evaluating all possible state transitions becomes intractable in complex environments, and limited interpretability and high inference latency constrain deeper analysis of LLMs' world modeling capabilities.
Neural world modeling is a long-standing research topic with widespread applications across various fields, including reinforcement learning____, robotics____, and autonomous driving____, among others.
In recent years, LLMs trained on massive datasets have demonstrated zero-shot capabilities across a variety of tasks, including planning____, robotics____, analog design____, and more.
% Preliminary studies propose directly using LLMs as world models____, where the input consists of the state and action, and the LLM outputs the predicted next state. 
% However, due to the unreliability of LLM outputs and their limited interpretability, this approach can lead to the problem of accumulating errors. 
Preliminary studies propose directly using LLMs as world models____, by taking the state and action as input and predicting the next state, but the unreliability and limited interpretability of LLM outputs can lead to accumulating errors.
Moreover, some studies have shown that autoregressive models perform poorly in predicting action effects____.
Tree-planner____ instead proposes to constructing the possible action space using LLMs before executing.
Another line of work focuses on leveraging LLMs to construct symbolic world models____. 
For example, ____ uses LLMs to generate a PDDL domain model and relies on human feedback to correct errors. 
AgentGen____ synthesizes diverse PDDL domains, aiming to create high-quality planning data. 
____ propose to finetune LLMs for predicting precondition and effect of actions.
% ____ proposes novel evaluation metrics inspired by the Myhill-Nerode theorem to assess the implicit world models within generative models.
% 
Despite the growing interest in this research direction, there is currently a lack of a comprehensive benchmark in this area. 


% \vspace{-10pt}