\section{Preliminary} 
\label{sec:preliminary}

\paragraph{Problem Definition.} MAD, which integrates multiple agents for interactive communication to derive solutions, has demonstrated an effective approach in the application of LLMs, particularly in addressing complex logical reasoning and mathematical problems \citep{liang2023encouraging, chan2023chateval}. Given an input question $Q$ that requires an answer, a total of $M$ participating Language Model (LLM)-based agents engage in a multi-round debate, which is denoted as $A_i$, where $i \in \{1, 2, \ldots, M\}$.
% denoted as $A = \{ A_i | i = 1, 2, \ldots, M \}$, 
Given a total of $T$ debate rounds, each round of debate is denoted as $t \in \{1, 2, \ldots, T\}$. We define the output of each agent $A_i$ at round $t$ as $O_i^{t}$. We assume that the upper bound of the token cost for each agent's output is $C$. Our goal is to maximize answer accuracy while minimizing token consumption through optimizing the interaction patterns among the agents in multi-agent debate.

\paragraph{MAD-based Methods and Token Cost.}
(i) MAD \citep{liang2023encouraging} involves several steps. Initially, each agent is provided with a question and generates an individual response. These responses then form the new input context for each agent, leading to the generation of subsequent responses. This debate procedure is repeated over multiple rounds, with the final answer derived through majority voting. The token cost complexity is $Token^{MAD} = \mathcal{O}\left( MTQ+(M^2T+MT^2)C \right)$. (ii) S-MAD \citep{li2024improving} decreases token consumption by sparsifying the fully connected topology of information exchange among agents within the standard MAD framework. Let $P_r$ denote the probability of each edge being removed. The token cost complexity can be represented as $Token^{S-MAD} = \mathcal{O}\left( MTQ+(1-P_r)M^2TC \right)$. (iii) GD \citep{liu2024GroupDebate} reduces token consumption through a group discussion strategy. Let $N$ denote the number of groups and $S$ represent the number of inter-group debate stages. The token cost complexity is computed as $Token^{GD} = \mathcal{O}\left(MTQ+(\frac{M^2T} {N}+MSN)C\right)$.