\documentclass[a4paper]{article}

\usepackage[l2tabu, orthodox]{nag}
\usepackage[utf8]{inputenc}
\usepackage[pdfusetitle]{hyperref}
\usepackage{microtype}

% \usepackage{xcolor}
% \colorlet{linkcolour}{blue!50!black}
% \colorlet{urlcolour}{magenta}
% \hypersetup{colorlinks=true, linkcolor=linkcolour, citecolor=linkcolour, urlcolor=urlcolour,}

\usepackage{mathtools}
\DeclarePairedDelimiter\paren{\lparen}{\rparen}
% usage: \paren{expression}, sized variant \paren*{expression}, 
% specified size variant \paren{\big} (\Big, \bigg, \Bigg)
\mathtoolsset{
  showonlyrefs=true % or false in draft mode
}%use \equation{} everywhere; only the referenced ones are numbered

% \usepackage[backref=true, isbn=true, url=true, firstinits=true, maxnames=20, style=numeric,]{biblatex}
%%% example bibliography tweaking

\usepackage{amsmath,amssymb,amsthm}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{claim}{Claim}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{remark}
\newtheorem{example}{Example}
\newtheorem{observation}{Observation}
\newtheorem{remark}{Remark}

\usepackage{kbordermatrix}
\usepackage{thm-restate}
\usepackage{cleveref}

\renewcommand{\kbldelim}{(}% Left delimiter
\renewcommand{\kbrdelim}{)}% Right delimiter

\newif\iflong
\newif\ifshort
% comment the below line out for short version
\longtrue

\iflong
\else
\shorttrue
\fi

\usepackage{fullpage}
% \usepackage{showframe}

\usepackage{enumerate} %Gives ability, e.g., \begin{enumerate}[Case 1.]

\usepackage[noend]{algpseudocode}
\usepackage{algorithm}

\usepackage{xspace}

\newcommand{\cclass}[1]{\ensuremath{\mbox{\textup{#1}}}\xspace}
\newcommand{\BPP}{\cclass{BPP}}
\newcommand{\RP}{\cclass{RP}}
\newcommand{\coRP}{\cclass{coRP}}
\newcommand{\NP}{\cclass{NP}}
\newcommand{\coNP}{\cclass{coNP}}
\renewcommand{\P}{\cclass{P}}

\newcommand{\N}{\ensuremath{\mathbb N}\xspace}
\newcommand{\R}{\ensuremath{\mathbb R}\xspace}
\newcommand{\F}{\ensuremath{\mathbb F}\xspace}
\newcommand{\Q}{\ensuremath{\mathbb Q}\xspace}
% \renewcommand{\P}{\ensuremath{\mathcal{P}}\xspace}
\newcommand{\cI}{\ensuremath{\mathcal I}\xspace}
\newcommand{\cB}{\ensuremath{\mathcal B}\xspace}
\newcommand{\cC}{\ensuremath{\mathcal C}\xspace}
\newcommand{\cP}{\ensuremath{\mathcal P}\xspace}
\newcommand{\cL}{\ensuremath{\mathcal L}\xspace}
\newcommand{\cF}{\ensuremath{\mathcal F}\xspace}
\newcommand{\cT}{\ensuremath{\mathcal T}\xspace}
\newcommand{\cS}{\ensuremath{\mathcal S}\xspace}
\newcommand{\cM}{\ensuremath{\mathcal M}\xspace}
\newcommand{\cX}{\ensuremath{\mathcal X}\xspace}
\newcommand{\cQ}{\ensuremath{\mathcal Q}\xspace}

\newcommand{\bD}{\ensuremath{\mathbf D}\xspace}

\newcommand{\ar}{\ensuremath{r}}

\newcommand{\containment}{\NP~$\subseteq$~\coNP/poly\xspace}
\newcommand{\assumption}{\NP~$\nsubseteq$~\coNP/poly\xspace}

\newcommand{\Oh}{\mathcal{O}}

\newcommand{\MM}{\mathsf{MM}}

\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\rank}{rank}

\usepackage[usenames,dvipsnames]{color}

\usepackage{tikz}
\usetikzlibrary{arrows.meta,calc} 
\usetikzlibrary{decorations.pathreplacing}

\tikzset{vertex/.style={draw, circle, fill=white, inner sep=2.4pt}}
\tikzset{bvertex/.style={draw, circle, fill=black, inner sep=2.4pt}}
\tikzset{>={Latex[length=3mm, width=2mm]}}

\usepackage{todonotes}

\newcommand{\discuss}[1]{\textcolor{OliveGreen}{#1}}
\newcommand{\changed}[1]{\textcolor{red}{\textbf{$\boldsymbol{<}$#1$\boldsymbol{>}$}}}
\DeclareMathOperator{\Pf}{Pf}
\DeclareMathOperator{\GF}{GF}


\newcommand{\flowervertex}[2]{
  \foreach \angle in {30,120,...,300}
      \draw[fill=white] (#1, #2) -- ++(\angle:0.3) arc (\angle:\angle+30:0.5) -- cycle;
      \node[bvertex] (_) at (#1, #2) {};
}

\title{FPT algorithms over linear delta-matroids with applications}

\author{Eduard Eiben\thanks{ Dept. of Computer Science, Royal
    Holloway, University of London, \texttt{Eduard.Eiben@rhul.ac.uk}}
  \and
  Tomohiro Koana\thanks{Utrecht University \& Research Institute for Mathematical Sciences, Kyoto University, \texttt{tomohiro.koana@gmail.com}}
  \and
  Magnus Wahlström\thanks{Dept. of Computer Science, Royal Holloway,
    University of London, \texttt{Magnus.Wahlstrom@rhul.ac.uk}}}

\date{}

\begin{document}
\maketitle

\begin{abstract}
  Matroids, particularly linear matroids, have been a powerful tool for
  applications in parameterized complexity, both for algorithms and
  kernelization.   In particular, they have been instrumental in speeding up or replacing dynamic programming.
  Delta-matroids are a generalization of matroids that further
  encapsulates structures such as non-maximum matchings in general
  graphs and various path-packing and topological structures.
  There is also a notion of linear delta-matroids
  (represented by skew-symmetric matrices) which carries significant
  expressive power and enables powerful algorithms. 
  We investigate parameterized complexity aspects of problems defined over linear delta-matroids,
  or with delta-matroid constraints.
  Our initial analysis of basic intersection and packing problems
  reveals a different complexity landscape compared to the more familiar matroid case. 
  In particular, there is a stark contrast in complexity between
  the \emph{cardinality} parameter $k$ and the \emph{rank} parameterization $r$.
  For example, finding an intersection of size $k$ of three linear delta-matroids
  is W[1]-hard when parameterized by $k$, while far more general problems
  (such as finding a set packing of size $k$ that is feasible in a
  given linear delta-matroid) are FPT when parameterized by
  the rank $r$ of the delta-matroid.
  In fact, we extend the recent \emph{determinantal sieving}
  procedure of Eiben, Koana and Wahlström (SODA 2024)
  into a process that sieves a given polynomial for a monomial whose support is
  feasible in a given linear delta-matroid, parameterized by $r$.
  This is a direct generalization of determinantal sieving. 
  
  Second, we investigate a curious class of problems that turns out
  to be FPT parameterized by $k$, even on delta-matroids of unbounded rank.
  We begin with \textsc{Delta-matroid Triangle Cover} -- find a
  feasible set of size $k$ that can be covered by a vertex-disjoint
  packing of triangles (i.e., sets of size 3) out of a given triangle collection.
  For example, this allows us to find a packing of $K_3$'s and $K_2$'s
  in a graph with a maximum number of edges, parameterized above the
  matching number of the graph. As applications, we settle
  questions on the FPT status of \textsc{Cluster Subgraph} and
  \textsc{Strong Triadic Closure} parameterized above the matching number. 
\end{abstract}

\section{Introduction}

\emph{Matroids} are abstract structures that capture a notion of \emph{independence} across various domains -- from linearly independent
sets in vector spaces to
the spanning forests in graphs. Formally, a matroid is a pair
$M=(V,\cI)$ where $\cI \subseteq 2^V$, referred to as the \emph{independent sets},
is a collection of sets satisfying the axioms (i) $\emptyset \in \cI$; (ii) if $A \subseteq B$ and $B \in \cI$
then $A \in \cI$ and (iii) the \emph{independence augmentation axiom},
\[
  \text{ if } A, B \in \cI \text{ and } |A|<|B| \text{ then there exists } x \in B \setminus A \text{ such that } A \cup \{x\} \in \cI.
\]
%satisfies certain axioms modeled on linear independence.
A \emph{basis} of a matroid is a maximal independent set. A matroid can equivalently be defined as a pair $M=(V,\cB)$
where $\cB \subseteq 2^V$ is the set of \emph{bases} of $M$, subject to (i) $\cB \neq \emptyset$ and (ii)
the \emph{basis exchange axiom},
\[
  \text{ if } A, B \in \cB  \text{ and } x \in A \setminus B \text{ then there exists } y \in B \setminus A \text{ such that } A \Delta \{x,y\} \in \cB,
\]
where $\Delta$ denotes symmetric difference. 
The study of matroids goes back nearly a century (to
Whitney, 1935~\cite{whitney1935abstract}), and matroids find applications in practically all
parts of computer science. One of the central classes of matroids is \emph{linear matroids},
where the ground set $V$ represents vectors in a vector space over a field $\F$ 
and the independent sets are precisely the sets of vectors that are
linearly independent. This is usually represented via a matrix over $\F$
whose columns are labelled by $V$. 

A major benefit of matroids is the powerful set of algorithmic meta-results they offer. 
Famously, matroids are precisely those structures in which the basic
\emph{greedy algorithm} is guaranteed to find a max-weight solution.
Another pair of well-studied problems are \textsc{Matroid Intersection},
or the problem of finding a maximum-cardinality set in the
intersection of two matroids, and \textsc{Matroid Parity},
where the input is a matroid $M=(V,\cI)$ and a partition $\Pi$
of $V$ into pairs, and the task is to find an independent set $I$
of maximum cardinality that is a union of pairs from $\Pi$
(or equivalently, a basis $B$ of $M$ in which a minimum number of
pairs is broken, where a pair $p \in \Pi$ is broken by $B$
if $|p \cap B|=1$). \textsc{Matroid Intersection} can be solved
in polynomial time over any matroid, while \textsc{Matroid Parity}
is NP-hard in general but can be efficiently solved on a class of
matroids including linear matroids~\cite{Lovasz79,LP86}. 
For more on matroids, see Oxley~\cite{OxleyBook2};
see also Schrijver~\cite{SchrijverBook} for more algorithmic
applications, 

Linear matroids have also seen powerful applications in parameterized
complexity, which is the starting point of our current investigations.
This started with Marx~\cite{Marx09-matroid}, who gave FPT
algorithms for problems over linear matroids, with applications, based
on an algorithmic version of the representative sets theorem
(a.k.a.\ the \emph{two families theorem} of Bollobás \cite{bollobas1965generalized}, 
generalized to linear matroids by Lovász~\cite{lovasz1977flats}).
This was further developed by Fomin et al.~\cite{FominLPS16JACM},
who showed a faster version and observed several applications to
speeding up FPT dynamic programming algorithms; this strategy has
since become a standard building block in the FPT toolbox~\cite{CyganFKLMPPS15PCbook,FominLPS16JACM,FominLPS17}.
Linear matroids and representative sets have also been instrumental
in results for kernelization, particularly for graph separation
problems~\cite{KratschW14TALG,KratschW20JACM,Kratsch18vc,Wahlstrom22TALG}.
In another direction, the \emph{extensor coding}~\cite{BrandDH18}
and recent \emph{determinantal sieving}~\cite{EKW23} methods
can be used to filter an arbitrary polynomial (e.g., a multivariate
generating function) for monomials whose support forms a basis of a
provided linear matroid. This unifies and generalizes earlier monomial
sieving methods~\cite{BjorklundHKK17narrow,BjorklundKK16} and currently
represents the fastest, most general FPT algorithm for a range of
problems~\cite{EKW23}. 


\paragraph*{Delta-matroids.}
\emph{Delta-matroids} are generalizations of matroids, defined and
initially investigated by Bouchet in the 1980s~\cite{Bouchet87DMone}
although similar or equivalent structures were introduced
independently by other researchers; see Moffatt~\cite{Moffatt19deltamatroids}
for a survey. A delta-matroid is a pair $D=(V,\cF)$ where $\cF \subseteq 2^V$ 
is a collection of subsets referred to as \emph{feasible sets},
subject to the \emph{symmetric exchange axiom}: 
\[
\text{ for all } A, B \in \cF \text{ and } x \in A \Delta B \text{ there exists } y \in A \Delta B \text{ such that }
A \Delta \{x,y\} \in \cF.
\]
%where $\Delta$ denotes symmetric difference. 
Both the set of bases of a matroid, and the independent sets of a matroid, 
form the feasible sets of a delta-matroid.
We refer to these as the \emph{basis delta-matroid}
and \emph{independence delta-matroid} of the matroid.
For another prominent example, let $G=(V,E)$ be a graph
and let $\cF$ contain those subsets $F \subseteq V$
such that $G[F]$ has a perfect matching. Then $D=(V,\cF)$
forms a delta-matroid, the \emph{matching delta-matroid} of $G$.
As a special case, given a partition $\Pi$ of $V$ into pairs,
we define the \emph{pairing delta-matroid} $D_\Pi$ of $\Pi$ as 
the matching delta-matroid of the graph whose edge set is $\Pi$;
i.e., $F \subseteq V$ is feasible in $D_\Pi$ if and only if no pair is
broken in $F$.
This is frequently used in combinatorial constructions in this paper.
Like with matroids, there is also a notion of \emph{linear delta-matroid}. 
A matrix $A$ is \emph{skew-symmetric} if $A^T=-A$.
Let $A$ be a skew-symmetric matrix with rows and columns labelled by $V$,
and let $\cF=\{F \subseteq V \mid A[F] \text{ is non-singular}\}$. 
Then $D=(V,\cF)$ defines a delta-matroid 
(although this fact is not trivial);
we denote this delta-matroid by $D=\bD(A)$.
More generally, let $S \subseteq V$ and define
$
  \cF \Delta S = \{F \Delta S \mid F \in \cF\}.
$
Then $D \Delta S=(V,\cF \Delta S)$ defines a delta-matroid,
the \emph{twist} of $D$ by $S$. A \emph{linear delta-matroid}
is a delta-matroid which can be represented as
$D=\bD(A) \Delta S$ for some skew-symmetric matrix $A$ as above.
We will refer to this representation as a \emph{twist representation}.
Examples of linear delta-matroids include 
basis delta-matroids of linear matroids,
\emph{twisted matroids} (being the twists of basis delta-matroids of linear matroids)
and matching delta-matroids, the latter represented via the famous
\emph{Tutte matrix} of $G$. Again, see Moffatt~\cite{Moffatt19deltamatroids}
for more background and applications. 
In addition, slightly more generally, one may consider
\emph{projected} linear representations of delta-matroids,
which includes independence delta-matroids of linear matroids.
%These are generally not linear, since linear delta-matroids are \emph{even}:
%all their feasible sets have the same cardinality. 

Finally, as with matroids, delta-matroids enjoy a collection of algorithmic meta-results.
There is a signed generalization of the greedy algorithm that finds a max-weight
solution in a set system precisely when the set system defines a
delta-matroid, and while both the natural generalizations
\textsc{Delta-matroid Intersection} and \textsc{Delta-matroid Parity}
are NP-hard in general, both are tractable on linear delta-matroids.
More precisely, the problems are usually defined as follows. In
\textsc{Delta-matroid Intersection}, the task is to find \emph{any}
common feasible set $F$ in two delta-matroids $D_1$ and $D_2$, without
any requirement on $|F|$, and in \textsc{Delta-matroid Parity}
the task is to find a feasible set $F$ in a delta-matroid
such that a minimum number of pairs are broken in $F$.
Since delta-matroids are not normally closed under taking subsets, 
this definition of \textsc{Delta-matroid Intersection} is already
a non-trivial problem. These versions are tractable on linear delta-matroids
due to a result of Geelen et al.~\cite{GeelenIM03}, recently
improved to $\tilde O(n^\omega)$ time by Koana and Wahlström~\cite{KW24}.
The natural variant \textsc{Maximum Delta-matroid Intersection},
where $|F|$ is maximised, was raised by Kakimura and Takamatsu~\cite{KakimuraT14sidma},
and appears harder. It was only recently solved by a randomized algorithm~\cite{KW24}.
See the latter for a deeper overview on algorithms for linear delta-matroids.
In summary, although they are less known, delta-matroids (and
especially linear delta-matroids) generalize many of the useful
aspects of (linear) matroids.\iflong\footnote{As a side note, another frequent occurrence of
  matroids in combinatorial optimization is as a form of \emph{constraints},
  striking a good balance between expressive power and algorithmic
  tractability. See for example the work on approximation algorithms
  under matroid and knapsack constraints, and the \emph{matroid secretary} problem.
  Although not precisely the same, there are also a number of
  cases where delta-matroid constraints capture the boundary
  of polynomial-time tractable problems; cf.~the \textsc{Boolean edge CSP}  
  and \textsc{Boolean planar CSP} problems~\cite{KW24,KazdaKR19}
  and the discussion on \emph{general factor} in~\cite{KW24}.
}\fi

%There is also an interesting use of matroids as tractable or otherwise 
%well-behaved classes of constraints in combinatorial optimization. 
%\todo{examples of 'knapsack subject to matroid constraints', 'matroid
%secretary', etc.}
%This goes beyond the scope of the current paper, but there is a
%distinct, but very interesting, occurrence of delta-matroids as
%encoding tractable constraint classes in the \textsc{General Matching}
%problem and more generally in the Boolean \textsc{Graph CSP}
%and \textsc{Planar CSP} problems; see Kazda et al.~\cite{KazdaKR19}
%and the discussion in Koana and Wahlström~\cite{KW24}.

%\paragraph*{Matroids in parameterized complexity.}

In this paper, we investigate parameterized complexity aspects of problems on linear and
projected linear delta-matroids, in the same vein as the work on linear matroids
surveyed above. The representative sets theorem over linear matroids 
has recently been extended to linear delta-matroids, with some initial applications
to kernelization~\cite{Wahlstrom24SODA}, but the applicability of
delta-matroid tools for FPT algorithms has so far not been pursued. 
Our purpose is to push this forward. We investigate
upper and lower bounds for parameterizations of natural NP-hard
problems over linear delta-matroids, on the way generalizing the
method of determinantal sieving. As a consequence, we settle open
questions on above-matching parameterizations of \textsc{Cluster Subgraph}
and \textsc{Strong Triadic Closure}~\cite{abs-2001-06867,GolovachHKLP20}.

\iffalse
\subsection{Minimal background on delta-matroids}

We refer to Section~\ref{sec:prelim} for more background on
delta-matroids and delta-matroid terminology, but let us highlight a
few items here to support the presentation of results. First, in
addition to the standard representation of a linear delta-matroid
$V=(V,\cF)$ as $D=\bD(A) \Delta S$ mentioned above, there is an equivalent
representation via \emph{contraction}, introduced by Koana and Wahlström~\cite{KW24}.
For a delta-matroid $D = (V, \cF)$, its contraction by a set $S \subseteq V$ is   
defined as 
\begin{align*}
  D / S = (V \setminus S, \{ F \setminus S \mid F \in \mathcal{F}, S \subseteq F \}),
\end{align*}
A contraction representation of $D$ is obtained by starting with a skew-symmetric matrix $A$ indexed by $V \cup T$, where $T$ is an auxiliary set.
In this representation, we express the delta-matroid as $D = \bD(A) / T$, 
so that the resulting delta-matroid has ground set $V$.
The contraction representation is often more convenient
for algorithmic purposes.
A twist representation can be converted into a contraction representation 
in polynomial time, and vice versa \cite{KW24}.
We also consider \emph{projected} linear delta-matroids
$D=D'|X$ where $D'=(V \cup X,\cF')$ is a linear delta-matroid;
this is a strict generalization of linear delta-matroids.
In fact, Koana and Wahlström~\cite{KW24} showed that every
projected linear delta-matroid can be represented as an
\emph{elementary} projection where $|X| \leq 1$. Thus, any algorithmic
result on linear delta-matroids generally carries over to projected
linear delta-matroids, simply by guessing the behaviour of the
projected set $X$. 

Let $M=(V,\cI)$ be a linear matroid represented by a
matrix $A$ of full row rank, and let $\cB \subseteq 2^V$ be its set of
bases. Let
\[
  A' = \kbordermatrix{
    & T & V \\
    T&O & A \\
    V&-A^T & O \\
  }
\]
be a skew-symmetric matrix indexed by $V \cup T$. 
Then the \emph{basis delta-matroid} $D=(V,\cB)$ of $M$ 
is $D=\bD(A')/T$ and the \emph{independence delta-matroid}
$D=(V,\cI)$ of $M$ is $D=\bD(A')|T$.
A \emph{twisted matroid} is a twisting $D \Delta S$ of the basis
delta-matroid of a matroid.

One of the most basic building blocks is the following. 
For $n \in \N$, let $V=\{u_i,v_i \mid i \in [n]\}$
and let $\Pi=\{\{u_i,v_i\} \mid i \in [n] \}$ be
a partition of $V$ into pairs. 
The \emph{pairing delta-matroid} $D = (V, \cF)$ associated with $\Pi$ is defined as the matching delta-matroid of the graph whose edges correspond exactly to the pairs in $\Pi$.  
Intuitively, a set $F \subseteq V$ is feasible if and only if it does not ``break'' any pair -- meaning that for each pair $\{u_i, v_i\}$, either both $u_i$ and $v_i$ are included in $F$ or both are excluded.  
This construction naturally arises in combinatorial problems.

Of particular interest in our applications is the \emph{dual matching delta-matroid}.
Given a graph $G=(V,E)$, a set $S \subseteq V$ is feasible in this delta-matroid
if and only if the induced subgraph $G-S$ (obtained by removing the vertices in $S$ from $G$) has a perfect matching. 

For more examples, we refer to Moffatt~\cite{Moffatt19deltamatroids}
and Koana and Wahlström~\cite{KW24}.

Let us also define the basic problems more precisely.
In \textsc{Delta-matroid Intersection}, the input is two
delta-matroids $D_1=(V,\cF_1)$ and $D_2=(V,\cF_2)$,
and the task is to find a set $F \in \cF_1 \cap \cF_2$.
In \textsc{Delta-matroid Parity}, the input is a delta-matroid
$D=(V,\cF)$ and a partition $\Pi$ of $V$ into pairs, and the task is
to find a set $F \in \cF$ that minimizes the number of pairs $p \in
\Pi$ broken by $F$ (i.e., with $|p \cap F|=1$).  
Note that these are equivalent: On the one hand, \textsc{Delta-matroid Parity}
is the intersection of $D$ and the pairing delta-matroid of $\Pi$;
on the other hand, \textsc{Delta-matroid Intersection}
reduces to \textsc{Delta-matroid Parity} by a disjoint union operation
on $D_1$ and $D_2$. Hence, both are intractable in general
but solvable in $O(n^{\omega+1})$ time deterministically~\cite{GeelenIM03}
and, very recently, $O(n^\omega)$ time randomly~\cite{KW24}.

In addition, one may impose a condition that $|F|$ is maximum.
This is a more difficult problem, raised by Kakimura and Takamatsu~\cite{KakimuraT14sidma},
only recently solved in randomized time $O(n^{\omega+1})$ on linear
and projected linear delta-matroids~\cite{KW24}.

\subsubsection{Contraction representation and other recent results}

A significant part of our results rely upon recent advancements on
representation of linear delta-matroids by Koana and
Wahlström~\cite{KW24}.

...

\fi
%%% old 1.1 block

\subsection{Delta-matroid intersection and packing problems}

\label{sec:ourresults}

In this paper, we investigate the parameterized complexity aspects of problems over linear delta-matroids.
We show two classes of results; we begin with delta-matroid versions
of basic intersection and packing problems. 


More precisely, we introduce the \textsc{$q$-Delta-matroid Intersection}
and \textsc{Delta-matroid Set Packing} problems,
generalizing \textsc{Delta-matroid Intersection} and \textsc{Delta-matroid Parity}.
Among others, we consider the following problems.
\begin{itemize}
\item \textsc{$q$-Delta-matroid Intersection}: Given $q$ 
  delta-matroids $D_1=(V,\cF_1)$, \ldots, $D_q=(V,\cF_q)$
  and an integer $k \in \N$, is there a set of cardinality $k$ that is
  feasible in $D_i$ for every $i \in [q]$?
\item \textsc{$q$-Delta-matroid Parity}: Given a delta-matroid $D=(V,\cF)$, 
  a partition $\cP$ of $V$ into blocks of size $q$, and an integer $k
  \in \N$, is there a feasible set that is the union of $k$ blocks
  from $\cP$? 
\item \textsc{Delta-matroid Set Packing}: Given a delta-matroid
  $D=(V,\cF)$, an arbitrary partition $\cP$ of $V$, and an integer $k
  \in \N$, is there a feasible set $F$ in $D$, $|F|=k$, which is the union of
  blocks from $\cP$?
\end{itemize}
We use the following shorthand: \textsc{DDD Intersection}
refers to \textsc{3-Delta-matroid Intersection};
\textsc{DDM Intersection} refers to the special case of
\textsc{3-Delta-matroid Intersection} where $D_3$ is a matroid;
and \textsc{D$\Pi$M Intersection} refers to the case where $D_1$ 
is an arbitrary delta-matroid, $D_2$ is a pairing
delta-matroid for a
partition $\Pi$, and $D_3$ is a matroid. In other words,
%partition $\Pi$ (i.e., a delta-matroid where a set is feasible if and only if it is a union of pairs), and $D_3$ is a matroid. In other words,
\textsc{D$\Pi$M Intersection} is \textsc{Delta-matroid Parity}
with an additional matroid constraint. Throughout, we assume that all
matroids and delta-matroids are linear (or projected linear)
with representations given in the input over some common field $\F$. 

The classical problems \textsc{Linear Delta-matroid Intersection} and 
\textsc{Linear Delta-matroid Parity} correspond to the special case
where $q=2$ and where there is in addition no cardinality constraint $k$
on the solution, and was solved by Geelen et al.~\cite{GeelenIM03}.
For the variant where $q=2$ and the cardinality constraint is present,
only a randomized algorithm is known~\cite{KW24} (even if replace the
constraint $|F|=k$ by asking $|F|$ to be maximized). 

The matroid versions of these problems are all FPT for linear matroids
when parameterized by the solution cardinality $k$ for any $q$ (and more
generally FPT when parameterized by $k+q$)~\cite{Marx09-matroid,BrandDH18,EKW23}.
The most powerful method here, at least for representations over
fields of characteristic 2, is the recent method of
\emph{determinantal sieving}~\cite{EKW23}, which forms the inspiration 
for our algorithms; see below.
In particular, for matroids represented over a field of characteristic 2,
\textsc{$q$-Matroid Intersection} can be solved in $O^*(2^{(q-2)k})$ time,
\textsc{$q$-Matroid Parity} in $O^*(2^{qk})$ time, and
\textsc{Matroid Set Packing} in $O^*(2^k)$ time~\cite{EKW23}.
Results over other fields are somewhat slower, but still single-exponential.

For delta-matroids, we find a more intricate picture. The \emph{rank}
of a delta-matroid $D=(V,\cF)$ is the cardinality of a largest
feasible set in $D$. We consider the above problems both parameterized
by $k$ and by the rank $r$. For matroids, this distinction makes no
difference -- given a linear matroid $M=(V,\cI)$ of rank $r$ and a
parameter $k$, via a \emph{truncation} operation one can construct
a linear matroid $M'=(V,\cI')$ of rank $k$, which retains all
independent sets of cardinality at most $k$ from $M$. 
Furthermore, this can be done efficiently with a maintained linear
representation~\cite{Marx09-matroid,LokshtanovMPS18TALG}. 
Thus, over matroids, parameterizing by $r$ is no easier than
parameterizing by $k$. However, for delta-matroids, the distinction
turns out to be highly significant. 

Parameterized by $k$, we find a dividing line where some problems,
including \textsc{DDM Intersection}, are FPT given representations
over a common field $\F$, whereas others, including \textsc{DDD Intersection},
are W[1]-hard. This holds even for very simple delta-matroids; see
Section~\ref{sec:part1}.

\begin{restatable}{theorem}{dmhardness} \label{hard:ddd}
  \textsc{DDD Intersection} is W[1]-hard parameterized by
  $k$ even for linear delta-matroids represented over the same field. 
\end{restatable}

On the other hand, if we parameterize by the rank $r$, then we recover
all positive cases from the matroid case, and with matching running times.

\begin{restatable}{theorem}{dmponefpt} \label{tractable:rank}
  The following problems are FPT over linear delta-matroids and
  matroids provided as representations over a common field. 
  \begin{itemize}
  \item \textsc{DDD Intersection} parameterized by $r=\rank(D_3)$,
    with a running time of $O^*(2^r)$ over characteristic 2
    and $O^*(2^{O(r)})$ otherwise
  \item More generally, \textsc{$q$-Delta-matroid Intersection} parameterized by $r$,
    where $r$ is the maximum rank of $D_i$, $i \geq 3$, with a running
    time of $O^*(2^{(q-2)r})$ over characteristic 2
    and $O^*(2^{O(qr)})$ otherwise
  \item \textsc{$q$-Delta-matroid Parity} and \textsc{Delta-matroid Set Packing}
    parameterized by the rank $r$ of the delta-matroid,
    with a running time of $O^*(2^r)$ over characteristic 2
    and $O^*(2^{O(r)})$  in general
  \item \textsc{DDM Intersection} and \textsc{D$\Pi$M Intersection},
    parameterized by the cardinality $k$, 
    with a running time of $O^*(2^k)$ over characteristic 2
    and $O^*(2^{O(k)})$  in general
  \item More generally, \textsc{Delta-matroid Intersection} and
    \textsc{Delta-matroid Parity} with an additional $q-2$ matroid constraints,
    parameterized by $q$ and the cardinality $k$, 
    with a running time of $O^*(2^{(q-2)k})$ over characteristic 2
    and $O^*(2^{O(qk)})$ in general
  \end{itemize}
  In all cases, the algorithms are randomized, and use polynomial
  space over characteristic 2 but exponential space otherwise.
\end{restatable}

Finally, we note a problem of \textsc{Colorful Delta-matroid Matching}
that will be used later in the paper. In this problem, we are
given a delta-matroid $D=(V,\cF)$, a graph $G=(V,E)$ 
with edges colored in $k$ colors, and an integer $k \in \N$.
The task is to find a feasible set $F \in \cF$, $|F|=2k$
such that $G[F]$ has a matching with $k$ distinct edge colors. 
This reduces to \textsc{D$\Pi$M Intersection}.

\begin{restatable}{theorem}{dmmatching} \label{corollary:colorful-matching}
  \textsc{Colorful Delta-matroid Matching} on linear delta-matroids
  can be solved in $O^*(2^k)$ time if the representation is over a
  field of characteristic~2 and in $O^*(2^{O(k)})$ time otherwise.
\end{restatable}

\paragraph*{Sieving methods and a brief technical comment.}
Due to the amount of required preliminaries, we leave all technical
discussion to Section~\ref{sec:part1}, but we want to highlight 
the main technical contribution from this part of our paper. 
As mentioned, the main technical tool in the fastest FPT algorithms
for problems over linear matroids is \emph{determinantal sieving}.
Let $P(X)$ be a multivariate polynomial over a field $\F$, and assume
that $P(X)$ is given as black-box evaluation access with the ability
to also evaluate $P(X)$ over an extension field of $\F$ as needed.
The \emph{support} of a monomial $m$ in $P(X)$ is the set of variables
with non-zero degree in $m$. 
The main technical statement is then as follows.

\begin{theorem}[Basis sieving~\cite{EKW23}]
  \label{theorem:basis-sieve}
  Let $P(X)$ be a polynomial of degree $d$ over a set of variables $X$ over a field $\F$ of characteristic 2
  and let $M=(X,\cI)$ be a linear matroid of rank $k$ represented over $\F$. 
  There is a randomized algorithm that in time $O^*(d2^k)$ and
  polynomial space tests if the monomial expansion of $P(X)$ contains
  a term which is multilinear of degree $k$ and whose support is a basis of $M$. 
\end{theorem}

For a quick illustration, let $A_1, A_2 \in \F^{k \times V}$ be two
linear matroids of rank $k$ over a ground set $V$. Recall that
by the Cauchy-Binet formula,
$\det A_1A_2^T=\sum_{B \in \binom{V}{k}} \det A_1[B] \det A_2[B]$.
By scaling the columns of $A_1$ by a set of variables, we can define
a polynomial $P(X)$ that ``enumerates'' (or, is a generating function for)
common bases of $A_1$ and $A_2$. Given a third linear matroid
represented by a matrix $A_3 \in \F^{k \times V}$, 
Theorem~\ref{theorem:basis-sieve} combined with the Schwartz-Zippel lemma
then immediately gives an $O^*(2^k)$-time
algorithm for \textsc{3-Matroid Intersection} for linear matroids
represented over a common field.

We show a generalization of Theorem~\ref{theorem:basis-sieve} to
linear delta-matroids. This generalization comes without any
additional cost to the running time (up to the polynomial factor)
compared to the above. For technical simplicity we assume that $P(X)$
is homogeneous and that the delta-matroid is given with a so-called
\emph{sparse representation}. Both of these assumptions can be made
without loss of generality; see Section~\ref{sec:dm-sieve}.

\begin{restatable}{theorem}{dmsieve} \label{theorem:sieving-char-2}
  Let $D=(V,\cF)$, $V=\{v_1,\ldots,v_n\}$ be a (projected) linear delta-matroid of rank $r$ given as a sparse representation $\bD(A) / T$,
  and let $P(X)$, $X=\{x_1,\ldots,x_n\}$ be a homogeneous polynomial of degree $k$
  given as black-box access over a field of characteristic 2 with at least $3k$ elements.
  In time $O^*(2^r)$ we can sieve for those terms of $P(X)$ which
  are multilinear and whose support $\{x_i \mid i \in I\}$
  in $X$ is such that $\{v_i \mid i \in I\}$ is feasible in $D$. 
\end{restatable}

Theorems~\ref{theorem:basis-sieve} and~\ref{theorem:sieving-char-2}
also have variants over fields of characteristic other than 2, except that
except (1) the process uses exponential space, (2) the process works
over an arithmetic circuit, not via black-box access, and (3) with
a higher running time of $O^*(2^{O(r)})$;
%the running time is higher, namely $O^*(2^{\omega r})$ where $\omega < 2.372$ is the
%matrix multiplication exponent in general, and $O^*(4^r)$ if the circuit is
%skew (see Section~\ref{sec:dm-sieve}).\todo{Same as above.}
see Section~\ref{sec:dm-sieve}.

Most of the algorithmic results in Theorem~\ref{tractable:rank} work
by applying Theorem~\ref{theorem:sieving-char-2} to a suitably
constructed polynomial $P(X)$ generalizing the Cauchy-Binet example
into linear delta-matroids. In fact, a good working summary of our
upper and lower bounds for packing and intersection problem is the
following: \emph{If a problem can be solved using Theorem~\ref{theorem:sieving-char-2} 
  then it is FPT; otherwise W[1]-hard.}
Thus, for example, \textsc{DDD Intersection} is hard with parameter $k$
since there is no guarantee that either of the three delta-matroids
$D_1$, $D_2$, $D_3$ has rank bounded by $k$, but it is FPT under the
parameter $r=\rank(D_3)$; and (e.g.) \textsc{DDM Intersection}
is FPT since the matroid layer can be truncated to rank $k$.
See Section~\ref{sec:part1} for details.

However, the second set of results in our paper is a counterpoint to
this, presenting a set of surprising algorithms that cannot be
explained purely algebraically. 

\subsection{Delta-matroid triangle cover with applications}
Inspired by applications in graph theory (see below) we study a
relaxation of \textsc{3-Delta-matroid Parity} which,
surprisingly, turns out to have an FPT algorithm.

Let $D=(V,\cF)$ be a linear delta-matroid and let
$\cT=\{T_1,\ldots,T_m\} \subseteq \binom{V}{3}$ be a set of
triples over $V$ (which we refer to as triangles).
We say that a set $S \subseteq V$ can be \emph{covered} by triangles
if there is a disjoint packing $\cT' \subseteq \cT$
such that $S \subseteq V(\cT')$. Note that we do not require
that $S$ is precisely partitioned by $\cT'$, i.e., we allow for
triangles $T_i \in \cT'$ such that $|T_i \cap S| \in \{1,2\}$.
Let $k \in \N$. The \textsc{Delta-matroid Triangle Cover} problem
asks to find a feasible set $F \in \cF$ with $|F|=k$ such that
$F$ can be covered by triangles. 

\begin{restatable}{theorem}{dmtc}
  \label{theorem:dmtc-fpt}
  \textsc{Delta-matroid Triangle Cover} can be solved in $O^*(k^{\Oh(k)})$ time
  over a linear delta-matroid.
\end{restatable}

Let us make a few observations. On the one hand, this problem generalizes
\textsc{Colorful Delta-matroid Matching}:
Let $(D=(V,\cF), G=(V,E), k)$ be an instance of the latter.
Introduce $k$ new vertices $z_1$, \ldots, $z_k$, and for
every edge $uv \in E$ of color $i$ create a triangle $\{u,v,z_i\}$.
Thus in particular \textsc{Delta-matroid Triangle Cover}
generalizes \textsc{Delta-matroid Parity}. 
On the other hand, the most obvious structural simplification
of \textsc{Delta-matroid Triangle Cover} would seem to be to assume 
that we are looking for a set $F \in \cF$ that is precisely
partitioned by triangles. However, this defines
\textsc{3-Delta-matroid Parity}, which is W[1]-hard.
Hence \textsc{Delta-matroid Triangle Cover} inhabits
an interesting spot of being ``just barely'' tractable.
Indeed, by the same reduction as above, the extended version
\textsc{Delta-matroid $K_4$ Cover}, where $F$ is to be covered
by sets of cardinality 4, generalizes
\textsc{3-Delta-matroid Parity} and is W[1]-hard.

Our algorithm for \textsc{Delta-matroid Triangle Cover} seamlessly combines combinatorial and algebraic techniques,
ultimately reducing the problem to \textsc{Colorful Delta-matroid Matching}.  
On the algebraic side, we introduce a new delta-matroid operation, termed \emph{$\ell$-projection},
which generalizes the classical matroid truncation operation.  
We believe that this operation is of independent interest, particularly for kernelization \cite{Wahlstrom24SODA}.
On the combinatorial side, we use arguments inspired by the sunflower lemma to effectively reduce triangles to pairs.  
For details, see Section~\ref{sec:triangle-cover}.

We give two graph-theoretic applications of this problem.
The first is to  \textsc{Cluster Subgraph}: given a graph $G$ and an integer $\ell$, the problem is to find a cluster subgraph (i.e., a graph whose connected components are cliques) with at least $\ell$ edges.
This problem is NP-hard \cite{ShamirST04}, but can be solved in $O^*(9^\ell)$ time \cite{GruttemeierK20}.
We consider a tighter \emph{above guarantee} parameterization of $\ell-\MM(G)$ where $\MM(G)$ is the matching number of $G$.
Note that a matching is a cluster subgraph. 
The parameterized complexity with respect to this parameterization is marked as an open question in a recent survey~\cite{abs-2001-06867}.

\begin{restatable}{theorem}{csam} \label{theorem:cs-am}
  \textsc{Cluster Subgraph} can be solved in $O^*(k^{O(k)})$ time for $k = \ell - \MM(G)$.
\end{restatable}

To see the connection to \textsc{Delta-matroid Triangle Cover},
consider the case that $G$ is $K_4$-free, and assume for simplicity
that $G$ has a perfect matching. Then the goal is to find a
packing of $K_2$'s and $K_3$'s with as many edges as possible, i.e.,
a collection of $K_3$'s whose deletion leaves a graph with a large
matching number.\footnote{
  We remark that the problem becomes polynomial-time solvable
  when the goal is to maximize the number of covered vertices \cite{HellK84}.
  However, the edge maximization problem is NP-hard as it generalizes the perfect triangle packing problem.
} We show that this reduces to solving
\textsc{Delta-matroid Triangle Cover} over the dual matching
delta-matroid of $G$. On the other hand, if $G$ contains a large
packing of $K_4$'s, then the instance is immediately positive. 
Thus, we may assume that $G$ has a set $K$ of $O(k)$ vertices
such that $G-K$ is $K_4$-free, and Theorem~\ref{theorem:cs-am}
follows by guessing its interaction with the solution. 

\begin{figure}
  \centering
  \begin{tikzpicture}
    \def\radius{1.2cm}
    \def\numvertices{7}
    \foreach \x in {1,...,\numvertices} {
      \pgfmathsetmacro{\angle}{360/\numvertices * (\x - 1)}
      \node[circle,fill,inner sep=2pt] (vertex\x) at (\angle:\radius) {};
    }
    \foreach \x in {1,...,\numvertices} {
      \pgfmathsetmacro{\angle}{360/\numvertices * (\x - 1)}
      \draw[very thick] (\angle:\radius) node[circle,fill,inner sep=2pt] {} -- (\angle + 360/\numvertices:\radius);
    }
    \foreach \x in {1,...,\numvertices} {
      \pgfmathtruncatemacro{\target}{mod(\x+1, \numvertices)+1}
      \pgfmathtruncatemacro{\source}{\x}
        \draw (vertex\source) to[bend left] (vertex\target);
    }
    % Label the vertices, if needed
    \foreach \x in {1,...,\numvertices} {
      \pgfmathsetmacro{\angle}{360/\numvertices * (\x - 1)}
      \node at (\angle:1.5cm) {\x};
    }
  \end{tikzpicture}
  \caption{A maximum strong set (thick lines) which does not induce a cluster graph.}
  \label{fig:strong-cycle}
\end{figure}

The second application is to \textsc{Strong Triadic Closure}, formalized by Sintos \cite{SintosT14}.
Let $G$ be a graph. We say that a 3-path $(u, v, w)$ is \emph{closed} if the edge $uw$ exists in $G$.
For a graph $G$, an edge set $S$ is said to be \emph{strong} if every 3-path in $S$ is closed. 
Note that the shortcut edge $uw$ is not required to exist in $S$, only in $G$.
Given $G=(V,E)$ and $\ell$, the \textsc{Strong Triadic Closure}
problem is to find a strong edge set $S \subseteq E$ with $|S| \geq \ell$. 
This is based on the notion of triadic closure \cite{granovetter1973strength}.
This problem is closely related to \textsc{Cluster Subgraph} as an edge set that induces a cluster graph is strong.
However, not every strong set induces a cluster subgraph (see Figure~\ref{fig:strong-cycle}).
It is known to be NP-hard~\cite{SintosT14} and FPT when parameterized by $\ell$~\cite{GolovachHKLP20,GruttemeierK20}
with the fastest algorithm running in $O^*(\ell^{O(\ell)})$ time~\cite{GruttemeierK20}.
Again, we consider the \emph{above matching} parameterization. 

Again, the $K_4$-free case turns out to be tractable: if $G$ is $K_4$-free, then the
solution $S$ must be a collection of strong paths and cycles, and
in fact it can be shown that there exists an optimum where $S$ consists
only of strong cycles and $K_2$'s. Similarly to above, we solve the
problem by looking for a feasible set $F$ in the dual matching
delta-matroid of $G$ such that $F$ can be covered by a collection of
strong cycles, possibly of length $\Omega(k)$. 
However, this is significantly more complicated -- as mentioned before, the \textsc{Delta-matroid $K_c$ Cover} is W[1]-hard for any $c \ge 4$.
Yet, we still manage to give a reduction to \textsc{Colorful Delta-matroid Matching}.
To that end, we identify several cases where a strong cycle can be replaced by triangles.
We then argue that when this triangle replacement fails, pairing constraints (which can be encoded in \textsc{Colorful Delta-matroid Matching}) are powerful enough to solve the problem.

\begin{restatable}{theorem}{tcamfpt} \label{theorem:tc-am-fpt}
  \textsc{Strong Triadic Closure} on $K_4$-free graphs can be solved in $O^*(k^{O(k)})$ time for $k = \ell - \MM(G)$.
\end{restatable}

Furthermore, as above, a $K_4$ in $G$ implies an immediate contribution to $\ell-\MM(G)$,
and we can find a small modulator $K$ such that $G-K$ is $K_4$-free.
Unfortunately, despite our efforts, achieving a full FPT algorithm remained elusive -- in fact, guided by the very precise 
structure of solutions where our algorithmic insights prove ineffective, we
show that the problem is W[1]-hard in general.
Intriguingly, the source of hardness contradicts our initial beliefs, stemming not from strong cycles but from an unexpectedly different aspect.

\begin{restatable}{theorem}{tcamhard}
  \textsc{Strong Triadic Closure} is W[1]-hard when parameterized by $k = \ell - \MM(G)$.
\end{restatable}

On the other hand, a counting argument shows that, although we do not get an FPT
algorithm, we can get a constant-factor FPT approximation simply by
packing a modulator $K$ and ignoring all edges $\delta(K)$ between
$K$ and $G-K$ in the solution, solving $G[K]$ (exhaustively)
and $G-K$ separately. That is, we give an algorithm that either
shows that $G$ has no strong edge set of size $\MM(G)+k$
or finds a strong edge set of size $\MM(G) + \Omega(k)$ in FPT time.
% Such a set cannot be found in polynomial time (unless P=NP), since
% ultimately, in very dense graphs, finding a maximum strong edge set
% implies finding a large clique, for which there are very strong
% inapproximability bounds. 

\begin{restatable}{theorem}{tcamapprox}
  There is $O^*(k^{O(k)})$-time algorithm for \textsc{Strong Triadic Closure} that approximates the above matching parameter $\ell - \MM(G)$ within a  factor 7.
\end{restatable}

In addition, we generalize a result of Golovach et al.~\cite{GolovachHKLP20}, who gave an FPT algorithm for maximum degree 4,  
by showing that \textsc{Strong Triadic Closure} can be solved in time $O^*(\Delta^{O(k)})$
for graphs of maximum degree $\Delta$:

\begin{restatable}{theorem}{tcamdelta}
  \textsc{Strong Triadic Closure} can be solved in $O^*(\Delta^{O(k)})$ time, where $\Delta$ is the maximum degree of $G$ and $k = \ell - \MM(G)$.
\end{restatable}

Lastly, we remark that the question whether \textsc{Strong Triadic Closure} is FPT has
been asked repeatedly in the literature~\cite[Open Problem 5.4]{abs-2001-06867},~\cite{GolovachHKLP20}.
The connection to \textsc{Cluster Subgraph} has also been observed~\cite{GruttemeierK20}. 
We settle the status of both of these problems. 

\subsection{Further related work}

Given a graph, the problem of finding a vertex-disjoint $\{K_2, K_3\}$-packing that covers the maximum number of vertices can be solved in polynomial time \cite{HellK84}.
In contrast, maximizing the number of edges in such a packing, as discussed above, is NP-hard since it generalizes the problem of partitioning a graph into triangles.
Moreover, this problem is APX-hard and admits a $3/2$-approximation \cite{ChataignerMWY09}.

There is a range of algebraic algorithms for FPT problems, beyond
what has been surveyed here. Much of this is surveyed by Eiben et
al.~\cite{EKW23}. For instance, the basic \textsc{$q$-Set Packing}
and \textsc{$q$-Dimensional Matching} problems parameterized by cardinality
are currently best solved by reduction to algebraic sieving arguments~\cite{BjorklundHKK17narrow},
and it is an intriguing question what the best possible running time
for these problems is.

Other intriguing connections for delta-matroids includes their
generalization to \emph{jump systems}~\cite{BouchetC95}, with 
connections to discrete convexity theory (cf.~\cite{murota2021note}).
In another direction, Bouchet generalized delta-matroids into
\emph{multi-matroids}, with their own notion of representations~\cite{BouchetMMI,BouchetMMIV}.

For further related work, we again refer to Moffatt~\cite{Moffatt19deltamatroids} and
Koana and Wahlström~\cite{KW24}.

%%% For git merge purposes -- changed to "\iffalse" to drop 1.4 from the paper
\iffalse
\subsection{Technical overview}

\paragraph*{Part 1: Intersection and packing problems.}

The FPT results in the first part of the paper (Section~\ref{sec:part1}) 
build entirely on algebraic methods, honed to apply to linear delta-matroids.
We base the algorithm on the method of determinantal sieving,
recently defined by Eiben et al.~\cite{EKW23}. 
This result leads to e.g., an FPT algorithm for \textsc{DDD-Intersection} parameterized by the rank (Theorem~\ref{tractable:rank}).

As mentioned, determinantal sieving is a method that filters a given
multivariate polynomial $P(X)$ in $n$ variables $X$ for monomials
$m=\prod_{i \in I} x_i$, $I \subseteq [n]$ whose support $I$ forms a
basis of a given linear matroid $M=(X,\cI)$ over the variables,
represented by a matrix $A \in \mathbb{F}^{k \times X}$.
Intuitively, the method multiplies each monomial by the determinant of the submatrix corresponding to its support, so that non-basis monomials are filtered out.
More precisely, for a homogeneous polynomial 
$P(X) = \sum_{m} P(m) \cdot m$,
where the sum is over all monomials and $P(m)$ denotes the coefficient of $m$, one can construct the polynomial
\begin{align*}
  Q(X) = \sum_{m} P(m) \cdot \det A[\cdot, m] \cdot m,
\end{align*}
where $\det A[\cdot, m]$ denotes the determinant of the submatrix formed by taking the corresponding copies of the columns of $A$.  
The computation of $Q(X)$ can be performed as a summation of $2^k$ evaluations of $P(X)$.
Since $\det A[\cdot, m]$ vanishes whenever $m$ is not multilinear, evaluating $Q(X)$ at randomly chosen coordinates (via the Schwartz-Zippel lemma) recovers the desired basis sieving result (Theorem~\ref{theorem:basis-sieve}).  

To generalize the result to the delta-matroid setting, we need a few
ingredients.
The first concept involves a  \emph{sparse representation} 
of linear delta-matroids of bounded rank.
Let $D=\bD(A)\Delta S$ be a linear delta-matroid of rank $r$ in twist representation.
Initially, it may be that the matrix $A$ has $\Theta(n^2)$ non-zero entries.
However, by pivoting $A$ by $F_{\max} \Delta S$, where $F_{\max}$ is a maximum feasible of $A$, one obtains a matrix $A^*$ such that $D = \bD(A^*) \Delta F_{\max}$ and $A^*[V \setminus F_{\max}] = O$.
When transformed into its contraction representation as described by Koana and Wahlström~\cite{KW24}, 
the linear delta-matroid $D=(V,\cF)$ of rank $r$ can be represented as
$D=\bD(A)/T$ where $A$ has the form
\[
  A = \kbordermatrix{
    & T & V \\
    T&B & C \\
    V& -C^T & O \\
  }
\]
where $|T|=r$.
This structure is similar to the contraction representation 
of a linear matroid surveyed above, except that $A[T]$ would
generally be non-zero. Consequently, the Pfaffian $\Pf A[T \cup F]$
for $F \subseteq V$ for a matrix of this form can be written as
\begin{equation}
  \Pf A[T \cup F] = \sum_{U \subseteq T: |U|=|T|-|F|} \sigma_{U} \Pf B[U] \cdot \det C[T \setminus U, F],
\end{equation}
where $\sigma_U = \pm 1$ is the sign of a permutation.
This is the consequence of a formula on the Pfaffian of the sum of two skew-symmetric matrices (see Lemma~\ref{lemma:sum-pf}).
We believe that this sparse representation is of independent interest.  
They seem to be useful, particularly when analyzing algorithmic runtimes with respect to the rank.

Note that each summand is a product of a Pfaffian and a determinant rather than a product of two Pfaffians (cf. Lemma~\ref{lemma:det-pf}).  
Additionally, the Pfaffian part does not depend on $F$.  
Thus, a delta-matroid sieving can be achieved by invoking determinantal sieving for each subset $U$ of $T$.  
More precisely, we apply the determinantal sieving method of Eiben et al.~\cite{EKW23} on the matrix $C[T \setminus U, \cdot]$, which transforms each monomial $m$ into $\det C[T \setminus U, m] \cdot m$.  
By summing the results with weights $\sigma_U \Pf B[U]$ over all $U \subseteq T$, we obtain the desired polynomial, where each monomial $m$ is turned into $\Pf A[T \cup m] \cdot m$.
This summation can be computed in 
$O^*(\sum_{U \subseteq T}  \binom{|T|}{|U|} \, 2^{|T| - |U|}) = O^*(3^r)$
time, where $r=|T|$, in the case of characteristic 2.

We improve over this observation to obtain a $O^*(2^r)$-time algorithm for characteristic~2, matching the running time of determinantal sieving.
For characteristic~2, the sign term of the Pfaffian disappears and $\Pf A[T \cup F]$ is simply the sum of perfect matchings in the support graph of $A[T \cup F]$ (with suitable coefficients from the entries of $A$).  
For this case, the algorithm follows from multilinear sieving: introduce a variable $y_t$ for every $t \in T$, and let 
$$
P_T(Y)=\sum_{U \subseteq T: |U|=r-k} \prod_{t \in U} y_t \Pf B[U].
$$  
Our key finding is that this expression can be evaluated as the square root of a specific coefficient of the characteristic polynomial of $B$, effectively compressing the summation over subsets into a single coefficient extraction.  
(This approach appears to be limited to characteristic~2.)  
We can now sieve for the coefficient of $\prod_{t \in T} y_t$ in $P_T(Y) \cdot P(X)$ with a suitable evaluation of $X$, as in~\cite{EKW23}.  
(The details are omitted here.)
This results in the summation over all bipartitions $(U, T \setminus U)$ of $T$, thereby avoiding the need to compute $\Pf B[U]$ for every subset $U \subseteq T$ explicitly.

Over other characteristics, we use the exterior algebra and extensor coding to compute $\det A[T \cup F]$ rather than $\Pf A[T \cup F]$.  
Via this approach we obtain $\det C[T \setminus U, F] \cdot \det C[T \setminus U', F]$ for all subsets $U, U' \subseteq T$ of size $|T| - |F|$, which can be incorporated into the terms of 
$$
\det A[T \cup F] = (\Pf A[T \cup F])^2.
$$  
This method, while conceptually similar, incurs a higher cost in both exponential space and running time.

Note that in the case that $|F|=|T|$, so that $F$ (if feasible) is a basis of $D$, the formula reduces to just a determinant computation.  
This agrees with the fact that the maximum feasible sets of a delta-matroid form the bases of a matroid, providing a consistency check for our generalization.

It appears that, at least for FPT purposes, feasible sets of linear delta-matroids of bounded rank are ``close'' in expressive power to bounded-rank linear matroids.  
(However, we are not aware of a formal version of this statement.  
For example, trying to pad a sought-after feasible set $F$ into a basis of $D$ and then modifying $D$ so that $F$ becomes a basis fails to work, since it does not retain the difference between sets that are feasible in $D$ and subsets of bases of $D$.)

Given such a sieving procedure, we only need a suitable polynomial to sieve over.  
Using the results of Koana and Wahlström~\cite{KW24}, it is easy to construct a skew-symmetric matrix whose Pfaffian is an enumerating polynomial (in the terminology of~\cite{EKW23}) for solutions to an instance of \textsc{Delta-matroid Intersection}, given that the input delta-matroids are represented over the same field.  
All FPT results in Section~\ref{sec:part1} follow from this construction.



\paragraph*{Part 2: Triangle cover and applications.}

The algorithm for \textsc{Delta-matroid Triangle Cover}
is a mixture of algebraic and combinatorial arguments.

On the algebraic side, we introduce a delta-matroid operation we refer
to as \emph{$\ell$-projection}, which we believe to be of independent interest, possibly for kernelization \cite{Wahlstrom24SODA}. Let $D=(V,\cF)$ be a delta-matroid,
$X \subseteq V$, and $\ell \in \N$. The $\ell$-projection of $D$ to $V'=V \setminus X$
is the delta-matroid $D|_\ell X = (V', \cF')$ where 
$\cF'=\{F \setminus X \mid F \in \cF, |F \cap X|=\ell\}$. 
Essentially, a set $F'$ is feasible in $D|_\ell X$ if it can be extended to a feasible set in $D$ by adding a subset of $X$ with $\ell$ elements.
We show that linear delta-matroids are preserved by $\ell$-projection.
This operation generalizes $k$-truncation of linear matroids:
If $D=\bD(A)/T$ is the contraction representation of the basis
delta-matroid of a linear matroid $M$ using the simple construction
surveyed above, then $D'=\bD(A)|_k T$ is the basis delta-matroid of
the $k$-truncation of $M$. 
Note, however, an intricacy: there is no ``truncation'' operation for delta-matroids.
The $\ell$-projection can be achieved  by delta-sum or, alternatively using the Ishikawa-Wakayama formula.

Our algorithm for \textsc{Delta-matroid Triangle Cover} is based on the following sunflower-like argument.
Let $D=(V,\cF)$, $\cT$ and $k$ be an input instance
and let $v \in V$. If there is a packing of $\Omega(k)$ triangles
in $\cT$ which pairwise intersect only in $v$ (we say that $v$ is \emph{flowery} if this holds), then for any feasible
set $F \in \cF$ with $v \in F$, and any triangle cover $\cT'$ of $F-v$,
unless $v \in V(\cT')$ already, we will always be able to select a
triangle $T \in \cT$ from the packing to cover $v$ disjoint from $\cT'$. 
By applying $\ell$-projection, this allows us to effectively exclude a flowery vertex from the delta-matroid feasibility condition and retain it purely for the triangle-packing structure.
This simplification reduces triangles to colored pairs -- that is, a triangle $\{u,v,w\}$ where $v$ is flowery and removed from $D$ by $\ell$-projection reduces to an edge $\{u,w\}$ of ``color'' $v$.
This makes the problem computationally more tractable:
While \textsc{3-Delta-matroid Parity} is NP-hard and W[1]-hard with respect to solution size (Theorem~\ref{hard:ddd}),
\textsc{2-Delta-matroid Parity} is polynomial-time solvable and its colorful variant being FPT parameterized by solution size (Theorem~\ref{corollary:colorful-matching}).
Note that this argument is not applicable to the \textsc{3-Delta-matroid Parity} problem,
as in that version, we cannot freely add a triangle $T$ to the solution where only one vertex of $T$ is in the feasible set. 
%as other elements of $T$ may intersect $\mathcal{T}'$.

For non-flowery vertices $v$, we consider a graph $G_v$ from all triangles covering $v$,
where every triangle $\{u,v,w\} \in \cT$ becomes an edge $\{u,w\}$ in $G_v$.
Then, $G_v$ has a small vertex cover $S_v$ of size $O(k)$ by the definition of flowery vertices.
In any feasible solution $F \in \cF$ with $v \in F$,
the triangle $T$ covering $v$ must intersect $S_v$.
As $S_v$ has size $O(k)$, we can guess one vertex $s(v) \in S_v$ such that if  $v$ is covered by a triangle $T$, then $s(v) \in T$.

However, to implement the above ideas, we must determine whether $v \in F$, 
which we cannot afford to guess.
To address this, we formalize these ideas using randomized arguments as follows.
We guess
the number of triangles $p \leq k$ that the solution will use 
and color-code the ground set into $p \times 3$ colors as $V=V_{1,1} \cup \ldots \cup V_{p,3}$.
With a success probability of $1/k^{O(k)}$, it can be assumed that the triangle cover of the solution takes the form 
$\cT'=\{T_1,\ldots,T_p\}$ where $T_i \in V_{i,1} \times V_{i,2} \times V_{i,3}$
for every $i \in [p]$.
For each $(i, j) \in [p] \times [3]$, we guess whether $v_{i,j}$ is part of the feasible set, and whether it is flowery.
For non-flowery vertices $v_{i,j}$, we randomly select an element $s(v)$ from the aforementioned small vertex cover $S_v$.
This results in one of the seven
``behaviours'' of $T_i$, as shown in Figure~\ref{fig:triangle-cover}.
For example in (5) of Figure~\ref{fig:triangle-cover}, for $T_i=\{v_{i,1},v_{i,2},v_{i,3}\}$,
we may guess that $v_{i,1}$ is flowery and $v_{i,2}$, $v_{i,3}$ are not,
that $s(v_{i,2})=v_{i,3}$ and $s(v_{i,3})=v_{i,2}$, and that $T_i \subseteq F$.
For each of the behaviors, 
we then proceed with the following principles.
If one of the vertices in $T_i$, say the vertex in $V_{i,j}$, is flowery and is part of $F$,
then the identity of $F \cap V_{i,j}$ in the solution does not matter.
In this case, we replace $D$ by the $1$-projection of $D$ to $V \setminus V_{i,j}$ and reduce $k$.
This way the triplet constraint simplifies to a pairing constraint, which can be formulated using the $s(v)$-information.
In the aforementioned example of (5), we simply add a pairing constraint between $v_{i,2}$ and $v_{i,3}$.
Moreover, when the triangle does not involve any flower vertex, we can also use the $s(v)$-information to coordinate choices of triangles.
For example, if $s(v_{i,1}) = v_{i,3}$, $s(v_{i,2}) = v_{i,1}$, $s(v_{i,3}) = v_{i,2}$, and $T_i \subseteq T$, then we apply $3$-projection to $V \setminus V_{i}$ (this corresponds to (6) in Figure~\ref{fig:triangle-cover}).
As shown in Section~\ref{sec:triangle-cover}, in each of the seven behaviors, 
the triangle are reduced to pairs.
This allows us to reduce the problem to
\textsc{Colorful Delta-matroid Matching}, which is shown to be FPT in
Section~\ref{sec:part1}.
Essentially, our method integrates a combinatorial argument, inspired by the sunflower lemma, with the algebraic delta-matroid $\ell$-projection approach, to reduce to the algebraically solvable  \textsc{Colorful Delta-Matroid Matching} problem.

\paragraph*{Cluster subgraph.}
Let $(G=(V,E),\ell)$ be an input to \textsc{Cluster Subgraph}. As discussed
above, we first show that if $G$ is $K_4$-free and has a perfect matching,
then \textsc{Cluster Subgraph} with parameter $\ell-\MM(G)$
reduces to \textsc{Delta-matroid Triangle Cover}
where the delta-matroid is the dual matching delta-matroid of $G$.

Furthermore, let $K$ be a $K_4$ in $G$. Then $K$ contains 6 edges, and
even if using $K$ in a solution reduces the matching number of $G-K$
by the maximum value 4, we still count a gain of +2 over $\MM(G)$. 
Thus, we greedily pack vertex-disjoint $K_4$'s. Either we find $k/2$
of them, and the input is a yes-instance, or we have a modulator $K \subseteq V(G)$
with $|K| \leq 2k$ such that $G-K$ is $K_4$-free.
We can guess the interaction of $K$ with the solution.
We can also handle the case that $\MM(G)<n/2$ by an alternative
construction of linear delta-matroid $D$ (effectively, we 
add vertices to $G$ such that $G$ gets a perfect matching, construct
the dual matching delta-matroid of the padded graph, then apply
$\ell$-projection to the padding vertices with a suitable value of $\ell$
to get a linear delta-matroid over ground set $V$ that captures the
matching aspect of the solution in $G$). 

\paragraph*{Strong triadic closure.}
The case of \textsc{Strong Triadic Closure} is similar in spirit to
the above, but significantly more involved in the details.
First, let $(G=(V,E),\ell)$ be an input instance and assume that $G$
is $K_4$-free and has a perfect matching. Then a strong edge set $S
\subseteq E$ can contain at most two edges incident to every vertex $v \in V$,
since otherwise the closed neighbourhood of $v$ in $S$ contains a $K_4$ in $G$. 
Thus any solution is a collection of strong paths and cycles
(i.e., paths and cycles whose edge set is a strong edge set). 
By reconfiguration arguments, we can further show that the solution
can be assumed to be a collection $\cC=\{C_1,\ldots,C_p\}$ of strong cycles 
such that $V(\cC)$ has rank at least $2k$ in the dual matching
delta-matroid $D$ of $G$, and such that $|V(C)| \leq 6k$. 
That is, we have reduced our problem to finding a feasible set $F$ in
a linear delta-matroid $D$ that can be covered by a collection of strong cycles. 

To solve this problem, we aim for a solution with a maximum number of
triangles in $\cC$. In particular, every strong cycle $C$ can be
decomposed into a set of triangles that covers all but at most two vertices of $C$. 
If $F \subseteq V(\cC)$ is the optimal feasible set in $D$, and if 
$C$ can be decomposed into triangles in a way that does not lose any
vertex of $V(C) \cap F$, then we may assume that $\cC$ contains these
triangles instead of $C$. Similarly, if $v \in V(C) \cap F$ is
flowery, then it is safe to ``lose'' $v$ in the decomposition 
and replace $C$ in $\cC$ by a new triangle covering $v$. 

Repeating such reconfiguration arguments combined with color-coding
gives us enough of a structural handle on the solution that we can
reduce the problem to \textsc{Colorful Delta-matroid Matching}
by arguments generalizing the case of \textsc{Delta-matroid Triangle Cover}
above. Thus \textsc{Strong Triadic Closure} parameterized above matching
is FPT on $K_4$-free graphs. 

As above, we can then greedily pack a modulator $K$ such that
$|K|=O(k)$ and $G-K$ is $K_4$-free. However, for the W[1]-hardness
proof, we find (surprisingly) that the algorithm fails to extend to
this case due to some very specific troublesome interactions between
the modulator $K$ and a packing of strong cycles or triangles in $G-K$. 
We leverage these interactions to produce a proof of W[1]-hardness.

Finally, a counting argument shows that, although we do not get an FPT
algorithm, we can get a constant-factor FPT approximation simply by
packing a modulator $K$ and ignoring all edges $\delta(K)$ between
$K$ and $G-K$ in the solution, solving $G[K]$ (exhaustively)
and $G-K$ separately. That is, we give an algorithm that either
shows that $G$ has no strong edge set of size $\MM(G)+k$
or finds a strong edge set of size $\MM(G) + \Omega(k)$ in FPT time.
Such a set cannot be found in polynomial time (unless P=NP), since
ultimately, in very dense graphs, finding a maximum strong edge set
implies finding a large clique, for which there are very strong
inapproximability bounds. 
\fi

\iflong
\paragraph*{Organization.} Section~\ref{sec:prelim} covers additional
background material on delta-matroids. Section~\ref{sec:part1}
contains the first set of results, regarding intersection and packing
problems as well as the delta-matroid sieving result.
Section~\ref{sec:triangle-cover} gives the FPT algorithm for
\textsc{Delta-matroid Triangle Cover}, and
Sections~\ref{sec:cs-above-matching} and~\ref{sec:tcam}
cover the \textsc{Cluster Subgraph} and \textsc{Strong Triadic Closure}
applications, respectively. We wrap up in Section~\ref{sec:conclusions}
with conclusions and open question.
\fi

\iflong
\paragraph*{Preliminaries.} 
A parameterized problem is a collection of pairs $(I, k)$, where $I$ is the instance and $k$ is called the parameter, and
is \emph{fixed-parameter tractable} (in FPT) if there is an \emph{FPT algorithm}.
Such an algorithm solves every instance $(I,k)$ in $f(k) \cdot |I|^{O(1)}$ time, for some computable function $f$.
This complexity is denoted by $O^*(f(k))$.
For a comprehensive exposition on parameterized complexity, refer to the book of Cygan et al.~\cite{CyganFKLMPPS15PCbook}.

We will use standard notation for graphs as outline in Diestel \cite{DiestelBook}.
By default, graphs are assumed to be undirected. 
For a graph $G$, we often denote the sets of its vertices and edges by $V(G)$ and~$E(G)$, respectively.
The vertex and edge sets of a graph $G$ are denoted by $V(G)$ and $E(G)$, respectively, with $n = |V(G)|$ and $m = |E(G)|$.
We may abbreviate an edge $\{ u, v \}$ in an undirected graph as $uv$.
Similarly, in directed graphs, an arc from $u$ to $v$ is represented by $(u, v)$ or $uv$.
A vertex $v\in V(G)$ has neighborhood $N_G(v)=\{u \mid uv \in E\}$, and degree
$|N_G(v)|$.
For any subset $V'\subseteq V$, $G[V']$ denotes the \emph{induced subgraph} of $G$ on $V'$, that is, $G[V']=(V',E')$ where $E' = \{ uv \in E \mid u \in V, v \in V \}$.
The notation $G - V'$ refers to the induced subgraph $G[V \setminus V']$.

Given a matrix $A$ and a set of rows $S$ and columns $T$, $A[S, T]$ denotes the submatrix containing rows $S$ and columns $T$.
If $S$ or $T$ contains all rows or columns, respectively, we use the notation $A[\cdot, T]$ and $A[S, \cdot]$, where $\cdot$ denotes all rows or columns.
The $n \times m$ zero matrix and the $n \times n$ identity matrix is denoted by $O_{n \times m}$ and $I_{n}$, with the subscripts omitted when the dimension are clear from context.

We assume that each field in this paper comprises $\exp(n^{O(1)})$ many elements.
Under this assumption, all arithmetic operations can be performed in polynomial time.
The Schwartz-Zippel lemma \cite{Schwartz80,Zippel79} states that a polynomial $P(X)$ of total degree $d$ or less becomes nonzero with a probability of at least $1 - d / |\F|$, when evaluated at uniformly random elements from $\F$, provided $P(X)$ is not identically zero.

\fi

\section{Background on delta-matroids} \label{sec:prelim}

\subsection{Preliminaries on delta-matroids and skew-symmetric matrices} 

For two sets $A, B$, we let $A \Delta B=(A \setminus B) \cup (B \setminus A)$ denote their symmetric difference. 

\paragraph*{Delta-matroids.}

A \emph{delta-matroid} is defined as a pair $D=(V,\cF)$ where $V$ is a ground set
and $\cF \subseteq 2^V$ is a collection of \emph{feasible sets}.
The feasible sets adhere to the following \emph{symmetric exchange axiom}:
\[
\forall A, B \in \cF,\, x \in A \Delta B\, \exists y \in A \Delta B :
A \Delta \{x,y\} \in \cF.
\]
For a delta-matroid $D=(V,\cF)$, let $V(D)=V$ and $\cF(D)=\cF$. 
A delta-matroid is called \emph{even} if all feasible sets share the same
parity. 
This implies $x \neq y$ in the application of the
symmetric exchange axiom, although this is not required generally.

Delta-matroids also feature operations that extend beyond traditional matroid theory. Specifically, the \emph{twisting} of a delta-matroid 
$D=(V,\cF)$ by $S \subseteq D$ is the delta-matroid
$D \Delta S = (V, \cF \Delta S)$, where
$
  \cF \Delta S = \{F \Delta S \mid F \in \cF\}.
$
The \emph{dual} delta-matroid of $D$ is $D \Delta V(D)$.
The \emph{deletion} of $S \subseteq V$
yields a delta-matroid
$
  D \setminus S = (V \setminus S, \{F \in \cF \mid F \subseteq V \setminus S\}),
$
and the \emph{contraction} by $S$ is defined by
$
  D/S = (D \Delta S) \setminus S = (V \setminus S, \{F \setminus S \mid F \in \cF, S \subseteq F\}).
$

\paragraph*{Skew-symmetric matrices.}

A square matrix $A$ is \emph{skew-symmetric} if $A = -A^T$.
Particularly, when $A$ is over a field of characteristic 2, it is assumed
 that it has zero diagonal, unless stated otherwise.
For a skew-symmetric matrix $A$, whose rows and columns are indexed by a set $V = [n]$, 
the \emph{support graph} of $A$ is the graph $G=(V, E)$ where $E$ is the set of edges $\{ uv \mid A[u, v] \ne 0 \}$.
A fundamental tool in the analysis of skew-symmetric matrices is the
\emph{Pfaffian}, which is defined for a skew-symmetric matrix $A$ as follows:
\begin{align*}
  \Pf A = \sum_{M} \sigma(M) \prod_{e \in M} A[u, v],
\end{align*}
where the sum is taken over all perfect matchings of $A$'s support graph, and 
$\sigma(M) \in \{1,-1\}$ denotes the sign of the permutation:
\[
  \begin{pmatrix}
    1 & 2 & \cdots n - 1 & n \\
    v_1 & v_1' & \cdots v_{n/2} & v_{n/2}'
  \end{pmatrix},
\]
where $M = \{ v_i v_i' \mid i \in [n/2] \}$ with $v_i < v_i'$ for all $i \in [n/2]$.
It is well-known that $\det A=(\Pf A)^2$, so $\Pf A \neq 0$
if and only if $A$ is non-singular.
However, directly working with the Pfaffian is often more effective. 
The Pfaffian can be computed in $O(n^3)$ steps by an elimination procedure.
Alternatively, a combinatorial algorithm for computing the Pfaffian is also known \cite{MahajanSV99}.
The Pfaffian generalizes the notion of determinants as follows.
\begin{lemma} \label{lemma:det-pf}
For an $n \times n$-matrix $M$, it holds that
\begin{align*}
 \det M = (-1)^{\binom{n}{2}} \Pf \begin{pmatrix}
  O & M \\
  -M^T & O \\
\end{pmatrix}.
\end{align*}
\end{lemma}

An essential operation on skew-symmetric matrices is \emph{pivoting}.
Given a skew-symmetric matrix $A \in \F^{n \times n}$ and $S \subseteq [n]$ where $A[S]$ is non-singular,
reorder the rows and columns of
$A$ such that 
\[
  A =
  \begin{pmatrix}
    B & C \\
    -C^T& D
  \end{pmatrix},
\]
where $A[S]=B$.
Then the \emph{pivoting} of $A$ by $S$ is defined as
\[
  A * S =
  \begin{pmatrix}
    B^{-1} & B^{-1}C \\
    C^T B^{-1}& D + C^T B^{-1} C
  \end{pmatrix},
\]
which is again a skew-symmetric matrix.

\begin{lemma}[\cite{tucker1960combinatorial}] \label{lemma:tucker}
It holds that for any $X \subseteq [n]$, 
\begin{align*}
  \det (A*S)[X] = \frac{\det A[X \Delta S]}{\det A[S]},
\end{align*}
thereby,
$(A*S)[X]$ is non-singular if and only if $A[X \Delta S]$ is
non-singular. 
\end{lemma}

Additionally, we mention a formula regarding the Pfaffian of the sum of two skew-symmetric matrices:
\begin{lemma}[{\cite[Lemma 7.3.20]{Murota99}}]
  \label{lemma:sum-pf}
  For two skew-symmetric matrices $A_1$ and $A_2$, both indexed by $V$,
  it holds that
  \[ \Pf \, (A_1 + A_2) = \sum_{U \subseteq V} \sigma_U \Pf A_1[U] \cdot \Pf A_2[V \setminus U], \]
  where $\Pf A_i[\emptyset] = 1$ for $i = 1, 2$ and $\sigma_U \in \{1,-1\}$ is a sign of the permutation
  \begin{align*}
    \begin{pmatrix}
      1 & 2 & \cdots & |U| & |U| + 1 & \cdots & |V| - 1 & |V| \\
      u_1 & u_2 & \cdots & u_{|U|} & v_1 & \cdots & v_{|V \setminus U|-1} & u_{|V \setminus U|}
    \end{pmatrix},
  \end{align*}
  where $u_i$ and $v_i$ are the $i$-th largest elements of $U$ and $V \setminus U$, respectively.
\end{lemma}

Finally, we note a generalization of the Cauchy-Binet formula to skew-symmetric matrices:

\begin{restatable}[Ishikawa-Wakayama formula \cite{IshikawaW95}]{lemma}{iwformula}
  \label{lemma:cauchy-binet-ss}
  For a skew-symmetric $2n \times 2n$-matrix $A$ and a $2k \times 2n$-matrix $B$ with $k \le n$, we have
  \begin{align*}
    \Pf B A B^T = \sum_{U \in \binom{[2n]}{2k}} \det B[\cdot, U] \Pf A[U].
  \end{align*}
\end{restatable}

\subsection{Representations and operations on linear delta-matroids}
\label{ssec:rep}

We review some recent results on representations of, and operations
over, linear delta-matroids due to Koana and Wahlström~\cite{KW24}.

The traditional notion of linear representation of a delta-matroid is as follows. 
For a skew-symmetric matrix $A \in \F^{V \times V}$ over a field $\F$, define
\[ \cF = \{ X \subseteq V \mid A[X] \text{ is nonsingular} \}. \]
Then, $(V, \cF)$, which is denoted by $\bD(A)$, is a delta-matroid. 
We say that a delta-matroid $D = (V, \cF)$ is \emph{representable} over $\F$ if there is a skew-symmetric matrix $A \in \F^{V \times V}$ and a twisting set $X \subseteq V$ such that $D = \bD(A) \Delta X$.
If $A[X]$ is nonsingular, or equivalently $\emptyset \in \cF$, we say that $D$ is \emph{directly representable} over~$\F$.
Note that a directly representable delta-matroid $D$ can be represented without a twisting set $X$, as $\bD(A) \Delta X = \bD(A * X)$ by Lemma~\ref{lemma:tucker}.
We will say that $D$ is \emph{directly represented} by $A$ if $D = \bD(A)$.
We also refer to a representation $D=\bD(A) \Delta S$ as a \emph{twist representation}.

Another linear representation, of equivalent expressive power,
recently proposed by Koana and Wahlström~\cite{KW24}
is a \emph{contraction representation}, where a delta-matroid $D=(V,\cF)$
is represented as $D=\bD(A)/T$ where $A$ is a skew-symmetric matrix over ground set $V \cup T$. 
These forms have equivalent power, in the sense that one can convert between
twist representations and contraction representations of a given
linear delta-matroid in matrix multiplication time~\cite{KW24}. 
However, contraction representations are easier to work with for
certain algorithmic purposes.
Given a twist representation $D = \bD(A) \Delta S$, we get a contraction representation $D = \bD(A^*) / T$, where $A^*$ is indexed by $V \cup T$ (with $T$ being a set of size~$|S|$), defined as follows:
\begin{align} \label{eq:twist-to-contraction}
  A^* = \kbordermatrix{
    & T & S & V \setminus S \\
    T & A[S] & -I & -A[S, V \setminus S] \\
    S & I & O & O \\
    V \setminus S & A[V \setminus S, S] & O & A[V \setminus S] \\
  }.
\end{align}

Let $D=(V,\cF)$ be a delta-matroid and $X \subseteq V$.
The \emph{projection} $D|X$ of $D$ to $V \setminus X$ 
is a delta-matroid $D'=(V \setminus X, \cF')$
where $\cF'=\{F \setminus X \mid F \in \cF\}$. 
A \emph{projected linear delta-matroid}
is a delta-matroid $D=(V,\cF)$ represented as a projection
$D=D'|X$ from some linear delta-matroid $D'$ on a ground set $V \cup X$.
A projected linear delta-matroid is not necessarily even,
hence they generalize linear delta-matroids.
Every projected linear delta-matroid can be represented
as an \emph{elementary projection} $D=D'|\{x\}$ for a single element $x$,
and such a representation can be constructed from a general
projected linear representation in randomized polynomial time.\footnote{Koana
  and Wahlström consider \emph{$\varepsilon$-approximate representations}
  for a tweakable error parameter $\varepsilon$. Since we are not
  optimizing the polynomial factors in the running times in this
  paper, we set $\varepsilon = 1/2^{\Omega(n)}$ and construct a representation
  in polynomial time that is inexact only with exponentially small failure probability.}

In particular, let $A \in \F^{r \times V}$ be a linear representation
of a matroid $M$ and construct the matrix
\[
  A_D= \kbordermatrix {
    & T & V \\
    T & 0 & A \\
    V & -A^T & 0
  }.
\]
Then, by Lemma~\ref{lemma:det-pf}, $D_B=\bD(A_D)/T$ represents the set of bases of $M$ (i.e.,
the feasible sets of $D_B$ are precisely the bases of $M$)
and $D_I=\bD(A_D)|T$ represents the independent sets of $M$. 

Let $D_1=(V,\cF_1)$ and $D_2=(V,\cF_2)$ be delta-matroids on a shared
ground set. The \emph{union} $D_1 \cup D_2$ is the delta-matroid $D=(V,\cF)$
with feasible set
\[
  \cF=\{F_1 \cup F_2 \mid F_1 \in \cF_1, F_2 \in \cF_2, F_1 \cap F_2 = \emptyset\}.
\]
Note the disjointness condition (which is distinct from the more
familiar union operation over matroids). Similarly,
the \emph{delta-sum} $D_1 \Delta D_2$ is the delta-matroid $D=(V,\cF)$
with feasible set
\[
  \cF=\{F_1 \Delta F_2 \mid F_1 \in \cF_1, F_2 \in \cF_2\}.
\]
Bouchet~\cite{Bouchet89dam} and Bouchet and Schwärzler~\cite{BouchetS98deltasum}
showed that $D_1 \cup D_2$ and $D_1 \Delta D_2$ are delta-matroids.
Koana and Wahlström~\cite{KW24} showed that the union and delta-sum
of linear delta-matroids, represented over a common field,
are also linear, and linear representations can be constructed
in randomized polynomial time.


%%if statements needed
\iflong
  \input{sec3_long.tex}
\else
  \input{sec3_short.tex}
\fi
\iflong
  \input{sec456_long.tex}
\else
  \input{sec456_short.tex}
\fi



\section{Conclusion}
\label{sec:conclusions}

In this work, we have initiated the studies into the parameterized complexity of problems formulated over linear delta-matroids.
We have found a striking contrast to matroids:
the choice of parameterization, whether by rank or by cardinality, influences problem complexity. 
Specifically, when parameterized by rank, solutions applicable to matroids can be extended to delta-matroids,
with more involved algorithms but ultimately with the same $f(k)$-factor in the running time.
In particular, the recent method of determinantal sieving can be lifted to linear delta-matroids parameterized by rank.
It is at the moment unclear precisely how powerful this result is, but it does solve the basic
intersection and packing problems over linear delta-matroids.
In contrast, when parameterized by cardinality, we stumble on hardness in seemingly simple delta-matroids.
This underscores the significant impact of the absence of a truncation operation on their complexity.
Yet, we demonstrated that the \textsc{Delta-matroid Triangle Cover} remains FPT for delta-matroids of unbounded rank.
This was achieved by leveraging a novel algebraic technique, namely $\ell$-contraction, complemented by sunflower-like combinatorial approaches.
It is worth noting that the $\ell$-contraction operation is an extension of the matroid truncation.
Employing delta-matroid perspectives, we resolve two open questions in the literature, \textsc{Cluster Subgraph} and \textsc{Strong Triadic Closure} parameterized by above matching.
These problems have close connections to \textsc{Delta-matroid Triangle Cover}.
Specifically, when the graph is $K_4$-free and has a perfect matching, \textsc{Cluster Subgraph} parameterized by above matching is a special case of \textsc{Delta-matroid Triangle Cover}.
We were then able to give FPT algorithms for both problems on $K_4$-free graphs.
However, to our surprise, the complexity differs in the presence of $K_4$'s.
While we successfully extended the FPT algorithm to accommodate the general case for \textsc{Cluster Subgraph}, our attempts for an algorithmic solution for \textsc{Strong Triadic Closure} were met with difficulties, which led us to a W[1]-hardness proof.

Let us conclude this study with several open questions.
The first question concerns the potential for enhancing our algorithm for \textsc{Delta-matroid Triangle Cover}, which currently runs in $O^*(k^{O(k)})$ time, to $O^*(2^{O(k)})$ time.
The algebraic approach, and the simpler color-coding approach, in conjunction with dynamic programming, typically achieves single-exponential time complexity.
However, we have not been able to improve our algorithm to $O^*(2^{O(k)})$ time.
Another area worth investigating is the derandomization of our algorithm.
Our algorithm is ``heavily'' randomized, as we repeatedly apply $\ell$-projection, which relies on the Schwartz-Zippel lemma.
Given that the matroid truncation can be performed deterministically \cite{LokshtanovMPS18TALG}, it is reasonable to believe that the deterministic implementation of $\ell$-projection can be achieved.
On the other hand, a more fundamental obstacle is to derandomize the algorithm for \textsc{Colorful Delta-matroid Matching}. 

Lastly, let us discuss future research directions regarding kernelization.
Over matroids, The representative sets approach yields a polynomial kernel for, e.g., \textsc{3-Matroid Intersection} and \textsc{3-Matroid Parity}.
For delta-matroids with the rank parameter, the picture looks promising using the delta-matroid representative sets statement of Wahlström~\cite{Wahlstrom24SODA}.
Under a cardinality parameter, things look more difficult. In particular, it is completely unclear to us whether \textsc{DMM Intersection} has a polynomial kernel parameterized by cardinality. 
We remark that \textsc{Cluster Subgraph} and \textsc{Strong Triadic  Closure} do not have polynomial kernels even when parameterized by $\ell$ \cite{GolovachHKLP20,GruttemeierK20}, under standard complexity assumptions.
More broadly, is there a further application of representative sets statement of Wahlström~\cite{Wahlstrom24SODA}? 
Though there is no truncation operation for delta-matroids, it is plausible that the $\ell$-projection unlocks the potential of delta-matroid representative sets approach for kernelization.




%===========OLD Proof, to be removed after the new one is finished typing ===========
%\begin{proof}
%We reduce from \textsc{Multicolored Clique}, which is formally defined as follows.
%Given a graph $G$, where $V(G)$ is partitioned into $k$ independent sets $V_1 \cup \cdots \cup V_k$, the problem is to find a clique of size $k$.
%To facilitate the reduction, we transform $G$ into a directed graph $G'$ by substituting every edge $(u,w)$ in $G$ with two bidirectional arcs $(u,w)$ and $(w,u)$, and by adding a self-loop at each vertex in $G$. 
%Then, the \textsc{Multicolored Clique} problem translates to identifying a set of vertices $K$ in $G'$, with $|K|=k$, such that
%every vertex in $G'[K]$ has both in-degree and out-degree $k$.
%We will construct a graph $H$, where $V(H)$ is partitioned into 5 parts $P \cup P' \cup Q \cup Q' \cup R$.
%
%\paragraph*{Construction.}
%We start with sets $P$ and $P'$.
%For every arc $a = v_i v_j$ in $G'$, we introduce a pair of vertices $p_a$ and $p_a'$, and connect them with an edge.
%Let $P = \{ p_a \mid a \in E(G') \}$ and $P' = \{ p_a' \mid a \in E(G') \}$.
%For each $i, j \in [k]$, and for each arc $a= v_i v_j \in E(G')$ with $v_i \in V_i$ and $v_j \in V_j$, we add an edge between $p_a$ and $p_{a}'$, in addition, we add an edge between $p_a$ and $p_{a'}$ for all arcs $a'$ that satisfy one of the following conditions: (i) $a' = v_{i+1} v_j$ where $v_{i+1} \in V_{i+1}$ or (ii) $a' = v_i v_{j+1}$ where $v_{j+1} \in V_{j+1}$.
%Here, we employ a cyclic notation, where $V_1 = V_{k+1}$.
%
%Let $Q$ and $Q'$ be defined as $Q = \{ q_{i,j,1}, q_{i,j,2} \mid i, j \in [k] \}$ and $Q' = \{ q_{i,j,1}', q_{i,j,2}' \mid i, j \in [k] \}$, respectively.
%For each $i, j \in [k]$, we introduce edges $q_{i,j,1} q_{i,j,1}'$, $q_{i,j,2} q_{i,j,2}'$, and $q_{i,j,1} q_{i,j,2}$.
%Furthermore, for each $i, j \in [k]$, and for each arc $a = v_i v_j$ with $v_i \in V_i$ and $v_j \in V_j$, we add edges so that $\{ q_{i,j,1}, q_{i,j,2}, p_a \}$ forms a triangle.
%
%Finally, let $R = \{ r_{i,j,1}, r_{i,j,2} \mid i,j \in [k] \}$.
%We add edges so that $R$ forms a clique.
%We also add an edge between each $p \in P$ and each $r \in R$. 
%For each $i, j \in [k]$, we add edges from $r_{i,j,1}$ to $q_{i,j,1}$, $q_{i,j,2}$, $q_{i+1,j,1}$, $q_{i+1,j,2}$, and from $r_{i,j,2}$ to $q_{i,j,1}$, $q_{i,j,2}$, $q_{i,j+1,1}$, $q_{i,j+1,2}$.
%
%\paragraph{Correctness.}
%
%We claim that there is a clique of size $k$ in $G$ if and only if $G'$ has a strong edge set of size $\ell = |P| + 6k^2 + \binom{2k^2}{2}$.
%Note that this yields a parameterized reduction:
%As $H$ contains a matching of size $|P|$, where each in $P$ is matched to the corresponding vertex in $P'$, the above matching parameter is thus at most $\ell - |P| = 6k^2 + \binom{2k^2}{2}$.\todo{even perfect matching in $Q\cup Q'$ and $R$ are quite obvious so parameter is even at most $3k^2+\binom{2k^2}{2}$}
%
%$(\Rightarrow)$
%Let $K = \{ v_1, \dots, v_k \}$ with $v_i \in V_i$ be a clique of size $k$ in $G$.
%It follows that, for all $i, j \in [k]$, the arc $v_i v_j$ exists in $G'$.
%We can construct a strong edge set $S$ of size $\ell$ as follows:
%\begin{itemize}
%  \item For every $i, j \in [k]$, include a triangle $\{ p_{v_i v_j}, q_{i,j,1}, q_{i,j,2} \}$.
%  \item For each vertex $p_a \in P$ not involved in these triangles, include the edge $p_ap_a'$.
%  \item Include all $\binom{2k^2}{2}$ edges inside $R$.
%  \item For each $i, j \in [k]$, add edges connecting $r_{i,j,1}$ to $p_{v_i v_j}$ and $p_{v_{i+1} v_j}$, as well as edges connecting $r_{i,j,2}$ to $p_{v_{i} v_{j}}$ and $p_{v_{i} v_{j+1}}$.
%\end{itemize}
%Note that $3k^2 + (|P| - k^2) + \binom{2k^2}{2} + 4k^2 = \ell$ edges have been included into $S$.
%
%We verify that the set $S$ indeed satisfies the triadic closure property.
%First, observe that the edges $p_a p_{a}'$ are isolated within $S$, i.e., neither $p_a$ nor $p_{a}'$ is incident with any other edge of $S$.
%Consequently, we identify four types of $P_3$'s in $S$:
%\begin{itemize}
%  \item 
%  $(r, p_{v_i v_j}, r')$ for $r, r' \in R$ is closed, as $R$ is a clique.
%  \item
%  $(r_{i-1,j,1}, p_{v_i v_j}, q_{i,j,\beta})$, $(r_{i,j-1,2}, p_{v_i v_j}, q_{i,j,\beta})$, and $(r_{i,j,\alpha}, p_{v_i v_j}, q_{i,j,\beta})$ for $\alpha, \beta \in \{ 1, 2 \}$ are closed because
%  our construction adds edges connecting $r_{i-1,j,1}$ to $q_{i,j,\beta}$, $r_{i,j-1,2}$ to $q_{i,j,\beta}$, and $r_{i,j,\alpha}$ to $q_{i,j,\beta}$.
%  \item
%  $(p_{v_i v_j}, r_{i, j, 1}, p_{v_{i+1} v_j})$ and $(p_{v_i v_j}, r_{i, j, 2}, p_{v_{i} v_{j+1}})$ are closed our construction adds edges from $p_{v_i v_j}$ to $p_{v_{i+1} v_j}$ and $p_{v_i v_{j+1}}$.
%  \item
%  $(p, r, r')$ for $p \in P$ and $r, r' \in R$ is closed as there is an edge between each $p \in P$ and $r \in R$.
%\end{itemize}
%Therefore, $S$ is a strong set of size $\ell$.
%
%$(\Leftarrow)$
%Let $S$ be a strong edge set of size $\ell$.
%We show that $G$ contains a clique of size $k$. 
%We first show that $S$ consists of the following edges:
%\begin{itemize}
%  \item $|P| + 2k^2$ edges from $G' - R$.
%  \item All $\binom{2k^2}{2}$ edges within $R$.
%  \item For every $r \in R$, there are exactly two edges connecting $r$ to vertices in $P$.
%\end{itemize}
%Since $G' - R$ is $K_4$-free, and the half of vertices in $G' - R$ has degree 1, $|E(G' - R) \cap S| \le |P| + 2k^2$.
%The equality holds precisely when for every vertex $v \in P \cup Q$ satisfies one of the following conditions: either $vv' \in S$ or $v$ has degree 2 in $E(G' - R) \cap S$. \todo{Should we elaborate why?}
%
%We classify vertices $r$ in $R$ into two types: (i) those for which there exists $qr \in S$ with $q \in Q$ and (ii) those for which no such edge $qr \notin S$ exists.
%For a vertex $r$ of type (i), suppose that $q_{i,j,z} r$ for $z \in \{ 1, 2 \}$ is part of $S$.
%As $q_{i,j,z}$ is adjacent to four distinct vertices in $Q$\todo{in $R$?}, namely $r_{i,j,1}, r_{i,j,2}, r_{i-1,j,1}$, and $r_{i,j-1,2}$, the number of edges directly connecting $r$ with vertices in $R$, and included in $S$, is limited to three.
%This is because the inclusion of any edge $r_{i,j,1} r$ into $S$, with $r$ being none of these four vertices, would break the triadic closure property, specifically on the path $(r, r_{i,j,1}, q_{i,j,1})$.
%For a vertex of type (ii), there can at most two connections to vertices in $P$ included in $S$.
%This is because if three edges incident with $r$ belong to $S \setminus E(R)$, these edges' endpoints (with $r$ exclude) must constitute a triangle to adhere to the triadic closure property.
%However, by the construction of $H$, these triangles in $G' - R$ are specifically of the form $\{ p_a, q_{i,j,1}, q_{i,j,2} \}$.
%
%% Focusing on a vertex $r = r_{i,j,1}$, there are two types of triangles within its neighborhood: one formed with $\{ p_e, q_{i,j,1}, q_{i,j,2} \}$ and the other with $\{ p_e, q_{i+1,j,1}, q_{i+1,j,2} \}$.
%
%Let $x$ be the number of vertices $r$ of type (i).
%We evaluate the total count of edges in $S$ incident with $R$ in terms of $x$.
%The edges hat have exactly one endpoint in $R$ amount to at most $3x + 2(2k^2 - x)$, as the vertices of type (i) three connections outside $R$, and the other $2k^2 - x$ vertices of type (ii) have at most two connections outside $R$.
%For the edges that have both endpoints in $R$, the hand-shaking lemma gives us an upper bound:
%\begin{align*}
%  \frac{1}{2} (3x + (2k^2 - 1)(2k^2 - x)) = \binom{2k^2}{2} - x(k^2 - 2).
%\end{align*}
%Putting all together, the total number of edges in $S$ is bounded from above by
%\begin{align*}
%  (|P| + 2k^2) + (3x + 2(2k^2 - x)) + \left( \binom{2k^2}{2} - x(k^2 - 2) \right) = |P| + 6k^2 + \binom{2k^2}{2} - x(k^2 - 5).
%\end{align*}
%\todo{Should 5 be 3 in the above equation?}
%Given that this must be no smaller than $\ell$, it follows that $x = 0$.
%Furthermore, the number of edges outside $R$, within $R$, and bridging across $R$, are exactly $|P| + 2k^2$, $\binom{2k^2}{2}$, and $4k^2$, respectively.
%As every vertex $r$ in $R$ has type (ii), it follows that there are exactly edges in $S$ connecting $r$ to $P$, establishing our initial claim regarding the composition of $S$.
%
%Consider the vertex $r_{i,j,1}$ and its neighborhood outside $R$, which consists of $P$ along with four vertices $q_{i,j,1}, q_{i,j,2}, q_{i+1,j,1}, q_{i+1,j,2}$ from $Q$.
%We have established that $r_{i,j,1}$ has two connections to vertices in $P$ in $S$.
%We further show that these two connections are to $p_{v_i v_j}$ and $p_{v_{i+1} v_j'}$, where $v_i \in V_i$, $v_{i+1} \in V_i$\todo{$V_{i+1}$?}, and $v_j, v_j' \in V_j$.
%Assume, for the sake of contradiction, that $r_{i,j,1} p_a \in S$, where the tail and head of $a$ are in $V_{i'}$ and $V_{j'}$, respectively, with $(i', j') \ne (i, j)$ and $(i', j') \ne (i+1, j)$.
%Then, $p_a p_a'$ is not part of $S$, as there is no edge between $r_{i,j,1}$ and $p_a'$.
%As $S$ contains $|P| + 2k^2$ edges from $G' - R$, it follows that $p_a$ has degree two in $S$, i.e., the edges $\{ p_a, b_{i',j',1} \}$ and $\{ p_a, b_{i',j',2} \}$ belong to $S$.\todo{I would like some polishing here: (1) You mean ``degree outside of $R$'' (2) Argue why these edges cannot be some edge $p_ap_b$ (3) Some $b$, $c$ are not yet changed here to $p, q, r$}
%However, this results in a contradiction, as it creates a path $(c_{i,j,1}, p_a, b_{i',j',1})$ that violates triadic closure.
%Consequently, it follows that for each $i, j \in [k]$, there is precisely one arc $a_{i,j} \in V_i \times V_j$ such that $r_{i,j,1} p_a$ belong to $S$, and therefore, the triangle $\{ p_{a_{i,j}}, q_{i,j,1}, q_{i,j,2} \}$ are part of $S$.
%In particular, there are $k$ self-loops $l_i$ on $v_i \in V_i$ for $i \in [k]$ such that the triangle $\{ a_{l_i}, b_{i,i,1}, b_{i,i,2} \}$ is part of $S$.
%
%We claim that $\{ v_i \mid i \in [k] \}$ forms a clique in $G$.
%Note that for each $i, j \in [k]$, the vertex $r_{i,j,1}$ incident with two edges: one connecting to $p_{a_{i,j}}$ and the other to $p_{a_{i+1,j}}$.
%Similarly, $r_{i,j,2}$ is connected to $p_{a_{i,j}}$ and $p_{a_{i,j+1}}$.
%The paths $(p_{a_{i,j}}, r_{i,j,1}, p_{a_{i+1,j}})$ and $(p_{a_{i,j}}, r_{i,j,2}, p_{a_{i,j+1}})$ satisfy triadic closure, which implies the existence of edges $p_{a_{i,j}} p_{a_{i+1,j}}$ and $p_{a_{i,j}} p_{a_{i,j+1}}$, for each $i, j \in [k]$.
%By our construction of $H$, it follows that the tails of $a_{i,j}$ and $a_{i+1,j}$ coincide, and similarly the heads of $a_{i,j}$ and $a_{i,j+1}$ coincide as well.
%Consequently, $a_{i,j} = v_i v_j$ for every $i, j \in [k]$, establishing that all pairwise edges $v_i v_j$ are present in $G$.
%
%\end{proof}



\bibliographystyle{abbrv}
\bibliography{all}


\end{document}


% 
