\section{Introduction}
\label{sec:intro}

% To insert a figure: \input{figs/template}
% Or table: \input{tables/template}

Image recognition techniques, particularly those based on deep learning, are promising for real-world applications; however, image classification using deep learning models is said that less robust to diverse real-world degradations than that of human visual perception~\cite{ImageNet-C,ImRobust}.
Consequently, as the accuracy of image recognition models improves, the goal of increasing classification robustness with respect to real-world degradations is widely seen as one of the central challenges in image recognition with deep learning.

One promising technique for improving the robustness of image recognition models for classification task is \textit{data augmentation} such as Mixup~\cite{Mixup} and CutMix~\cite{CutMix}.
Using these data augmentation methods, we can increase image counts while reducing overfitting, thus potentially improving robustness.
The proposed data augmentation method known as PixMix~\cite{PIXMIX} extends training datasets with images taken from mixing sets by combining real and synthetic images. The mixing is done both additively or multiplicatively. PixMix pipeline achieves improvements in both robustness and image classification accuracy.

% \figPIXMIXmoireConcept
\input{figs/en-fig1}

The mixing set images used by PixMix include mathematically generated Fractal arts and feature visualizations (FVis).
Fractal arts are collected on DeviantArt. The images are visually diverse.
FVis are collected using OpenAI Microscope. The images are created from convolutional neural networks (CNNs) such as AlexNet~\cite{alexnet}, VGGNet~\cite{vggnet}, and ResNet~\cite{resnet}, which are basically pre-trained on ImageNet~\cite{distill,FVis}.

However, the use of these images entail at least three practical disadvantages:
i) Some of the human-designed digital patterns and generative arts
are protected by copyright, and thus commercial use of PixMix
data augmentation remains questionable.
ii) Generating FVis requires multiple CNNs trained on large image
datasets, and is thus a high-cost operation for assembling images into datasets.
iii) For both Fractal arts and FVis the number of
images that may be feasibly assembled into a mixing set is limited in practice.

All of these problems may be eliminated by auto-generating data augmentation images from mathematical formulas.
For example, in the methods named ``Formula-driven Supervised Learning (FDSL)'', we can construct large-scale image datasets which are privacy-safe, copyright-free, and light collecting images costs. 
In FDSL context, Fractal DataBase (FractalDB), Shader, etc. are proposed.
As discussed in the original PixMix paper,
tests to characterize the performance of
FractalDB and Shader, indicate that formula-generated images are less effective
for improving robustness than Fractal arts and FVis.

Thus, the aim of the present study is to devise a strategy for
constructing a formula-generated mixing set that can improve the robustness of the classification model compared to fractal / FVis.
Our data augmentation procedure is the same as that of PixMix,
but here we propose a novel family of data augmentation
images that promise improved robustness:
\textit{interference-fringe images}, for which 
we use the term \textit{MoireDB} images and their constructed \textit{Moir\'e DataBase (MoireDB)}.

The idea of generating Moir\'e images
is motivated by the hypothesis that using
illusory images for data augmentation should
tend to increase robustness; this hypothesis, in
turn, is based on the close relation between Moir\'e images
and optical illusions, as well as on the known fact that
deep learning models trained on such images exhibit
increased robustness.
We construct MoireDB, a repository
of automatically formula-generated Moir\'e images, and
use it as an auxiliary image set to create a PixMix augmented
dataset; then we train a deep learning model on this
augmented dataset and measure the robustness of its image classification.

Our proposal of MoireDB offers several key advantages, including the following.
\begin{itemize}
      \item The use of MoireDB for data augmentation
            improves robustness with respect to real-world degradations.
      \item MoireDB contains only formula-generated images,
            eliminating copyright problems and making the database
            suitable for commercial use.
      \item The images constituting MoireDB are auto-generated, reducing
            the cost of assembling images into datasets.
\end{itemize}