
\documentclass[10pt,conference]{IEEEtran} 

\IEEEoverridecommandlockouts


\input{macro}


\begin{document}

\title{Combining Language and App UI Analysis for the Automated Assessment of Bug Reproduction Steps
\thanks{The first two authors contributed equally to this work.}
}

\author{
	
	\IEEEauthorblockN{Junayed Mahmud\IEEEauthorrefmark{1}, Antu Saha\IEEEauthorrefmark{2}, Oscar Chaparro\IEEEauthorrefmark{2}, Kevin Moran\IEEEauthorrefmark{1}, Andrian Marcus\IEEEauthorrefmark{3}}
    \IEEEauthorblockA{\IEEEauthorrefmark{1}\textit{University of Central Florida (USA)}, 
		    \IEEEauthorrefmark{2}\textit{William \& Mary (USA)},
		    \IEEEauthorrefmark{3}\textit{George Mason University (USA)}
		    \\ \href{mailto:}{junayed.mahmud@ucf.edu}, \href{mailto:}{asaha02@wm.edu}, \href{mailto:}{oscarch@wm.edu},
		    \href{mailto:}{kpmoran@ucf.edu},
		    \href{mailto:}{amarcus7@gmu.edu}
    }
}

\maketitle

%page numbering
\thispagestyle{plain}
\pagestyle{plain}

\begin{abstract}
Bug reports are essential for developers to confirm software problems, investigate their causes, and validate fixes. Unfortunately, reports often miss important information or are written unclearly, which can cause delays, increased issue resolution effort, or even the inability to solve issues. One of the most common components of reports that are problematic is the steps to reproduce the bug(s) (S2Rs), which are essential to replicate the described program failures and reason about fixes. Given the proclivity for deficiencies in reported S2Rs, prior work has proposed techniques that assist reporters in writing or assessing the quality of S2Rs. However, automated understanding of S2Rs is challenging, and requires linking nuanced natural language phrases with specific, semantically related program information. Prior techniques often struggle to form such language $\leftrightarrow$ program connections -- due to issues in language variability and limitations of information gleaned from program analyses. 

To more effectively tackle the problem of S2R quality annotation, we propose a new technique called \tool, which leverages the language understanding capabilities of LLMs to identify and extract the S2Rs from bug reports and \rev{map} them to GUI interactions in a program state model derived via dynamic analysis. We compared \tool to a related state-of-the-art approach and we found that \tool annotates S2Rs 25.2\% better (in terms of F1 score) than the baseline. 
Additionally, \tool suggests more accurate missing S2Rs than the baseline (by 71.4\% in terms of F1 score). 
\looseness=-1
\end{abstract}




\input{1_intro}

\input{motivating_example}

\input{2_approach}

\input{3_prompt_development}

\input{4_empirical_evaluation}

\input{4_results}

\input{5_related_work}



\input{threats_to_validity}

\input{6_conclusions}

\input{7_acknowledgements}

\balance

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
