\PassOptionsToPackage{dvipsnames}{xcolor}

\documentclass[10pt,sigconf,nonacm,screen]{acmart}
\usepackage[all]{nowidow}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{acronym}
% \usepackage[cachedir=.minted]{minted}
% step 1
% \usepackage[finalizecache,cachedir=.]{minted}
% step 2
\usepackage[frozencache,cachedir=.]{minted}

\usepackage[capitalise,noabbrev]{cleveref}

\crefformat{section}{\S#2#1#3}
\crefformat{subsection}{\S#2#1#3}
\crefformat{subsubsection}{\S#2#1#3}
\crefrangeformat{section}{\S#3#1#4 to~\S#5#2#6}
\crefrangeformat{subsection}{\S#3#1#4 to~\S#5#2#6}
\crefrangeformat{subsubsection}{\S#3#1#4 to~\S#5#2#6}

\setlength{\marginparwidth}{2cm}

\newcommand{\sysname}{\textrm{LLM4FaaS}}

\hyphenation{ ma-the-in-der }

\begin{document}

\author{Minghe Wang}
\affiliation{%
    \institution{TU Berlin \& ECDF}
    \city{Berlin}
    \country{Germany}}
\email{mw@3s.tu-berlin.de}
\orcid{0009-0001-3780-5828}

\author{Tobias Pfandzelter}
\affiliation{%
    \institution{TU Berlin \& ECDF}
    \city{Berlin}
    \country{Germany}}
\email{tp@3s.tu-berlin.de}
\orcid{0000-0002-7868-8613}

\author{Trever Schirmer}
\affiliation{%
    \institution{TU Berlin \& ECDF}
    \city{Berlin}
    \country{Germany}}
\email{ts@3s.tu-berlin.de}
\orcid{0000-0001-9277-3032}

\author{David Bermbach}
\affiliation{%
    \institution{TU Berlin \& ECDF}
    \city{Berlin}
    \country{Germany}}
\email{db@3s.tu-berlin.de}
\orcid{0000-0002-7524-3256}

\title{LLM4FaaS: No-Code Application Development using LLMs and FaaS}
\keywords{Large Language Models, Function-as-a-Service, No-Code Development}

\begin{abstract}
        Large language models (LLMs) are powerful tools that can generate code from natural language descriptions.
        While this theoretically enables non-technical users to develop their own applications, they typically lack the expertise to execute, deploy, and operate generated code.
        This poses a barrier for such users to leverage the power of LLMs for application development.
        
        In this paper, we propose leveraging the high levels of abstraction of the Function-as-a-Service (FaaS) paradigm to handle code execution and operation for non-technical users.
        FaaS offers function deployment without handling the underlying infrastructure, enabling users to execute LLM-generated code without concern for its operation and without requiring any technical expertise.
        We propose \sysname{}, a novel no-code application development approach that combines LLMs and FaaS platforms to enable non-technical users to build and run their own applications using only natural language descriptions.
        Specifically, \sysname{} takes user prompts, uses LLMs to generate function code based on those prompts, and deploys these functions through a FaaS platform that handles the application's operation.
        \sysname{} also leverages the FaaS infrastructure abstractions to reduce the task complexity for the LLM, improving result accuracy.
        
        We evaluate \sysname{} with a proof-of-concept implementation based on \emph{GPT-4o} and an open-source FaaS platform, using real prompts from non-technical users.
        Our evaluation based on these real user prompts demonstrates the feasibility of our approach and shows that \sysname{} can reliably build and deploy code in 71.47\% of cases, up from 43.48\% in a baseline without FaaS.
\end{abstract}

\maketitle

\input{sections/1_introduction.tex}
\input{sections/2_background.tex}
\input{sections/3_architecture.tex}
\input{sections/4_evaluation.tex}
\input{sections/5_related_work.tex}
\input{sections/6_conclusion.tex}

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography.bib}

\input{sections/7_appendix.tex}

\balance
\end{document}
