\section{Related Work}
\textbf{Retrieval Augmented Generation in AI for Science}.
Retrieval augmented generation~(RAG) is a technique that integrates relevant information retrieved from external databases into a model's training and inference processes. While originally developed to aid Large Language Model (LLM)-based AI agents, their applicability has readily extended to scientific tasks, spanning LLMs for material discovery____, drug design____, and organic synthesis____. Utilizing RAG in various forms  has also shown ground-breaking advances in scientific discovery beyond LLM agents; one of the most prominent examples is AlphaFold model____, whose sequence alignment module is effectively a RAG module. 
RAG is also found helpful in molecular machine learning tasks such as molecular generation____, drug design____, and protein function prediction____. Despite success in certain molecular learning tasks, there remains the opportunity for a principled RAG design strategy tailored to small molecules % the small molecule scope 
that can perform robust and informative structural matching. 


\textbf{Neural Graph Matching}.
Graph matching is a combinatorial optimization problem that matches the nodes of multiple graphs by maximizing the edge-wise and node-wise affinities. Neural graph matching was developed to tackle the computational challenges of the NP-hard quadratic assignment problem____ and the vulnerability of predefined affinity metrics____.
Among all design choices, a family of linear-matching  methodologies____ incorporates graph neural networks____ to embed edge affinities into node embeddings that are then used in a differentiable node-matching layer using, e.g., ____ or a simpler Softmax for assignment. 
This paper explores and validates the applicability of neural graph matching in the context of retrieval-augmented mass spectra generation in molecular machine learning.