\documentclass[twoside]{article}

\usepackage{caption}
\usepackage[accepted]{aistats2025}

\usepackage[round]{natbib}
\usepackage{bbding}
\usepackage{balance}
\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

\usepackage{comment}
\usepackage{twemojis}

\newcommand{\rev}[1]{#1}
\newenvironment{revsec}{}{} %\color{blue}
\usepackage{sidecap}
\usepackage{wrapfig}
\usepackage{bm}
\usepackage{url}

\input{preamble}


\begin{document}

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
\runningtitle{\raisebox{-0.3ex}{\scriptsize{x}}CEBRA for explainable attribution maps in time-series data}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
\runningauthor{Schneider, González Laiz, Filippova, Frey, Mathis}

\twocolumn[


\aistatstitle{Time-series attribution maps with regularized contrastive learning}

\aistatsauthor{Steffen Schneider$^{1}$ \And Rodrigo González Laiz$^{1}$ \And Anastasiia Filippova$^{1}$}
\vspace{6pt}
\aistatsauthor{Markus Frey$^{1}$ \And Mackenzie Weygandt Mathis$^{1}$}

\vspace{5pt}

\aistatsaddress{ $^{1}$EPFL \Envelope mackenzie.mathis@epfl.ch \& steffen.schneider@helmholtz-munich.de} ]

\input{content}
\bibliography{bibliography}
\bibliographystyle{unsrtnat}


\onecolumn
\appendix
\input{content_supplement}

\section{Checklist}


% %%% BEGIN INSTRUCTIONS %%%
%The checklist follows the references. For each question, choose your answer from the three possible options: Yes, No, Not Applicable.  You are encouraged to include a justification to your answer, either by referencing the appropriate section of your paper or providing a brief inline description (1-2 sentences). 
%Please do not modify the questions.  Note that the Checklist section does not count towards the page limit. Not including the checklist in the first submission won't result in desk rejection, although in such case we will ask you to upload it during the author response period and include it in camera ready (if accepted).

%\textbf{In your paper, please delete this instructions block and only keep the Checklist section heading above along with the questions/answers below.}
% %%% END INSTRUCTIONS %%%


 \begin{enumerate}

 \item For all models and algorithms presented, check if you include:
 \begin{enumerate}
   \item A clear description of the mathematical setting, assumptions, algorithm, and/or model: Yes, we provide rigorous definitions of the data-generating process, as well as the notions of identifiability used in Section~\ref{sec:attribution-maps}.
   \item An analysis of the properties and complexity (time, space, sample size) of any algorithm: Although we do not perform a theoretical analysis on time and space complexity, we include a detailed comparison of model runtimes in Appendix C, which is also referenced in the results.
   \item (Optional) Anonymous source code, with specification of all dependencies, including external libraries. Yes, source code is provided as part of the supplementary material. Note that the implementation of our method is based on the publicly available CEBRA codebase \citep{schneider2023cebra}, and the source code provided is a fully functional fork of that repository with our changes added. We also provide demo notebooks for reproducing the key experiments of the paper.
 \end{enumerate}


 \item For any theoretical claim, check if you include:
 \begin{enumerate}
   \item Statements of the full set of assumptions of all theoretical results: Yes, the assumptions are stated in the theorems in section~\ref{sec:identifiability}. Where applicable, the assumptions reference the definitions given in Section~\ref{sec:attribution-maps}.
   \item Complete proofs of all theoretical results: Yes, the proofs are attached in full in Appendix A. For Theorem 2, we also provide a proof sketch in the main paper. 
   \item Clear explanations of any assumptions: Yes, we motivate our assumptions with the typical structure of scientific time-series datasets.
 \end{enumerate}


 \item For all figures and tables that present empirical results, check if you include:
 \begin{enumerate}
   \item The code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL). %[Yes/No/Not Applicable]
        Yes, we provide demo notebooks for both the synthetic verification experiments and the synthetic RatInABox experiments. 
   \item All the training details (e.g., data splits, hyperparameters, how they were chosen). Yes, we provide most of these details in Section~\ref{sec:methods}, and further expand in Appendix B.
         \item A clear definition of the specific measure or statistics and error bars (e.g., with respect to the random seed after running experiments multiple times). Yes, we report 95\% confidence intervals computed in 10 seeds of data-generating processes. Full statistical results for the validation experiments are provided in Appendix~\ref{app:statistical_analysis}.
         \item A description of the computing infrastructure used. (e.g., type of GPUs, internal cluster, or cloud provider). Yes, we provide this briefly here. The experiments were mostly conducted on a compute cluster with V100 GPUs. Typically, multiple experiments can be loaded onto a single GPU. Attribution map computation can be performed on both CPU and GPU at an acceptable speed.%[Yes/No/Not Applicable]
 \end{enumerate}

 \item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets, check if you include:
 \begin{enumerate}
   \item Citations of the creator If your work uses existing assets.
        Yes, we outline this in Appendix E and cite in the main text where appropriate.
   \item The license information of the assets, if applicable. 
        Yes, we outlined this in Appendix E. All assets used were published under open source licenses before.
   \item New assets either in the supplemental material or as a URL, if applicable.
        Not applicable (except for code, see Appendix E).
   \item Information about consent from data providers/curators.
        Not applicable, as simulated and/or previously open sourced data was used exclusively.
   \item Discussion of sensible content if applicable, e.g., personally identifiable information or offensive content. 
        Not applicable.
 \end{enumerate}

 \item If you used crowd-sourcing or conducted research with human subjects, check if you include:
 \begin{enumerate}
   \item The full text of instructions given to participants and screenshots.
        Not applicable.
   \item Descriptions of potential participant risks, with links to Institutional Review Board (IRB) approvals if applicable. 
        Not applicable.
   \item The estimated hourly wage paid to participants and the total amount spent on participant compensation. 
        Not applicable.
 \end{enumerate}

 \end{enumerate}



\end{document}
