\section{\RR{Pre-Interview Questionnaire}}
\label{appendix:questionnaire}

\RR{The pre-interview questionnaire asked the interviewees to share the information of their favorite VR story projects through the following questions:}

\RR{\begin{itemize}
    \item What was your favorite self-created animated VR story about?
    \item How many months did it take to complete your favorite animated VR story?
    \item What motivated your choice of VR as the medium for your favorite animated VR story?
    \item Did your favorite animated VR story have a target audience? If so, who was your target audience?
    \item What is the online link to your favorite animated VR story? Input “None” if not applicable. 
\end{itemize}
}

\section{\RR{Interview Guide}}
\label{appendix:guide}

\RR{The following questions served as a semi-structured interview guide to gather insights addressing the two research questions on creation processes (RQ1) and challenges (RQ2).}

\subsection{\RR{Creation Processes (\textbf{RQ1})}}


\RR{\begin{itemize}
    \item Thinking about your favorite VR story, what were the main stages you went through from scratch to final completion? Please describe the stages chronologically and illustrate with intermediate results by sharing your screen.
    \item For each stage you mentioned:
    \begin{itemize}
        \item What were your primary objectives in this stage?
        \item What specific tasks did you perform in this stage?
        \item What were the inputs and outputs of this stage? 
        \item What tools and software did you use in this stage?
        \item How did using VR and computer animation technologies make this stage different from other storytelling forms (e.g., 2D/3D animation, 2D/3D games, micro-films)?
        \item Compared to other storytelling forms, what unique considerations did you have in this stage for your animated VR story?
    \end{itemize}
    \item Regarding the overall workflow you described:
    \begin{itemize}
        \item How did you decide on the sequence of the stages? 
        \item Were there times you needed to revisit earlier stages? What prompted this?
        \item Did you work on multiple stages in parallel? What led to that decision?
    \end{itemize}
\end{itemize}
}


\subsection{\RR{Challenges (\textbf{RQ2})}}
\RR{For each stage that you mentioned:}

\RR{\begin{itemize}
    \item What unique difficulties did you encounter when integrating multiple story elements within VR and computer animation constraints? Please provide concrete examples.
    \item Compared to other storytelling forms, which tasks were particularly time-consuming? Why?
    \item Compared to other storytelling forms, what features or ideas did you initially plan but couldn't implement in the final VR story? Why?
    \item Compared to other storytelling forms, what were your most frustrating moments? Why?
    \item Earlier you mentioned [specific difference or unique consideration] compared to other storytelling forms. What challenges emerged from this?
\end{itemize}
}

\section{Project Details of Interviewees}
\label{appendix:project}
Table \ref{tab:project_details} summarizes the project details mentioned by our participants during the interview study.
\input{src/tables/project_details_revised}