\begin{comment}\section*{Author Contributions}
\label{author_contributions}


\begin{enumerate}
\item Sarah Luger and Heather Frase led the alignment of the paper's accuracy, thoroughness, appropriate narrative specificity, and formatting across author contributions. 
\item Fazl Barez, Heather Frase, Sarah Luger contributed to managing the report's writing and copyediting in collaboration with Peter Mattson and Joaquin Vanschoren.
\item Mala Kumar contributed to portions in Section 4 (4.1 and 4.2) and in Section 8 Exploratory Studies and Research related to the expression-of-interest projects.
\item Marisa Ferrara Boston, James Ezick, and Mike Kuniavsky (Workstream 5) wrote Section 2.3: Understanding and Using Results.
\item Tarunima Prabhakar (Tattle Civic Tech), Dr. Declan Grabb and Dr. Nina Vasan (Brainstorm Solutions), Tajuddeen Gwadabe (Masakhane), and Nicolas Miailhe (PRISM Eval) contributed to Section 11: Exploratory Studies and Research material related to the expression-of-interest projects.
\item Workstream 7 wrote Section~\ref{subsec:bias}: Chris Knotz (CommonGround), Adina Williams (FAIR, Meta), Alicia Parrish (Google), Vassil Tashev (Simon Fraser University), Shachi H. Kumar (Intel Labs), Saurav Sahay (Intel Labs), Bhatipriya Radharapu (FAIR, Meta), and Heather Frase.
\item Workstream 8 wrote Section~\ref{subsec:multimodal}: with primary authors Alicia Parrish (Google) and Paul Röttger (Bocconi University).  Kenneth Fricklas (Turaco) provided significant input and editing.
\item Heather Frase and Eleonora Presani (Meta) wrote Section 3 and sections 2.1.2 on Hazard categories. 
\item Eleonora Presani, James Noh (AWS), and Peter Mattson also contributed to Section 1.
\item Sean McGregor (UL Research Institutes) wrote Section 10 and produced appendix A and its figures.
\item Wiebke Hutiri (Sony AI) Sarah Luger, Heather Frase, James Ezick wrote Section 6.
\item Shaona Ghosh (NVIDIA), Ryan Tsang (MLCommons), and Kurt Bollacker (MLCommons) wrote the Response Evaluator Section 5.
\item Sarah Luger, Heather Frase, and Mala Kumar wrote Section 7 
\item Sarah Luger and Mala Kumar wrote Section 8.2.
\item Jonathan Bennion prepared the prompt figures in Section 4.3 and Heather Frase wrote the text.
\item Heather Frase worked on figures for the prompts and grading/scoring in Section 6.
\item Quentin Feuillade (Montixi) provided the PRISM Eval content.
\end{enumerate}

\section*{Acknowledgments}
\end{comment}