\section{Conclusion}\label{sec:8-conclusion}
This survey provides the first comprehensive exploration of integrating LLMs into MRS, a topic at the intersection of robotics and artificial intelligence that is rapidly gaining prominence. Unlike general robotics or multi-agent systems, MRS pose unique challenges and opportunities due to their reliance on physical embodiment and real-world interaction. This paper highlights how LLMs can address these challenges, offering novel possibilities for collective intelligence and collaboration in MRS.

We introduced a structured framework for understanding the role of LLMs in MRS, spanning high-level task allocation and planning, mid-level motion planning, low-level action execution, and human intervention. This framework reflects the diverse functionalities enabled by LLMs, including decomposing complex tasks, coordinating multi-robot multi-task scenarios, and facilitating seamless human-robot interaction. Additionally, we reviewed applications of MRS across various domains, from household tasks to construction, formation control, target tracking, and game/competition, demonstrating the versatility and transformative potential of LLMs in these systems.

The significance of integrating LLMs into MRS lies in their ability to augment individual and collective intelligence, enabling robots to operate autonomously and collaboratively in increasingly complex environments. As LLMs continue to demonstrate their potential in everyday applications, their adoption in robotics promises to unlock new possibilities for innovation and efficiency in MRS.

Looking ahead, both immediate and long-term perspectives present exciting opportunities for research and development. In the near term, addressing challenges such as benchmarking, reasoning capabilities, and real-time performance will be critical for bridging the gap between lab-based simulations and real-world applications. Long-term prospects include leveraging LLMs to enable more complex missions, such as disaster response, space exploration, and large-scale autonomous operations, expanding the boundaries of what MRS can achieve.

We hope this survey will be a valuable resource for researchers, providing an overview of current progress, identifying gaps, and highlighting opportunities for future exploration. By advancing our understanding of LLMs in MRS, we aim to inspire innovation and foster collaboration across disciplines, accelerating the transition from theoretical studies to practical deployments that benefit society.

\section{Acknowledgment} 
We thank the authors of~\cite{liu_bolaa_2023, chen_scalable_2024, wang_dart-llm_2024, yu_co-navgpt_2023} for granting permission to include their original, unaltered figures in this paper.

\section{Competing interests}
The authors declare that they have no competing interests. 