\documentclass{ws-jmrr}
\usepackage{caption}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx,adjustbox}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{color}
\usepackage{afterpage}
\usepackage{makecell} %thicker lines I think
\newcommand*{\vcenteredhbox}[1]{\begingroup
\setbox0=\hbox{#1}\parbox{\wd0}{\box0}\endgroup}
\usepackage{colortbl}
\usepackage{float}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{wrapfig,lipsum,booktabs}
\usepackage{subfig}
\usepackage{units}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage{multirow}
%\usepackage[sort,compress,super]{cite}
%\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
%\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
\usepackage{xspace}
\newcommand*{\eg}{\emph{e.g.}\@\xspace}
\newcommand*{\Eg}{\emph{E.g.}\@\xspace}
\newcommand*{\ie}{\emph{i.e.}\@\xspace}
\newcommand*{\etc}{\emph{etc.}\@\xspace}
\newcommand*{\etal}{\emph{et al.}\@\xspace}
\newcommand*{\cf}{\emph{cf.}\@\xspace}
\newcommand*{\vs}{\emph{vs.}\@\xspace}

\usepackage[acronym]{glossaries}
\newacronym{cao}{CAO}{central airway obstruction}
\newacronym{sam}{SAM}{Segment Anything Model}



\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{} % Clear all headers and footers
\fancyhead[R]{\thepage} % Page number at the top right
\fancyhead[L]{} % No section names in the header
\renewcommand{\sectionmark}[1]{} % Prevents section headers from appearing
\renewcommand{\headrulewidth}{0pt} % Removes the horizontal line at the top
\fancypagestyle{plain}{ % Also applies to first page if needed
    \fancyhf{}
    \fancyhead[R]{\thepage}
    \renewcommand{\headrulewidth}{0pt}
}

\fancypagestyle{firstpage}{
    \fancyhf{} % Clear all headers and footers
}

\AtBeginDocument{
    \renewcommand{\thefootnote}{\fnsymbol{footnote}}
    \footnotetext{Preprint of an article accepted for publication in the Journal of Medical Robotics Research, 2025. Copyright World Scientific Publishing Company [https://worldscientific.com/worldscinet/jmrr].}
    \renewcommand{\thefootnote}{\arabic{footnote}} % Reset footnote numbering for later use
}


\begin{document}

\thispagestyle{firstpage} % Apply the custom style to only page 1

% COMMENTS
\definecolor{antiquefuchsia}{rgb}{0.57, 0.36, 0.51}
\definecolor{teal}{rgb}{0.196, 0.659, 0.620}
\definecolor{TEE_blue}{RGB}{0,128,255}
\definecolor{nicepurple}{RGB}{150,50,200}
\newcommand{\MS}[1]{\textcolor{nicepurple}{[MS: #1]}}
\newcommand{\NY}[1]{\textcolor{TEE_blue}{[NY: #1]}}
\newcommand{\TW}[1]{\textcolor{teal}{[TW: #1]}}
\newcommand{\PS}[1]{\textcolor{green}{[PS: #1]}}
\newcommand{\JG}[1]{\textcolor{red}{[JG: #1]}}
\newcommand{\AKZ}[1]{\textcolor{olive}{[Axel: #1]}} 
\newcommand{\AK}[1]{\textcolor{cyan}{[Alan: #1]}}




\title{Autonomous Vision-Guided Resection of Central Airway Obstruction}


\author{Mariana E. Smith$^{a*}$, Nural Yilmaz$^{a*}$, Tanner Watts$^{b}$, Paul Maria Scheikl$^{a}$, Jiawei Ge$^{a}$, Anton Deguet$^{c}$, Alan Kuntz$^{b}$, and Axel Krieger$^{a}$ }

\address{$^{*}$ Both authors contributed equally to this work.}

\address{$^a$Department of Mechanical Engineering, Johns Hopkins University, Baltimore, MD 21211, USA\\
E-mail: msmit458@jhu.edu}

\address{$^b$Robotics Center and Kahlert School of Computing, University of Utah, Salt Lake City, UT 84112, USA}

\address{$^c$Malone Center for Engineering and Healthcare, Johns Hopkins University, Baltimore, MD 21211, USA}

%\thanks{This material is based upon work supported in part by the xxx under grant numbers xxx and xxx. Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the xxx.}

\maketitle

\begin{abstract}


Existing tracheal tumor resection methods often lack the precision required for effective airway clearance, and robotic advancements offer new potential for autonomous resection. We present a vision-guided, autonomous approach for palliative resection of tracheal tumors. This system models the tracheal surface with a fifth-degree polynomial to plan tool trajectories, while a custom Faster R-CNN segmentation pipeline identifies the trachea and tumor boundaries. The electrocautery tool angle is optimized using handheld surgical demonstrations, and trajectories are planned to maintain a 1\,mm safety clearance from the tracheal surface. We validated the workflow successfully in five consecutive experiments on ex-vivo animal tissue models, successfully clearing the airway obstruction without trachea perforation in all cases (with more than 90\% volumetric tumor removal). These results support the feasibility of an autonomous resection platform, paving the way for future developments in minimally-invasive autonomous resection.

\end{abstract}

\keywords{tumor resection, autonomous surgery, vision-based control}

\begin{multicols}{2}
\input{01-Introduction}
\input{02-Methods}
\input{03-Experiments}
\input{04-Discussion}
\input{Acknowledgements}

\bibliographystyle{ws-jmrr}
\bibliography{CAO-References.bib}


\end{multicols}
\end{document}
