\begin{itemize}

\item \textbf{Blackbox AI Agent} \cite{blackbox} is building coding agent to transform the way we build software.

\item \textbf{Gru(2024-12-08)} \cite{gru} builds different agents to solve different software engineering problems. But all Grus are built with the same principles: Clear Problem Domain, Dedicated Tools and Direct Value Delivery.

\item \textbf{Globant Code Fixer Agent} \cite{globant} is an independent and intelligent software entities designed to transform business operations.

\item \textbf{devlo} \cite{devlo} boosts user's productivity by handling development tasks, freeing user to focus on innovation and ship products faster.

\item \textbf{OpenCSG Starship Agentic Coder} \cite{opencsgstarship} is a multi-agent collaborative and scalable environment to empower user in building the next generation of intelligent applications.

\item \textbf{Bytedance MarsCode Agent} \cite{marscode} is a novel framework that leverages LLMs to automatically identify and repair bugs in software code.

\item \textbf{Alibaba Lingma Agent} \cite{repounderstander} understands the whole software repository to achieving automatic software engineering.

\item \textbf{Kodu-v1} \cite{kodu} implements a VS Code extension that adapts to user's skill level, helping user bring ideas to life faster than ever before. 

\item \textbf{OpenHands + CodeAct v2.1} \cite{wang2024openhands} is a platform for the development of powerful and flexible AI agents that interact with the world in similar ways to those of a human developer: by writing code, interacting with a command line, and browsing the web.

\item \textbf{PatchKitty-0.9}: It may have been developed concurrently with our work and is reportedly designed by researchers from UC Santa Barbara. While it was claimed to be open-source in its SWE-bench Lite submission, no repository or related links have been released yet.

\item \textbf{Composio SWE-Kit (2024-10-30)} \cite{composio} helps user connect AI agents to external tools like Gmail, GitHub, Salesforce, etc. It’s like a bridge between user's AI and the tools it needs to get work done.

\item \textbf{Moatless Tools} \cite{moatless} is a hobby project where the authors experiment with some ideas they have about how LLMs can be used to edit code in large existing codebases. They believe that rather than relying on an agent to reason its way to a solution, it is crucial to build good tools to insert the right context into the prompt and handle the response.

\item \textbf{AutoCodeRover-v2.0} \cite{zhang2024autocoderover} is an automated approach for solving Github issues to autonomously achieve
program improvement, where LLMs are combined with sophisticated code search capabilities, ultimately leading to a program modification or patch.

\item \textbf{Agentless-1.5} \cite{xia2024agentless} is an agentless approach to automatically resolve software development issues. Compared to the verbose and complex setup of agent-based approaches, it employs a simplistic three-phase process of localization, repair, and patch validation, without letting the LLM decide future actions or operate with complex tools.

\item \textbf{RepoGraph} \cite{ouyang2024repograph} is a plug-in module that manages a repository-level structure for modern AI software engineering solutions.

\item \textbf{HyperAgent} \cite{phan2024hyperagent} is a novel generalist multi-agent system that addresses a broad spectrum of SE tasks across multiple programming languages by emulating the workflows of human developers.

\item \textbf{SWE-agent} \cite{yang2024swe}: is a system that facilitates LM agents to autonomously use computers to solve software engineering tasks. SWE-agent’s custom agent-computer interface
(ACI) significantly enhances an agent’s ability to create and edit code files, navigate entire repositories, and execute tests and other programs.
\end{itemize}