\section{Related Work}
\subsection{Advanced Reasoning and Instruction Following in IR}
In the last couple years, retrieval systems have started to move beyond simple phrase-based semantic matching, to more complex information retrieval tasks. This has included a focus on new benchmarks: such as reasoning in retrieval ____, instruction-following capabilities ____, and retrieval for RAG systems ____.

On the modeling side, we have seen a surge of interest in models that can understand the meaning behind the user's query rather than doing phrase-based matching: this includes models like Instructor ____, TART ____, GritLM ____, FollowIR ____, Gecko ____, and Promptriever ____. These models typically use instruction-based data in their training data, so that they learn to adapt to new user instructions. However, for \modelnamepretty\ we do not provide any instruction-based training data and fine-tune from the base (non-instruct) version of the LMs -- despite this our model shows SOTA ability in these tasks.

\subsection{Reasoning Language Models}
Reasoning language models were introduced by OpenAI with their o1 model ____. These models showed significantly improved performance on tasks that needed reasoning, such as math, logic, and programming. Since their release, many others have trained similar style models, including Google's Gemini Flash Thinking and Deepseek's R1. Notably, R1 is the only reasoning model that provides reasoning chains through APIs and is the only open-weights model.

Other than the impressive performance gains of these models, one additional feature is that models can quickly learn to emulate stronger models through basic supervised fine-tuning, rather than the more complex reinforcement learning pipelines that are typically used. We take advantage of this capability to train \modelnamepretty\ using a simple training process.

There have also been a flurry of works in the open-source space, both before and after o1 on reasoning language models and systems, focused on reproduction ____, calibration and confidence ____, agentic capabilities with explainable reasoning traces ____, and much more. We expect this line of work to continue and continued collaboration will likely improve these models in retrieval as well.