\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts
\overrideIEEEmargins  

\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{lipsum}
\usepackage{xcolor}
\usepackage{soul}
\usepackage{doi}
\usepackage{caption}
\usepackage{multirow}

\usepackage{hyperref} 
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}
\usepackage[all]{hypcap}

\usepackage{notoccite}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
% updated with editorial comments 8/9/2021



\begin{document}

\title{ARENA: Adaptive Risk-aware and Energy-efficient NAvigation for Multi-Objective 3D Infrastructure Inspection with a UAV}

\author{David-Alexandre Poissant$^{1-2}$, Alexis Lussier Desbiens$^{1}$, François Ferland$^{2}$, and Louis Petit$^{1-2}$% <-this % stops a space
\thanks{This research was funded by Alliance grant number 2601-2600-703 and CRIAQ between Université de Sherbrooke, Hydro-Québec and DRONE VOLT®.}% <-this % stops a space
\thanks{$^{1}$The authors are with the Createk Design Lab, University of Sherbrooke, Sherbrooke, J1K 2R1, Canada, {\tt\small [david-alexandre.poissant, louis.petit, alexis.lussier.desbiens]@usherbrooke.ca} }
\thanks{$^{2}$The authors are with the Intelligent Interactive Integrated Interdisciplinary Robot Lab (IntRoLab), University of Sherbrooke, Sherbrooke, J1K 2R1, Canada, {\tt\small françois.ferland@usherbrooke.ca} }}


% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2021}%
%{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}

%\IEEEpubid{0000--0000/00\$00.00~\copyright~2024 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

\maketitle

\begin{abstract}
Autonomous robotic inspection missions require balancing multiple conflicting objectives while navigating near costly obstacles. Current multi-objective path planning (MOPP) methods struggle to adapt to evolving risks like localization errors, weather, battery state, and communication issues. This letter presents an Adaptive Risk-aware and Energy-efficient NAvigation (ARENA) MOPP approach for UAVs in complex 3D environments. Our method enables online trajectory adaptation by optimizing safety, time, and energy using 4D NURBS representation and a genetic-based algorithm to generate the Pareto front. A novel risk-aware voting algorithm ensures adaptivity. Simulations and real-world tests demonstrate the planner's ability to produce diverse, optimized trajectories covering 95\% or more of the range defined by single-objective benchmarks and its ability to estimate power consumption with a mean error representing 14\% of the full power range. The ARENA framework enhances UAV autonomy and reliability in critical, evolving 3D missions.
\end{abstract}

\begin{keywords}
Motion and Path Planning, Autonomous Vehicle Navigation, Aerial Systems: Applications, Optimization and Optimal Control, Robust/Adaptive Control
\end{keywords}

\input{0_introduction}
\input{1_related_works}
\input{fundamentals}
\input{risk_aware_path_planning}
\input{4_results}
\input{5_conclusion}

\section*{Acknowledgments}
The authors would like to thank Hydro-Québec and DRONE VOLT® for supplying the LineDrone. The authors
also thank Sophie Stratford and Marc-Antoine Leclerc for their help during outdoor flights.

\bibliographystyle{IEEEtran}
\bibliography{main}

\end{document}


\typeout{get arXiv to do 4 passes: Label(s) may have changed. Rerun}