\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{lipsum} 
\usepackage{cite}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}

\usepackage{times}
\usepackage{microtype}
\usepackage{epsfig}
\usepackage[table,xcdraw]{xcolor}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{float}
\usepackage{placeins}
\usepackage{color, colortbl}
\usepackage{stfloats}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{xstring}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{url}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{tcolorbox}
\usepackage[hang,flushmargin]{footmisc}
\usepackage{color}
\usepackage{tikz}
\usepackage{bbding}
\usepackage{makecell}

\usepackage[edges]{forest}
\definecolor{hidden-draw}{RGB}{20,68,106}
\definecolor{hidden-pink}{RGB}{255,245,247}
\newcommand{\etal}{{\emph{et al.}}}
\definecolor{lightgrey}{gray}{0.92}
\definecolor{light\_double\_grey}{gray}{0.95}
\definecolor{lightred}{RGB}{251,49,153}
\definecolor{lightorange}{RGB}{244,230,217}

\definecolor{LightRed}{rgb}{1,0.92,0.92}
\definecolor{LightBlue}{rgb}{0.9,0.94,1}
\definecolor{LightGreen}{rgb}{0.9,1.0,0.88}
\newcommand{\lightgraytext}[1]{\textcolor[rgb]{0.5,0.5,0.5}{#1}}

\newcommand{\paratitle}[1]{\vspace{1.5ex}\noindent\textbf{#1}}
\newcommand{\definedsection}[1]{\noindent\textit{#1}\vspace{1.0ex}}

\newcommand{\thickcline}[1]{%
    \arrayrulecolor{black}\Xcline{#1}{2.5pt}%
    \arrayrulecolor{black}% 
}

\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

\newcommand{\zekun}[1]{\textcolor{blue}{[Zekun: #1]}}
\newcommand{\issue}[1]{\textcolor{red}{[Issue: #1]}}


\begin{document}


\title{Survey on AI-Generated Media Detection: From Non-MLLM to MLLM}

\author{Yueying Zou, Peipei Li*, Zekun Li, Huaibo Huang, Xing Cui, \\Xuannan Liu, Chenghanyu Zhang, Ran He,~\IEEEmembership{Fellow,~IEEE}
        % <-this % stops a space  
\thanks{Yueying Zou, Peipei Li, Xing Cui, and Xuannan Liu are with the School of Artificial Intelligence, and Chenghanyu Zhang is with the School of Science, all at Beijing University of Posts and Telecommunications, Beijing 100876, China. E-mail: {zouyueying2001, lipeipei, cuixing, liuxuannan, zhangchenghanyu}@bupt.edu.cn.}
\thanks{Zekun Li is with the School of Computer Science, University of California, Santa Barbara, USA. E-mail: zekunli@cs.ucsb.edu.}
\thanks{Huaibo Huang and Ran He are with the State Key Laboratory of Multimodal Artificial Intelligence Systems, CASIA, New Laboratory of Pattern Recognition, CASIA, and School of Artificial Intelligence, University of Chinese Academy of Sciences, Beijing 100190, China. E-mail: {huaibo.huang, rhe}@cripac.ia.ac.cn.}
\thanks{Peipei Li* is the corresponding author. E-mail: lipeipei@bupt.edu.cn.}
}
%\thanks{Huaibo Huang and Ran He are with the State Key Laboratory of Multimodal Artificial Intelligence Systems, CASIA, New Laboratory of Pattern Recognition, CASIA, and School of Artificial Intelligence, University of Chinese Academy of Sciences, Beijing 100190, China. E-mail: {huaibo.huang, rhe}@cripac.ia.ac.cn.}

% <-this % stops a space

%\thanks{Manuscript received April 19, 2021; revised August 16, 2021.}

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2021}%
{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}

%\IEEEpubid{0000--0000/00\$00.00~\copyright~2021 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

\maketitle

\begin{abstract}
The proliferation of AI-generated media poses significant challenges to information authenticity and social trust, making reliable detection methods highly demanded. Methods for detecting AI-generated media have evolved rapidly, paralleling the advancement of Multimodal Large Language Models (MLLMs). Current detection approaches can be categorized into two main groups: Non-MLLM-based and MLLM-based methods. The former employs high-precision, domain-specific detectors powered by deep learning techniques, while the latter utilizes general-purpose detectors based on MLLMs that integrate authenticity verification, explainability, and localization capabilities. Despite significant progress in this field, there remains a gap in literature regarding a comprehensive survey that examines the transition from domain-specific to general-purpose detection methods. This paper addresses this gap by providing a systematic review of both approaches, analyzing them from single-modal and multi-modal perspectives. We present a detailed comparative analysis of these categories, examining their methodological similarities and differences. Through this analysis, we explore potential hybrid approaches and identify key challenges in forgery detection, providing direction for future research. Additionally, as MLLMs become increasingly prevalent in detection tasks, ethical and security considerations have emerged as critical global concerns. We examine the regulatory landscape surrounding Generative AI (GenAI) across various jurisdictions, offering valuable insights for researchers and practitioners in this field.

\end{abstract}

\begin{IEEEkeywords}
AI-generated Media detection, MLLM, deep learning, literarture survey
%\LaTeX, paper, template, typesetting.
\end{IEEEkeywords}

\input{section/intro}
\input{section/back}
\input{section/mllm}
\input{section/non-mllm}
\input{section/eval}
\input{section/regulation}
\input{section/future}
\input{section/concl}




%\vfill
\bibliographystyle{unsrt}
\bibliography{mainbib}
\end{document}
\renewcommand\refnamesec{参考文献} 
% \bibliographystylesec{plain}
\bibliographystylesec{unsrt}
\bibliographysec{引用}

%\newpage

% \section{Biography Section}
% If you have an EPS/PDF photo (graphicx package needed), extra braces are
%  needed around the contents of the optional argument to biography to prevent
%  the LaTeX parser from getting confused when it sees the complicated
% \begin{IEEEbiographynophoto}{John Doe}
% Use $\backslash${\tt{begin\{IEEEbiographynophoto\}}} and the author name as the argument followed by the biography text.
% \end{IEEEbiographynophoto}
% \end{document}



