% \documentclass[conference]{IEEEtran}
\documentclass[letterpaper, 10 pt, conference]{ieeeconf}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage[letterpaper, top=57pt, left=52pt, right=52pt, bottom=57pt]{geometry}
% \usepackage[noproof]{amsthm}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{dsfont}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\input{macro}
    
\begin{document}

\title{Safety Verification of Stochastic Systems under Signal Temporal Logic Specifications
}

\author{Liqian Ma, Zishun Liu, Hongzhe Yu, and Yongxin Chen% <-this % stops a space
\thanks{The authors are with Georgia Institute of Technology, Atlanta, GA 30332 
        {\tt\small \{mlq\}\{zliu910\}\{hyu419\}\{yongchen\}@gatech.edu}}%
}


% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% }

\maketitle

\begin{abstract}
We study the verification problem of stochastic systems under signal temporal logic (STL) specifications. We propose a novel approach that enables the verification of the probabilistic satisfaction of STL specifications for nonlinear systems subject to both bounded deterministic disturbances and stochastic disturbances. Our method, referred to as the STL erosion strategy, reduces the probabilistic verification problem into a deterministic verification problem with a tighter STL specification. The degree of tightening is determined by leveraging recent results on bounding the deviation between the stochastic trajectory and the deterministic trajectory. Our approach can be seamlessly integrated with any existing deterministic STL verification algorithm. Numerical experiments are conducted to showcase the efficacy of our method.
\end{abstract}

% \begin{IEEEkeywords}

% \end{IEEEkeywords}

\input{sec_introduction}    
\input{sec_problem_formulation}
\input{sec_method}
\input{sec_experiments}

\section{Concluding remarks}
We propose an STL erosion strategy for probabilistic STL verification of discrete-time nonlinear stochastic systems under sub-Gaussian disturbances. Leveraging a tight bound on stochastic deviation, our strategy reduces the probabilistic verification problem into a deterministic verification problem with a tightened STL specification which can then be solved using existing deterministic methods. Our strategy is validated on both linear and nonlinear systems. Control synthesis for nonlinear stochastic systems with probabilistic STL specifications based on our results is an interesting and promising future direction.

\bibliographystyle{IEEEtran}        
\bibliography{references}

\end{document}
