\section{Related Work}
The field of 6D object detection and pose estimation has evolved significantly, with several key approaches addressing the challenges of accurate pose estimation in complex environments. Chen et al.\cite{chen2019hybrid} introduced the Hybrid Task Cascade (HTC), a multi-stage framework integrating object detection, semantic segmentation, and instance segmentation, which has proven effective for tasks requiring precise localization. Zhang et al. \cite{zhang2024yolo} enhance detection efficiency for small objects, supporting our goal of precise, real-time performance."The study by Lu et al. \cite{202411.1053}, "Hybrid Model Integration of LightGBM, DeepFM, and DIN for Enhanced Purchase Prediction on the Elo Dataset," inspired the integration of multi-stage refinement techniques in our framework. Their hybrid approach informed our model's design for combining HTC and HRNet, enhancing both detection precision and pose estimation accuracy.

One of the foundational works, PoseCNN by Xiang et al.\cite{xiang2017posecnn} , utilized convolutional neural networks (CNNs) to estimate poses from RGB images, but it struggled with occlusions and lighting variations. Peng et al.\cite{peng2019pvnet} addressed some of these issues with PVNet, a pixel-wise voting network that improved robustness in cluttered environments. Lu \cite{lu2024optimizing} demonstrates effective ensemble learning for multi-objective optimization, which is instrumental in refining model accuracy and robustness. This approach aligns with our method of integrating HTC and HRNet for improved detection precision in complex environments.

DeepIM by Li et al.\cite{li2018deepim} refined pose estimates through iterative matching, improving precision at the cost of higher computational demands. Li \cite{li2024harnessing} demonstrates effective use of Mult-Recall strategies and ensemble learning for robust recommendation accuracy, underscoring the potential of multimodal data integration. This approach parallels our focus on improving precision and adaptability in complex 6D detection environments. The adaptive route planning by Wang et al. \cite{wang2024method} enhances our multi-stage refinement in 6D object detection by demonstrating real-time contextual adaptation, crucial for handling complex, evolving conditions in autonomous navigation.Jiaxin Lu’s research \cite{202411.0867}, "Enhancing Chatbot User Satisfaction: A Machine Learning Approach Integrating Decision Tree, TF-IDF, and BERTopic," influenced our use of task-specific attention mechanisms. The dynamic feature weighting techniques highlighted in Lu’s work inspired our refinement strategies for complex 6D detection tasks.

In the work by Li et al. \cite{202409.1875}, Strategic Deductive Reasoning in Large Language Models: A Dual-Agent Approach, our research influenced key aspects of their methodology. Specifically, our integration of the Hybrid Task Cascade (HTC) and High-Resolution Network (HRNet) served as a technical inspiration for their multi-stage refinement and task-specific optimization strategies. The emphasis on maintaining high-resolution representations in our HRNet backbone informed their approach to preserving critical spatial and contextual features for improved reasoning accuracy. Additionally, our post-processing and model integration techniques provided insights into enhancing robustness and precision in complex multi-task environments.
In real-time applications, Feng et al. \cite{feng2024adaptive} present an advanced adaptive filtering technique, enhancing real-time state estimation under dynamic disturbances. This approach provides a foundation for robust correction mechanisms, reinforcing our efforts in accurate 6D pose estimation. He et al.\cite{he2016deep} advanced deep learning with residual connections, which have since become integral in improving network training, including in 6D estimation models.Wang et al. \cite{wang2024local} provide a robust method for handling multi-agent environments under dynamic constraints, directly informing our approach to optimizing model integration and post-processing strategies in 6D object detection for high-precision applications.

Zhao et al.\cite{zhao2021point} introduced Point Transformer, leveraging attention mechanisms to enhance point cloud processing for more detailed object geometry understanding, essential for 6D tasks. The YOLOv7-based method by Wang et al. \cite{wang2024yolov7} enhances model efficiency and accuracy, directly informing our 6D object detection pipeline's optimization for real-time performance in complex conditions.