
\bibitem{abraham2013scuba}
Abraham L, Allen J, Barykin O, Borkar V, Chopra B, Gerea C, Merl D, Metzler J, Reiss D, Subramanian S, Wiener JL.: Scuba: diving into data at facebook. Proceedings of the VLDB Endowment. \textbf{6}(11), 1057--1067 (2013)

%\bibitem{baulier1998datablitz}
%Baulier J, Bohannon P, Gogate S, Joshi S, Gupta C, Khivesera A, Korth HF, McIlroy P, Miller J, Narayan PP, Nemeth M. DataBlitz: A High Performance Main-Memory Storage Manager. In: Proceedings of the 24th International Conference on Very Large Data Bases. 1998, p. 701

\bibitem{barber2011blink}
Barber R, Bendel P, Czech M, Draese O, Ho F, Hrle N, Idreos S, Kim MS, Koeth O, Lee JG, Li TT, Lohman G, Morfonios , Mueller R, Murthy K, Pandis I, Qiao L, Raman V, Szabo S, Sidle R, Stolze K.: Blink: Not Your Father\textquotesingle s Database!. In: Proceedings of International Workshop on Business Intelligence for the Real-Time Enterprise, pp. 1--22 (2011)


\bibitem{farber2012sap}
F\"{a}rber F, Cha SK, Primsch J, Bornh\"{o}vd C, Sigg S, Lehner W.: SAP HANA database: data management for modern business applications. ACM Sigmod Record.  \textbf{40}(4), 45--51 (2012)

\bibitem{lindstrom2013ibm}
Lindstr\"{o}m J, Raatikka V, Ruuth J, Soini P, Vakkila K.: IBM solidDB: In-Memory Database Optimized for Extreme Speed and Availability. IEEE Data Eng. Bull. \textbf{36}(2), 14--20 (2013)

\bibitem{PeletonLink}
Peloton Database Management System. Published at http://pelotondb.org (2019)

\bibitem{pavlo2017self}
Pavlo A, Angulo G, Arulraj J, Lin H, Lin J, Ma L, Menon P, Mowry TC, Perron M, Quah I, Santurkar S, Tomasic A, Toor S, Aken D.V, Wang Z, Wu Y, Xian R, Zhang T.: Self-Driving Database Management Systems. In: Proceedings of Biennial Conference on Innovative Data Systems Research (CIDR) (2017)

\bibitem{larson2011sql}
Larson P\r{A}, Clinciu C, Hanson EN, Oks A, Price SL, Rangarajan S, Surna A, Zhou Q.: SQL server column store indexes. In: Proceedings of the ACM SIGMOD International Conference on Management of Data, pp. 1177--1184 (2011)

%\bibitem{lehman1986study}
%Lehman TJ, Carey MJ. A study of index structures for main memory database management systems. In: Proceedings of the 12th International Conference on Very Large Data Bases. 1986, 294--303

\bibitem{zhang2016reducing}
Zhang H, Andersen DG, Pavlo A, Kaminsky M, Ma L, Shen R.: Reducing the storage overhead of main-memory OLTP databases with hybrid indexes. In: Proceedings of the International Conference on Management of Data, pp. 1567--1581 (2016)

\bibitem{ongaro2011fast}
Ongaro D, Rumble SM, Stutsman R, Ousterhout J, Rosenblum M.: Fast crash recovery in RAMCloud. In: Proceedings of the Twenty-Third ACM Symposium on Operating Systems Principles (SOSP), pp. 29--41 (2011)

%\bibitem{lehman1987recovery}
%Lehman TJ, Carey MJ. A recovery algorithm for a high-performance memory-resident database system. In: Proceedings of the 1987 ACM SIGMOD international conference on Management of data. 1987, 104--117

\bibitem{diaconu2013hekaton}
Diaconu C, Freedman C, Ismert E, Larson PA, Mittal P, Stonecipher R, Verma N, Zwilling M.: Hekaton: SQL server's memory-optimized OLTP engine. In: Proceedings of the ACM SIGMOD International Conference on Management of Data, pp. 1243--1254 (2013)

\bibitem{lee2001single}
Lee I, Yeom HY.: A single phase distributed commit protocol for main memory database systems. In: Proceedings of Parallel and Distributed Processing Symposium, (2002)

\bibitem{debrabant2014prolegomenon}
DeBrabant J, Arulraj J, Pavlo A, Stonebraker M, Zdonik S, Dulloor S.: A prolegomenon on OLTP database systems for non-volatile memory. In International Workshop on Accelerating Data Management Systems Using Modern Processor and Storage Architectures. (2014)

\bibitem{driskill2010latest}
Driskill-Smith A.: Latest advances and future prospects of STT-RAM. In Non-Volatile Memories Workshop. 11--13, (2010)

\bibitem{mandelman2002challenges}
Mandelman JA, Dennard RH, Bronner GB, DeBrosse JK, Divakaruni R, Li Y, Radens CJ.: Challenges and future directions for the scaling of dynamic random-access memory (DRAM). IBM Journal of Research and Development. \textbf{46}(2.3), 187--212 (2002)

\bibitem{melnik2010dremel}
Melnik S, Gubarev A, Long JJ, Romer G, Shivakumar S, Tolton M, Vassilakis T.: Dremel: interactive analysis of web-scale datasets. Proceedings of the VLDB Endowment.  \textbf{3}(1--2) 330--339 (2010)

\bibitem{plattner2011sanssoucidb}
Plattner H.: SanssouciDB: An In-Memory Database for Processing Enterprise Workloads. In: Proceedings of Datenbanksysteme f\"{u}r Business, Technologie und Web (BTW), pp. 2--21 (2011)

\bibitem{sikka2012efficient}
Sikka V, F\"{a}rber F, Lehner W, Cha SK, Peh T, Bornh\"{o}vd C.: Efficient transaction processing in SAP HANA database: the end of a column store myth. In: Proceedings of the 2012 ACM SIGMOD International Conference on Management of Data, pp. 731--742 (2012)

\bibitem{arulraj2017build}
Arulraj J, Pavlo A.: How to Build a Non-Volatile Memory Database Management System. In: Proceedings of the 2017 ACM International Conference on Management of Data, pp. 1753--1758 (2017)

\bibitem{qureshi2009scalable}
Qureshi MK, Srinivasan V, Rivers JA.: Scalable high performance main memory system using phase-change memory technology. In: Proceedings of the 36th annual international symposium on Computer architecture (ISCA), pp. 24--33 (2009)

\bibitem{andrei2017sap}
Andrei M, Lemke C, Radestock G, Schulze R, Thiel C, Blanco R, Meghlan A, Sharique M, Seifert S, Vishnoi S, Booss D, Peh T, Schreter I, Thesing W, Wagle  M, Willhalm T.: SAP HANA adoption of non-volatile memory. Proceedings of the VLDB Endowment. \textbf{10}(12), 1754--1765 (2017)

\bibitem{raoux2008phase}
Raoux S, Burr GW, Breitwisch MJ, Rettner CT, Chen YC, Shelby RM, Salinga M, Krebs D, Chen SH, Lung HL, Lam CH.: Phase-change random access memory: A scalable technology. IBM Journal of Research and Development. \textbf{52}(4.5), 465--479 (2008)

\bibitem{strukov2008missing}
Strukov DB, Snider GS, Stewart DR, Williams RS.: The missing memristor found. nature. \textbf{453}(7191), 80--83 (2008)


\bibitem{arulraj2015let}
Arulraj J, Pavlo A, Dulloor SR.: Let's talk about storage \& recovery methods for non-volatile memory database systems. In: Proceedings of the 2015 ACM SIGMOD International Conference on Management of Data, pp. 707--722 (2015)

\bibitem{chang2012limits}
Chang J, Ranganathan P, Mudge T, Roberts D, Shah MA, Lim KT.: A limits study of benefits from nanostore-based future data-centric system architectures. In: Proceedings of the 9th conference on Computing Frontiers (CF), pp. 33--42 (2012)

\bibitem{hagmann2002real}
Hagmann R, Skeen MD.: Real-time query optimization in a decision support system. United States patent US 6,338,055. (2002)

\bibitem{shim2002past}
Shim JP, Warkentin M, Courtney JF, Power DJ, Sharda R, Carlsson C.: Past, present, and future of decision support technology. Decision support systems. \textbf{33}(2), 111--126 (2002)

\bibitem{chaudhuri2001database}
Chaudhuri S, Dayal U, Ganti V.: Database technology for decision support systems. Computer. \textbf{34}(12), 48--55 (2001)

\bibitem{DBRanking}
DB-Engines Ranking of Relational DBMS. Published at https://db-engines.com/en/ranking (2019)

\bibitem{council2008tpc}
Transaction Processing Performance Council. TPC-H benchmark specification. Published at http://www.tpc.org/tpch/ (2008)

\bibitem{perez2010non}
Perez T, De Rose CA.: Non-volatile memory: Emerging technologies and their impacts on memory systems. Technical Report, Porto Alegre (2010)

\bibitem{wang2013low}
Wang KL, Alzate JG, Amiri PK.: Low-power non-volatile spintronic memory: STT-RAM and beyond. Journal of Physics D: Applied Physics. \textbf{46}(7), 074003 (2013)


\bibitem{mittal2016survey}
Mittal S, Vetter JS.: A survey of software techniques for using non-volatile memories for storage and main memory systems. IEEE Transactions on Parallel and Distributed Systems. \textbf{27}(5), 1537--1550 (2016)

\bibitem{arulraj2016write}
Arulraj J, Perron M, Pavlo A.: Write-behind logging. Proceedings of the VLDB Endowment. \textbf{10}(7), 337--348 (2016)

\bibitem{oukid2014sofort}
Oukid I, Booss D, Lehner W, Bumbulis P, Willhalm T.: SOFORT: A hybrid SCM-DRAM storage engine for fast data recovery. In: Proceedings of the Tenth International Workshop on Data Management on New Hardware (DaMoN), (2014)

\bibitem{chatzistergiou2015rewind}
Chatzistergiou A, Cintra M, Viglas SD.: Rewind: Recovery write-ahead system for in-memory non-volatile data-structures. Proceedings of the VLDB Endowment. \textbf{8}(5), 497--508 (2015)

\bibitem{huang2012register}
Huang Y, Liu T, Xue CJ.: Register allocation for write activity minimization on non-volatile main memory for embedded systems. Journal of Systems Architecture. \textbf{58}(1), 13--23 (2012)

\bibitem{oukid2015instant}
Oukid I, Lehner W, Kissinger T, Willhalm T, Bumbulis P.: Instant Recovery for Main Memory Databases. In: Proceedings of Biennial Conference on Innovative Data Systems Research (CIDR), (2015)

\bibitem{chakrabarti2014atlas}
Chakrabarti DR, Boehm HJ, Bhandari K.: Atlas: Leveraging locks for non-volatile memory consistency. In: Proceedings of ACM International Conference on Object Oriented Programming Systems Languages \& Applications (OOPSLA), pp. 433--452 (2014)

\bibitem{zhang2015study}
Zhang Y, Swanson S.: A study of application performance with non-volatile main memory. In: Proceedings of 31st IEEE Symposium on Mass Storage Systems and Technologies (MSST), pp. 1--10 (2015)

\bibitem{viglas2014write}
Viglas SD.: Write-limited sorts and joins for persistent memory. Proceedings of the VLDB Endowment. \textbf{7}(5), 413--424 (2014)

\bibitem{zhou2009durable}
Zhou P, Zhao B, Yang J, Zhang Y.: A durable and energy efficient main memory using phase change memory technology. In: Proceedings of the 36th annual international symposium on Computer architecture (ISCA), pp. 14--23 (2009)

\bibitem{3DXPoint}
Intel and Micron. 3D XPoint Technology. https://www.micron.com/products/advanced-solutions/3d-xpoint-technology (2019)

\bibitem{intel2016architecture}
Intel. 2016. Architecture Instruction Set Extensions Programming Reference. Intel Corporation. (2016)

\bibitem{dulloor2014system}
Dulloor SR, Kumar S, Keshavamurthy A, Lantz P, Reddy D, Sankaran R, Jackson J.: System software for persistent memory. In: Proceedings of the Ninth European Conference on Computer Systems, p. 15 (2014)

\bibitem{githubPMFS}
Intel. 2014. Linux-pmfs. https://github.com/linux-pmfs/pmfs (2014)

\bibitem{kimura2015foedus}
Kimura H.: FOEDUS: OLTP Engine for a Thousand Cores and NVRAM. In: Proceedings of the ACM SIGMOD International Conference on Management of Data, pp. 691--706 (2015)

\bibitem{huang2014nvram}
Huang J, Schwan K, Qureshi MK.: NVRAM-aware logging in transaction systems. Proceedings of the VLDB Endowment. \textbf{8}(4), 389--400 (2014)

\bibitem{schindler2002track}
Schindler J, Griffin JL, Lumb CR, Ganger GR.: Track-Aligned Extents: Matching Access Patterns to Disk Drive Characteristics. In: Proceedings of the 1st USENIX Conference on File and Storage Technologies (FAST), pp. 259--274 (2002)

\bibitem{debrabant2013anti}
DeBrabant J, Pavlo A, Tu S, Stonebraker M, Zdonik S.: Anti-caching: A new approach to database management system architecture. Proceedings of the VLDB Endowment. \textbf{6}(14), 1942--1953 (2013)

%\bibitem{garcia1992main}
%Garcia-Molina H, Salem K. Main memory database systems: An overview. IEEE Transactions on knowledge and data engineering, 1992, 4(6): 509--16

\bibitem{pelley2014memory}
Pelley S, Chen PM, Wenisch TF.: Memory persistency. In: Proceedings of ACM/IEEE 41st International Symposium on Computer Architecture (ISCA), pp. 265--276 (2014)

\bibitem{momjian2001postgresql}
Momjian B.: PostgreSQL: introduction and concepts. New York: Addison-Wesley (2001)

\bibitem{neumann2015fast}
Neumann T, M\"{u}hlbauer T, Kemper A.: Fast serializable multi-version concurrency control for main-memory database systems. In: Proceedings of the ACM SIGMOD International Conference on Management of Data, pp. 677--689 (2015)

\bibitem{PostgreSQLDocBook}
The PostgreSQL Global Development Group. PostgreSQL 9.0.22 Documentation. (2015)

\bibitem{gao2011pcmlogging}
Gao S, Xu J, He B, Choi B, Hu H.: PCMLogging: reducing transaction logging overhead with PCM. In: Proceedings of the 20th ACM international conference on Information and knowledge management, pp. 2401--2404 (2011)

\bibitem{son2017log}
Son Y, Kang H, Yeom HY, Han H.: A log-structured buffer for database systems using non-volatile memory. In: Proceedings of the Symposium on Applied Computing (SAC), pp. 880--886 (2017)

\bibitem{annavaram2001data}
Annavaram M, Patel JM, Davidson ES. Data prefetching by dependence graph precomputation. In: Procedings of 28th Annual International Symposium on Computer Architecture (ISCA), pp. 52--61 (2001)

\bibitem{jung2006helper}
Jung C, Lim D, Lee J, Solihin Y.: Helper thread prefetching for loosely-coupled multiprocessor systems. In: Proceedings of 20th International Parallel and Distributed Processing Symposium, (2006)

\bibitem{pelley2013storage}
Pelley S, Wenisch TF, Gold BT, Bridge B.: Storage management in the NVRAM era. Proceedings of the VLDB Endowment. \textbf{7}(12), 121--132 (2013)

\bibitem{johnson2009shore}
Johnson R, Pandis I, Hardavellas N, Ailamaki A, Falsafi B.: Shore-MT: a scalable storage manager for the multicore era. In: Proceedings of the 12th International Conference on Extending Database Technology: Advances in Database Technology (EDBT), pp. 24--35 (2009)

\bibitem{fang2011high}
Fang R, Hsiao HI, He B, Mohan C, Wang Y.: High performance database logging using storage class memory. In: Proceedings of IEEE 27th International Conference on Data Engineering, pp. 1221--1231 (2011)

\bibitem{wang2014scalable}
Wang T, Johnson R.: Scalable logging through emerging non-volatile memory. Proceedings of the VLDB Endowment. \textbf{7}(10), 865--876 (2014)

\bibitem{RankingLink}
DB-Engines Ranking - popularity ranking of database management systems. Published at https://db-engines.com/en/ranking (2019)

\bibitem{kamruzzaman2011inter}
Kamruzzaman M, Swanson S, Tullsen DM.: Inter-core prefetching for multicore processors using migrating helper threads. In: Proceedings of the sixteenth international conference on Architectural support for programming languages and operating systems (ASPLOS), pp. 393--404 (2011)

\bibitem{kim2002design}
Kim D, Yeung D.: Design and evaluation of compiler algorithms for pre-execution. In: Proceedings of the 10th international conference on Architectural support for programming languages and operating systems (ASPLOS), pp. 159--170  (2002)
%\vspace*{-1mm}

\bibitem{collins2001dynamic}
Collins JD, Tullsen DM, Wang H, Shen JP.: Dynamic speculative precomputation. In: Proceedings of the 34th annual ACM/IEEE international symposium on Microarchitecture (MICRO), pp. 306--317 (2001)

\bibitem{hirofuchi2019preliminary}
Hirofuchi T, Takano R.: The Preliminary Evaluation of a Hypervisor-based Virtualization Mechanism for Intel Optane DC Persistent Memory Module. arXiv preprint arXiv:1907.12014, (2019)

\bibitem{peng2019system}
Peng IB, Gokhale MB, Green EW.: System evaluation of the Intel optane byte-addressable NVM. In: Proceedings of the International Symposium on Memory Systems (MEMSYS), pp. 304--315 (2019)

\bibitem{yang2019empirical}
Yang J, Kim J, Hoseinzadeh M, Izraelevitz J, Swanson S.: An Empirical Guide to the Behavior and Use of Scalable Persistent Memory. arXiv preprint arXiv:1908.03583, (2019)

\bibitem{izraelevitz2019basic}
Izraelevitz J, Yang J, Zhang L, Kim J, Liu X, Memaripour A, Soh YJ, Wang Z, Xu Y, Dulloor SR, Zhao J.: Basic performance measurements of the intel optane DC persistent memory module. arXiv preprint arXiv:1903.05714, (2019)

\bibitem{xu2016nova}
Xu J, Swanson S.: NOVA: A Log-structured File System for Hybrid Volatile/Non-volatile Main Memories. In 14th Conference on File and Storage Technologies (FAST), pp. 323--338, (2016)

\bibitem{PMDKLib}
pmem.io: PMDK. Published at https://pmem.io/pmdk/ (2019)