\pdfoutput=1
\documentclass[journal]{IEEEtran}
%\documentclass[]{report}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{fontawesome5}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}

\input{prefix}

\begin{document}

\title{Some Insights of Construction of Feature Graph to Learn Pairwise Feature Interactions with Graph Neural Networks}

% \author{Phaphontee Yamchote, Saw Nay Htet Win, Chainarong Amornbunchornvej,~\IEEEmembership{Member,~IEEE},  Thanapon Noraset
%         % <-this % stops a space
%         }
\author{Phaphontee Yamchote, Saw Nay Htet Win, Chainarong Amornbunchornvej,  Thanapon Noraset
        % <-this % stops a space
        }

%\author{Thanapon Noraset}
%
%
%\author{IEEE Publication Technology,~\IEEEmembership{Staff,~IEEE,}
%        % <-this % stops a space
%}

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2021}%
{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}

%\IEEEpubid{0000--0000/00\$00.00~\copyright~2021 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.
%\setcounter{page}{0}
%\begin{figure*}[t]
%	\tableofcontents
%\end{figure*}

\maketitle


\begin{abstract}
	Feature interaction is crucial in predictive machine learning models, as it captures the relationships between features that influence model performance. In this work, we focus on pairwise interactions and investigate their importance in constructing feature graphs for Graph Neural Networks (GNNs). Rather than proposing new methods, we leverage existing GNN models and tools to explore the relationship between feature graph structures and their effectiveness in modeling interactions. Through experiments on synthesized datasets, we uncover that edges between interacting features are important for enabling GNNs to model feature interactions effectively. We also observe that including non-interaction edges can act as noise, degrading model performance. Furthermore, we provide theoretical support for sparse feature graph selection using the Minimum Description Length (MDL) principle. We prove that feature graphs retaining only necessary interaction edges yield a more efficient and interpretable representation than complete graphs, aligning with Occam's Razor.
	Our findings offer both theoretical insights and practical guidelines for designing feature graphs that improve the performance and interpretability of GNN models.
\end{abstract}

\begin{IEEEkeywords}
Feature interaction, Graph neural network
\end{IEEEkeywords}

\input{contentFile.tex}

\end{document}


